### 2018-05-23 figures.tex

\documentclass[draft,linenumbers]{agujournal}
\draftfalse
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
\begin{document}

 \begin{figure}[h]
     \centering
     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
     \caption{Response to Nitrogen Deposition}
     \label{Cax CO2 response}
  \end{figure}

 \begin{figure}[h]
     \centering
     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI_y1.png}
     \caption{Response to Nitrogen Deposition}
     \label{ORNL CO2 response}
  \end{figure}

 \begin{figure}[h]
     \centering
     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI_y1.png}
     \caption{Response to Nitrogen Deposition}
     \label{BCI CO2 response}
  \end{figure}

 \begin{figure}[h]
     \centering
     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
     \caption{Response to Nitrogen Deposition}
     \label{CAX state}
  \end{figure}

 \begin{figure}[h]
     \centering
     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
     \caption{Response to Nitrogen Deposition}
     \label{Ndep}
  \end{figure}

\end{document}

### 2018-05-23 matlab/CLM5_FACE_senstest_output_ELEV_apr18.m
added five altered scripts to get up to speed with master

@@ -32,6 +32,7 @@ params  = {'slatop','froot:leaf','stem:leaf','ncosts',...
 root = 'CLM5_1x1pt_Br-cax_ens_PI'
 ending = '_ens_transient_ELEV_'
 sitenames = {'1x1pt_Br-cax' 'bci_0.1x0.1_v4.0i' '1x1pt_US-ORN'}
+figsitenames = {'CAX','BCI','ORN'}
 
 ichoose = 1:4  % loop roun iterations for each parameter
 vchoose = [1 2 3 5 13]; % variables.
@@ -72,6 +73,7 @@ set(0,'DefaultAxesColorOrder',colord)
 
 for s=schoose
   root = strcat('CLM5_',char(sitenames(s)),ending,'PI')
+  figroot = strcat('CLM5_',char(figsitenames(s)),'_')
   clear('var_arrayl','var_array2');
   for pp =[5]
 
@@ -159,9 +161,10 @@ if(v==13);var_arrayl(p,(i),v) = var_arrayl(p,(i),v)./var_arrayl(p,(i),3);end
     end
     l=legend(params(pchoose))
     set(l,'position', [0.7551    0.2585    0.1393    0.0729])
+    set(l,'fontsize',7)
     if(pr==1)
       wysiwyg
-      fnm = strcat(figdir,'/frac_deviation_p',char(num2str(pchoose(1))),root,'_y',char(num2str((pchoose(1)))),'.png') 
+      fnm = strcat(figdir,'/frac_deviation_p',char(num2str(pchoose(1))),figroot,'_y',char(num2str((pchoose(1)))),'.png') 
       print('-dpng', '-r100', char(fnm) )
     end
     
@@ -252,10 +255,11 @@ if(v==13);var_arrayl(p,(i),v) = var_arrayl(p,(i),v)./var_arrayl(p,(i),3);end
        end
 
        l=legend(params(pchoose))
+       set(l,'fontsize',7)
        set(l,'position', [0.7551    0.2585    0.1393    0.0729])
        if(pr==1)
 	 wysiwyg
-	 fnm = strcat(figdir,'/frac_deviation_CO2_response_',char(num2str(pchoose(1))),root,'_y',char(num2str((pchoose(1)))),'.png') 
+	 fnm = strcat(figdir,'/frac_deviation_CO2_response_',char(num2str(pchoose(1))),figroot,'_y',char(num2str((pchoose(1)))),'.png') 
 	 print('-dpng', '-r100', char(fnm) )
        end
        

### 2018-05-23 Workflow.tex

\documentclass{article}
\usepackage{hyperref}
\usepackage[dvipsnames]{color}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}

\begin{document}
\title{Workflow for CLM5 paper}
\maketitle{}

Scripts for these analyses should all be found in the public github repo at:
{\color{blue}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS}}}\\

The scripts are specific to usage on the NCAR Cheyenne machine...

{\color{blue} Scripts, on github are in blue}\\
{\color{cyan} Paths on scratch are in cyan}\\
{\color{red}  .nc files are in red}\\
{\color{magenta} Figures on github are in magenta}\\

\section{Obtain CLM5 code}
cd /glade/p/work/rfisher/git/\\
(this is 'gitpath' from now on)

git clone https://github.com/ESCOMP/ctsm 
\section{Parameters}

\subsection{Make new ensemble param files with high leaf allocation }

Use PFT script

{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/FACE_clm5_ens_oaat.m}}}\\ 
commit b6ba94

Using the basefile. \\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/clm5_params.c171017_highleaf.nc}}} \\
commit c539e89

Output location..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi11.nc }}} \\
commit 54f863b

Copy default file into '00'..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi00.nc }}} \\
commit 61517c5

\section{Spin-up}
\subsection{Make 1850 spin-up default runs from scratch}
Make single point case from scratch for each site by running this script (based on Charlie's FATES version) three times once for each site (that could be improved on!)..:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/fromscratch_CTSM_make_site_spinup_case_defaults}}}\\
commit 38ced1

The resulting cases are \\
{\color{cyan}{\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.1x1pt_US-ORN}}}\\
\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.bci_0.1x0.1_v4.0i}}\\
\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.1x1pt_Br-cax}}}\\

These are all run out to year 321.

\subsection{Make 1850 spin-up ensembles for each site}

Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
commit 2bb89c5

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI**}}\\
\small{\url{CLM5_1x1pt_Br-cax_ens_PI**}}\\
\small{\url{CLM5_1x1pt_US-ORN_ens_PI*}}}\\

These are all run to y201.

\subsection{Make 1850 spin-up defaults for each site}

Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_defaults}}}\\
commit 0d2894a

Case names are:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_PI00}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI00}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_PI00}}}

\subsection{Analyse output of spinup cases}

Run analysis script on spinup ensemble output (at 1850 CO2). \\
{\color{blue}\small{\url{/glade/p/work/rfisher/git/CLM5_FACE_EXPTS/matlab/CLM5_FACE_senstest_output_mar18.m}}}\\
commit 7043e8b

Output here\\
{\color{magenta}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS/tree/master/matlab/figures\frac_deviation_p1CLM5_1x1pt_Br-cax_ens_PI_y1.png}}}\\
commit 1f9a57c

\section{Transients}

\subsection{Make transient defaults}
Use this script. It employs the 20tr default CO2 ramping, but with an 1850 climate. .\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_defaults}}}\\
commit 36ced2a

Case names are:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_PI00}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_PI00}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_PI00}}}

\subsection{Make transient ensemble}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ensembles}}}\\
commit 7b8bca7

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_PI**}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_PI**}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_PI**}}}

\subsection{Analyse transient ensemble}

\section{Elevated CO$_{2}$ transient runs}
\subsection{Make elevated CO$_{2}$ file}
Use this script\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/modify_co2_drivers.m}}}\\
commit 0d4fb93

Which generates\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/co2_datm_ORNL_ELEV.nc}}}\\
commit 0d4fb93

\subsection{Make elevated CO$_{2}$ defaults}
Use this script:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults}}}\\
commit 8e6a3a5

Which uses these user$\_$nl$\_$datm files:\\
{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_1x1pt_Br-cax}}}\\
{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_bci_0.1x0.1_v4.0i}}}\\
{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_1x1pt_US-ORN}}}\\
commit 4cb99a4

And this user$\_$datm.streams.txt.co2tseries.20tr file\\
{\color{blue}\small{\url{gitpath/co2_files/user_datm.streams.txt.co2tseries.20tr}}}\\
commit 41d4aeb

Which references the previous co2 file\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/co2_datm_ORNL_ELEV.nc}}}\\
commit 0d4fb93

The case names are:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI00}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI00}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI00}}}

\subsection{Make elevated CO$_{2}$ ensemble}
Use this script:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_ensembles}}}\\
commit ffe1b6b

The case names are like:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI**}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI**}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI**}}}

\subsection{Analyze elevated CO2 runs}
Use this script:\\
{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_apr18.m}}}\\
commit ec202a0

Creates these figures:
{\color{magenta}\small{\url{gitpath/matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}}}\\
commit 23ce56e

\section{Make runs with LUNA off}

\subsection{Make ensemble spinup runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
commit 88b6508

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI15*}}\\
\small{\url{CLM5_1x1pt_Br-cax_ens_PI15*}}\\
\small{\url{CLM5_1x1pt_US-ORN_ens_PI15*}}}\\

\subsection{Make default spinup runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_defaults_lunaoff}} }\\
commit 88b6508

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI00_noluna}}\\
\small{\url{CLM5_1x1pt_Br-cax_ens_PII00_noluna}}\\
\small{\url{CLM5_1x1pt_US-ORN_ens_PII00_noluna}}}\\

\subsection{Make transient runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ensembles}} }\\
commit acdee08

The case names are like:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI15*}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI15*}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI15*}}}

\subsection{Make default transient runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_defaults_lunaoff}} }\\
commit b26f487

\section{Correct the CO2 fertilization to be all year}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/modify_co2_drivers_allyear.m }} }\\
commit 723bbf3

File is here: 
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/mco2_datm_ORNL_ELEV_allyear.nc }} }\\
commit 8604dd9

\subsection{Restart ensemble transients with new CO2 file}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_ensembles}}}\\
commit 9f8c771

\subsection{Restart default transients with new CO2 file}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults}}}\\
commit cc6e00e

\subsection{Restart LUNA defaults transients with new CO2 file}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults_lunaoff}}}\\
commit cc6e00e

\subsection{Analyze new elevated CO2 runs}
Use this script:\\
{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_apr18.m}}}\\
commit  7eab468

Creates these figures:
{\color{magenta}\small{\url{gitpath/matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}}}\\
commit 23ce56e

\section{Add additional sites}

\section{Add parameters to look at open/closed N cycle}

\section{Modify range of FlexCN parameters to completely stop C:N flexibility?}

\section{Investigate differences in driving data for single sites vs reanalysis}

\section{Add Nitrogen fertilization experiments}

\end{document}

### 2018-05-23 main.tex

\documentclass[draft,linenumbers]{agujournal}
\draftfalse
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
\begin{document}

\authors{Rosie Fisher\affil{1},
Will Wieder\affil{1},
Ben Sanderson\affil{1},
Charlie Koven\affil{2},
Keith Oleson\affil{1},
Chonggang Xu\affil{3},
Ashehad Ali\affil{1},
Josh Fisher\affil{4},
Mingjie Shi\affil{4},
Katie Dagon\affil{1},
Danica Lombardozzi\affil{1},
Anthony Walker\affil{5},
S\"onke Zaehle\affil{6},
Gordon Bonan\affil{1},
David Lawrence\affil{1}
}

\title{Nitrogen cycling CLM5: controls on responses to environmental forcing}
\author{rosie fisher}
\date{April 2018}

\affiliation{1}{National Center for Atmospheric Research, Table Mesa Drive, Boulder, Colorado, USA}
\affiliation{2}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
\affiliation{3}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
\affiliation{4}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
\affiliation{5}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
\affiliation{6}{Max Planck Institute for Biogeochemistry, Jena, Germany}
\correspondingauthor{Rosie Fisher}{rfisher@ucar.edu}

\begin{keypoints}
\item A suite of modifications to the Nitrogen cycle representation were added to the Community Land Model (version 5).
\item New features of the N model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of Nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios, dependant on N costs and distance from the target C:N ratio, and dynamic leaf resorption of Nitrogen. 
\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycle of Carbon and Nitrogen. We assess the dependence of the model state to parameter uncertainty, and also the responses of the system to higher CO2, Nitrogen and temperature.
\item The model responses to CO2 and N fertilization are linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment. Responses to temperature depend upon x,y,z. 
\item The model runs are conducted at representative sites in temperate, tropical and boreal systems.

\end{keypoints}

\section{to do}
1.	Write up developments of N cycle. LUNA, FUN, FLEXCN. 
2.	Develop sensitivity methods
t

\begin{abstract}
= enter abstract here =
\end{abstract}

\section{Introduction}

Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models. In the CMIP5 analysis, only one model, the Community Land Model (v4.0) included an active nitrogen (N) cycle, and this model suggested that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}, ).  This result raises the question of whether the inclusion of Nitrogen cycling necessarily generates such substantial limitations controls on the fertilization response of the biosphere.

Substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models. Unlike representations of, say, photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.

Subsequent to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{zaehle2010},  \cite{goll2012}, \cite{smith2014}, \cite{wang2007}). Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the theoretical framework for and parameterization of nutrient cycling processes (\cite{zaehle2014}).

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}.  Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release. An overview of model developments is presented by \cite{lawrence2018}, a descriptions of the global Nitrogen cycling by Wieder et al. (in prep), the crop model (Lombardozzi et al. in prep), the land use components (Lawrence P et al. in prep), land surface hydrology (Swenson et al. in prep) and plant hydraulics (Kennedy et al. in prep).

The CLM5 model includes numerous major update of the nitrogen cycling representation subsequent to CLM4.0 and CLM4.5. The introduction of these new dynamic features into the CLM5 model has implications for parameter uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale implications of the full model update.  In this paper, we describe the new components of the biogeochemical cycling of nitrogen, their impacts upon the model system state and responses to environmental forcing.

\subsection{Model Updates}
The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{Thornton}, modifications to the biogeochemical cycling included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).  The new model contains updates to three different elements of the Nitrogen cycle, including: 
1)	The LUNA module, (Leaf Utilization of Nitrogen for Assimilation), described by \cite{xu2012} and \cite{ali2016}. 
2)	The FUN module (Fixation and Uptake of Nitrogen) described by \cite{fisher2010fun} \cite{brzostek2014} and \cite{shi2016}. 
3)	Part of the `FLEXCN' N cycle implementation, described by \cite{ghimire2016}.

\subsubsection{LUNA}
The LUNA model implements the prediction of $V_{c,max}$ and $J_{max}$ from environmental conditions.   The model determines the rate constants that are consistant with a co-limitation of photosynthesis by both electron capture and carboxylation processes. In this fashion, $V_{c,max}$ is largely independent of PFT, and is primarily controlled by the N per unit leaf area ($N_{area}$) and environmental conditions. The capacity of the LUNA model to represent appropriate responses to temperature and CO2 was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

\subsubsection{FUN}
The FUN model operates on the principle that for the primary sources of Nitrogen acquisition by plants (active uptake, retranslocation and symbiotic N fixation) there is a cost in terms of carbon. Further, the cost of each of these uptake pathways is variable in time and space, and thus plants will likely take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for growth can be deployed to acquire more N. The central element of FUN is a simulataneous equation which assumes, firstly, that the carbon available for growth ($C_{growth}$) is the total carbon available ($C_{avail}$), minus that used for Nitrogen uptake, $C_{nuptake}$:

\begin{equation}
C_{growth}=C_{avail}-C_{nuptake}
\end{equation}

Further, the Nitrogen acquired from the environment must equal that which is deployed in the growing plant tissues. The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$) which applies only to tissues that are constructed:

\begin{equation}
C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
\end{equation}

Thus the Carbon that is expended on N uptake is determined by:
\begin{equation}
C_{nuptake} =\frac{C_{avail}}{ ( (1.0+f_{gr})*(CN_{target} / N_{cost}) + 1) }
\end{equation}

The average Nitrogen cost ($N_{cost}$) is derived from the simultaneous uptake of N across all uptake streams, while the target CN ratio ($CN_{target}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools. The FUN model is documented \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}{here}

\subsubsection{FLEXCN}
\cite{ghimire2016} implemented a suite of modifications to the Nitrogen cycling including variable tissue C:N ratio, prognostic Vcmax (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm. In CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen. In this new algorithm, Nitrogen supply is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates. This new allocation code we term the FLEXCN module, and it is documented in ***.

\subsubsection{Adaptation of FUN to flexible CN ratios}
The FUN model was originally conceived with a static tissue CN ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target CN ratio. In CLM5, one of the aims of the development process was to incorporate a varying CN ratio. To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required). The degree to which tissue CN content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments in light of limitations (including \cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content. When N is expensive in the environment, less C is spent on uptake, increasing CN ratios where N is scarce. When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high.

\subsubsection{Justification for model developments}
Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability  emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}). For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage,  but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization \cite{ainsworth2007} and changing temperature \cite{hikosaka2005}, irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}). Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).

\subsubsection{Content of this paper}

\section{Methods}

Four 
We ran the default model configuration to equilibrium, and

\section{Results}

\section{Discussion}

\subsubsection{Discourse on how target CN ratio could be optimized but isn't}
\cite{thomas2014}, \cite{friedlingstein1999}, \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{anten2011}, \cite{vanwijk2003}

\subsubsection{discuss geographical spread of predictions for N and CO2 fertilizaion}
\cite{hickler2008}

\subsubsection{Discourse on whether nitrogen fixation is realistic in CLM5}
The 'fraction of fixers' ($frac\_fixers$) is a proxy for the community level composition of Nitrogen fixing plants. Technically, because plant density is not represented in the CLM5, this parameter really represents the fraction of net assimilated carbon ($gpp$ - $R_{auto}$) that can be used for symbiotic N fixation. Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would response to changes in environmental conditions and the relative competitive abilities of N fixing plants. In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (references**), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).  The results included here indicate that the

 \cite{thomas2013global} find that the CLM4.0 model responds very strongly to nutrient additions, and is heavily limited by nitrogen availability in the present day.

\section{Conclusions}

\section{Figures.}

Figure 1: Sensitivity analysis of state of CLM at site to N parameters @ORNL
Figure 2: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @ORNL
Figure 3: Sensitivity analysis of state of CLM at site to N parameters @BCI
Figure 4: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @BCI
Figure 5: Sensitivity analysis of state of CLM at site to N parameters @CAX
Figure 6: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @CAX
Figure 7: Operation of the flexCN ratio module. 
Figure 8: Operation of the leaf N retranslocation module
Figure 9: Details on the N uptake in FUN
Figure 10: Some sort of Will figure.

Also add LUNA on and off simulations to figures.

\nocite{*} 
\bibliography{aguCLM5_bibtex}

\section{Appendix 1: Technical Documentation of Nitrogen Cycle}
\subsubsection{Modifications to allow variable C:N ratio}
The total cost of N uptake is calculated based on the assumption that carbon is partitioned to each stream in proportion to the inverse of the cost of uptake. So, more expensive pathways receive less carbon. Earlier versions of FUN (Fisher et al., 2010)) utilized a scheme whereby plants only took up N from the cheapest pathway. Brzostek et al. (2014) introduced a scheme for the simultaneous uptake from different pathways. Here we calcualate a ‘conductance’ to N uptake (analagous to the inverse of the cost function conceptualized as a resistance term) N_{conductance} ( gN/gC) as:

N_{conductance,f}=  \sum{(1/N_{cost,x})}

From this, we then calculate the fraction of the carbon allocated to each pathway as

C_{frac,x} = \frac{1/N_{cost,x}}{N_{conductance}}

These fractions are used later, to calculate the carbon expended on different uptake pathways. Next, the N acquired from each uptake stream per unit C spent (N_{exch,x}, gN/gC) is determined as

N_{exch,x} = \frac{C_{frac,x}}{N_{cost,x}}

We then determine the total amount of N uptake per unit C spent (N_{exch,tot}, gN/gC) as the sum of all the uptake streams.

N_{exch,tot} = \sum{N_{exch,x}}

and thus the subsequent overall N cost is

N_{cost,tot} = 1/{N_{exch,tot}}

Retranslocation is determined via a different set of mechanisms, once the N_{cost,tot} is known

\end{document}

### 2018-05-24 matlab/make_namelist_files.m
added namelist matlab file

@@ -1,0 +1,32 @@
+template = '/glade/p/cesm/cseg/inputdata/lnd/clm2/paramdata/clm5_params.c171017.nc'
+
+varnames(1:16) = {'null'}
+
+varnames(17:21)={'k_nitr_max','denitrif_respiration_coefficient','denitrif_respiration_exponent','denitrif_nitrateconc_coefficient','denitrif_nitrateconc_exponent'}
+
+defaults(17:21) = [ 0.1 0.1 1.3 1.15 0.57]
+dir = 'namelists'
+
+mult(:,17)=[0.1 0.5 2  4] %perecm. 1 is the default. 
+mult(:,18)=[0.1 0.5 2  4] 
+mult(:,19)=[0.1 0.5 2  4] 
+mult(:,20)=[0.1 0.5 2  4] 
+mult(:,21)=[0.1 0.5 2  4] 
+
+formatspec = '%.4f'
+vchoose = 17:21
+for i=1:4
+for vr=vchoose
+
+  new_var =  defaults(vr) .* mult(i,vr)
+  namefile = strcat(dir,'/','namelist','PI',char(num2str(vr)),char(num2str(i)))
+  copyfile('namelists/defaultnl',namefile)
+  textnl = strcat(char(varnames(vr)),char('='),char(num2str(new_var)))
+  textnl
+
+  matrix1={char(varnames(vr))}
+  matrix2=new_var
+
+  writetable(cell2table([matrix1 num2cell(matrix2)]),namefile,'writevariablenames',0)
+end
+end

### 2018-06-29 figures.tex

@@
 \begin{document}
+ \begin{figure}[h]
+     \centering
+     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_CAX__y1.png}
+     \caption{Caxiuana State Parameter Sensitivity}
+     \label{CAX state}
+  \end{figure}
+
+   \begin{figure}[h]
+     \centering
+     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_BCI__y1.png}
+     \caption{BCI State Parameter Sensitivity}
+     \label{BCI state}
+  \end{figure}
+
+ \begin{figure}[h]
+     \centering
+     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_ORN__y1.png}
+     \caption{Oak Ridge State Parameter Sensitivity}
+     \label{ORN state}
+  \end{figure}
 
@@
      \caption{Response to Nitrogen Deposition}
-     \label{Cax CO2 response}
+     \label{Caxiuana CO2 response Parameter Sensitivity }
   \end{figure}
@@
      \caption{Response to Nitrogen Deposition}
-     \label{ORNL CO2 response}
+     \label{Oak Ridge CO2 response Parameter Sensitivity}
   \end{figure}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI_y1.png}
+     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_BCI__y1.png}
      \caption{Response to Nitrogen Deposition}
-     \label{BCI CO2 response}
-  \end{figure}
-
- \begin{figure}[h]
-     \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
-     \caption{Response to Nitrogen Deposition}
-     \label{CAX state}
-  \end{figure}
-
- \begin{figure}[h]
-     \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
-     \caption{Response to Nitrogen Deposition}
-     \label{Ndep}
+     \label{BCI CO2 response Parameter Sensitivity}
   \end{figure}

### 2018-06-29 main.tex

@@
 
-\title{Nitrogen cycling CLM5: controls on responses to environmental forcing}
+\title{Nitrogen cycling in the CLM5: controls on responses to environmental forcing}
 \author{rosie fisher}
@@
 
-\section{to do}
-1.	Write up developments of N cycle. LUNA, FUN, FLEXCN. 
-2.	Develop sensitivity methods
-t
-
-\begin{abstract}
-= enter abstract here =
-\end{abstract}
-
 \section{Introduction}
 
-Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models.
-In the CMIP5 analysis, only one model, the Community Land Model (v4.0) included an active nitrogen (N) cycle, and this model suggested that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}, ).
+Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}).
+In the CMIP5 multi-model intercomparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle.
+The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}).
 This result raises the question of whether the inclusion of Nitrogen cycling necessarily generates such substantial limitations controls on the fertilization response of the biosphere.
 
-Substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models.
-Unlike representations of, say, photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.
+Substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehledalmonech2011}).
+Unlike representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.
 
-Subsequent to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{zaehle2010},  \cite{goll2012}, \cite{smith2014}, \cite{wang2007}).
+Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{wang2007}, \cite{zaehle2010},  \cite{goll2012}, \cite{smith2014} ).
 Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the theoretical framework for and parameterization of nutrient cycling processes (\cite{zaehle2014}).
@@
 \subsection{Model Updates}
-The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{Thornton}, modifications to the biogeochemical cycling included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).
-The new model contains updates to three different elements of the Nitrogen cycle, including:
-1)	The LUNA module, (Leaf Utilization of Nitrogen for Assimilation), described by \cite{xu2012} and \cite{ali2016}. 
-2)	The FUN module (Fixation and Uptake of Nitrogen) described by \cite{fisher2010fun} \cite{brzostek2014} and \cite{shi2016}. 
-3)	Part of the `FLEXCN' N cycle implementation, described by \cite{ghimire2016}.
+The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).
+The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\
+
+1)	The LUNA model (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes ( \cite{xu2012} and \cite{ali2016}) \\
+2)  The FUN module (Fixation and Uptake of Nitrogen), which simulates the dynamics of Nitrogen acquisition from the environment \cite{fisher2010fun} \cite{brzostek2014} and \cite{shi2016}.\\
+3)	The `FLEXCN' N cycle implementation, which is used primarily to allow variation in tissue C:N ratios, and is adapted from \cite{ghimire2016}.\\
+
+More details on each of these new components is given below.
+A full technical description of the CLM5 is an appendix to \cite{lawrence2018} and is also available online at \url{https://escomp.github.io/ctsm-docs}
 
 \subsubsection{LUNA}
-The LUNA model implements the prediction of $V_{c,max}$ and $J_{max}$ from environmental conditions.
-The model determines the rate constants that are consistant with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
-In this fashion, $V_{c,max}$ is largely independent of PFT, and is primarily controlled by the N per unit leaf area ($N_{area}$) and environmental conditions.
-The capacity of the LUNA model to represent appropriate responses to temperature and CO2 was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
+
+Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
+There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization \cite{ainsworth2007} and changing temperature \cite{hikosaka2005}, irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
+Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
+
+The LUNA model predicts the optimal balance of of $V_{c,max}$ and $J_{max}$ for the prevailing time-averaged environmental conditions.
+The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
+In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area ($N_{area}$) and environmental conditions.
+The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
 
 \subsubsection{FUN}
-The FUN model operates on the principle that for the primary sources of Nitrogen acquisition by plants (active uptake, retranslocation and symbiotic N fixation) there is a cost in terms of carbon.
+Under circumstances where insufficient N exists to match all of the carbon assimiliated for a given C:N ratio, the CLM4.0 model reduced the gross photosynthetic flux down to the level at which growth could be supported by the assmilated N.
+This process occurred after the calculation of stomatal condutance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistancies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{dekauwe2014}, \cite{walker2014}, \cite{bonan2012}).
+One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions.
+DISCUSS OTHER MODELS, OCN, ETC.
+
+The FUN model operates on the principle that for the primary sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescing tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon.
 Further, the cost of each of these uptake pathways is variable in time and space, and thus plants will likely take up Nitrogen from pathways which are 'cheaper' to them.
@@
 
+A further issue addressed by FUN is the preferential use of symbiotic N fixation by leguminous plants when N is very limiting.
+The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity.
+\cite{wieder2015} illustrate the impacts of uncertainty in this function, but FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition, and temperature, which exerts a primary control over the enzymatic processes, following \cite{houlton2008}.
+
 \subsubsection{FLEXCN}
+Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability  emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}).
+For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage,  but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
+For this reason, we introduce a flexible tissue C:N ratio into the CLM5.
+
 \cite{ghimire2016} implemented a suite of modifications to the Nitrogen cycling including variable tissue C:N ratio, prognostic Vcmax (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
@@
 
-\subsubsection{Justification for model developments}
-Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability  emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}).
-For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage,  but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
-For this reason, we introduce a flexible tissue C:N ratio into the CLM5.
-
-Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
-There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization \cite{ainsworth2007} and changing temperature \cite{hikosaka2005}, irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
-Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
-
 \subsubsection{Content of this paper}
+In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters. We use\\
+1. A range of single site experiments \\
+2. One-at-a-time and global parameter uncertainty analyses\\
+3. CO$_{2}$ fertilization experiments and\\
+4. Nitrogen addition experiments.\\
+to fully investigate the performance of the model under steady state, transient and step-change modifications.
+We thus illustrate how model parametric choices might affect overall behaviour in larger-scale experiments,  provide guidance on which elements of the model system require greatest attention in future model development and testing cycles.
+Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including a removal of N fixation capacity and a return to fixed C:N ratios.
+We further investigate the consequences of turning off the LUNA model and returning to a fixed leaf N/$V_{c,max}$ relationship as defined by \cite{ghimire2016}.
 
@@
 
-Four 
-We ran the default model configuration to equilibrium, and
+\subsection{Sites}
+For this experiment we use three different model evaluation sites.
+Two in tropical forests (Caxiuana and Barro Colorado Island) one in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment) and the Barrow flux tower, as an example of a boreal system.
+
+\subsection{Simulation Protocol}
+We spun up the default version of the model for 400 years at each site, using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available meterological inputs.
+400 years was sufficient to bring all carbon and nitrogen pools into equilibrium.
+Subsequent to this, parameters we perturbed (see below), and a second spin-up conducted for 180 years at each site.
+Using the state at the end of the perturbed spin-ups, we ran transient simulations for each site and parameter combination starting in 1850 and invoking transient CO$_{2}$ and Nitrogen deposition.
+We ran additional CO$_{2}$ and Nitrogen deposition experiments by increasing the CO$_{2}$ concentration to 550ppm.
+In the case of the Oak Ridge FACE experiment, we only increased CO$_{2}$ in the growing season, in line with the experimental protocol applied there and using the observed CO$_{2}$ timeseries.
+For all other sites, we increased CO$_{2}$ to 550ppm year round starting in 1997**.
+For Nitrogen, we added 5kgN/$m^{2}$/year, also starting in 1997, so that the results can be compared with those reported by Wieder et al.
+(in prep).
+
+\subsection{Perturbed Physics Ensemble (PPE)}
+The CLM5 has a very large number of parameters to which the C and N cycles are potentially sensitive.
+Here we focused on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
+We choose a larger set of parameters for one-a-time analyses, followed by a smaller set of these to which we subject the model to global sensitivity tests.
+Knowledge of model structure as well as previous investigations of the CLM model informs our choice of parameters here.
+Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model.
+However, we also report responses of the model state to perturbations.
+Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al.
+in prep) and into the plant hydraulics parmeterizations (Kennedy et al in prep).
+
+In the first instance, we conduct one-at-a-time (OAAT) perturbations of the parameters around a mean state, to allow a straightforward visualization and interpretation of the results.
+For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation).
+
+\subsubsection{Parameter Choices}
+The parameters included in our sensitivity analysis (Table \ref{table_parameters}) are: 1) \textbf{Specific Leaf Area}, \emph{`SLA'}, which is a critical parameter determining the area of canopy (in m$^{2}$) derived from one gram of leaf biomass.
+This parameter is known to be one of the critical elements of the leaf economics spectrum, and is well-characterized by the TRY database (\cite{kattge2011}).
+2) \textbf{Fine root are per unit leaf area}, \emph{`froot\_leaf'}, the ratio of fine root biomass to leaf area.
+Unlike in previous versions of the CLM, fine root biomass now affects the capacity of the model to acquire both water and nutrients, thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth.
+3) \textbf{Stem to leaf are ratio}, \emph{`stem\_leaf'}, reflects the balance between accumulation of stem and woody biomass, and leaf area, and thus has large impacts both on the leaf area index achievable for a given productivity, and conversely on the equilibrium woody biomass stored.
+4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.
+These six parameters were previously defined by \cite{brzostek2014}.
+Here and in CLM5 in general, we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly.
+These parameters include \emph{ekc\_active} and  \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and  \emph{akn\_active} for arbuscular mycorrhizal fungi, and  \emph{kc\_nonmyc} and  \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
+5) \textbf{The fraction of carbon that can be used for fixation}, \emph{fracfixers}.
+This parameter is the fraction of the assimilation ($gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants.
+6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{grperc}, was revised substantially in CLM5 based on Atkin O.
+(pers comm.) from 0.3 to 0.11.
+Here we test the impacts of this change on overall carbon budgets.
+7) \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}.
+CLM5 uses the concept of a 'target' leafcn ratio, around which flexibility is allowed, given the environmental costs of Nitrogen.
+LeafCN ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$) 8)  \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}.
+This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, assessed in this set of experiments.
+9) \emph{The intercept of the Atkin 2015 leaf maintenance respiration model}, \textbf{lmr\_intercept}.
+The new leaf maintenance respiration model, proposed by \cite{atkin2015} and adopted by CLM5, has it's intercept as a PFT dependant parameter, with variation between PFTs reported by \cite{atkin2015} and used herein.
+Here we test the impacts of the uncertainty in the observations of this parameter.
+10) \textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is another ecosystem-level property defined by the FUN model and allowing the proportions of the fungal symbiotic pathways to be modified.
+This capacity was not fully developed in the CLM5 development process and so here we assess its importance.
+11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in Appendix 1.
+
+\subsubsection{Parameters not used}
+We conducted a set of tests on the parameters of the denitrification model, to assess the impacts of differing rates of N loss on the system.
+However, analysis has shown that the rates of denitrification in CLM5 are very low, overall (compared to plant uptake and immobilization) and hence these parameters had very little impact.
+We did not conduct tests on leaf lifespan, since for evergreen trees in CLM5, leaf lifespan and specific leaf area have very similar impacts on overall achievable leaf area.
+Parameters related to the hydrological cycle and to the plant hydraulics scheme are the subject of assessment in Dagon et al.
+(in prep).
+We did not investigate any parameters relating to the crop model (Lombardozzi et al.
+in prep) nor to plant harvest, fire, dust, soil biogeochemical cycling (which is largely unaltered from CLM4.5, \cite{koven2013} apart from modifications presented by Lawrence et al.
+(in prep).
+
+\subsubsection{Parameter Ranges}
+1.
+\emph{Specific Leaf Area} and \emph{leafCN} are defined in the TRY database, and their distributions are log-normally distributed.
+Thus, we take the average log-normal standard deviation range from TRY and apply it across all PFTs.
+Thus the lower bounds are closer to the mean than the higher bounds (Table \ref{table_ranges}).
+The range of \emph{fracfixers} was set to vary across the whole logical range from 0 to 1.
+\emph{grperc} ranges from zero up to the range of the previously used CLM value (0.3).
+\emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}.
+The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood, thus we used ranges that span the response limits of these parameters, particularly for  \emph{cn\_flex\_c} which we explored across four orders of magnitude to investigate the impacts of leafCN flexibility.
+The allometric parameters ***
+
+\subsection{Analysis}
+We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the transient simulations ** for the years 2001-2003.
+For some parameter settings, the constraint on growth caused by the parameter itself placed the system in a state close to death.
+Addition of N or CO$_{2}$ therefore causes either a very strong recovery of the plants, or no recovery at all, leading to large non-linearities in how parameters are linked with responses to environmental forcing.
+Therefore, for those parameter combinations which cause LAI to be less than 80\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plotting, so as to focus on parameter combinations that give realistic initial vegetation conditions.
 
 \section{Results}
+
+\subsection{Response of state to parameter perturbation}
+One notable feature of the sensitivity analysis here is the difference in the dominant parameter responses between different output variables, particularly as one moves along the C allocation cascade from GPP, to NPP and then LAI and vegetation carbon.
+Leafcn ratio, for example, is a dominant control over GPP, but is much less important for NPP, where the respiration intercept ($lmr\_{intercept}$), fine root ratio, and the fraction of fixers (by means of the carbon expenditure on respiration) impart a greater impact.
+For LAI, the allocation parameters, specific leaf  area and stem:leaf ratio have a much greater import.
+It is implicit, therefore, that even a very confident understanding of gross photosynthetic fluxes would be of limited importance in predictions of vegetation structure within this model framework.
+
+\subsection{Parametric control over CO$_{2}$ and Ndep responses}
+
+Figure ** shows the response of GPP, for each site, to CO$_{2}$ and nitrogen fertilization, respectively, for each one-at-a-time parameter modification.
+Figure ** shows the same for Net Primary Productivity, and Figure ** for LAI, and Figure ** for leaf N concentration.
+There is a large quantity of infomation contained in these figures, and, given that the analysis is partly meant to serve as a resource for users interested in the functionality of C and N cycling in the CLM5,  rather than describe the results exhaustively, we focus here on explaining the functionality behind a few of the major patterns observed.
+
+1.
+\textbf{Impact of fracfixers}.
+The results quite strongly illustrate that the fraction of fixers (blue line, Figures ** to **) is a highly dominant control over all aspects of the environmental response surfaces.
+For GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$.
+In low fixer systems, N uptake is highly expensive in the first instance, and is greatly alleviated by the increase in available N, leading to larger allocations to leaves, and thus higher leaf area and GPP.
+For NPP, this result is muted for the two (less N limited) tropical sites, particularly Caxiuaxa (Figure **)  because the increase in shaded leaves and in tissue C:N ratios (Figure **)  with fertilization increased the respiratory load on these systems.
+
+2. \textbf{Impact of Ncosts}
+Most of the variables and sites displayed a complex relationship between the Nitrogen costs and the response to CO$_{2}$ and N addition. 
+For the tropical sites, where default N limitation is low and responses to Ndep are smaller, the simulations with a higher cost of N uptake push the system into a space where N expenditure is higher and so the response to fertilization is more pronounced.
+Conversely, for all sites, decreasing the cost of N
+
+In general, the simulation with very high N uptake costs (large brown symbols) showed higher responses to N addition than the default case, on account of their large percentage expenditures on
+
+whereas their response to CO$_{2}$ addition was not systematically different from the default.
+
+This is not unexpected, but future implementations of the model would do well to focus on the dynamics of N fixer dynamics in space and time.
+For NPP, all sites showed the same responses, except for Caxiuana, where leaf N responded strongly to N deposition (Figure **) and so altered respiratory costs of the low N fixer state .
+For LAI, at all sites, the reduction in the uptake costs of N in  allows higher allocation to growing tissues.
+
+For the two tropical sites, (BCI and Caxiuana) the response surfaces are similar,
+
+\subsection{Response of Nitrogen fertilization to parameter perturbation}
 
@@
 
-\subsubsection{discuss geographical spread of predictions for N and CO2 fertilizaion}
+\subsubsection{Discuss geographical spread of predictions for N and CO2 fertilization}
 \cite{hickler2008}
@@
 
-\section{Figures.}
+\pagebreak
+\begin{table}
+\begin{center}
+\begin{tabular}{ |c|c|c| } 
+ \hline
+ Parameter Name & Meaning & Units & Min & Max \\
+  \hline
+ slatop & Specific Leaf Area & m2/g \\ 
+ froot\_leaf & Fine root:leaf ratio. & g/g \\
+ stem\_leaf  & Stem:leaf ratio. & g/g \\ 
+ n\_costs    & 6 parameters related to Nitrogen costs & g/g \\
+ fracfixers  & Fraction of C availble for fixation & - \\
+  grperc  & Fraction of C used for growth respiration  & - \\
+  leafcn  & Leaf C:N ratio. & g/g \\
+     medlyn\_slope  & Slope of stomatal conductance curve & g/g \\
+      lmr\_intercept & intercept of leaf N /respiration relationship &\\
+      frac\_ecto & Fraction of active uptake that is ectohycorrhizal & \\
+      cn\_flex\_a & Parameter of N flexibilty algorithm& \\
+      cn\_flex\_b & Parameter of N flexibilty algorithm & \\
+      cn\_flex\_c & Parameter of N flexibilty algorithm & \\
+\hline
+\end{tabular}
+\end{center}
+\caption{CLM5 Parameters subject to sensitivity analysis in this study. }
+\label{table_parameters}
+\end{table}
+
+ \begin{table}
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| } 
+ \hline
+ Parameter Name & s1 &s2 & s3 & s4\\
+  \hline
+ slatop & 0.6761 & 0.8769 &1.2785 &1.4791\\ 
+ froot\_leaf &  0.342 &0.671 &1.05 & 1.1\\
+ stem\_leaf  &  0.3057 &0.61 &1 &1\\ 
+ n\_costs    & 10$^{-2}$ &10$^{-2}$&  10$^{1}$& 10$^{2}$\\
+ fracfixers  & 0 &0.5 & 2 & 4 \\
+  grperc  &  0 & 0.5& 2 & 3\\
+  leafcn  &0.7413 & 0.8932 & 1.1970  & 1.349\\
+
+     medlyn\_slope  &0.5258 & 0.7191 & 1.1057  &1.29\\
+      lmr\_intercept & 0.8539 & 0.9531 & 1.1512& 1.25\\
+      frac\_ecto & 0 &0.5& 1 & 1 \\
+      cn\_flex\_a &0 &0.5 & 2  &4\\
+      cn\_flex\_b & 0.1 & 0.5 & 2  & 4\\
+      cn\_flex\_c &0.001 & 0.1 & 10 & 100\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Multipliers on the original values in CLM5 used in this sensitivity analysis .}
+\label{table_ranges}
+\end{table}
+
+\section{Figures}
 

### 2018-06-29 Workflow.tex
Updates from ShareLaTeX

@@
 
-Scripts for these analyses should all be found in the public github repo at:
+Scripts for these analyses should all be found in the public github repo at:\\
 {\color{blue}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS}}}\\
 
-The scripts are specific to usage on the NCAR Cheyenne machine...
+The scripts are specific to usage on the NCAR Cheyenne machine...\\
 
@@
 
-\section{Add additional sites}
+\section{Add parameters to look at open/closed N cycle}
+\subsection{Make namelist files witht this matlab script}
+{\color{blue}\small{\url{gitpath/matlab/make_namelist_files.m}}}\\
+commit 532c208
 
-\section{Add parameters to look at open/closed N cycle}
+Files are in here:
+{\color{magenta}\small{\url{gitpath/matlab/namelists/**}}}\\
+commit c7633e9
+
+\subsection{Make spinup runs for namelist parameters}
+{\color{blue}\small{\url{gitpath//case_scripts/make_site_spinup_ensembles}}}\\
+commit faa49a9
+
+\subsection{Analyze spinup results for namelist parameters} 
+matlab script modified
+{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_mar18.m
+}}}\\
+e04d71
+
+\subsection{Make transient runs for namelist parameters}
+
+{\color{blue}\small{\url{gitpath//case_scripts/case_scripts/make_site_transient_ELEV_ensembles}}}\\
+commit 96fafee
+
+\subsection{Analyze transient runs for namelist parameters}
+
+{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_mar18.m}}}\\
+commit fe1a59
 
@@
 
+\subsection{Modified ranges of flex\_cn\_c parameters}
+
+{\color{blue}\small{\url{gitpath/matlab/FACE_clm5_ens_oaat.m 
+}}}\\
+commit b3f67e9
+
+Parameter files here:
+Output location..\\
+{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi13*.nc }}} \\
+commit 823385e
+
+\subsection{Set up ensemble runs just for parameter 13}
+{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
+b3f67e9
+
+\subsection{Set up transient runs just for parameter 13}
+{\color{blue}\small{\url{gitpath//case_scripts/case_scripts/make_site_transient_ELEV_ensembles}}}\\
+b3f67e9
+
+\subsection{Analyze new runs}
+
+\section{Add Nitrogen fertilization experiments}
+\subsection{Made new +5gCm2 file}
+
+\subsection{Make transient NDEP ensembles}
+{\color{blue}\small{\url{gitpath/case_scripts/make_site_transient_NDEP_ensembles}}}\\ 
+commit 2e395ef
+
+\subsection{Make transient NDEP defaults}
+{\color{blue}\small{\url{gitpath/case_scripts/make_site_transient_NDEP_defaults}}}\\ 
+commit 2e395ef
+
+\section{Add additional sites}
+added Barrow
+
+\section{add temperature experiments}
+
+\section{add reduced form global sensitivity analysis}
+
+ \section{Figures}
+ \subsection{Make Will figure}
+ \subsection{Make initial LAI figure}
+ \subsection{Make 3D figures of parameter space}
+
 \section{Investigate differences in driving data for single sites vs reanalysis}
 
-\section{Add Nitrogen fertilization experiments}
-
 \end{document}

### 2018-06-29 main.tex
Updates from ShareLaTeX

@@
 For the tropical sites, where default N limitation is low and responses to Ndep are smaller, the simulations with a higher cost of N uptake push the system into a space where N expenditure is higher and so the response to fertilization is more pronounced.
-Conversely, for all sites, decreasing the cost of N
-
-In general, the simulation with very high N uptake costs (large brown symbols) showed higher responses to N addition than the default case, on account of their large percentage expenditures on
-
-whereas their response to CO$_{2}$ addition was not systematically different from the default.
+Conversely, for all sites, decreasing the cost of N uptake increase Leaf Area Index
 

### 2018-10-04 matlab/CLM5_FACE_senstest_output_ELEV_apr18.m
new case scripts

@@ -4,7 +4,7 @@
 % force matlab to write graphics correctly.... 
 %opengl neverselect
 					   
-% If we re-run this script, we need to close all the existing figures and void the existing variables 
+% If we re-run this script, we need to close all the existing figures and void the ing variables 
 mkfigs=1
 
 if(mkfigs ==1)
@@ -14,14 +14,14 @@ mkfigs=1
 end
 figdir = 'figures'
 u=3600*24*365'
-vars         = ({'GPP','NPP','TLAI','QVEGT','TOTVEGC','BTRANMN','COST_NFIX','COST_NRETRANS','NPP_NUPTAKE','NACTIVE','NFIX','NRETRANS','LEAFN'})
-namevars     = ({'Gross Production (gCm^{-2}y^{-1})','Net Production (gCm^{-2}y^{-1})','Leaf Area Index (m^{2}m^{-2})','QVEGT','Total Veg Carbon (Kgm^{-2})' ,'BTRANMN' ,'COST NFIX','COST RETRANS' ,'N uptake cost (gCm^{-2}y^{-1})','NACTIVE','NFIX','NRETRANS','Leaf Nitrogen (g m^{-2})'})
+vars         = ({'GPP','NPP','TLAI','QVEGT','TOTVEGC','BTRANMN','COST_NFIX','NPP_NUPTAKE','NPP_NFIX','NPP_NACTIVE','NPP_NRETRANS','NRETRANS','LEAFN'})
+namevars     = ({'Gross Production (gCm^{-2}y^{-1})','Net Production (gCm^{-2}y^{-1})','Leaf Area Index (m^{2}m^{-2})','QVEGT','Total Veg Carbon (Kgm^{-2})' ,'BTRANMN' ,'COST NFIX','NPP Nuptake' ,'N uptake cost (gCm^{-2}y^{-1})','NACTIVE','NFIX','NRETRANS','Leaf Nitrogen (g m^{-2})'})
 miny            = [3550  1400    6    1500    15700             1       10           10           0.4              10         10         5    0]
 maxy            = [3350  1100    7    1500    19000             1       10           10              0.4           10         10       5      3]
-unit_conversion = [u      u      1      u      1/1000             1       1            1               u              u         u        u  1]
[…]

### 2018-10-04 matlab/CLM5_FACE_senstest_output_mar18.m

@@ -1,5 +1,6 @@
 
 
+
 %%%%%%%%%%%%%%%%%%% LOOKING AT NCL FILES IN Matlab %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % force matlab to write graphics correctly.... 
 %opengl neverselect
@@ -32,16 +33,16 @@ root = 'FUN_sens_ens_transient_PI'
 root = 'FUN_sens_ens_D_PI'
 root = 'FUN_sens_ens_transient_D_PI'
 root = 'CLM5_1x1pt_Br-cax_ens_PI'
-sitenames = {'1x1pt_Br-cax' 'bci_0.1x0.1_v4.0i' '1x1pt_US-ORN'}
+sitenames = {'1x1pt_Br-cax' 'bci_0.1x0.1_v4.0i' '1x1pt_US-ORN' '1x1pt_US-Brw' '1x1pt_US-NR1' '1x1pt_US_Ha1','1x1pt_US-Me2'}
 
 ichoose = 1:4  % loop roun iterations for each parameter
 vchoose = [1 2 3 5 9]; % variables.
-ychoose = 80; %choose year
+ychoose = 100; %choose year
 mchoose = 8 %6:9 %choose month
 fileend = '.nc'
  pr=1 %are we printing figures?
 anom=0; %are we doing anomolies?
-schoose = 1:3
+schoose = 1:7
 fleafcn=0
 fmrf=0
 arc=1 %in archive? 
@@ -81,7 +82,7 @@ for s=schoose
     if(pp==2);pchoose=[ 4 5 6 7];end
     if(pp==3);pchoose=[ 11 12 13 ];end
     if(pp==4);pchoose=[ 8 9 14];end 
-    if(pp==5);pchoose=[ 1:14,18:21];end   
+    if(pp==5);pchoose=[ 1:11];end   
     if(pp==6);pchoose=[18:21];end
 
     for p=pchoose %parameter loop. 
@@ -102,7 +103,11 @@ for s=schoose
 		 for m=mchoose								
 		    filen = strcat(dirname,'.clm2.h0.',num2str(y,'%04d'),'-',num2str(m,'%02d'),'.nc');
 		    filename = strcat(dir_clm,filen)
+                    if(exist(filename))
 		    rawvar = ncread(filename,char(vars(v)));
+                    else
+                    vm(m)=nan;
+                    end
[…]

### 2018-10-04 Workflow.tex
update of paper files

@@
 \title{Workflow for CLM5 paper}
-\maketitle{}
 
@@
 
-\section{Elevated CO$_{2}$ transient runs}
-\subsection{Make elevated CO$_{2}$ file}
+\subsection{Elevated CO2}
+\subsection{Make elevated CO2 file}
 Use this script\\
@@
 
-\subsection{Make elevated CO$_{2}$ defaults}
+\subsection{Make elevated CO2 defaults}
 Use this script:\\
@@
 
-\subsection{Make elevated CO$_{2}$ ensemble}
+\subsection{Make elevated CO2 ensemble}
 Use this script:\\

### 2018-10-04 figures.tex
update of paper files

@@
 \usepackage{hyperref}
+\usepackage[export]{adjustbox}
+\addtolength{\oddsidemargin}{-.875in}
+\addtolength{\evensidemargin}{-.875in}
 
@@
 \journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
+
 \begin{document}
+
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_CAX__y1.png}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_p1CLM5_CAX__y1.png}
      \caption{Caxiuana State Parameter Sensitivity}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_BCI__y1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_p1CLM5_BCI__y1.png}
      \caption{BCI State Parameter Sensitivity}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_ORN__y1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_p1CLM5_ORN__y1.png}
      \caption{Oak Ridge State Parameter Sensitivity}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
-     \caption{Response to Nitrogen Deposition}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
+     \caption{Caxiuana CO2 response}
      \label{Caxiuana CO2 response Parameter Sensitivity }
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI_y1.png}
-     \caption{Response to Nitrogen Deposition}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI_y1.png}
+     \caption{Oak Ridge CO2 response}
      \label{Oak Ridge CO2 response Parameter Sensitivity}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_BCI__y1.png}
-     \caption{Response to Nitrogen Deposition}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_CO2_response_1CLM5_BCI__y1.png}
+     \caption{BCI CO2 response}
      \label{BCI CO2 response Parameter Sensitivity}
@@
 
+ \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_GPP.png}
+     \caption{GPP CO2 and N respones}
+     \label{GPP CO2 and N respones}
+  \end{figure}
+
+ \begin{figure}[h]
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_NPP.png}
+     \caption{NPP CO2 and N respones}
+     \label{NPP CO2 and N respones}
+  \end{figure}
+
+ \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_LAI.png}
+     \caption{LAI CO2 and N respones}
+     \label{LAI CO2 and N respones}
+  \end{figure}
+
 \end{document}

### 2018-10-04 main.tex

--- 2018-05-23 main.tex
@@
 \documentclass[draft,linenumbers]{agujournal}
+
 \draftfalse
+
 \usepackage{hyperref}
-
 \hypersetup{
@@
 filecolor=magenta,
-    urlcolor=cyan,
-}
+urlcolor=cyan}
 
 \journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
+
 \begin{document}
@@
 Anthony Walker\affil{5},
-S\"onke Zaehle\affil{6},
 Gordon Bonan\affil{1},
@@
 \affiliation{1}{National Center for Atmospheric Research, Table Mesa Drive, Boulder, Colorado, USA}
+
 \affiliation{2}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
+
 \affiliation{3}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
+
 \affiliation{4}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
+
 \affiliation{5}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
-\affiliation{6}{Max Planck Institute for Biogeochemistry, Jena, Germany}
+
 \correspondingauthor{Rosie Fisher}{rfisher@ucar.edu}
@@
 \begin{keypoints}
-\item A suite of modifications to the Nitrogen cycle representation were added to the Community Land Model (version 5).
-\item New features of the N model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of Nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios, dependant on N costs and distance from the target C:N ratio, and dynamic leaf resorption of Nitrogen.
+
+\item A suite of modifications to the Nitrogen cycle representation were added to the Community Land Model, version 5 (CLM5).
+
+\item New features of the N model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of Nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios and dynamic leaf resorption of Nitrogen.
+
 \item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycle of Carbon and Nitrogen.
-We assess the dependence of the model state to parameter uncertainty, and also the responses of the system to higher CO2, Nitrogen and temperature.
-\item The model responses to CO2 and N fertilization are linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment.
-Responses to temperature depend upon x,y,z.
-\item The model runs are conducted at representative sites in temperate, tropical and boreal systems.
+We assess the dependence of the both the model state, and also the responses of the system to higher CO2, Nitrogen and temperature to parameter uncertainty.
+Model runs are conducted at representative individual sites in temperate, tropical and boreal systems.
+
+\item The model responses to CO2 and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment.
+
+\item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.
 
@@
 
-\section{to do}
-1.	Write up developments of N cycle. LUNA, FUN, FLEXCN. 
-2.	Develop sensitivity methods
-t
-
-\begin{abstract}
-= enter abstract here =
-\end{abstract}
-
 \section{Introduction}
 
-Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models.
-In the CMIP5 analysis, only one model, the Community Land Model (v4.0) included an active nitrogen (N) cycle, and this model suggested that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}, ).
-This result raises the question of whether the inclusion of Nitrogen cycling necessarily generates such substantial limitations controls on the fertilization response of the biosphere.
+Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}).
+In the CMIP5 multi-model intercomparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle.
+The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}).
+Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}).
+Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehledalmonech2011}).
 
-Substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models.
-Unlike representations of, say, photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.
+Unlike representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.
+Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{zaehle2014}).
+Few, if any studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model.
 
-Subsequent to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{zaehle2010},  \cite{goll2012}, \cite{smith2014}, \cite{wang2007}).
-Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the theoretical framework for and parameterization of nutrient cycling processes (\cite{zaehle2014}).
-
-The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}.
+The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}).
 Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release.
@@
 The CLM5 model includes numerous major update of the nitrogen cycling representation subsequent to CLM4.0 and CLM4.5.
-The introduction of these new dynamic features into the CLM5 model has implications for parameter uncertainty and responses to environmental forcing.
+The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing.
 Wieder et al.
-(in prep) report the global-scale implications of the full model update.
+(in prep) report the global-scale implications of the full model update and note that the CLM( responds more strongly to CO2 and less strongly to N fertilization than its predecessors.
 In this paper, we describe the new components of the biogeochemical cycling of nitrogen, their impacts upon the model system state and responses to environmental forcing.
@@
 \subsection{Model Updates}
-The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{Thornton}, modifications to the biogeochemical cycling included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).
-The new model contains updates to three different elements of the Nitrogen cycle, including:
-1)	The LUNA module, (Leaf Utilization of Nitrogen for Assimilation), described by \cite{xu2012} and \cite{ali2016}. 
-2)	The FUN module (Fixation and Uptake of Nitrogen) described by \cite{fisher2010fun} \cite{brzostek2014} and \cite{shi2016}. 
-3)	Part of the `FLEXCN' N cycle implementation, described by \cite{ghimire2016}.
+
+The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).
+The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\
+
+1) The `LUNA' model (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}) \\
+
+2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics of Nitrogen acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}).\\
+
+3) The `FLEXCN' N cycle implementation, which is used here primarily to allow variation in tissue C:N ratios, and is adapted from \cite{ghimire2016}.\\
+
+More details on each of these new components is given below.
+A full technical description of the CLM5 is an appendix to \cite{lawrence2018} and is also available online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html).}
 
 \subsubsection{LUNA}
-The LUNA model implements the prediction of $V_{c,max}$ and $J_{max}$ from environmental conditions.
-The model determines the rate constants that are consistant with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
-In this fashion, $V_{c,max}$ is largely independent of PFT, and is primarily controlled by the N per unit leaf area ($N_{area}$) and environmental conditions.
-The capacity of the LUNA model to represent appropriate responses to temperature and CO2 was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
+
+Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
+There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}) and changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
+Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
+
+The LUNA model predicts the optimal balance of of $V_{c,max}$ and $J_{max}$ for the prevailing time-averaged environmental conditions.
+The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
+In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area ($N_{area}$, which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and environmental conditions.
+The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
 
 \subsubsection{FUN}
-The FUN model operates on the principle that for the primary sources of Nitrogen acquisition by plants (active uptake, retranslocation and symbiotic N fixation) there is a cost in terms of carbon.
-Further, the cost of each of these uptake pathways is variable in time and space, and thus plants will likely take up Nitrogen from pathways which are 'cheaper' to them.
-Where N is scarce in the environment, carbon that would otherwise have been used for growth can be deployed to acquire more N.
-The central element of FUN is a simulataneous equation which assumes, firstly, that the carbon available for growth ($C_{growth}$) is the total carbon available ($C_{avail}$), minus that used for Nitrogen uptake, $C_{nuptake}$:
+
+Under circumstances where insufficient N exists to match all of the carbon assimilated for a given C:N ratio, the CLM4.0 model reduced the gross photosynthetic flux down to the level at which growth could be supported by the assimilated N.
+This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).
+
+One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}).
+The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon.
+Further, the cost of each of these uptake pathways is variable in both time and space.
+FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them.
+Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N.
+The central element of FUN is a simultaneous equation which assumes, firstly, that ($C_{avail}$), the total carbon available (after maintenance respiration and replenishment of stores are accounted for) is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).
 
 \begin{equation}
-C_{growth}=C_{avail}-C_{nuptake}
+C_{growth}=C_{avail}+C_{nuptake}
 \end{equation}
 
-Further, the Nitrogen acquired from the environment must equal that which is deployed in the growing plant tissues.
+Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal that which is deployed in the growing plant tissues.
 The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$) which applies only to tissues that are constructed:
@@
 \begin{equation}
-C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
+N_{uptake}=C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
 \end{equation}
@@
 Thus the Carbon that is expended on N uptake is determined by:
+
 \begin{equation}
@@
 
+A further issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}).
+The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{wieder2015} illustrate the impacts of uncertainty in this function) but FUN considers fixation rates to be an emergent property of plant assimilation rates (providing carbon for fixation), the relative costs of environmental N acquisition, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
+Fixation is preferred when the costs of N acquisition are higher than that of fixation.
+Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground, in productive areas, N fixation is preferred and soil N builds up much quicker.
+In low productivity areas, carbon is required to fix N from a 'col start' initial state, impacting the survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.
+
 \subsubsection{FLEXCN}
-\cite{ghimire2016} implemented a suite of modifications to the Nitrogen cycling including variable tissue C:N ratio, prognostic Vcmax (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
-In CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen.
-In this new algorithm, Nitrogen supply is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates.
-This new allocation code we term the FLEXCN module, and it is documented in ***.
+
+Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}).
+For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage, but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
+For this reason, we introduce a flexible tissue C:N ratio into the CLM5.
+
+\cite{ghimire2016} implemented a suite of modifications to the CLM Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
+In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen (although the alternative schemes implemented by \cite{ghimire2016} are available as options).
+In this new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates.
+This new allocation code we term the FLEXCN module.
 
 \subsubsection{Adaptation of FUN to flexible CN ratios}
+
 The FUN model was originally conceived with a static tissue CN ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target CN ratio.
@@
 When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high.
+This new algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.
 
-\subsubsection{Justification for model developments}
-Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability  emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}).
-For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage,  but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
-For this reason, we introduce a flexible tissue C:N ratio into the CLM5.
+\subsection{Content of this paper}
 
-Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
-There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization \cite{ainsworth2007} and changing temperature \cite{hikosaka2005}, irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
-Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
+In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters. We use\\
 
-\subsubsection{Content of this paper}
+1. A range of single site experiments \\
+
+2. One-at-a-time and global parameter uncertainty analyses\\
+
+3. CO$_{2}$ fertilization experiments and\\
+
+4. Nitrogen addition experiments.\\
+
+to fully investigate the performance of the model under steady-state and step-change modifications in forcing.
+Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect overall behaviour in larger-scale experiments, including the controls on global carbon fertilization `$\beta$' values.
+We also hope to highlight areas of process and parameter uncertainty that appear to require greater attention in future model development and testing cycles.
+
+Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible.
+We further investigate the consequences of turning off the LUNA model and returning to the fixed leaf N/$V_{c,max}$ relationship from \cite{ghimire2016}.
 
@@
 
-Four 
-We ran the default model configuration to equilibrium, and
+\subsection{Sites}
+
+For this experiment we utilized six different experimental sites.
+Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest).
+
+\subsection{Simulation Protocol}
+
+We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (where the turnover of slower cycling carbon and nitrogen pools is accelerated, see (/cite{lawrence}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level meteorological inputs.
+400 years was sufficient to bring all carbon and nitrogen pools into equilibrium.
+Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years at each site, with accelerated mode turned off.
+Using the state at the end of the perturbed parameter spin-ups, we ran transient simulations for each site and parameter combination starting in 1850 and invoking transient CO$_{2}$ and Nitrogen deposition until the end of the observed meteorological forcing.
+We ran additional CO$_{2}$ and Nitrogen deposition experiments by increasing the CO$_{2}$ concentration to 550ppm.
+For the Oak Ridge FACE site, we only increased CO$_{2}$ in the growing season, in line with the experimental protocol applied there and using the observed CO$_{2}$ time-series.
+For all other sites, we increased CO$_{2}$ to 550ppm year round starting in 1997**.
+For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1997, so that the results can be compared with those reported by Wieder et al.
+(in prep).
+
+\subsection{Perturbed Physics Ensemble (PPE)}
+
+The CLM5 has a very large number of parameters to which the C and N cycles are potentially sensitive.
+Here we focused on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
+We choose a larger set of parameters for one-at-a-time analyses, followed by a smaller set of these to which we subject the model to global sensitivity tests.
+Knowledge of model structure as well as previous investigations of the CLM model informs our choice of parameters.
+Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model.
+However, we also report responses of the model state to perturbations.
+Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al.
+in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here.
+
+In the first instance, we conduct one-at-a-time (OAAT) perturbations of the parameters around a mean state, to allow a straightforward visualization and interpretation of the results.
+For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation).
+
+\subsubsection{Parameter Choices}
+
+The parameters included in our sensitivity analysis (Table \ref{table_parameters}) are: \\
+
+1) \textbf{Specific Leaf Area}, \emph{`SLA'}, which is a critical parameter determining the area of canopy (in m$^{2}$) derived from one gram of leaf biomass.
+This parameter is known to be one of the primary elements of the leaf economics spectrum (\cite{wright2004}), and is well-characterized by the TRY database (\cite{kattge2011}).\\
+
+2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, the ratio of fine root biomass to leaf area.
+Unlike in previous versions of the CLM, fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth.
+\\
+
+3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, reflects the balance between accumulation of stem and woody biomass, and leaf area.
+Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions.\\
+
+4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.
+These six parameters were previously defined by \cite{brzostek2014}.
+Here and in CLM5 in general, we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}.
+These parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
+\\
+
+5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`'fracfixers'}.
+This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.
+\\
+
+6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{grperc}, was revised substantially in CLM5 based on Atkin O.
+(pers comm.) from 0.3 to 0.11.
+Here we test the impacts of this change on overall carbon budgets.\\
+
+7) \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}.
+CLM5 uses the concept of a 'target' leafcn ratio, around which flexibility is allowed, given the environmental costs of Nitrogen.
+LeafCN ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$) \\
+
+8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}.
+This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, assessed in this set of experiments.
+\\
+
+9) \textbf{Respiration intercept}, \emph{lmr\_intercept}.
+The new leaf maintenance respiration model, proposed by \cite{atkin2015} and adopted by CLM5, has it's intercept as a PFT dependant parameter, with variation between PFTs reported by \cite{atkin2015} and used herein.
+Here we test the impacts of the uncertainty in the observations of this parameter.
+\\
+
+10) \textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is another ecosystem-level property defined by the FUN model and allowing the proportions of the fungal symbiotic pathways to be modified.
+Default parameters are as provided by \cite{shi2016}.
+This capacity was not fully explored in the CLM5 development process and so here we assess its importance.
+\\
+
+11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in Appendix 1, which determine the response of tissue CN ratios to depletion of N in both the environment and the plant tissues themselves.
+
+\subsubsection{Parameters not used}
+
+We conducted a set of tests on the parameters of the denitrification model, to assess the impacts of differing rates of N loss on the system.
+However, recent analysis has shown that the rates of overall denitrification in CLM5 are very low, overall (compared to plant uptake and immobilization \textbf{I've forgotten why this is?
+Can someone remind me}) and hence these parameters had very little impact.
+We did not conduct tests on leaf lifespan, since for evergreen trees in CLM5, leaf lifespan and specific leaf area have very similar impacts on overall achievable leaf area, and for deciduous trees, leaf lifespan is primarily determined by phenological triggers.
+We did not investigate any parameters relating to the crop model (Lombardozzi et al.
+in prep) nor to plant harvest, fire, dust, soil biogeochemical cycling (the latter being is largely unaltered from CLM4.5, \cite{koven2013} apart from modifications presented by \cite{lawrence}.
+
+\subsubsection{Parameter Ranges}
+
+\emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY database, and their distributions are log-normally distributed.
+Thus, we take the average log-normal standard deviation range from TRY and apply it across all PFTs.
+Thus the lower bounds are closer to the mean than the higher bounds (Table \ref{table_ranges}).
+The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across the whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake).
+\emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM value (0.3).
+\emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}.
+The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood, thus we used ranges that span the response limits of these parameters, particularly for \emph{cn\_flex\_c} which we explored across four orders of magnitude to investigate the impacts of leafCN flexibility.
+The allometric parameters ***.
+Nitrogen parameters.
+
+\subsection{Analysis}\\
+We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the fertilized simulations for the years 2001-2003.
+For some parameter settings, the constraint on growth placed the system in a state close to death.
+Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in how parameters are linked with responses to environmental forcing.
+Therefore, for those parameter combinations which cause LAI to be less than 80\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plots, so as to focus on the responses of parameter combinations that give realistic initial vegetation conditions.
 
@@
 
+The analysis conducted involved 6 sites, with 13 parameters perturbed to 4 different values, the each of those with an increase CO2, N and control run.
+Including the default simulations, this is 936 different run configurations.
+In this analysis, we look at the parameteric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon) first for their impacts on the system state and then on the responses to carbon and nitrogen limitation.
+
+\subsection{Parametric control over system state}
+Figure *** shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Caxiuana site.
+Analagous figures for the other five sites are in the SI figures 1-5 ***.
+
+One notable feature of the sensitivity analysis here is the difference in the dominant parameter responses between different output variables, particularly as one moves along the C allocation cascade from GPP, to NPP and then LAI.
+Leaf CN ratio, (\emph{leafcn}) for example, is a dominant control over GPP, but is much less important for NPP, where the respiration intercept ($lmr\_{intercept}$), fine root ratio, and the fraction of fixers (by means of the carbon expenditure on respiration) impart a greater impact.
+For leaf area index (LAI), the allocation parameters, specific leaf area, root:leaf and stem:leaf ratio, have a much greater import, than almost all the parameters that affect NPP, given the linear and fixed allocation scheme.
+We here discuss the causality of how each individual parameter affects the major output variables, and how these responses relate to model structure, particularly in cases where the response outputs are not intuitive.
+
+Impact of \emph{slatop}\\
+At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, (Figure **c) and on leaf N concentrations (Figure **e), since it is used directly in the calculation of both of these quantities.
+At most sites, however, this change in LAI is not translated into increases in GPP and NPP.
+For GPP, the response to SLA is often negative, despite huge increases in LAI, since increasing SLA (thinner leaves) decreases the total N per unit area of leaf, and so reduces the photosynthetic capacity, in particular of the leaf area that is in the sunlight.
+Thus, reductions in sunlit V$_{c,max}$ and J$_{max}$ can contribute to this outcome.
+
+Further, increasing LAI, in CLM5, can generate an under-storey of highly shaded leaves, with the possibility of those leaves being in negative carbon balance,  SLA driven increases in LAI can in principle also reduce NPP for respiratory reasons.
+
+Impact of \emph{froot\_leaf}\\
+Low values of root:leaf ratio (\emph{froot\_leaf}) tended to increase NPP, presumably driven by lower respiratory costs.
+LAI also increases, which can be due to both lower maintenance costs and turnover demand for roots leaving more carbon for leaves.
+GPP does not increase at higher LAI sites, reflecting different amounts of leaf area saturation.
+
+Impact of \emph{stem\_leaf}\\
+Decreases in stem:leaf ratio (stem\_leaf) below their default condition typically resulted in unrealistically high LAI.
+They also caused a decline in leaf N per unit area (figure **e) , and an increase in photosynthetic capacity (not shown).
+Litter with a high C content (woody material) has the capacity to “lock up” large amounts of soil Nitrogen in soil decomposing pools, and to require large fluxes of mineral N for decomposition.
+Thus, removal of this woody matter appears to alleviate some N limitation, and subsequently increasing leaf N content and thus photosynthetic capacity.
+In contrast to the root:leaf ratio response, GPP is strongly increased, apparently driven primarily by changes in N limitation and not
+
+This response is not demonstrated by the root:leaf ratio, since it has a similar C:N ratio as the leaf tissue.
+There are therefore large increases in GPP associated with low stem:leaf ratios, with smaller increases in NPP, given the higher respiratory load of the extra Nitrogen.
+
+Impact of \emph{Ncosts}\\
+Somewhat more complex responses include that of leaf N costs, whereby the carbon allocated to N uptake (figure 1d ***) increases when N costs reach the lowest values assessed here, as well as (intuitively) when they go up.
+This issue is traceable to a feature in the solutions FUN which 'over' allocates carbon to active uptake, given the lower costs of this pathway, which then cannot be used if stocks are too low.
+Plants are known to release carbon as exudates into soils to stimulate the uptake of nutrients, hence this 'wasted' carbon may reflect a plausible real-life process, but the relationship between the representation here and real-world optimization of this process presents a need for further research.
+
+Impact of \emph{fracfixers}\\
+\emph{Fracfixers} has a relatively moderate impact on the overall system state.
+NPP is moderately decreased with a lower fraction of N fixing capacity, resulting in higher N\_uptake expenditure, very slight increases in LAI and very slight response of GPP to that change.
+There are no response of CN ratio.
+These systems are only subject to transiently ramping CO2, and are thus closer to equilibrium than the step-change fertilizations, and this suggests that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.
+
+Impact of \emph{leafcn}\\
+Changing the target leaf C:N ratio \emph{leafcn} over one standard deviation from the mean TRY values had substantial impacts on all the output variables.
+In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that productivity was almost zero at Oak Ridge.
+Low C:N ratios (high leaf N content) uniformly increased GPP, with somewhat dampened impacts on NPP, reflecting decreasing carbon use efficiency, and further dampened impacts on LAI, reflecting increased costs of N acquisition with increasing N demand.
+No optimum value for leafCN ratio was found, in contrast with the standard expectation that over-investment in tissue N might at some point result in declining NPP, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
+We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations.
+
+Impact of \emph{grperc}.\\
+The response to modifications in the growth respiration coefficient in line with first-order expectations, where the higher numbers (which correspond to the previously used value of 0.33 in CLM4.5) resulted in substantial NPP reductions of ~30\% and corresponding changes in LAI.
+GPP was typically unaffected.
+
+Impact of \emph{medlyn\_slope}\\
+As expected, the change in the stomatal efficiency parameter of the Medlyn 2001 stomatal conductance model impacts GPP primarily In contrast to most of other parameters here, there were some instances where a clear optimum condition was found.
+Too-low values of this parameter results in not enough stomatal opening and photosynthesis, but too-high values result in to much water loss.
+This is illustrated for the two tropical sites in figure **a and figure **a.
+There was a marked difference in responses between sites, reflecting different humidity regimes.
+Values which caused a large decline in photosynthesis at Oak ridge, for example, did not have a large impact in the tropics, indicating that too much stomatal closure was more prohibitive there.
+The large response to \emph{leafcn} at this site also suggests that there is significant carboxylation limitation, potentially on account of lower temperatures.
+At this site, greater stomatal opening always caused more photosynthesis, indicating a lack of impact of hydraulic limitation (as inferred from the decline at high values at the tropical sites, where high atmospheric demand can cause low water potentials and stomatal closure even in high rainfall conditions [Kennedy et al.
+in prep].
+
+Impact of \emph{lmr\_intercept}\\
+The variance in the reported intercepts in \cite{atkin2016} have important consequences for ecosystem productivity, and deserve further investigation.
+These upper ranges of this parameter were among the some of the lowest NPP and LAI simulations for all sites, which, given that the range here is taken from published values, suggests a genuinely high range of uncertainty associated with the leaf maintenance respiration algorithm.
+Of course, the values provided by Atkin et al.
+were for individual PFTs, so the full range of uncertainty probed here is not necessarily relevant to each PFT.
+For example, the higher values of the parameter pertain to grasses, and thus are likely outside of the range of plausible values for the tree PFTs used here.
+
+Impact of \emph{frac\_ecto\_fungi}\\
+The fraction of ectomycorhizzal vs arbuscular fungi had limited impact here, suggesting that the parameterization of the varying costs between these uptake pathways did not have a large impact on N uptake.
+This is somewhat surprising, and suggests that this difference is overridden by other parts of the dynamics parameterized in FUN.
+
+Impact of \emph{cn\_flex\_a}\\
+\emph{cn\_flex\_a} was positively associated with leaf N content, and hence GPP, and NPP and LAI.
+This is in line with expectations, since this term is used as the intercept; above which an increase in the cost of N will reduce expenditure on N uptake.
+(see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details.
+Thus, low values of \emph{cn\_flex\_a}  results in more common reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.
+The threshold response to this parameter (only the very highest values have an impact) results from a lower bound (0.50) on the degree to which N cost can reduce uptake, suggesting that the sites here were operating in the region of this limit.
+
+Impact of \emph{cn\_flex\_b}\\
+\emph{cn\_flex\_b} is slightly negatively associated with leaf N concentration, GPP;, NPP and LAI, but positively associated with NPP spent on carbon uptake.
+This is a somewhat confusing outcome, since, in principle the responses of \emph{cn\_flex\_a} and \emph{cn\_flex\_b} should be the same way around, as increases in both increase the amount of carbon spent on N uptake.
+There is some indication that at low values of this parameter, C allocated to N uptake might end up being too high, thus creating some of the 'wasted' carbon expenditure described in the ( see Ncosts section), but this potentially merits investigation (notwithstanding the point the this algorithm is a placeholder for a more evidence-based depiction of stoichiometric flexibility).
+
+Impact of \emph{cn\_flex\_c}\\
+\emph{cn\_flex\_c} has the greatest impact of the three CN flexibility parameters.
+Notably, high values of this calibration parameter allow significant increases in leaf N content, which are, counterintuituively, associated with decreases in expenditure on N uptake, but little chqnge in LAI, implying increases in N uptake for less cost.
+Potentially, decreasing the C allocated to N uptake in a given timestep (when N is cheap) might be in some cases avoiding 'wasted' carbon.
+Further investigation of this phenomenon would be merited.
+
+\subsection{Parametric control over CO$_{2}$ and Ndep responses}
+To illustrate the relationships between N and CO$_{2}$ fertilization, and to reduce the dimensionality of the output, figures ** to ** illustrate the response for GPP (figure **), NPP (figure **), LAI (figure **) and vegetation carbon (figure **) for five sites each.
+Here we summarize the impact of each parameter on the output patterns.
+
+Impact of \emph{slatop}\\
+Despite it’s very large impact on the system state, via modifications to the LAI predictions, specific leaf area has limited impact on the overall response to fertilization, compared to other parameters, and the responses to both N and CO$_{2}$ fertilization differ between sites, potentially reflecting the different limiting conditions at each site.
+
+Impact of \emph{froot\_leaf}\\
+\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values, which were generally associated with limited response to N fertilization.
+This might reflect either a lack of demand for N, or an inability to respond to the fertilization itself through lack of root matter.
+
+Impact of \emph{stem\_leaf}\\
+Low values of the stem/leaf ratio also caused very limited responses to N fertilization at some sites, and CO$_{2}$ fertilization at others.
+These very low values were associated with very high LAI numbers, which potentially were already subject to light limitation; so as not to profit too much from the addition of extra N or CO$_{2}$.
+**think about the LAI vs fertilization response figure**
+
+Impact of \emph{N\_costs}\\
+As described previously, too-low N costs in the environment tend to increase the amount of carbon wasted during N uptake (on the attempted acquisition of N which is not there).
+The model calibration at present appears to be in an interim point, but **we probably need to reduce the bounds of the sensitivity analysis to make this all a bit less shocking**
+
+Impact of \emph{frac\_fixers}\\
+The fraction of N fixers in the system has a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
+In line with expectations, for GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$.
+For NPP, this result is muted for the two (less N limited) tropical sites, particularly Caxiuaxa (Figure **) because the increase in shaded leaves and in tissue C:N ratios (Figure **) with fertilization increases the respiratory load on these systems.
+Note that NPP reported here is the total C available for growth and N uptake, and therefore a lower NPP (as in figure **a) can still be reconciled with a higher LAI (figure **a).
+
+In CLM5, each PFT except crops has the same N fixation fraction (0.25), but this analysis clearly suggests that much greater emphasis should be afforded to the dynamics of changing N fixation capacity under transient atmospheric conditions.
+
+One interesting feature of this response is that, despite not exhibiting much difference in N limitation in the present day (given the muted response of the system state to changes in \emph{frac\_fixers}), simulations with low (or no) fixer capacity had typically very high responses to N fertilization.
+** this needs an explanation from looking at the N response plots on their own...**
+
+Impact of \emph{leafcn}\\
+ High target C:N ratio simulations exhibited a degree of N limitation in the control condition, which was strongly alleviated by N fertilization.
+Similarly, these scenarios had lower responses to CO$_{2}$, since in many cases, insufficient N was available to expand the canopy, leading to declines in leaf N concentration and a less strong (but still positive) fertilization response.
+
+Impact of \emph{grperc}\\
+Of all the variables tested, \emph{grperc} had one of the lowest impacts on fertilization responses, since the degree to which growth is 'taxed' is not responsive to any of the physiological processes upstream that depend on fertilization.
+
+Impact of \emph{medlyn\_slope}\\
+For many sites, the response of the system to CO$_{2}$ fertilization with low values of the \emph{medlyn\_slope} parameter was higher than the cut-off imposed for our visualization (>80\% of control LAI).
+For example figure si** indicates that the lowest values had a x4 increase in GPP under double CO2 scenarios with the lowest values of this parameter, illustrative of very low leaf internal CO$_{2}$ concentrations where stomatal opening is heavily restricted.
+
+Impact of \emph{lmr\_intercept}\\
+Like \emph{grperc}, the respiration intercept has only a straightforward linear impact on the model state, and while the very high values of this parameter were in some cases completely prohibitive of growth, they do not impact sensitivity to fertilization in any notable fashion.
+
+Impact of \emph{frac\_ectomy\_fungi}\\
+In line with its limited impact on the system state, \emph{frac\_ectomy\_fungi} had no discernible impact on fertilization responses.
+
+Impact of \emph{cn\_flex\_a},  \emph{cn\_flex\_b} and \emph{cn\_flex\_c}\\
+**These will all need looking at once a more careful set of runs are done**
+Low values of \emph{cn\_flex\_b}, where the leafCN concentration can increase in response to fertilization, showed very high responses to n fertilization, and low responses to CO2 fertilization (where N content of leaves declined).
+
 \section{Discussion}
+The responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).
+In all of these activities, the carbon cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
+At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes, without necessarily changing the control state to a large extent.
+As expected, parameters the relate most strongly to the acquisition \emph{frac\_fixers}, \emph{Ncosts} and use \emph{leafcn}, \emph{cn\_flex\_b}, have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
+
+\subsection{Limitations of step-change experiments}
+Here we propose to investigate the parameter space of the model using step-change experiments.
+However, such experiments necessarily involve unrealistically rapid changes in environmental boundary condition, and cannot necessarily be considered as analagous to either real-world or transient future scenario conditions.
+We assessed the response to the ramping historical CO2, to illustrate how parameter responses differed between the rapid and slow responses.
+The results are in figure SI **
+
+\subsection{Limitations from single sites}
+The analysis presented here deliberately does not focus on the relative fit of the model to the data collected at the individual sites.
+The fit at individual sites of the globally adjusted model cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, which is beyond the scope of this paper.
+Instead, we chose a set of individual sites, rather than global analysis, to allow for the 936 model integrations that comprise our sensitivity tests, and to probe a set of alternative vegetation and climate conditions.
+This study is incomplete, as are all investigations of the climatic and parameteric space of land surface models.
+
+\subsection{Discussions of parameterization issues}
+During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
+We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces,  3) the tendency of the modeled ecosystem to die under past low-CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting pattern that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the different data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
+Thus, the parameterization used in the default version of the model is a less objective calibration than is ideal, and a limited set of parameters were tuned 'by hand' to avoid excessively spurious values for some model states.
+Given the sensitivity of the fertilization responses in the CLM5, we advocate careful interpretation of the outcome of the model, particularly with regard to the climate feedback strength.
+
+\subsection{How to reduce model uncertainty}
+Many model development efforts are motivated under the auspices of 'reducing uncertainty'.
+Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 are partly motivated by the need to move the model behaviour closer to several key observable features of the system.
+Notably, the dynamics of leaf Nitrogen content with respect to environmental scarcity (\cite{zaehle2014}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen.
+The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.
+We argue that all of these development generate model emergent behaviour that is closer in essence to reality.
+However, in doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantititive understanding.
+
+The development of the CLM5 also slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
+In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable, as well as accelerate the rejection of model structures that are unrealistic.
+Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
+
+This implies that a critical control in the CLM, and in coupled simulations with the CESM for CMIP are dependant on the representation of Nitrogen fixations and its dynamics.
 
 \subsubsection{Discourse on how target CN ratio could be optimized but isn't}
+
 \cite{thomas2014}, \cite{friedlingstein1999}, \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{anten2011}, \cite{vanwijk2003}
 
-\subsubsection{discuss geographical spread of predictions for N and CO2 fertilizaion}
+\subsubsection{Discuss geographical spread of predictions for N and CO2 fertilization}
+
 \cite{hickler2008}
@@
 \subsubsection{Discourse on whether nitrogen fixation is realistic in CLM5}
-The 'fraction of fixers' ($frac\_fixers$) is a proxy for the community level composition of Nitrogen fixing plants.
+
+The 'fraction of fixers' (\emph{frac\_fixers}) is a proxy for the community level composition of Nitrogen fixing plants.
 Technically, because plant density is not represented in the CLM5, this parameter really represents the fraction of net assimilated carbon ($gpp$ - $R_{auto}$) that can be used for symbiotic N fixation.
@@
 
-\section{Figures.}
+\pagebreak
+\begin{table}
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| } 
+ \hline
+ Parameter Name & Meaning & Units & Min & Max \\
+  \hline
+ slatop & Specific Leaf Area & m2/g \\ 
+ froot\_leaf & Fine root:leaf ratio. & g/g \\
+ stem\_leaf  & Stem:leaf ratio. & g/g \\ 
+ n\_costs    & 6 parameters related to Nitrogen costs & g/g \\
+ fracfixers  & Fraction of C availble for fixation & - \\
+  grperc  & Fraction of C used for growth respiration  & - \\
+  leafcn  & Leaf C:N ratio. & g/g \\
+     medlyn\_slope  & Slope of stomatal conductance curve & g/g \\
+      lmr\_intercept & intercept of leaf N /respiration relationship &\\
+      frac\_ecto & Fraction of active uptake that is ectohycorrhizal & \\
+      cn\_flex\_a & Parameter of N flexibilty algorithm& \\
+      cn\_flex\_b & Parameter of N flexibilty algorithm & \\
+      cn\_flex\_c & Parameter of N flexibilty algorithm & \\
+\hline
+\end{tabular}
+\end{center}
+\caption{CLM5 Parameters subject to sensitivity analysis in this study. }
+\label{table_parameters}
+\end{table}
+
+ \begin{table}
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| } 
+ \hline
+ Parameter Name & s1 &s2 & s3 & s4\\
+  \hline
+ slatop & 0.6761 & 0.8769 &1.2785 &1.4791\\ 
+ froot\_leaf &  0.342 &0.671 &1.05 & 1.1\\
+ stem\_leaf  &  0.3057 &0.61 &1 &1\\ 
+ n\_costs    & 10$^{-2}$ &10$^{-2}$&  10$^{1}$& 10$^{2}$\\
+ fracfixers  & 0 &0.5 & 2 & 4 \\
+  grperc  &  0 & 0.5& 2 & 3\\
+  leafcn  &0.7413 & 0.8932 & 1.1970  & 1.349\\
+
+     medlyn\_slope  &0.5258 & 0.7191 & 1.1057  &1.29\\
+      lmr\_intercept & 0.8539 & 0.9531 & 1.1512& 1.25\\
+      frac\_ecto & 0 &0.5& 1 & 1 \\
+      cn\_flex\_a &0 &0.5 & 2  &4\\
+      cn\_flex\_b & 0.1 & 0.5 & 2  & 4\\
+      cn\_flex\_c &0.001 & 0.1 & 10 & 100\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Multipliers on the original values in CLM5 used in this sensitivity analysis .}
+\label{table_ranges}
+\end{table}
+
+\section{Figures}
 
 Figure 1: Sensitivity analysis of state of CLM at site to N parameters @ORNL
+
 Figure 2: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @ORNL
+
 Figure 3: Sensitivity analysis of state of CLM at site to N parameters @BCI
+
 Figure 4: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @BCI
+
 Figure 5: Sensitivity analysis of state of CLM at site to N parameters @CAX
+
 Figure 6: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @CAX
+
 Figure 7: Operation of the flexCN ratio module.
+
 Figure 8: Operation of the leaf N retranslocation module
+
 Figure 9: Details on the N uptake in FUN
+
 Figure 10: Some sort of Will figure.
+
+\section{To Do}
+Add data from the CO2 experiment to figures.
+
+Add a figure of the transient responses through time to CO2 and N...
+
+Look at the transient response to ramping CO2 and see if its different to the step response
 
@@
 \nocite{*}
+
 \bibliography{aguCLM5_bibtex}
 
-\section{Appendix 1: Technical Documentation of Nitrogen Cycle}
-\subsubsection{Modifications to allow variable C:N ratio}
-The total cost of N uptake is calculated based on the assumption that carbon is partitioned to each stream in proportion to the inverse of the cost of uptake.
-So, more expensive pathways receive less carbon.
-Earlier versions of FUN (Fisher et al., 2010)) utilized a scheme whereby plants only took up N from the cheapest pathway.
-Brzostek et al.
-(2014) introduced a scheme for the simultaneous uptake from different pathways.
-Here we calcualate a ‘conductance’ to N uptake (analagous to the inverse of the cost function conceptualized as a resistance term) N_{conductance} ( gN/gC) as:
-
-N_{conductance,f}=  \sum{(1/N_{cost,x})}
-
-From this, we then calculate the fraction of the carbon allocated to each pathway as
-
-C_{frac,x} = \frac{1/N_{cost,x}}{N_{conductance}}
-
-These fractions are used later, to calculate the carbon expended on different uptake pathways. Next, the N acquired from each uptake stream per unit C spent (N_{exch,x}, gN/gC) is determined as
-
-N_{exch,x} = \frac{C_{frac,x}}{N_{cost,x}}
-
-We then determine the total amount of N uptake per unit C spent (N_{exch,tot}, gN/gC) as the sum of all the uptake streams.
-
-N_{exch,tot} = \sum{N_{exch,x}}
-
-and thus the subsequent overall N cost is
-
-N_{cost,tot} = 1/{N_{exch,tot}}
-
-Retranslocation is determined via a different set of mechanisms, once the N_{cost,tot} is known
-
 \end{document}

### 2018-10-29 main.tex
Updates from ShareLaTeX

@@
 \item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.
+
+\item This implies that a critical control in the CLM, and in coupled simulations with the CESM for CMIP are dependant on the representation of Nitrogen fixations and its dynamics.
 
@@
 \section{Discussion}
-The responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).
-In all of these activities, the carbon cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
-At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes, without necessarily changing the control state to a large extent.
-As expected, parameters the relate most strongly to the acquisition \emph{frac\_fixers}, \emph{Ncosts} and use \emph{leafcn}, \emph{cn\_flex\_b}, have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
+The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).
+In all of these activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
+At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes.
+In particular, as expected, parameters that relate most strongly to the acquisition of N, \emph{frac\_fixers}, \emph{Ncosts} and use \emph{leafcn}, \emph{cn\_flex\_b}, have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
+It has long been understood that uncertainties in the representation of Nitrogen are large(\cite{zaehle2014}) .
+In the CMIP6 generation of land surface models, it is expected that a much larger number of model submissions will contain active representations of N limitation.
+We do not yet know how CLM5 compares to others in it's carbon cycle feedback strength, but we propose that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained.
 
-\subsection{Limitations of step-change experiments}
-Here we propose to investigate the parameter space of the model using step-change experiments.
+\subsection{Limitations}
+\subsubsection{Limitations of step-change experiments}
+Here investigate the parameter space of the model using step-change experiments.
 However, such experiments necessarily involve unrealistically rapid changes in environmental boundary condition, and cannot necessarily be considered as analagous to either real-world or transient future scenario conditions.
@@
 
-\subsection{Limitations from single sites}
+\subsubsection{Limitations from single site analyses}
 The analysis presented here deliberately does not focus on the relative fit of the model to the data collected at the individual sites.
@@
 
-\subsection{Discussions of parameterization issues}
+\subsubsection{Limitations of parameter calibration issues}
 During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
@@
 
-\subsection{How to reduce model uncertainty}
+\subsection{Future Investigations}
+\subsubsection{Temperature and water responses}
+Here we probe the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5.
+In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization.
+The full response of the model has considerably more dimensions than these two transient properties.
+In particular, here we have not considered responses to temperature and moisture changes.
+The suite of parameters than in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implictly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.
+Properties more closely linked to water availability are associated with the hydrological components of the model as well as the PHS plant hydraulics scheme and will be investigated further by [Dagon et al in prep.
+and Kennedy et al., in prep.]
+
+\subsubsection{Filtering parameter space with observations}
+The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
+In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of availble and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
+Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
+
+\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
 Many model development efforts are motivated under the auspices of 'reducing uncertainty'.
@@
 We argue that all of these development generate model emergent behaviour that is closer in essence to reality.
-However, in doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantititive understanding.
-
-The development of the CLM5 also slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
-In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable, as well as accelerate the rejection of model structures that are unrealistic.
-Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
-
-This implies that a critical control in the CLM, and in coupled simulations with the CESM for CMIP are dependant on the representation of Nitrogen fixations and its dynamics.
+However, in doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.
 
@@
 \subsubsection{Discourse on whether nitrogen fixation is realistic in CLM5}
-
 The 'fraction of fixers' (\emph{frac\_fixers}) is a proxy for the community level composition of Nitrogen fixing plants.

### 2018-10-29 workflow2.tex
Updates from ShareLaTeX

--- 2018-05-23 Workflow.tex
@@
 \begin{document}
-\title{Workflow for CLM5 paper}
-\maketitle{}
+\title{Final Workflow for CLM5 paper}
 
-Scripts for these analyses should all be found in the public github repo at:
+Scripts for these analyses should all be found in the public github repo at:\\
 {\color{blue}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS}}}\\
 
-The scripts are specific to usage on the NCAR Cheyenne machine...
+The scripts are specific to usage on the NCAR Cheyenne machine...\\
 
@@
 
-\subsection{Make new ensemble param files with high leaf allocation }
-
-Use PFT script
+Make ensemble parameter files using
 
 {\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/FACE_clm5_ens_oaat.m}}}\\
-commit b6ba94
 
 Using the basefile. \\
-{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/clm5_params.c171017_highleaf.nc}}} \\
-commit c539e89
+{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/clm5_params.c171017.nc}}} \\
 
 Output location..\\
-{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi11.nc }}} \\
-commit 54f863b
+{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_defpft_pi11.nc }}} \\
 
 Copy default file into '00'..\\
-{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi00.nc }}} \\
-commit 61517c5
+{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_defpft_pi00.nc }}} \\
 
@@
 \subsection{Make 1850 spin-up default runs from scratch}
-Make single point case from scratch for each site by running this script (based on Charlie's FATES version) three times once for each site (that could be improved on!)..:\\
+Make single point case from scratch for each site by running this script (based on Charlie's FATES version) 'n' times once for each site (that could be improved on!)..:\\
 {\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/fromscratch_CTSM_make_site_spinup_case_defaults}}}\\
-commit 38ced1
-
-The resulting cases are \\
-{\color{cyan}{\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.1x1pt_US-ORN}}}\\
-\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.bci_0.1x0.1_v4.0i}}\\
-\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.1x1pt_Br-cax}}}\\
-
-These are all run out to year 321.
-
-\subsection{Make 1850 spin-up ensembles for each site}
-
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
-commit 2bb89c5
-
-Case names are along the lines of:\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI**}}\\
-\small{\url{CLM5_1x1pt_Br-cax_ens_PI**}}\\
-\small{\url{CLM5_1x1pt_US-ORN_ens_PI*}}}\\
-
-These are all run to y201.
-
-\subsection{Make 1850 spin-up defaults for each site}
-
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_defaults}}}\\
-commit 0d2894a
-
-Case names are:\\
-{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_PI00}}}\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI00}}}\\
-{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_PI00}}}
-
-\subsection{Analyse output of spinup cases}
-
-Run analysis script on spinup ensemble output (at 1850 CO2). \\
-{\color{blue}\small{\url{/glade/p/work/rfisher/git/CLM5_FACE_EXPTS/matlab/CLM5_FACE_senstest_output_mar18.m}}}\\
-commit 7043e8b
-
-Output here\\
-{\color{magenta}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS/tree/master/matlab/figures\frac_deviation_p1CLM5_1x1pt_Br-cax_ens_PI_y1.png}}}\\
-commit 1f9a57c
-
-\section{Transients}
-
-\subsection{Make transient defaults}
-Use this script. It employs the 20tr default CO2 ramping, but with an 1850 climate. .\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_defaults}}}\\
-commit 36ced2a
-
-Case names are:\\
-{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_PI00}}}\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_PI00}}}\\
-{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_PI00}}}
-
-\subsection{Make transient ensemble}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ensembles}}}\\
-commit 7b8bca7
-
-Case names are along the lines of:\\
-{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_PI**}}}\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_PI**}}}\\
-{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_PI**}}}
-
-\subsection{Analyse transient ensemble}
-
-\section{Elevated CO$_{2}$ transient runs}
-\subsection{Make elevated CO$_{2}$ file}
-Use this script\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/modify_co2_drivers.m}}}\\
-commit 0d4fb93
-
-Which generates\\
-{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/co2_datm_ORNL_ELEV.nc}}}\\
-commit 0d4fb93
-
-\subsection{Make elevated CO$_{2}$ defaults}
-Use this script:\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults}}}\\
-commit 8e6a3a5
-
-Which uses these user$\_$nl$\_$datm files:\\
-{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_1x1pt_Br-cax}}}\\
-{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_bci_0.1x0.1_v4.0i}}}\\
-{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_1x1pt_US-ORN}}}\\
-commit 4cb99a4
-
-And this user$\_$datm.streams.txt.co2tseries.20tr file\\
-{\color{blue}\small{\url{gitpath/co2_files/user_datm.streams.txt.co2tseries.20tr}}}\\
-commit 41d4aeb
-
-Which references the previous co2 file\\
-{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/co2_datm_ORNL_ELEV.nc}}}\\
-commit 0d4fb93
-
-The case names are:\\
-{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI00}}}\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI00}}}\\
-{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI00}}}
-
-\subsection{Make elevated CO$_{2}$ ensemble}
-Use this script:\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_ensembles}}}\\
-commit ffe1b6b
-
-The case names are like:\\
-{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI**}}}\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI**}}}\\
-{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI**}}}
-
-\subsection{Analyze elevated CO2 runs}
-Use this script:\\
-{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_apr18.m}}}\\
-commit ec202a0
-
-Creates these figures:
-{\color{magenta}\small{\url{gitpath/matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}}}\\
-commit 23ce56e
-
-\section{Make runs with LUNA off}
-
-\subsection{Make ensemble spinup runs with LUNA off}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
-commit 88b6508
-
-Case names are along the lines of:\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI15*}}\\
-\small{\url{CLM5_1x1pt_Br-cax_ens_PI15*}}\\
-\small{\url{CLM5_1x1pt_US-ORN_ens_PI15*}}}\\
-
-\subsection{Make default spinup runs with LUNA off}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_defaults_lunaoff}} }\\
-commit 88b6508
-
-Case names are along the lines of:\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI00_noluna}}\\
-\small{\url{CLM5_1x1pt_Br-cax_ens_PII00_noluna}}\\
-\small{\url{CLM5_1x1pt_US-ORN_ens_PII00_noluna}}}\\
-
-\subsection{Make transient runs with LUNA off}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ensembles}} }\\
-commit acdee08
-
-The case names are like:\\
-{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI15*}}}\\
-{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI15*}}}\\
-{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI15*}}}
-
-\subsection{Make default transient runs with LUNA off}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_defaults_lunaoff}} }\\
-commit b26f487
-
-\section{Correct the CO2 fertilization to be all year}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/modify_co2_drivers_allyear.m }} }\\
-commit 723bbf3
-
-File is here: 
-{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/mco2_datm_ORNL_ELEV_allyear.nc }} }\\
-commit 8604dd9
-
-\subsection{Restart ensemble transients with new CO2 file}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_ensembles}}}\\
-commit 9f8c771
-
-\subsection{Restart default transients with new CO2 file}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults}}}\\
-commit cc6e00e
-
-\subsection{Restart LUNA defaults transients with new CO2 file}
-Use.\\
-{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults_lunaoff}}}\\
-commit cc6e00e
-
-\subsection{Analyze new elevated CO2 runs}
-Use this script:\\
-{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_apr18.m}}}\\
-commit  7eab468
-
-Creates these figures:
-{\color{magenta}\small{\url{gitpath/matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}}}\\
-commit 23ce56e
-
-\section{Add additional sites}
-
-\section{Add parameters to look at open/closed N cycle}
-
-\section{Modify range of FlexCN parameters to completely stop C:N flexibility?}
-
-\section{Investigate differences in driving data for single sites vs reanalysis}
-
-\section{Add Nitrogen fertilization experiments}
 

### 2018-11-05 matlab/CLM5_FACE_senstest_output_mar18.m
changed case scripts

@@ -1,270 +1,295 @@
-
-
-
 %%%%%%%%%%%%%%%%%%% LOOKING AT NCL FILES IN Matlab %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% force matlab to write graphics correctly.... 
+% force matlab to write graphics correctly....
 %opengl neverselect
-					   
-% If we re-run this script, we need to close all the existing figures and void the existing variables 
-mkfigs=1
 
-if(mkfigs ==1)
-close all
-clear all
-mkfigs=1
+% If we re-run this script, we need to close all the existing figures and void the existing variables
+mkfigs = 1
+
+if (mkfigs == 1)
+  close all
+  clear all
+  mkfigs = 1
 end
 figdir = 'figures'
-u=3600*24*365'
-vars         = ({'GPP','NPP','TLAI','QVEGT','TOTVEGC','BTRANMN','COST_NFIX','COST_NRETRANS','NPP_NUPTAKE','NACTIVE','NFIX','NRETRANS',})
-namevars     = ({'Gross Production (gCm^{-2}y^{-1})','Net Production (gCm^{-2}y^{-1})','Leaf Area Index (m^{2}m^{-2})','QVEGT','Total Veg Carbon (gm^{-2})' ,'BTRANMN' ,'COST NFIX','COST RETRANS' ,'N uptake cost (gCm^{-2}y^{-1})','NACTIVE','NFIX','NRETRANS'})
+u = 3600 * 24 * 365'
+vars = ({'GPP', 'NPP', 'TLAI', 'QVEGT', 'TOTVEGC', 'BTRANMN', 'COST_NFIX', 'COST_NRETRANS', 'NPP_NUPTAKE', 'NACTIVE', 'NFIX', 'NRETRANS','LEAFN'})
+namevars = ({'Gross Production (gCm^{-2}y^{-1})', 'Net Production (gCm^{-2}y^{-1})', 'Leaf Area Index (m^{2}m^{-2})', 'QVEGT', 'Total Veg Carbon (gm^{-2})', 'BTRANMN', 'COST NFIX', 'COST RETRANS', 'N uptake cost (gCm^{-2}y^{-1})', 'NACTIVE', 'NFIX', 'NRETRANS','N_{area}'})
 %maxy            = [3050  1400    6    1500    15700             1       10           10           0.4              10         10         5]
-maxy            = [2350  1100    7    1500    19000             1       10           10              0.4           10         10       5]
[…]

### 2018-11-05 SIfigs.tex
Updates from ShareLaTeX

--- 2018-10-04 figures.tex
@@
      \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_GPP.png}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_GPP1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
      \caption{GPP CO2 and N respones}
@@
 \begin{figure}[h]
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_NPP.png}
-     \caption{NPP CO2 and N respones}
-     \label{NPP CO2 and N respones}
-  \end{figure}
-
- \begin{figure}[h]
      \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_LAI.png}
-     \caption{LAI CO2 and N respones}
-     \label{LAI CO2 and N respones}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_TOTVEGC1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
+     \caption{Total vegetation carbon CO2 and N respones}
+     \label{Total vegetation carbon CO2 and N respones}
   \end{figure}

### 2018-11-06 main.tex

@@
 \section{Results}
-
 The analysis conducted involved 6 sites, with 13 parameters perturbed to 4 different values, the each of those with an increase CO2, N and control run.

### 2018-11-07 figures.tex

@@
      \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_GPP.png}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_GPP1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2001.png}
      \caption{GPP CO2 and N respones}
-     \label{GPP CO2 and N respones}
+     \label{GPP CO2 and N respones 2001}
   \end{figure}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_NPP.png}
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_GPP1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
+     \caption{GPP CO2 and N respones}
+     \label{GPP CO2 and N respones 2006}
+  \end{figure}
+
+ \begin{figure}[h]
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_NPP1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
      \caption{NPP CO2 and N respones}
@@
      \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_LAI.png}
-     \caption{LAI CO2 and N respones}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_TLAI1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2001.png}
+     \caption{LAI CO2 and N respones 2001}
      \label{LAI CO2 and N respones}
@@
 
+   \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_TLAI1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
+     \caption{LAI CO2 and N respones 2006}
+     \label{LAI CO2 and N respones}
+  \end{figure}
+
+\begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_LEAFN1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
+     \caption{Leaf Nitrogen  CO2 and N respones}
+     \label{Leaf Nitrogen CO2 and N respones}
+  \end{figure}
+
 \end{document}

### 2018-11-07 figures.tex
Updates from ShareLaTeX

--- 2018-06-29 figures.tex
@@
 \journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
+
 \begin{document}
+
 \begin{figure}[h]
-     \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_CAX__y1.png}
-     \caption{Caxiuana State Parameter Sensitivity}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1.png}
+     \caption{Caxiuana Forest State}
      \label{CAX state}
@@
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_BCI__y1.png}
-     \caption{BCI State Parameter Sensitivity}
-     \label{BCI state}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
+     \caption{Metolius Forest State}
+     \label{MET state}
+  \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1.png}
+     \caption{Niwot Forest State}
+     \label{NR1 state}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1.png}
+     \caption{Harvard Forest State}
+     \label{HVF state}
+  \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \caption{Harvard Forest Nitrogen Fertilization}
+     \label{HRV ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \caption{Harvard Forest CO2 Fertilization}
+     \label{HRV celev}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1.png}
+     \caption{Caxiuana Forest Nitrogen Fertilization}
+     \label{CAX ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1.png}
+     \caption{Caxiuana Forest CO2 Fertilization}
+     \label{CAX celev}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \caption{Metolius Forest Nitrogen Fertilization}
+     \label{MET ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1.png}
+     \caption{Metolius Forest CO2 Fertilization}
+     \label{MET celev}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1.png}
+     \caption{Niwot Forest Nitrogen Fertilization}
+     \label{NR1 ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
+     \caption{Niwot Forest CO2 Fertilization}
+     \label{NR1 celev}
   \end{figure}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_ORN__y1.png}
-     \caption{Oak Ridge State Parameter Sensitivity}
-     \label{ORN state}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
+     \caption{GPP CO2 and N respones}
+     \label{GPP CO2 and N respones 2001}
   \end{figure}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
-     \caption{Response to Nitrogen Deposition}
-     \label{Caxiuana CO2 response Parameter Sensitivity }
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_NPP1__p2012.png}
+     \caption{NPP CO2 and N respones}
+     \label{NPP CO2 and N respones 2001}
+  \end{figure}
+
+ \begin{figure}[h]
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_NPP1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
+     \caption{NPP CO2 and N respones}
+     \label{NPP CO2 and N respones}
   \end{figure}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI_y1.png}
-     \caption{Response to Nitrogen Deposition}
-     \label{Oak Ridge CO2 response Parameter Sensitivity}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TLAI1__p2012.png}
+     \caption{LAI CO2 and N respones}
+     \label{LAI CO2 and N respones 2001}
   \end{figure}
@@
      \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_CO2_response_1CLM5_BCI__y1.png}
-     \caption{Response to Nitrogen Deposition}
-     \label{BCI CO2 response Parameter Sensitivity}
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
+     \caption{VEGC CO2 and N respones}
+     \label{VEGC CO2 and N respones 2001}
+  \end{figure}
+
+   \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
+     \caption{Leaf Nitrogen CO2 and N respones}
+     \label{Leaf N CO2 and N respones 2001}
   \end{figure}

### 2018-11-09 SIfigs.tex
Updates from ShareLaTeX

--- 2018-06-29 figures.tex
@@
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=35pc]{matlab/figures/frac_deviation_p1CLM5_BCI__y1.png}
-     \caption{BCI State Parameter Sensitivity}
-     \label{BCI state}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
+     \caption{Metolius Forest State}
+     \label{MET state}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1.png}
+     \caption{Harvard Forest State}
+     \label{HVF state}
+  \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \caption{Harvard Forest Nitrogen Fertilization}
+     \label{HRV ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \caption{Harvard Forest CO2 Fertilization}
+     \label{HRV celev}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1.png}
+     \caption{Caxiuana Forest Nitrogen Fertilization}
+     \label{CAX ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1.png}
+     \caption{Caxiuana Forest CO2 Fertilization}
+     \label{CAX celev}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \caption{Metolius Forest Nitrogen Fertilization}
+     \label{MET ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1.png}
+     \caption{Metolius Forest CO2 Fertilization}
+     \label{MET celev}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1.png}
+     \caption{Niwot Forest Nitrogen Fertilization}
+     \label{NR1 ndep}
+  \end{figure}
+
+   \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
+     \caption{Niwot Forest CO2 Fertilization}
+     \label{NR1 celev}
   \end{figure}

### 2018-11-09 main.tex

--- 2018-10-29 main.tex
@@
 For some parameter settings, the constraint on growth placed the system in a state close to death.
-Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in how parameters are linked with responses to environmental forcing.
+Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in the relationship between parameters and  with  to environmental forcing.
 Therefore, for those parameter combinations which cause LAI to be less than 80\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plots, so as to focus on the responses of parameter combinations that give realistic initial vegetation conditions.
@@
 \section{Results}
-
-The analysis conducted involved 6 sites, with 13 parameters perturbed to 4 different values, the each of those with an increase CO2, N and control run.
-Including the default simulations, this is 936 different run configurations.
-In this analysis, we look at the parameteric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon) first for their impacts on the system state and then on the responses to carbon and nitrogen limitation.
+The analysis conducted involved four sites, with 13 parameters perturbed to 4 different values, each of those with an increase CO$_{2}$, N and control run.
+Including the default simulations, this is 672 different run configurations.
+In this analysis, we look at the parametric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon), first at the impacts of parameteric variation on the system state and then on the responses to carbon and nitrogen fertilization.
 
 \subsection{Parametric control over system state}
-Figure *** shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Caxiuana site.
-Analagous figures for the other five sites are in the SI figures 1-5 ***.
+Figure \ref{NWRstate} shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Niwot Ridge site.
+The other site state variation figures are presented in the Supplementary Information (Figures 1-3).
 
-One notable feature of the sensitivity analysis here is the difference in the dominant parameter responses between different output variables, particularly as one moves along the C allocation cascade from GPP, to NPP and then LAI.
-Leaf CN ratio, (\emph{leafcn}) for example, is a dominant control over GPP, but is much less important for NPP, where the respiration intercept ($lmr\_{intercept}$), fine root ratio, and the fraction of fixers (by means of the carbon expenditure on respiration) impart a greater impact.
-For leaf area index (LAI), the allocation parameters, specific leaf area, root:leaf and stem:leaf ratio, have a much greater import, than almost all the parameters that affect NPP, given the linear and fixed allocation scheme.
-We here discuss the causality of how each individual parameter affects the major output variables, and how these responses relate to model structure, particularly in cases where the response outputs are not intuitive.
+In general, we found that leaf are indicies were often low for these sites, relative to both observations at the sites, and to the magnitude of LAI where gas exchange fluxes are typically known  to saturate as a function of intercepting leaf area (4-5 m$^{2}$ m$^{-2}$.
+This is a surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}.
+A seperate set of simulations (not shown here) conducted with higher allocation to leaves (lower stem and root allocation parameters) indicated that, with the resulting higher LAI's, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAIs.
+In the default set of simulations the same parameters dominate the responses for differnt variables, given the intrinsic correlation between LAI and GPP, but for saturating LAI conditions, a greater disparity between the dominant parameters is obtained.
+
+Here we assess the impacts of each parameter on the overall output both at the Niwot Ridge site shown and the other sites in the SI figures.
 
 Impact of \emph{slatop}\\
-At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, (Figure **c) and on leaf N concentrations (Figure **e), since it is used directly in the calculation of both of these quantities.
-At most sites, however, this change in LAI is not translated into increases in GPP and NPP.
-For GPP, the response to SLA is often negative, despite huge increases in LAI, since increasing SLA (thinner leaves) decreases the total N per unit area of leaf, and so reduces the photosynthetic capacity, in particular of the leaf area that is in the sunlight.
-Thus, reductions in sunlit V$_{c,max}$ and J$_{max}$ can contribute to this outcome.
+At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI,  and on leaf N concentrations (Figure \ref{NR1 stat}), since it is used directly in the calculation of both of these quantities.
+Whether this impact on LAI is translated into an impact on GPP depends on the degree of saturated of productivity at the default LAI.
 
-Further, increasing LAI, in CLM5, can generate an under-storey of highly shaded leaves, with the possibility of those leaves being in negative carbon balance,  SLA driven increases in LAI can in principle also reduce NPP for respiratory reasons.
+For GPP, the impact of SLA is always positive, but the degree of it's impact shifts between sites, as it integrates a trade-off between an increase in LAI and the deployment of thinner leaves (lower carbon content per unit area) and subsequent decreases the total N per unit area of leaf reducing photosynthetic capacity
+
+The impact on NPP however, can be negative (e.g.
+Metolius), since the CLM5 can generate an under-storey of shaded leaves, and with it the possibility of those leaves being in negative carbon balance.
+SLA driven increases in LAI can in principle also reduce NPP for respiratory reasons.
+At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the intrisic trade-offs between costs and benefits of thinner leaves.
 
 Impact of \emph{froot\_leaf}\\
-Low values of root:leaf ratio (\emph{froot\_leaf}) tended to increase NPP, presumably driven by lower respiratory costs.
-LAI also increases, which can be due to both lower maintenance costs and turnover demand for roots leaving more carbon for leaves.
-GPP does not increase at higher LAI sites, reflecting different amounts of leaf area saturation.
+Root:leaf ratio (\emph{froot\_leaf}) a direct effect on LAI, via changes in allocation fractions to different tissues, an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, and (newly introduced in CLM5) also affects the uptake of both nitrogen and water via it's impact on N uptake costs and hydraulic resistance terms respectively.
+The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.
+Both types of effect are seen here, with the Niwot Ridge and Metolius site suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios, but the Harvard forest and Caxiuana sites showing increases.
 
 Impact of \emph{stem\_leaf}\\
-Decreases in stem:leaf ratio (stem\_leaf) below their default condition typically resulted in unrealistically high LAI.
-They also caused a decline in leaf N per unit area (figure **e) , and an increase in photosynthetic capacity (not shown).
-Litter with a high C content (woody material) has the capacity to “lock up” large amounts of soil Nitrogen in soil decomposing pools, and to require large fluxes of mineral N for decomposition.
-Thus, removal of this woody matter appears to alleviate some N limitation, and subsequently increasing leaf N content and thus photosynthetic capacity.
-In contrast to the root:leaf ratio response, GPP is strongly increased, apparently driven primarily by changes in N limitation and not
+Low values of stem:leaf ratio typically resulted in much higher LAI at all sites, as well as declines in vegetation carbon, as expected from the lower stem carbon allocation.
+At Niwot Ridge, there is an optimum stem carbon observed, but this is due to the impacts of a too-high leaf area index, which actually appears to cause declines in NPP when LAI is above the default.
+At all other sites, the impact of increasing leaf allocation is to increase productivity.
 
-This response is not demonstrated by the root:leaf ratio, since it has a similar C:N ratio as the leaf tissue.
-There are therefore large increases in GPP associated with low stem:leaf ratios, with smaller increases in NPP, given the higher respiratory load of the extra Nitrogen.
+Decreases in stem:leaf ratio (stem\_leaf) below their default condition also caused a decline in leaf N per unit area (figure **e).  The causes of this behaviour are still not understood **
 
 Impact of \emph{Ncosts}\\
-Somewhat more complex responses include that of leaf N costs, whereby the carbon allocated to N uptake (figure 1d ***) increases when N costs reach the lowest values assessed here, as well as (intuitively) when they go up.
-This issue is traceable to a feature in the solutions FUN which 'over' allocates carbon to active uptake, given the lower costs of this pathway, which then cannot be used if stocks are too low.
-Plants are known to release carbon as exudates into soils to stimulate the uptake of nutrients, hence this 'wasted' carbon may reflect a plausible real-life process, but the relationship between the representation here and real-world optimization of this process presents a need for further research.
+The first-order impact of N$_{costs}$ is to increase the respiration losses associated with nitrogen uptake.
+Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of N$_{costs}$ on the system state with the exception of the Niwot Ridge site, which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low.
+Other sites showed more muted responses in the vegetation state.
 
 Impact of \emph{fracfixers}\\
-\emph{Fracfixers} has a relatively moderate impact on the overall system state.
-NPP is moderately decreased with a lower fraction of N fixing capacity, resulting in higher N\_uptake expenditure, very slight increases in LAI and very slight response of GPP to that change.
-There are no response of CN ratio.
-These systems are only subject to transiently ramping CO2, and are thus closer to equilibrium than the step-change fertilizations, and this suggests that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.
+Increasing \emph{Fracfixers} affects the maximum amount of carbon that plants can pay to symbiotically uptake Nitrogen.
+In all cases here, it has a relatively moderate and positive impact on the overall system state, in particular at the Niwot and Harvard forest sites.
+As is the case for N$_{costs}$, with high fixing capacity N\_uptake expenditure is lower, N uptake high, and NPP is moderately increased, resulting in increases in LAI and a slight response of GPP to that change.
+These systems are only subject to transiently ramping CO$_{2}$$, and are thus closer to equilibrium than the step-change fertilizations, suggesting that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.
 
@@
 Changing the target leaf C:N ratio \emph{leafcn} over one standard deviation from the mean TRY values had substantial impacts on all the output variables.
-In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that productivity was almost zero at Oak Ridge.
-Low C:N ratios (high leaf N content) uniformly increased GPP, with somewhat dampened impacts on NPP, reflecting decreasing carbon use efficiency, and further dampened impacts on LAI, reflecting increased costs of N acquisition with increasing N demand.
-No optimum value for leafCN ratio was found, in contrast with the standard expectation that over-investment in tissue N might at some point result in declining NPP, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
+As expected, higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.
+In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge.
+
+For NPP, the corresponding  decreases in respiration rates at high CN somewhat ameliorated this effect, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result.
+At Niwot, an optimum leaf CN ratio was apparent in the range of the default values, indicating the possibility that other sites might be operating distant from their theoretical optima according to the model, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
 We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations.
 
+In general, for a given change in NPP, the relative change in LAI was lower than for other parametric modification, reflecting the increased spending on nitrogen uptake inherent in the high $N_{area}$ simulations.
+Interestingly, despite a lower NPP for low CN values at Niwot Ridge, LAI values were higher, for reasons not yet understood ***
+
 Impact of \emph{grperc}.\\
-The response to modifications in the growth respiration coefficient in line with first-order expectations, where the higher numbers (which correspond to the previously used value of 0.33 in CLM4.5) resulted in substantial NPP reductions of ~30\% and corresponding changes in LAI.
-GPP was typically unaffected.
+The response to modifications in the growth respiration coefficient in line with first-order expectations, where the higher numbers (which correspond to the previously used value of 0.33 in CLM4.5) resulted in  NPP reductions, corresponding changes in LAI, impacting GPP at the largest values.
+The impact varied between sites, reflecting different partitioning of the carbon budget into maintenance respiration, nitrogen uptake and growth.
 
 Impact of \emph{medlyn\_slope}\\
-As expected, the change in the stomatal efficiency parameter of the Medlyn 2001 stomatal conductance model impacts GPP primarily In contrast to most of other parameters here, there were some instances where a clear optimum condition was found.
-Too-low values of this parameter results in not enough stomatal opening and photosynthesis, but too-high values result in to much water loss.
-This is illustrated for the two tropical sites in figure **a and figure **a.
-There was a marked difference in responses between sites, reflecting different humidity regimes.
-Values which caused a large decline in photosynthesis at Oak ridge, for example, did not have a large impact in the tropics, indicating that too much stomatal closure was more prohibitive there.
-The large response to \emph{leafcn} at this site also suggests that there is significant carboxylation limitation, potentially on account of lower temperatures.
-At this site, greater stomatal opening always caused more photosynthesis, indicating a lack of impact of hydraulic limitation (as inferred from the decline at high values at the tropical sites, where high atmospheric demand can cause low water potentials and stomatal closure even in high rainfall conditions [Kennedy et al.
-in prep].
+The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties.
+At the Caxiuana site, the very lowest values caused a complete death of the ecosystem.
+It is unclear why this occured, since the impacts can be direct, on the carbon balance itself (not enough GPP to cover costs), but also on the leaf temperature (via surface energy balance issues) and hence the respiration rates.
+It is hard to discern the exact reasons, given the feedback inherent between all of these properties, but no impact on the carbon balance is found at intermediate levels.
+
+At Niwot, a clear optimum was found in the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough.
+At Metolius and Harvard forest, only lower value affected the photosynthetic rates, indicating co-limitation by another process, and the absence of impacts of excessive plant transpiration on the system (note that \cite{kennedy} and \cite{dagon} discuss surface energy and plant hydrodynamic functioning of the model in much greater depth).
 
 Impact of \emph{lmr\_intercept}\\
-The variance in the reported intercepts in \cite{atkin2016} have important consequences for ecosystem productivity, and deserve further investigation.
-These upper ranges of this parameter were among the some of the lowest NPP and LAI simulations for all sites, which, given that the range here is taken from published values, suggests a genuinely high range of uncertainty associated with the leaf maintenance respiration algorithm.
-Of course, the values provided by Atkin et al.
-were for individual PFTs, so the full range of uncertainty probed here is not necessarily relevant to each PFT.
-For example, the higher values of the parameter pertain to grasses, and thus are likely outside of the range of plausible values for the tree PFTs used here.
+The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization.
+For the high leaf area simulations (now shown **), the impacts of maintenance respiration parameterization were much higher, reflecting the higher amount of respiring tissue in those simulations.
+At Caxiuana, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.
 
@@
 Impact of \emph{cn\_flex\_a}\\
-\emph{cn\_flex\_a} was positively associated with leaf N content, and hence GPP, and NPP and LAI.
-This is in line with expectations, since this term is used as the intercept; above which an increase in the cost of N will reduce expenditure on N uptake.
+\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake.
 (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details.
-Thus, low values of \emph{cn\_flex\_a}  results in more common reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.
-The threshold response to this parameter (only the very highest values have an impact) results from a lower bound (0.50) on the degree to which N cost can reduce uptake, suggesting that the sites here were operating in the region of this limit.
+Thus, low values of \emph{cn\_flex\_a}  results in reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.
+For the range investigated here, this parameter had limited impact on the system state.
 
 Impact of \emph{cn\_flex\_b}\\
-\emph{cn\_flex\_b} is slightly negatively associated with leaf N concentration, GPP;, NPP and LAI, but positively associated with NPP spent on carbon uptake.
-This is a somewhat confusing outcome, since, in principle the responses of \emph{cn\_flex\_a} and \emph{cn\_flex\_b} should be the same way around, as increases in both increase the amount of carbon spent on N uptake.
-There is some indication that at low values of this parameter, C allocated to N uptake might end up being too high, thus creating some of the 'wasted' carbon expenditure described in the ( see Ncosts section), but this potentially merits investigation (notwithstanding the point the this algorithm is a placeholder for a more evidence-based depiction of stoichiometric flexibility).
+\emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, but has limited impacts on the system state here.
+At low values, there is some indication at Niwot Ridge that it might increase LAI and vegetation carbon, indicating a decrease in the N uptake expenditure and in leaf N concentrations.
+The limited impact at other sites suggests that Ncosts are rarely high enough to trigger reductions in N expenditure (notwithstanding the point the this algorithm is a placeholder for a more evidence-based depiction of stoichiometric flexibility).
 
 Impact of \emph{cn\_flex\_c}\\
-\emph{cn\_flex\_c} has the greatest impact of the three CN flexibility parameters.
-Notably, high values of this calibration parameter allow significant increases in leaf N content, which are, counterintuituively, associated with decreases in expenditure on N uptake, but little chqnge in LAI, implying increases in N uptake for less cost.
-Potentially, decreasing the C allocated to N uptake in a given timestep (when N is cheap) might be in some cases avoiding 'wasted' carbon.
-Further investigation of this phenomenon would be merited.
+ Low values of \emph{cn\_flex\_c} affects the degree to whch plants attempte to maintain tissue CN values near to their target when environmental costs are varying.
+For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.
 
-\subsection{Parametric control over CO$_{2}$ and Ndep responses}
-To illustrate the relationships between N and CO$_{2}$ fertilization, and to reduce the dimensionality of the output, figures ** to ** illustrate the response for GPP (figure **), NPP (figure **), LAI (figure **) and vegetation carbon (figure **) for five sites each.
-Here we summarize the impact of each parameter on the output patterns.
+\subsection{CO$_{2}$ and Ndep responses}
+
+To illustrate the relationships between N and CO$_{2}$ fertilization, and to reduce the dimensionality of the output, figures \ref{GPP CO2 and N respones 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the response for GPP, NPP and, LAI respectively.
+Vegetation carbon and $N_{area}$ are in the supplemental infomation (\ref{VEGC CO2 and N respones 2001}, \ref{Leaf N CO2 and N respones 2001}).
+For Niwot Ridge, figures showing the N and CO$_{2}$ responses are also shown (Figure \ref{NR1 ndep} and \ref{NR1 celev}), while the corresponding figures for the other sites are in the SI.
+
+\subsection{Differences across sites}
+The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.
+The default range of CO$_{2}$ fertilization impacts on NPP, for example, ranged from 1.22 to 1.66 at Harvard and Caxiuana respectively, with Niwot having a lower range (1.28) and Metolius higher (1.62).
+For Nitrogen responses, Caxiuana had a negative response (0.96), Niwot had almost no response (0.98) while Harvard and Metolius were positive (1.25 and 1.09 respectively).
+Impacts on GPP were all positive, from  (albeit lower for the Caxiuana and Niwot sites) through increases in leaf N content, and impacts on LAI were all positive, reflecting lower expenditure on N uptake (1.05 to 2.09).
+Interestingly, the highest impacts on LAI were at Niwot, which has almost no change in NPP, reflecting a large initial expenditure on nitrogen uptake ameliorated by fertilization.
+
+\subsection{Impact of parameters}
+Across the four sites, parametric impacts were very varied, when compared to the relatively consistant impact of parameters on the system state.
+A majority of parameters show no consistant relationship with CO$_{2}$ or N deposition responses, and the range of their impacts vary over an order of magnitude between sites.
+We describe here the major features of how parameteric variation controls the behaviour of the model under these fertilization scenarios.
 
 Impact of \emph{slatop}\\
-Despite it’s very large impact on the system state, via modifications to the LAI predictions, specific leaf area has limited impact on the overall response to fertilization, compared to other parameters, and the responses to both N and CO$_{2}$ fertilization differ between sites, potentially reflecting the different limiting conditions at each site.
 
 Impact of \emph{froot\_leaf}\\
-\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values, which were generally associated with limited response to N fertilization.
-This might reflect either a lack of demand for N, or an inability to respond to the fertilization itself through lack of root matter.
+Despite it’s  impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses.
+At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fetilization is lower since the feedback from LAI to GPP is less proncounced.
 
-Impact of \emph{stem\_leaf}\\
-Low values of the stem/leaf ratio also caused very limited responses to N fertilization at some sites, and CO$_{2}$ fertilization at others.
-These very low values were associated with very high LAI numbers, which potentially were already subject to light limitation; so as not to profit too much from the addition of extra N or CO$_{2}$.
-**think about the LAI vs fertilization response figure**
+Impact of \emph{stem\_leaf}\\}
+Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuana and Harvard Forest.
+Initial LAI was larger for these cases for these sites, explaining the decease in fertilization response as above.
+However, at Niwot Ridge, CO$_{2}$ fertilization was larger with lower \emph{stem\_leaf} ratios.
+At Niwot, however, a larger control LAI was associated with lower $N_{area}$, GPP and NPP, suggesting a high nitrogen limitation associated with these states.
+Nonetheless, fertilization responses of GPP here are very high, (x1.8) in response to quite small changes in $N_{area}$ (~2\%)  and LAI (~22\%) (Figure \ref{NPP CO2 and N respones 2001}).
+Further, Nitrogen increases LAI substantially, buft only has a very limited impact on GPP and NPP at Niwot Ridge, suggesting that the fluxes might be limited by water availability, and thus explaining the very high CO$_{2}$ sensitivity.
 
 Impact of \emph{N\_costs}\\
-As described previously, too-low N costs in the environment tend to increase the amount of carbon wasted during N uptake (on the attempted acquisition of N which is not there).
-The model calibration at present appears to be in an interim point, but **we probably need to reduce the bounds of the sensitivity analysis to make this all a bit less shocking**
+Given that they impact on the respiration expended on N uptake and the availability of carbon for tissue growth, the impact of Nitrogen costs is most clear for NPP and LAI (Figure \ref{NPP CO2 and N respones 2001}, \ref{LAI CO2 and N respones 2001}}).
+At Caxiuana, Niwot and Harvard forest, high initial N costs resulted in stimulation of the N fertilization response, as expected from the alleviation of the higher N limitation in these cases.
+There was also a slight increase in the CO$_{2}$ fertilization response (maybe on account of extra carbon alleviating some of the N limitation in those systems).
+For very low values, at Caxiuana and Niwot, N responses increased, counter-intituvely.
+At both these sites, at very low N costs, plants appear to make leaves with higher N content (as per the algorithm which allows a higher $N_{area}$ to be achieved at low N cost).
+However, the total amount of N in the system remains the same with low N costs, and so these cases do not have higher leaf area.
+When more total N is added, the leaf area can expand, and thus the N concentration declines.
 
 Impact of \emph{frac\_fixers}\\
+While nitrogen uptake costs were important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available, in many cases.
 The fraction of N fixers in the system has a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
 In line with expectations, for GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$.
-For NPP, this result is muted for the two (less N limited) tropical sites, particularly Caxiuaxa (Figure **) because the increase in shaded leaves and in tissue C:N ratios (Figure **) with fertilization increases the respiratory load on these systems.
-Note that NPP reported here is the total C available for growth and N uptake, and therefore a lower NPP (as in figure **a) can still be reconciled with a higher LAI (figure **a).
 
-In CLM5, each PFT except crops has the same N fixation fraction (0.25), but this analysis clearly suggests that much greater emphasis should be afforded to the dynamics of changing N fixation capacity under transient atmospheric conditions.
+For NPP, this result is muted for the Caxiuana and Metolius sites, and indeed at Caxiuana the low N fixer cases had a very negative (-8\%) NPP response (despite a positive 7.6\% GPP response), on account of the model increasing the N content of the leaves by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}).
+Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.
 
-One interesting feature of this response is that, despite not exhibiting much difference in N limitation in the present day (given the muted response of the system state to changes in \emph{frac\_fixers}), simulations with low (or no) fixer capacity had typically very high responses to N fertilization.
-** this needs an explanation from looking at the N response plots on their own...**
+At Niwot Ridge and Metolius, the changing fixation capacity had a very large impact on carbon fertilization, but not on nitrogen response, particularly for GPP.
+At Metolius, this impact persisted, but at Niwot the fixation capacity did have an impact on LAI.
+Interestingly, many variables increased NPP at this site, but only \emph{fracfixers} increased LAI, suggesting a limitation in the total available N suppressed the impact of all the other case where NPP was increased, and also that water limitation prevented any greater assimilation of carbon.
 
 Impact of \emph{leafcn}\\
- High target C:N ratio simulations exhibited a degree of N limitation in the control condition, which was strongly alleviated by N fertilization.
-Similarly, these scenarios had lower responses to CO$_{2}$, since in many cases, insufficient N was available to expand the canopy, leading to declines in leaf N concentration and a less strong (but still positive) fertilization response.
+The impact of target \emph{leafcn} on fertilization responses was complex (recalling that this was a significant control over the system state at all sites).
+t Caxiuana, high values \emph{leafcn} (low leaf  $N_{area}$) had much stronger responses to N fertilization, indicating the dominant impact of the carboxylation capacity at this tropical site.
+At Niwot Ridge, low values of \emph{leafcn} (low leaf  $N_{area}$) had much stronger responses to CO$_{2}$ fertilization (the strongest responses at this site for GPP), potentially since the low stomatal conductance at this water-limited site could benefit most from additional CO$_{2}$ where carboxylation capacity was high.
 
 Impact of \emph{grperc}\\
-Of all the variables tested, \emph{grperc} had one of the lowest impacts on fertilization responses, since the degree to which growth is 'taxed' is not responsive to any of the physiological processes upstream that depend on fertilization.
+\emph{grperc}, or the degree to which growth is 'taxed' is constant and not responsive to any of the physiological processes upstream that depend on fertilization.
+Thus, its effects are all likely due to indirect impacts on the initial state.
+Two main impacts were observed.
+At Niwot Ridge low values of \emph{grperc} caused higher responses to CO$_{2}$ fertilization for GPP and NPP, but not LAI.
+These values corresponded to high initial LAI (on account of larger amounts of C being available for N uptake) and therefore allowed greater leverage of the increasing CO$_{2}$.
+At Harvard forest, high \emph{grperc} had caused low levels of LAI, resulting in increased capacity for CO$_{2}$ fertilization at that site.
 
 Impact of \emph{medlyn\_slope}\\
-For many sites, the response of the system to CO$_{2}$ fertilization with low values of the \emph{medlyn\_slope} parameter was higher than the cut-off imposed for our visualization (>80\% of control LAI).
-For example figure si** indicates that the lowest values had a x4 increase in GPP under double CO2 scenarios with the lowest values of this parameter, illustrative of very low leaf internal CO$_{2}$ concentrations where stomatal opening is heavily restricted.
+At Niwot Ridge, low values of the Medlyn Slope parameter (with high water use efficiency) were able to respond very strongly to increasing Nitrogen.
+These cases also had higher GPP in the initial state, and their higher WUE enabled them to take advantage of the increasing N availability.
+These cases also ended up with higher vegetation carbon.
+
+Surpriingly, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization reponse, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.
 
 Impact of \emph{lmr\_intercept}\\
-Like \emph{grperc}, the respiration intercept has only a straightforward linear impact on the model state, and while the very high values of this parameter were in some cases completely prohibitive of growth, they do not impact sensitivity to fertilization in any notable fashion.
+In general the impacts of the maintenance respiration intercept were small, and limited to the impact of large values increasing fertilization responses to CO$_{2}$ and N, potentially on account of the low LAI and thus greater feedback between LAI and GPP.
 
@@
 Impact of \emph{cn\_flex\_a}, \emph{cn\_flex\_b} and \emph{cn\_flex\_c}\\
+As with it's impact on the state, there was limited impact of \emph{cn\_flex\_a} on the fertilization responses, indicating that this parametr is
+
+Impact of \emph{cn\_flex\_b}
+Low values of \emph{cn\_flex\_b}, where the leafCN concentration can increase in response to fertilization, showed higher responses to n fertilization, particularly at Harvard Forest.
+Small values of  \emph{cn\_flex\_b} should  decrease the amount of carbon spent on N uptake when NCosts are high, thus, a relaxation of the cost of N will increase the amount spent on uptake, and subsequently $N_{leaf}$ and all fluxes downstream.
+Thus, these changes result from the relaxation of an over-suppression of leaf N content.
+
+These cases also showed increases in leaf N content for N fertilization, LAI, GPP and NPP.
+Given that small values of  \emph{cn\_flex\_b} should in fact decrease the amount f carbon spent on N uptake, and thus they
+
+and \emph{cn\_flex\_c}\\
 **These will all need looking at once a more careful set of runs are done**
-Low values of \emph{cn\_flex\_b}, where the leafCN concentration can increase in response to fertilization, showed very high responses to n fertilization, and low responses to CO2 fertilization (where N content of leaves declined).
 
@@
 Given the sensitivity of the fertilization responses in the CLM5, we advocate careful interpretation of the outcome of the model, particularly with regard to the climate feedback strength.
+
+\subsubsection{Limitations of the CLM5 revealed by this parameter analysis
+
+1. Lack of optimization of the leaf N area.
 
@@
 
-Figure 1: Sensitivity analysis of state of CLM at site to N parameters @ORNL
+\begin{figure}[h]
+     \caption{Sensitivity of system state to parameter perturbation at Niwot Ridge site.}
+     \label{NWRstate}
+  \end{figure}
+Figure 1: Sensitivity analysis of state of CLM at site to N parameters.
 
@@
 
-Figure 10: Some sort of Will figure.
-
 \section{To Do}
@@
 
+Make a figure with a bunch of extra outputs for the state at Niwot Ridge (CUE, fraction NPP nuptake, BTRAN)
 \nocite{*}
+
+Make some SI figures with the high leaf run (state for Niwot?)
+
+Make all on the same figure figure
+
+Slightly re-format the fert figures. (dashes, GPP label)
+Check the CN figures. They don't match the NWT figures.
 

### 2018-11-13 main.tex

@@
 
-These cases also showed increases in leaf N content for N fertilization, LAI, GPP and NPP.
-Given that small values of  \emph{cn\_flex\_b} should in fact decrease the amount f carbon spent on N uptake, and thus they
-
-and \emph{cn\_flex\_c}\\
-**These will all need looking at once a more careful set of runs are done**
+Impact of \emph{cn\_flex\_c}\\
+The impact of \emph{cn\_flex\_c} (the paremter controlling the degree to which plant CN ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon at Harvard Forest and Niwot Ridge (and Caxiuana, to a smaller extent).
+These impacts do not mirror those on leaf area index or NPP.
+At Harvard Forest and Caxiuana, there is a relative increase in the leaf N content, while at Metolius and Niwot, this parameter has a dominant control over the impacts of leaf N to CO$_{2}$ fertilization.
+The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high, shown clearly at Niwot Ridge and Metolius (e.g.
+Figure \ref{NR1 celev}, \ref{MET celev}).
+Of course, if N concentrations are allowed to decline too much in response to rising CO$_{2}$, then productivity declines, and other complext feedbacks occur, as shown at Harvard Forest and Niwot Ridge (Figure \ref{HRV celev}, \ref{CAX celev}).
+In terms of the overall impact, we can see that the degree of flexibility of the tissue CN ratio acts as an orthoganal control on fertilization response to the total fixation capacity.
 
@@
 At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes.
-In particular, as expected, parameters that relate most strongly to the acquisition of N, \emph{frac\_fixers}, \emph{Ncosts} and use \emph{leafcn}, \emph{cn\_flex\_b}, have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
-It has long been understood that uncertainties in the representation of Nitrogen are large(\cite{zaehle2014}) .
-In the CMIP6 generation of land surface models, it is expected that a much larger number of model submissions will contain active representations of N limitation.
-We do not yet know how CLM5 compares to others in it's carbon cycle feedback strength, but we propose that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained.
+In particular, as expected, parameters that relate most strongly to the acquisition of N, (\emph{frac\_fixers}, \emph{Ncosts}) and to the flexibility of plant tissues in their C:N ratio (\emph{leafcn}, \emph{cn\_flex\_b}, \emph{cn\_flex\_c}), have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
+It has long been understood that uncertainties in the representation of Nitrogen are large(\cite{zaehle2014}), but previous model-model comparison studies have struggled to illustrate the exact cause of differences between models predictions, given the complexity of the structural and parametric differences involved.
+The new CLM5 model has a number of explicit parameterizations, many of which were hard-coded or assumed to be 0 or 1 in previous versions of the model.
+This allows both a clear exploration of the sensitivities of the parameter space to critical assumptions, as shown here, as well as more comprehensively tuning by future studies and comparisons with relevant datasets.
+
+In the CMIP6 generation of land surface models, it is expected that a greater number of model submissions will contain active representations of N limitation.
+We propose here that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained for all models.
 
@@
 However, such experiments necessarily involve unrealistically rapid changes in environmental boundary condition, and cannot necessarily be considered as analagous to either real-world or transient future scenario conditions.
-We assessed the response to the ramping historical CO2, to illustrate how parameter responses differed between the rapid and slow responses.
-The results are in figure SI **
+Further, we expect that the carbon and nitrogen cycles mayb take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.
+
+To test the impact of time on the fertilization responses, we also analyzed the impact of fertilization at 5 and 50 years into the elevated CO$_{2}$ and N experiments.
+At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, result not shown) to those at 15 years, but at 5 years, in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent x~2.06 impact on NPP, compared to ~0.95 in the 2015 analysis at NVR), indicating that the results described above had already been subject to progressive N limitation and were closer to an equilibrated state.
 
@@
 The fit at individual sites of the globally adjusted model cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, which is beyond the scope of this paper.
-Instead, we chose a set of individual sites, rather than global analysis, to allow for the 936 model integrations that comprise our sensitivity tests, and to probe a set of alternative vegetation and climate conditions.
-This study is incomplete, as are all investigations of the climatic and parameteric space of land surface models.
 
-\subsubsection{Limitations of parameter calibration issues}
+Instead, we chose a set of individual sites, rather than global analysis to extensively probe a set of alternative vegetation and climate conditions under realistic forcing scenarios.
+We also do not make any assessment of how well the model output fits observations across this parameter space, since the relevant datasets with which to test alternative model configurations are, we argue, the set of globally-relevant data products described by \cite{lawrence2018}.
+Future studies will consider the impact of alternative parameterizations, informed by these site-level analyses, on the global performance of CLM5, but it is clear that a wide range of biogeochemical responses are possible, within the context of this model structure.
+
+\subsubsection{Parameter calibration of the default model, or absence thereof}
+
 During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
 We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces,  3) the tendency of the modeled ecosystem to die under past low-CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting pattern that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the different data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
-Thus, the parameterization used in the default version of the model is a less objective calibration than is ideal, and a limited set of parameters were tuned 'by hand' to avoid excessively spurious values for some model states.
-Given the sensitivity of the fertilization responses in the CLM5, we advocate careful interpretation of the outcome of the model, particularly with regard to the climate feedback strength.
+The parameterization used in the default version of the model utilized the mean trait values where possible, from trait datasets (\cite{lawrence2018}) in combination with targeted tuning of parameters   (including allocation parameters and Nitrogen costs)that were less well constrained by data, to avoid excessively spurious values for some model states.
 
-\subsubsection{Limitations of the CLM5 revealed by this parameter analysis
+We recognize that this less objective calibration method is than is ideal, and ongoing studies (\cite{dagon}) will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated in a recent study by \cite{fer2018}, .
 
-1. Lack of optimization of the leaf N area.
+The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
+In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
+Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
 
-\subsection{Future Investigations}
 \subsubsection{Temperature and water responses}
-Here we probe the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5.
+In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5.
 In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization.
 The full response of the model has considerably more dimensions than these two transient properties.
-In particular, here we have not considered responses to temperature and moisture changes.
+In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water.
 The suite of parameters than in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implictly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.
-Properties more closely linked to water availability are associated with the hydrological components of the model as well as the PHS plant hydraulics scheme and will be investigated further by [Dagon et al in prep.
-and Kennedy et al., in prep.]
-
-\subsubsection{Filtering parameter space with observations}
-The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
-In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of availble and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
-Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
+Properties more closely linked to water availability are associated with the hydrological components of the model as well as the PHS plant hydraulics scheme and will be investigated further by \cite{dagon} and \cite{kennedy}.
 
@@
 However, in doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.
+
+In this study, we found numerous instances where the model reduced carbon use efficiency by increasing tissue C:N ratios in response to forcing.
+In principle, there exists an optimal leaf CN ratio that maximizes leaf C export under a given set of conditions.
+Despite its optimal allocation of leaf N within the leaf (to photosynths
 

### 2018-11-13 main.tex

@@
 In principle, there exists an optimal leaf CN ratio that maximizes leaf C export under a given set of conditions.
-Despite its optimal allocation of leaf N within the leaf (to photosynths
-
-\subsubsection{Discourse on how target CN ratio could be optimized but isn't}
-
-\cite{thomas2014}, \cite{friedlingstein1999}, \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{anten2011}, \cite{vanwijk2003}
-
-\subsubsection{Discuss geographical spread of predictions for N and CO2 fertilization}
-
-\cite{hickler2008}
+Despite its optimal allocation of leaf N within the leaf (to photosynthetic enzymes), CLM5 does not include any scheme that ensures that overall tissue CN ratios conform to any physiological optima.
+Instead, a heuristic model adjusts tissue CN ratios round the target ratio (an input parameter).
+Numerous models have been proposed along these lines in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{a,nten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}).
+We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in the LUNA model (Xu.
+C.
+pers comm), but future nitrogen cycling developments might re-visit this topic.
 
@@
 In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (references**), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).
-The results included here indicate that the
 
-\cite{thomas2013global} find that the CLM4.0 model responds very strongly to nutrient additions, and is heavily limited by nitrogen availability in the present day.
+\subsection{Differences between sites}
+The overall response of the different sites to CO$_{2}$ and N fertilization, in terms of the total vegetation carbon, is to some extent a function of their initial limiting conditions.
+The Niwot Ridge site is limited by both N availablilty and water in the initial state, and thus demonstrates strong CO$_{2}$ and N responses.
 

### 2018-11-13 figures.tex
Updates from ShareLaTeX

@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1.png}
-     \caption{Caxiuana Forest State}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Caxiuan\~a National Forest}
      \label{CAX state}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
-     \caption{Metolius Forest State}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest State}
      \label{MET state}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
      \caption{Niwot Forest State}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
      \caption{Harvard Forest State}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
      \caption{Harvard Forest Nitrogen Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
      \caption{Harvard Forest CO2 Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1_2012.png}
      \caption{Caxiuana Forest Nitrogen Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1_2012.png}
      \caption{Caxiuana Forest CO2 Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth,left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p1_2012.png}
      \caption{Metolius Forest Nitrogen Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1_2012.png}
      \caption{Metolius Forest CO2 Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1_2012.png}
      \caption{Niwot Forest Nitrogen Fertilization}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
      \caption{Niwot Forest CO2 Fertilization}
@@
      \label{NPP CO2 and N respones 2001}
-  \end{figure}
-
- \begin{figure}[h]
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/OCT_CNdep_NPP1_CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p2006.png}
-     \caption{NPP CO2 and N respones}
-     \label{NPP CO2 and N respones}
   \end{figure}

### 2018-11-13 figures.tex

@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest State}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest}
      \label{MET state}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
-     \caption{Niwot Forest State}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Niwot Ridge.}
      \label{NR1 state}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
-     \caption{Harvard Forest State}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Harvard Forest}
      \label{HVF state}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Harvard Forest Nitrogen Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Harvard Forest}
      \label{HRV ndep}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Harvard Forest CO2 Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Harvard Forest}
      \label{HRV celev}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Caxiuana Forest Nitrogen Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 uears of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Caxiuana}
      \label{CAX ndep}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Caxiuana Forest CO2 Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Caxiuana }
      \label{CAX celev}
@@
      \includegraphics[width=1.3\textwidth,left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Metolius Forest Nitrogen Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Metolius Forest}
      \label{MET ndep}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Metolius Forest CO2 Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Metolius Forest}
      \label{MET celev}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Niwot Forest Nitrogen Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
      \label{NR1 ndep}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Niwot Forest CO2 Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
      \label{NR1 celev}

### 2018-11-13 SIfigs.tex
Updates from ShareLaTeX

--- 2018-10-04 figures.tex
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_p1CLM5_CAX__y1.png}
-     \caption{Caxiuana State Parameter Sensitivity}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Caxiuan\~a National Forest}
      \label{CAX state}
@@
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_p1CLM5_BCI__y1.png}
-     \caption{BCI State Parameter Sensitivity}
-     \label{BCI state}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Harvard Forest}
+     \label{HVF state}
  \end{figure}
@@
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_p1CLM5_ORN__y1.png}
-     \caption{Oak Ridge State Parameter Sensitivity}
-     \label{ORN state}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest}
+     \label{MET state}
  \end{figure}
@@
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}
-     \caption{Caxiuana CO2 response}
-     \label{Caxiuana CO2 response Parameter Sensitivity }
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Caxiuan\~a }
+     \label{CAX celev}
   \end{figure}
@@
  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI_y1.png}
-     \caption{Oak Ridge CO2 response}
-     \label{Oak Ridge CO2 response Parameter Sensitivity}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 uears of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Caxiuan\~a}
+     \label{CAX ndep}
   \end{figure}
@@
   \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/frac_deviation_CO2_response_1CLM5_BCI__y1.png}
-     \caption{BCI CO2 response}
-     \label{BCI CO2 response Parameter Sensitivity}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Harvard Forest}
+     \label{HRV celev}
   \end{figure}
@@
   \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_GPP.png}
-     \caption{GPP CO2 and N respones}
-     \label{GPP CO2 and N respones}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Harvard Forest}
+     \label{HRV ndep}
   \end{figure}
@@
   \begin{figure}[h]
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_NPP.png}
-     \caption{NPP CO2 and N respones}
-     \label{NPP CO2 and N respones}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Metolius Forest}
+     \label{MET celev}
   \end{figure}
@@
   \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/rel_LAI.png}
-     \caption{LAI CO2 and N respones}
-     \label{LAI CO2 and N respones}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth,left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Metolius Forest}
+     \label{MET ndep}
   \end{figure}

### 2018-11-13 figures.tex

--- 2018-11-13 figures.tex
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Caxiuan\~a National Forest}
-     \label{CAX state}
-  \end{figure}
-
- \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest State}
-     \label{MET state}
-  \end{figure}
-
- \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
-     \caption{Niwot Forest State}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Niwot Ridge.}
      \label{NR1 state}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
-     \caption{Harvard Forest State}
-     \label{HVF state}
-  \end{figure}
-
- \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Harvard Forest Nitrogen Fertilization}
-     \label{HRV ndep}
-  \end{figure}
-
-   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Harvard Forest CO2 Fertilization}
-     \label{HRV celev}
-  \end{figure}
-
-   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Caxiuana Forest Nitrogen Fertilization}
-     \label{CAX ndep}
-  \end{figure}
-
-   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Caxiuana Forest CO2 Fertilization}
-     \label{CAX celev}
-  \end{figure}
-
-   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth,left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Metolius Forest Nitrogen Fertilization}
-     \label{MET ndep}
-  \end{figure}
-
-   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Metolius Forest CO2 Fertilization}
-     \label{MET celev}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
+     \label{NR1 celev}
   \end{figure}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Niwot Forest Nitrogen Fertilization}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
      \label{NR1 ndep}
-  \end{figure}
-
-   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Niwot Forest CO2 Fertilization}
-     \label{NR1 celev}
   \end{figure}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
-     \caption{GPP CO2 and N respones}
+     \caption{Influence of parameteric variation on the impact of 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
+}
      \label{GPP CO2 and N respones 2001}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_NPP1__p2012.png}
-     \caption{NPP CO2 and N respones}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on net primary productivity (NPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
      \label{NPP CO2 and N respones 2001}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TLAI1__p2012.png}
-     \caption{LAI CO2 and N respones}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf area index (LAI), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
      \label{LAI CO2 and N respones 2001}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
-     \caption{VEGC CO2 and N respones}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
      \label{VEGC CO2 and N respones 2001}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
-     \caption{Leaf Nitrogen CO2 and N respones}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
      \label{Leaf N CO2 and N respones 2001}

### 2018-11-15 main.tex

@@
 \authors{Rosie Fisher\affil{1},
-Will Wieder\affil{1},
-Ben Sanderson\affil{1},
-Charlie Koven\affil{2},
+William Wieder\affil{1},
+Benjamin Sanderson\affil{2},
+Charles Koven\affil{3},
 Keith Oleson\affil{1},
-Chonggang Xu\affil{3},
+Chonggang Xu\affil{4},
 Ashehad Ali\affil{1},
-Josh Fisher\affil{4},
-Mingjie Shi\affil{4},
-Katie Dagon\affil{1},
-Danica Lombardozzi\affil{1},
-Anthony Walker\affil{5},
+Joshua Fisher\affil{5},
+Mingjie Shi\affil{5},
+Anthony Walker\affil{6},
 Gordon Bonan\affil{1},
@@
 
-\title{Nitrogen cycling in the CLM5: controls on responses to environmental forcing}
-\author{rosie fisher}
-\date{April 2018}
+\title{Parameteric controls on responses to biogeochemical forcing in the CLM5}
+\author{Rosie Fisher}
+\date{November 2018}
 
@@
 
-\affiliation{2}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
+\affiliation{2}{CERFACS, Toulouse, France}
 
-\affiliation{3}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
+\affiliation{3}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
 
-\affiliation{4}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
+\affiliation{4}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
 
-\affiliation{5}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
+\affiliation{5}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
+
+\affiliation{6}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
 
@@
 
-\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycle of Carbon and Nitrogen.
-We assess the dependence of the both the model state, and also the responses of the system to higher CO2, Nitrogen and temperature to parameter uncertainty.
+\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycling of Carbon and Nitrogen.
+We assess the dependence of the both the model state, and also the responses of the system to CO$_{2}$ and Nitrogen fertilization to parametric uncertainty.
 Model runs are conducted at representative individual sites in temperate, tropical and boreal systems.
 
-\item The model responses to CO2 and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment.
+\item The model responses to CO$_{2}$  and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, the costs of N in the environment and stoichiometric flexibility of plant tissues, all of which are subject to large empirical and theoretical uncertainty.
 
 \item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.
-
-\item This implies that a critical control in the CLM, and in coupled simulations with the CESM for CMIP are dependant on the representation of Nitrogen fixations and its dynamics.
-
 \end{keypoints}
@@
 Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}).
-In the CMIP5 multi-model intercomparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle.
-The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}).
-Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}).
-Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehledalmonech2011}).
+In the CMIP5 multi-model inter-comparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle.
+The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{friedlingstein2014}, \cite{arora2013}).
+Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent global N dynamics using alternative structural assumptions have been developed (\cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}).
+Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehle2011}).
 
 Unlike representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.
-Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{zaehle2014}).
-Few, if any studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model.
+Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{kauwe2013}, \cite{zaehle2014}, \cite{medlyn2015}).
+Few studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model  (c.f.
+\cite{meyerholt2018}).
 
 The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}).
-Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release.
-An overview of model developments is presented by \cite{lawrence2018}, a descriptions of the global Nitrogen cycling by Wieder et al.
-(in prep), the crop model (Lombardozzi et al.
-in prep), the land use components (Lawrence P et al.
-in prep), land surface hydrology (Swenson et al.
-in prep) and plant hydraulics (Kennedy et al.
+Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}.
+An overview of model developments is presented by \cite{lawrence2018}, a description of the global Nitrogen cycling by Wieder et al.
+(in prep), the crop model by(Lombardozzi et al.
+in prep), the land use components by (Lawrence P et al.
+in prep), land surface hydrology by (Swenson et al.
+in prep) and plant hydraulics by (Kennedy et al.
 in prep).
 
-The CLM5 model includes numerous major update of the nitrogen cycling representation subsequent to CLM4.0 and CLM4.5.
+CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4.0 and CLM4.5.
 The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing.
 Wieder et al.
-(in prep) report the global-scale implications of the full model update and note that the CLM( responds more strongly to CO2 and less strongly to N fertilization than its predecessors.
-In this paper, we describe the new components of the biogeochemical cycling of nitrogen, their impacts upon the model system state and responses to environmental forcing.
+(in prep) report the global-scale implications of the full model update and note that the CLM5 responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.
+In this paper, we describe the new components of the biogeochemical cycling, and investigate the features of the new model that control the model state and biogeochemical forcing responses in depth, using single-site experiments and parametric sensitivity analyses.
 
@@
 
-The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).
+The CLM5 nitrogen cycling builds on the implementation of carbon-nitrogen interactions in the CLM4.0(CN) model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).
 The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\
@@
 
-2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics of Nitrogen acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}).\\
+2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics and carbon economics of N acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}), and\\
 
-3) The `FLEXCN' N cycle implementation, which is used here primarily to allow variation in tissue C:N ratios, and is adapted from \cite{ghimire2016}.\\
+3) The `FLEXCN' N cycle implementation, adapted from \cite{ghimire2016}, which is used in CLM5 primarily to allow variation in tissue C:N ratios,  .\\
 
@@
 \subsubsection{LUNA}
-
 Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
@@
 The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
-In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area ($N_{area}$, which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and environmental conditions.
+In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area $N_{area}$, (which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and prevailing environmental conditions (CO$_{2}$, temperature, humidity, soil moisture).
 The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
@@
 The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon.
-Further, the cost of each of these uptake pathways is variable in both time and space.
+Further, the cost of each of each uptake pathway is variable in both time and space.
 FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them.
 Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N.
-The central element of FUN is a simultaneous equation which assumes, firstly, that ($C_{avail}$), the total carbon available (after maintenance respiration and replenishment of stores are accounted for) is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).
+Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency).
+The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).
 
@@
 
-Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal that which is deployed in the growing plant tissues.
-The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$) which applies only to tissues that are constructed:
+Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:
 
 \begin{equation}
-N_{uptake}=C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
+N_{uptake}=C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{plant}/(1.0+f_{gr})}
 \end{equation}
 
-Thus the Carbon that is expended on N uptake is determined by:
+The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$, which applies only to tissues that are constructed).
+Rearrangement of these constraints gives a predictive term for the amount of carbon that is expended on N uptake, as:
 
@@
 
-The average Nitrogen cost ($N_{cost}$) is derived from the simultaneous uptake of N across all uptake streams, while the target CN ratio ($CN_{target}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools.
-The FUN model is documented \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}{here}
+The average Nitrogen cost ($N_{cost}$) is derived from the uptake of N, weighted for it's cost across all uptake streams, while the plant CN ratio ($CN_{plant}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools.
+The FUN model is documented at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}
 
-A further issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}).
-The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{wieder2015} illustrate the impacts of uncertainty in this function) but FUN considers fixation rates to be an emergent property of plant assimilation rates (providing carbon for fixation), the relative costs of environmental N acquisition, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
-Fixation is preferred when the costs of N acquisition are higher than that of fixation.
-Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground, in productive areas, N fixation is preferred and soil N builds up much quicker.
-In low productivity areas, carbon is required to fix N from a 'col start' initial state, impacting the survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.
+One issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}).
+The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity ((\cite{wieder2015} illustrated the impacts of uncertainty in this function).
+FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compmared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
+Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation.
+Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground.
+In most cases, N fixation is preferred when soil N is low, and so and soil N builds up much quicker.
+In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the carbon use efficiency and survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.
 
 \subsubsection{FLEXCN}
-
-Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}).
-For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage, but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
+The results of the FACE model data comparison exercise illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).
+These changes were not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
 For this reason, we introduce a flexible tissue C:N ratio into the CLM5.
@@
 \cite{ghimire2016} implemented a suite of modifications to the CLM Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
-In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen (although the alternative schemes implemented by \cite{ghimire2016} are available as options).
-In this new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates.
-This new allocation code we term the FLEXCN module.
+In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, thus the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen.
+In the new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates.
+The allocation scheme retains a fixed-fraction approach to allocation in general, and thus is capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an issue which must be taken into account during simulations under high productivity conditions.
 
@@
 
-In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters. We use\\
+In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters.
+We use a range of single site experiments, combined with one-at-a-time parameter uncertainty analyses from the default state, as well as CO$_{2}$ fertilization experiments and Nitrogen addition experiments,
+to fully investigate the performance of the model under steady-state and step-change modifications in forcing.
 
-1. A range of single site experiments \\
-
-2. One-at-a-time and global parameter uncertainty analyses\\
-
-3. CO$_{2}$ fertilization experiments and\\
-
-4. Nitrogen addition experiments.\\
-
-to fully investigate the performance of the model under steady-state and step-change modifications in forcing.
 Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect overall behaviour in larger-scale experiments, including the controls on global carbon fertilization `$\beta$' values.
@@
 Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible.
-We further investigate the consequences of turning off the LUNA model and returning to the fixed leaf N/$V_{c,max}$ relationship from \cite{ghimire2016}.
+We thus consider that an exploration of the CLM5 parameter space is also equivalent to an exploration of a large part of the relevant model structural space.
 
@@
 For this experiment we utilized six different experimental sites.
-Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest).
+Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest) **tbd**.
 
 \subsection{Simulation Protocol}
+We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the turnover of slower cycling carbon and nitrogen pools is accelerated, see (\cite{lawrence2018}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level half-hourly meteorological inputs.
+400 years was sufficient to bring all carbon and nitrogen pools into equilibrium.
+Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years for each site, parameter and value combination, with accelerated mode turned off.
+Restarting from the state at the end of the perturbed parameter spin-ups, we ran 'control' transient simulations for each site and parameter combination starting in 1760 and invoking historical transient CO$_{2}$ and Nitrogen deposition until the end of the observed meteorological forcing.
 
-We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (where the turnover of slower cycling carbon and nitrogen pools is accelerated, see (/cite{lawrence}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level meteorological inputs.
-400 years was sufficient to bring all carbon and nitrogen pools into equilibrium.
-Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years at each site, with accelerated mode turned off.
-Using the state at the end of the perturbed parameter spin-ups, we ran transient simulations for each site and parameter combination starting in 1850 and invoking transient CO$_{2}$ and Nitrogen deposition until the end of the observed meteorological forcing.
-We ran additional CO$_{2}$ and Nitrogen deposition experiments by increasing the CO$_{2}$ concentration to 550ppm.
-For the Oak Ridge FACE site, we only increased CO$_{2}$ in the growing season, in line with the experimental protocol applied there and using the observed CO$_{2}$ time-series.
-For all other sites, we increased CO$_{2}$ to 550ppm year round starting in 1997**.
-For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1997, so that the results can be compared with those reported by Wieder et al.
-(in prep).
+For each site/parameter/value combination, we also ran one CO$_{2}$ and one Nitrogen deposition experiment.
+To allow comparison with \cite{wieder}, we increased CO$_{2}$ to 550ppm, starting in 1998.
+For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1998.
+In line with \cite{wieder}, we assessed the impact of changes after 15 years of fertilization in both cases.
 
 \subsection{Perturbed Physics Ensemble (PPE)}
+The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5.
+Here, we are focused on the recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization.
+Therefore,
+we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
+The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.
 
-The CLM5 has a very large number of parameters to which the C and N cycles are potentially sensitive.
-Here we focused on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
-We choose a larger set of parameters for one-at-a-time analyses, followed by a smaller set of these to which we subject the model to global sensitivity tests.
-Knowledge of model structure as well as previous investigations of the CLM model informs our choice of parameters.
 Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model.
@@
 in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here.
+We expect that further focused analysed will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.
 
-In the first instance, we conduct one-at-a-time (OAAT) perturbations of the parameters around a mean state, to allow a straightforward visualization and interpretation of the results.
+To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around a mean state, which allows more straightforward visualization and interpretation of the results.
 For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation).
+We recognize that a global parameter sensitivity analysis in this realm would also be of interest, and indeed have conducted many such analyses, but consider that assessment of the output of a global as well as a OAAT analysis of the entire domain of results presented here would reduce the opportunity for understanding the model parameter space in depth, as is the goal of our study.
 
@@
 
-The parameters included in our sensitivity analysis (Table \ref{table_parameters}) are: \\
+The parameters included in our sensitivity analysis are both detailed in (Table \ref{table_parameters}), and described in more depth below. \\
 
-1) \textbf{Specific Leaf Area}, \emph{`SLA'}, which is a critical parameter determining the area of canopy (in m$^{2}$) derived from one gram of leaf biomass.
-This parameter is known to be one of the primary elements of the leaf economics spectrum (\cite{wright2004}), and is well-characterized by the TRY database (\cite{kattge2011}).\\
+1) \textbf{Specific Leaf Area}, \emph{`SLA'}, is a critical parameter, common to almost all land surface models, which determines the area of canopy (in m$^{2}$) derived from one gram of leaf biomass.
+High SLA values correspond to thinner, 'cheaper' leaves, which are typically associated with shorter leaf lifespan and lower nitrogen per unit leaf area (\cite{wright2004}).\\
 
-2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, the ratio of fine root biomass to leaf area.
-Unlike in previous versions of the CLM, fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth.
+2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf area.
+Unlike in previous versions of the CLM, the fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth.
 \\
 
-3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, reflects the balance between accumulation of stem and woody biomass, and leaf area.
-Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions.\\
+3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody biomass, which in CLM5 is a constant fraction of leaf allocation.
+Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions.
+Woody biomass in CLM5 has no function and thus there is no cost to plants associated with low values of this parameter\\
 
 4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.
-These six parameters were previously defined by \cite{brzostek2014}.
-Here and in CLM5 in general, we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}.
-These parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
+These six parameters were defined and calibrated by \cite{brzostek2014}.
+In CLM5 we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}.
+The \emph(`n\_costs'}) parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
 \\
 
-5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`'fracfixers'}.
+5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}.
 This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.
@@
 
-6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{grperc}, was revised substantially in CLM5 based on Atkin O.
+6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{`grperc'}, was revised substantially in CLM5 based on methodolicaly issues with the observation of growth respiration from Atkin O.
 (pers comm.) from 0.3 to 0.11.
+CLM5 also adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significnatly increases maintenance respiration costs.
 Here we test the impacts of this change on overall carbon budgets.\\
@@
 8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}.
-This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, assessed in this set of experiments.
-\\
+This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, since it determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit (assuming no soil moisture lmitations).\\
 
@@
 
-10) \textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is another ecosystem-level property defined by the FUN model and allowing the proportions of the fungal symbiotic pathways to be modified.
-Default parameters are as provided by \cite{shi2016}.
-This capacity was not fully explored in the CLM5 development process and so here we assess its importance.
-\\
+10) \textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is an ecosystem-level property defined by the FUN model which allows the proportions of the fungal symbiotic pathways to be modified.
+Default parameters are as provided by \cite{shi2016}.\\
 
-11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in Appendix 1, which determine the response of tissue CN ratios to depletion of N in both the environment and the plant tissues themselves.
+11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CN technical note, determine the response of tissue CN ratios to depletion of N in both the environment and the plant tissues themselves.
 
@@
 We did not investigate any parameters relating to the crop model (Lombardozzi et al.
-in prep) nor to plant harvest, fire, dust, soil biogeochemical cycling (the latter being is largely unaltered from CLM4.5, \cite{koven2013} apart from modifications presented by \cite{lawrence}.
+in prep) nor to plant harvest, fire, dust, or soil decomposition, (the latter is largely unaltered from CLM4.5 as per \cite{koven2013}.
 
@@
 
-\emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY database, and their distributions are log-normally distributed.
-Thus, we take the average log-normal standard deviation range from TRY and apply it across all PFTs.
-Thus the lower bounds are closer to the mean than the higher bounds (Table \ref{table_ranges}).
+\emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY plant trait database(\url{www.try-db.org} and we take their values from the analysis by (\cite{kattge2011}) of their mean and distribution.
+For both parameters, values are log-normally distributed.
+We here take the PFT average log-normalised standard deviation range from TRY and apply it across all PFTs (thus the lower bounds are closer to the mean than the higher bounds, Table \ref{table_ranges}).
 The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across the whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake).
@@
 \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}.
-The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood, thus we used ranges that span the response limits of these parameters, particularly for \emph{cn\_flex\_c} which we explored across four orders of magnitude to investigate the impacts of leafCN flexibility.
-The allometric parameters ***.
-Nitrogen parameters.
+The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood.
+Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization (thus, these results should not be interpreted as covering a constrained range of variation).
+The same issues apply to $N_{costs}$.
 
 \subsection{Analysis}\\
-We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the fertilized simulations for the years 2001-2003.
+We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the fertilized simulations 15 years after fertilization began.
 For some parameter settings, the constraint on growth placed the system in a state close to death.
 Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in the relationship between parameters and  with  to environmental forcing.
-Therefore, for those parameter combinations which cause LAI to be less than 80\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plots, so as to focus on the responses of parameter combinations that give realistic initial vegetation conditions.
+Therefore, for those parameter combinations which cause LAI to be less than 60\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plots, so as to focus on the responses of parameter combinations that give more reasonable initial vegetation conditions.
 
 \section{Results}
-The analysis conducted involved four sites, with 13 parameters perturbed to 4 different values, each of those with an increase CO$_{2}$, N and control run.
+The analysis used four sites, with 13 parameters perturbed to 4 different values, each of those with an increase CO$_{2}$, N and control run.
 Including the default simulations, this is 672 different run configurations.
-In this analysis, we look at the parametric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon), first at the impacts of parameteric variation on the system state and then on the responses to carbon and nitrogen fertilization.
+In this analysis, we look at the parametric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon).
+In the first section, we look at the impacts of parametric variation on the system state and secondly on the responses to carbon and nitrogen fertilization.
+We do not focus on the goodness of fit to observations at these sites, because our aim is to discern the primary parametric controls over CO$_{2}$ and N fertilization responses, and including extensive data-fitting and calibration experiments would detract from this goal.
+
+\subsection{Overall system state}
+Figure \ref{NR1 state} shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Niwot Ridge site (which we choose to highlight as it shows some of the more complex parameter responses).
+The other site state variation figures are presented in the Supplementary Information (SI \ref{CAX state}, \ref{HVF state}, \ref{MET state}).
+
+In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
+(for Caxiuana, 3.0 v.s.
+5.3 \cite{fisher2007}, for Niwot Ridge, 1.7 vs.
+4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs.
+3.5 \cite{williams1996} and for Metolius, 0.5 (modeled) v.s.
+1.5 (\cite{spadavecchia2011}).
+This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect biases between the use of reanalysis products used to drive the model and site-specific meteorological data.
+
+To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
+These indicated that, with the resulting higher LAI's, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAIs.
+For example, at Harvard Forest, the high leaf allocation  simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.43, 1.66, 1.47 for GPP, NPP and LAI in the control vs.
+1.18, 1.17 and 1.22 for the high LAI case) and a lower N response as well (1.1, 1.26, 1.21 for the control GPP, NPP, LAI, and 1.05, 1.20 and 1.15 for the high LAI case).
+Similar results were found at all other sites.
+Thus, by virtue of their lower LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology (analysis tbd).
 
 \subsection{Parametric control over system state}
-Figure \ref{NWRstate} shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Niwot Ridge site.
-The other site state variation figures are presented in the Supplementary Information (Figures 1-3).
-
-In general, we found that leaf are indicies were often low for these sites, relative to both observations at the sites, and to the magnitude of LAI where gas exchange fluxes are typically known  to saturate as a function of intercepting leaf area (4-5 m$^{2}$ m$^{-2}$.
-This is a surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}.
-A seperate set of simulations (not shown here) conducted with higher allocation to leaves (lower stem and root allocation parameters) indicated that, with the resulting higher LAI's, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAIs.
-In the default set of simulations the same parameters dominate the responses for differnt variables, given the intrinsic correlation between LAI and GPP, but for saturating LAI conditions, a greater disparity between the dominant parameters is obtained.
-
-Here we assess the impacts of each parameter on the overall output both at the Niwot Ridge site shown and the other sites in the SI figures.
 
@@
 At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI,  and on leaf N concentrations (Figure \ref{NR1 stat}), since it is used directly in the calculation of both of these quantities.
-Whether this impact on LAI is translated into an impact on GPP depends on the degree of saturated of productivity at the default LAI.
 
-For GPP, the impact of SLA is always positive, but the degree of it's impact shifts between sites, as it integrates a trade-off between an increase in LAI and the deployment of thinner leaves (lower carbon content per unit area) and subsequent decreases the total N per unit area of leaf reducing photosynthetic capacity
+Whether this impact on LAI is translated into an impact on GPP depends on the degree of saturation of productivity at the control LAI.
+
+For GPP, the impact of SLA is always positive, given increasing LAI.
+There is a large impact at  but the degree of it's impact shifts between sites, as it integrates a trade-off between an increase in LAI and the deployment of thinner leaves (lower carbon content per unit area) and subsequent decreases the total N per unit area of leaf reducing photosynthetic capacity
 
@@
 SLA driven increases in LAI can in principle also reduce NPP for respiratory reasons.
-At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the intrisic trade-offs between costs and benefits of thinner leaves.
+At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the intrinsic trade-offs between costs and benefits of thinner leaves.
 
@@
 As is the case for N$_{costs}$, with high fixing capacity N\_uptake expenditure is lower, N uptake high, and NPP is moderately increased, resulting in increases in LAI and a slight response of GPP to that change.
-These systems are only subject to transiently ramping CO$_{2}$$, and are thus closer to equilibrium than the step-change fertilizations, suggesting that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.
+These systems are only subject to transiently ramping CO$_{2}$, and are thus closer to equilibrium than the step-change fertilizations, suggesting that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.
 
@@
 
+\nocite{*}
+\bibliography{aguCLM5_bibtex}
+\clearpage
+
 \pagebreak
@@
  \hline
- Parameter Name & Meaning & Units & Min & Max \\
+ Parameter Name & Meaning & Units \\
   \hline
@@
      medlyn\_slope  & Slope of stomatal conductance curve & g/g \\
-      lmr\_intercept & intercept of leaf N /respiration relationship &\\
-      frac\_ecto & Fraction of active uptake that is ectohycorrhizal & \\
+      lmr\_intercept & Intercept of leaf N /respiration relationship & \\
+      frac\_ecto & Fraction of active uptake that is ectohycorrhizal & - \\
       cn\_flex\_a & Parameter of N flexibilty algorithm& \\
@@
 
-\section{Figures}
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Niwot Ridge.}
+     \label{NR1 state}
+ \end{figure}
 
   \begin{figure}[h]
-     \caption{Sensitivity of system state to parameter perturbation at Niwot Ridge site.}
-     \label{NWRstate}
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
+     \label{NR1 celev}
   \end{figure}
-Figure 1: Sensitivity analysis of state of CLM at site to N parameters.
 
-Figure 2: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @ORNL
+  \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
+     \label{NR1 ndep}
+  \end{figure}
 
-Figure 3: Sensitivity analysis of state of CLM at site to N parameters @BCI
+ \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
+     \caption{Influence of parameteric variation on the impact of 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
+}
+     \label{GPP CO2 and N respones 2001}
+  \end{figure}
 
-Figure 4: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @BCI
+ \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_NPP1__p2012.png}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on net primary productivity (NPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{NPP CO2 and N respones 2001}
+  \end{figure}
 
-Figure 5: Sensitivity analysis of state of CLM at site to N parameters @CAX
+  \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TLAI1__p2012.png}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf area index (LAI), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{LAI CO2 and N respones 2001}
+  \end{figure}
 
-Figure 6: Sensitivity analysis of the response of CLM5 to CO2 \& Ndep fertilization @CAX
+  \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{VEGC CO2 and N respones 2001}
+  \end{figure}
 
-Figure 7: Operation of the flexCN ratio module.
-
-Figure 8: Operation of the leaf N retranslocation module
-
-Figure 9: Details on the N uptake in FUN
+  \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{Leaf N CO2 and N respones 2001}
+  \end{figure}
 
 \section{To Do}
-Add data from the CO2 experiment to figures.
+Make a figure with a bunch of extra outputs for the state at Niwot Ridge (CUE, fraction NPP nuptake, BTRAN)
 
-Add a figure of the transient responses through time to CO2 and N...
+Make SI figures with the high leaf run (state for Niwot?)
 
-Look at the transient response to ramping CO2 and see if its different to the step response
-
-Also add LUNA on and off simulations to figures.
-
-Make a figure with a bunch of extra outputs for the state at Niwot Ridge (CUE, fraction NPP nuptake, BTRAN)
-\nocite{*}
-
-Make some SI figures with the high leaf run (state for Niwot?)
-
-Make all on the same figure figure
-
-Slightly re-format the fert figures. (dashes, GPP label)
-Check the CN figures. They don't match the NWT figures.
-
-\bibliography{aguCLM5_bibtex}
+Site details
 

### 2018-11-15 main.tex
Updates from ShareLaTeX

@@
 
+\subsubsection{Low leaf area investigations}
 In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
@@
 Thus, by virtue of their lower LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology (analysis tbd).
+
+\subsubsection{Water and N limitation across sites}
+Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control (unfertilized) parameter ensemble.
+Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of prognostic leaf water potential, which impacts gas exchange processes in a similar manner to the $B_{tran}$ parameter in CLM4.5, (in that it is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$).
+Because leaf water potential changes diurnally, we now use the minimum term to calculate relative water stress between sites.
+For N limitation, the fraction of total NPP which is used to pay for nitrogen uptake, is a direct indicator of how N stress affects the carbon budget.
+
+The Caxiuana site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress.
+Niwot Ridge indicated substantial amounts of drought ($B_{tran,mn}$ ~0.5), and N limitation, spending 27\% of carbon on nutrient acquisition.
+Harvard Forest site had a moderate drought stress ($B_{tran,mn}$~90\%), and moderate N limitation (~15\% of assimilation spent on uptake).
+The Metolius site had some drought stress ($B_{tran,mn}$ ~86\%) and very little nutrient stress.
+At the sites with dought stress, it is evident that low values of the Medlyn parameter (which reduce stomatal conductance) relieved drought stress as expected from first principles, as do high C:N values, which reduce photosynthetic capacity, and thus also stomatal conductance.
+Lower root allocation increase drought stress, via reduction in root uptake capacity.
 
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/btran_npp.png}
+     \caption{Response of drought index ($B_{tran,mn}$ and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{btran state}
+ \end{figure}
+
+  \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}

### 2018-11-16 figures.tex
Updates from ShareLaTeX

@@
 
-  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{VEGC CO2 and N respones 2001}
-  \end{figure}
-
-  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{Leaf N CO2 and N respones 2001}
-  \end{figure}
-
 \end{document}

### 2018-11-19 SIfigs.tex

@@
 
+    \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{VEGC CO2 and N respones 2001}
+  \end{figure}
+
+  \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{Leaf N CO2 and N respones 2001}
+  \end{figure}
 \end{document}

### 2018-11-19 main.tex

@@
 Gordon Bonan\affil{1},
-David Lawrence\affil{1}
+David M. Lawrence\affil{1}
 }
@@
 in prep), the land use components by (Lawrence P et al.
-in prep), land surface hydrology by (Swenson et al.
-in prep) and plant hydraulics by (Kennedy et al.
-in prep).
+in prep), and plant hydraulics by (Kennedy et al., 2018).
 
@@
 Wieder et al.
-(in prep) report the global-scale implications of the full model update and note that the CLM5 responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.
+(in prep) report the global-scale carbon cycle implications of the full model update and note that the CLM5 carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.
 In this paper, we describe the new components of the biogeochemical cycling, and investigate the features of the new model that control the model state and biogeochemical forcing responses in depth, using single-site experiments and parametric sensitivity analyses.
@@
 One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}).
-The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon.
-Further, the cost of each of each uptake pathway is variable in both time and space.
+The FUN model operates on the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon.
+Further, the cost of each uptake pathway is variable in both time and space.
 FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them.
@@
 Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency).
-The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).
+The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for nitrogen uptake, ($C_{nuptake}$).
 
@@
 
-Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:
+Further, the nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:
 
@@
 The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity ((\cite{wieder2015} illustrated the impacts of uncertainty in this function).
-FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compmared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
+FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
 Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation.
 Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground.
-In most cases, N fixation is preferred when soil N is low, and so and soil N builds up much quicker.
+In most cases, N fixation is preferred when soil N is low, and so soil N builds up quickly.
 In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the carbon use efficiency and survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.
@@
 \subsubsection{FLEXCN}
-The results of the FACE model data comparison exercise illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).
+The results of FACE model-data comparison exercises illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).
 These changes were not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
@@
 
-The FUN model was originally conceived with a static tissue CN ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target CN ratio.
-In CLM5, one of the aims of the development process was to incorporate a varying CN ratio.
+The FUN model was originally conceived with a static tissue C:N ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target C:N ratio.
+In CLM5, one of the aims of the development process was to incorporate a varying C:N ratio.
 To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required).
-The degree to which tissue CN content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments in light of limitations (including \cite{zaehle2010},\cite{ghimire2016}).
+The degree to which tissue C:N content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments under N limited conditions (including \cite{zaehle2010},\cite{ghimire2016}).
 Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content.
-When N is expensive in the environment, less C is spent on uptake, increasing CN ratios where N is scarce.
+When N is expensive in the environment, less C is spent on uptake, increasing C:N ratios where N is scarce.
 When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high.
@@
 
-Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible.
+Because of the increased number of prognostic processes with the CLM5, many alternative representations of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible.
 We thus consider that an exploration of the CLM5 parameter space is also equivalent to an exploration of a large part of the relevant model structural space.
@@
 The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5.
-Here, we are focused on the recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization.
+For this manuscript, we focus on parameters included in recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization.
 Therefore,
@@
 in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here.
-We expect that further focused analysed will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.
+We expect that further focused analyses will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.
 
 To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around a mean state, which allows more straightforward visualization and interpretation of the results.
-For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation).
-We recognize that a global parameter sensitivity analysis in this realm would also be of interest, and indeed have conducted many such analyses, but consider that assessment of the output of a global as well as a OAAT analysis of the entire domain of results presented here would reduce the opportunity for understanding the model parameter space in depth, as is the goal of our study.
+For each parameter, for each site, we conduct four simulations, with two simulations each using parameter values above and  below the default value (as well as the default simulation).
+We recognize that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses, but consider that assessment of the output of a global as well as a OAAT analysis of the entire domain of results presented here would reduce the opportunity for understanding the model parameter space in depth, as is the goal of our study.
 
@@
 
-The parameters included in our sensitivity analysis are both detailed in (Table \ref{table_parameters}), and described in more depth below. \\
+The parameters included in our sensitivity analysis are both detailed in Table \ref{table_parameters}, and described in more depth below. \\
 
@@
 2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf area.
-Unlike in previous versions of the CLM, the fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth.
+Unlike in previous versions of the CLM, the fine root biomass in CLM5 affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth.
 \\
@@
 Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions.
-Woody biomass in CLM5 has no function and thus there is no cost to plants associated with low values of this parameter\\
+Woody biomass in CLM5 has no function and thus there is no cost to plants associated with low values of this parameter.\\
 
@@
 These six parameters were defined and calibrated by \cite{brzostek2014}.
-In CLM5 we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}.
+In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}.
 The \emph(`n\_costs'}) parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
@@
 5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}.
-This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.
+This parameter is the fraction of the assimilation (gross primary productivity - autotrophic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.
 \\
 
-6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{`grperc'}, was revised substantially in CLM5 based on methodolicaly issues with the observation of growth respiration from Atkin O.
-(pers comm.) from 0.3 to 0.11.
-CLM5 also adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significnatly increases maintenance respiration costs.
+6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{`grperc'}, was revised substantially in CLM5 to reflect new estimates that account for methodological issues with the observation of growth respiration (Atkin O., pers comm.) from 0.3 to 0.11.
+CLM5 also adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significantly increases maintenance respiration costs.
 Here we test the impacts of this change on overall carbon budgets.\\
@@
 7) \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}.
-CLM5 uses the concept of a 'target' leafcn ratio, around which flexibility is allowed, given the environmental costs of Nitrogen.
-LeafCN ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$) \\
+CLM5 uses the concept of a 'target' leaf C:N ratio, around which flexibility is allowed, given the environmental costs of Nitrogen.
+Leaf C:N ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$).
+\\
 
 8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}.
-This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, since it determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit (assuming no soil moisture lmitations).\\
+This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependent), is critically linked to the response of ecosystems to CO$_{2}$, since it determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit (assuming no soil moisture limitations).\\
 
@@
 
-11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CN technical note, determine the response of tissue CN ratios to depletion of N in both the environment and the plant tissues themselves.
+11,12,13) \textbf{Parameters a,b and c of the flexible C:N model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.
 
@@
 The Metolius site had some drought stress ($B_{tran,mn}$ ~86\%) and very little nutrient stress.
-At the sites with dought stress, it is evident that low values of the Medlyn parameter (which reduce stomatal conductance) relieved drought stress as expected from first principles, as do high C:N values, which reduce photosynthetic capacity, and thus also stomatal conductance.
-Lower root allocation increase drought stress, via reduction in root uptake capacity.
+
+These differences in the initial conditions at the sites have implications for the parameteric responses, as documented for each parameter in turn in the section below.
 
@@
 
-Whether this impact on LAI is translated into an impact on GPP depends on the degree of saturation of productivity at the control LAI.
-
-For GPP, the impact of SLA is always positive, given increasing LAI.
-There is a large impact at  but the degree of it's impact shifts between sites, as it integrates a trade-off between an increase in LAI and the deployment of thinner leaves (lower carbon content per unit area) and subsequent decreases the total N per unit area of leaf reducing photosynthetic capacity
-
-The impact on NPP however, can be negative (e.g.
-Metolius), since the CLM5 can generate an under-storey of shaded leaves, and with it the possibility of those leaves being in negative carbon balance.
-SLA driven increases in LAI can in principle also reduce NPP for respiratory reasons.
-At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the intrinsic trade-offs between costs and benefits of thinner leaves.
+For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area and leaves, reducing photosynthetic capacity.
+The impact of SLA on NPP can however be negative (e.g.
+Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance.
+At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.
 
 Impact of \emph{froot\_leaf}\\
-Root:leaf ratio (\emph{froot\_leaf}) a direct effect on LAI, via changes in allocation fractions to different tissues, an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, and (newly introduced in CLM5) also affects the uptake of both nitrogen and water via it's impact on N uptake costs and hydraulic resistance terms respectively.
+Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) affects the uptake of water via root biomass and 4) of nutrients via root-mediated n uptake costs.
 The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.
-Both types of effect are seen here, with the Niwot Ridge and Metolius site suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios, but the Harvard forest and Caxiuana sites showing increases.
+Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios.
+Harvard forest and Caxiuana both show increases in all variables.
+In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at the low productivity sites.
 
@@
 
-Decreases in stem:leaf ratio (stem\_leaf) below their default condition also caused a decline in leaf N per unit area (figure **e).  The causes of this behaviour are still not understood **
+Decreases in stem:leaf ratio (stem\_leaf) below their default condition also caused a decline in leaf N per unit area, potentially on account of the higher demand for N uptake (illustrated clearly for Niwot Ridge in Figure \ref{btran state}) when the LAI is higher.
 
 Impact of \emph{Ncosts}\\
-The first-order impact of N$_{costs}$ is to increase the respiration losses associated with nitrogen uptake.
-Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of N$_{costs}$ on the system state with the exception of the Niwot Ridge site, which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low.
+The first-order impact of $N_{costs}$ is to increase the respiration losses associated with nitrogen uptake.
+Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of $N_{costs}$ on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low.
 Other sites showed more muted responses in the vegetation state.
@@
 In all cases here, it has a relatively moderate and positive impact on the overall system state, in particular at the Niwot and Harvard forest sites.
-As is the case for N$_{costs}$, with high fixing capacity N\_uptake expenditure is lower, N uptake high, and NPP is moderately increased, resulting in increases in LAI and a slight response of GPP to that change.
-These systems are only subject to transiently ramping CO$_{2}$, and are thus closer to equilibrium than the step-change fertilizations, suggesting that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.
+As is the case for $N_{costs}$, with high fixing capacity N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that change.
 
 Impact of \emph{leafcn}\\
-Changing the target leaf C:N ratio \emph{leafcn} over one standard deviation from the mean TRY values had substantial impacts on all the output variables.
-As expected, higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.
-In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge.
+Changing the target leaf C:N ratio \emph{leafcn} over one standard deviation from the mean TRY values had substantial impacts on all the output variables at all sites.
+Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.
+In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).
+We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
 
-For NPP, the corresponding  decreases in respiration rates at high CN somewhat ameliorated this effect, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result.
-At Niwot, an optimum leaf CN ratio was apparent in the range of the default values, indicating the possibility that other sites might be operating distant from their theoretical optima according to the model, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
-We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations.
-
-In general, for a given change in NPP, the relative change in LAI was lower than for other parametric modification, reflecting the increased spending on nitrogen uptake inherent in the high $N_{area}$ simulations.
-Interestingly, despite a lower NPP for low CN values at Niwot Ridge, LAI values were higher, for reasons not yet understood ***
+For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result.
+At Niwot, an optimum leaf CN ratio was apparent.
+For NPP this was around the default value, but for LAI it was 0.5s.d.
+lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.
 
@@
 Impact of \emph{medlyn\_slope}\\
-The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties.
-At the Caxiuana site, the very lowest values caused a complete death of the ecosystem.
-It is unclear why this occured, since the impacts can be direct, on the carbon balance itself (not enough GPP to cover costs), but also on the leaf temperature (via surface energy balance issues) and hence the respiration rates.
-It is hard to discern the exact reasons, given the feedback inherent between all of these properties, but no impact on the carbon balance is found at intermediate levels.
+The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (including on leaf temperature and water stress).
+At the Caxiuana site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem.
+In figure
+\ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses).
+It is worth noting that these same simulations typically had extremely low photosynthesis, particularly at Caxiuana and Metolius.
 
 At Niwot, a clear optimum was found in the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough.
-At Metolius and Harvard forest, only lower value affected the photosynthetic rates, indicating co-limitation by another process, and the absence of impacts of excessive plant transpiration on the system (note that \cite{kennedy} and \cite{dagon} discuss surface energy and plant hydrodynamic functioning of the model in much greater depth).
+At Metolius and Harvard forest, only the lower values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.
+(note that \cite{kennedy} and \cite{dagon} discuss surface energy and plant hydrodynamic functioning of the model in greater depth).
 
@@
 The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization.
-For the high leaf area simulations (now shown **), the impacts of maintenance respiration parameterization were much higher, reflecting the higher amount of respiring tissue in those simulations.
 At Caxiuana, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.
@@
 Impact of \emph{cn\_flex\_a}\\
-\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake.
-(see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details.
+\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details).
 Thus, low values of \emph{cn\_flex\_a}  results in reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.
@@
 Impact of \emph{cn\_flex\_b}\\
-\emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, but has limited impacts on the system state here.
-At low values, there is some indication at Niwot Ridge that it might increase LAI and vegetation carbon, indicating a decrease in the N uptake expenditure and in leaf N concentrations.
-The limited impact at other sites suggests that Ncosts are rarely high enough to trigger reductions in N expenditure (notwithstanding the point the this algorithm is a placeholder for a more evidence-based depiction of stoichiometric flexibility).
+Increasing \emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, and also has limited impacts on the system state here.
 
 Impact of \emph{cn\_flex\_c}\\
- Low values of \emph{cn\_flex\_c} affects the degree to whch plants attempte to maintain tissue CN values near to their target when environmental costs are varying.
+ \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue CN values near to their target when environmental costs are varying.
 For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.
@@
 To illustrate the relationships between N and CO$_{2}$ fertilization, and to reduce the dimensionality of the output, figures \ref{GPP CO2 and N respones 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the response for GPP, NPP and, LAI respectively.
-Vegetation carbon and $N_{area}$ are in the supplemental infomation (\ref{VEGC CO2 and N respones 2001}, \ref{Leaf N CO2 and N respones 2001}).
-For Niwot Ridge, figures showing the N and CO$_{2}$ responses are also shown (Figure \ref{NR1 ndep} and \ref{NR1 celev}), while the corresponding figures for the other sites are in the SI.
+Vegetation carbon and $N_{area}$ are in the supplemental information (\ref{VEGC CO2 and N respones 2001}, \ref{Leaf N CO2 and N respones 2001}).
+For Niwot Ridge, figures showing the isolated N and CO$_{2}$ responses are also included here (Figure \ref{NR1 ndep} and \ref{NR1 celev}), while the corresponding figures for the other sites are in the SI.
 
@@
 The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.
-The default range of CO$_{2}$ fertilization impacts on NPP, for example, ranged from 1.22 to 1.66 at Harvard and Caxiuana respectively, with Niwot having a lower range (1.28) and Metolius higher (1.62).
-For Nitrogen responses, Caxiuana had a negative response (0.96), Niwot had almost no response (0.98) while Harvard and Metolius were positive (1.25 and 1.09 respectively).
-Impacts on GPP were all positive, from  (albeit lower for the Caxiuana and Niwot sites) through increases in leaf N content, and impacts on LAI were all positive, reflecting lower expenditure on N uptake (1.05 to 2.09).
-Interestingly, the highest impacts on LAI were at Niwot, which has almost no change in NPP, reflecting a large initial expenditure on nitrogen uptake ameliorated by fertilization.
+
+For the default parameterization, the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.
+For LAI (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) growth at the Metolius site.
+Leaf N concentrations also dropped the furthest at this site (0.92; CAX, 0.93; NWR, 0.93 HVF, 0.90 MET) suggesting that expenditure on N uptake might be limited by environmental costs at this site (thus leaf CN deviates further from the target).
+
+For Nitrogen responses, impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET) (albeit lower for the drought-stressed Caxiuana and Niwot sites).
+For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET) increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts as some sites, but impacts on LAI were all positive, reflecting lower costs of N uptake and a greater capacity to acquire N to build tissues and retain carbon for growth (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET)
 
@@
 Across the four sites, parametric impacts were very varied, when compared to the relatively consistant impact of parameters on the system state.
-A majority of parameters show no consistant relationship with CO$_{2}$ or N deposition responses, and the range of their impacts vary over an order of magnitude between sites.
-We describe here the major features of how parameteric variation controls the behaviour of the model under these fertilization scenarios.
+A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses, and the range of their impacts varied over an order of magnitude between sites.
+Thus, the parameter space of the model is complex with regard to these changing initial conditions and fertilization experiments.
+Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.
 
 Impact of \emph{slatop}\\
+ tbd. This one is making my brain hurt.
 
@@
 Despite it’s  impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses.
-At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fetilization is lower since the feedback from LAI to GPP is less proncounced.
+At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP is less pronounced.
 
 Impact of \emph{stem\_leaf}\\}
-Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuana and Harvard Forest.
-Initial LAI was larger for these cases for these sites, explaining the decease in fertilization response as above.
-However, at Niwot Ridge, CO$_{2}$ fertilization was larger with lower \emph{stem\_leaf} ratios.
-At Niwot, however, a larger control LAI was associated with lower $N_{area}$, GPP and NPP, suggesting a high nitrogen limitation associated with these states.
-Nonetheless, fertilization responses of GPP here are very high, (x1.8) in response to quite small changes in $N_{area}$ (~2\%)  and LAI (~22\%) (Figure \ref{NPP CO2 and N respones 2001}).
-Further, Nitrogen increases LAI substantially, buft only has a very limited impact on GPP and NPP at Niwot Ridge, suggesting that the fluxes might be limited by water availability, and thus explaining the very high CO$_{2}$ sensitivity.
+Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuana and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and\ref{LAI CO2 and N respones 2001}).
+Initial LAI was larger in these cases, explaining the decease in fertilization response as above.
+
+At Niwot Ridge, where the low values of \emph{stem\_leaf} were associated with very high LAI, low carbon use efficiency, high N expenditure and low $N_{area}$, CO$_{2}$ fertilization had a large impact on (the very low) GPP and NPP (reflecting the high water limitation for these high LAI cases, figure\ref{btran state}) but has no impact on LAI itself, (Figure \ref{NR1 celev}), reflecting the still high N limitation).
+Correspondingly, adding Nitrogen increases LAI substantially, but only has a very limited impact on GPP and NPP, given water limitation.
 
 Impact of \emph{N\_costs}\\
-Given that they impact on the respiration expended on N uptake and the availability of carbon for tissue growth, the impact of Nitrogen costs is most clear for NPP and LAI (Figure \ref{NPP CO2 and N respones 2001}, \ref{LAI CO2 and N respones 2001}}).
-At Caxiuana, Niwot and Harvard forest, high initial N costs resulted in stimulation of the N fertilization response, as expected from the alleviation of the higher N limitation in these cases.
-There was also a slight increase in the CO$_{2}$ fertilization response (maybe on account of extra carbon alleviating some of the N limitation in those systems).
-For very low values, at Caxiuana and Niwot, N responses increased, counter-intituvely.
-At both these sites, at very low N costs, plants appear to make leaves with higher N content (as per the algorithm which allows a higher $N_{area}$ to be achieved at low N cost).
-However, the total amount of N in the system remains the same with low N costs, and so these cases do not have higher leaf area.
+ \emph{N\_costs} primarily impact the respiration expended on N uptake and the availability of carbon for tissue growth thus their impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}}).
+
+ At Caxiuana, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
+There was also a slight increase in the CO$_{2}$ fertilization response at high \emph{N\_costs}, perhaps as this high N limitation was alleviated by extra assimilated carbon.
+
+ For very low \emph{N\_costs}, at Caxiuana and Niwot, N responses increased, counter-intuitively.
+In these cases, plants appear to make leaves with higher N content (Figure \ref{Leaf N CO2 and N respones 2001}), as per the algorithm which allows a higher $N_{area}$ to be achieved at low N cost.
+However, the total amount of N in the system (the balance of N inputs and outputs) remains the same with low N costs, and so these cases do not have higher leaf area.
 When more total N is added, the leaf area can expand, and thus the N concentration declines.
@@
 
-  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{VEGC CO2 and N respones 2001}
-  \end{figure}
-
-  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{Leaf N CO2 and N respones 2001}
-  \end{figure}
-
 \section{To Do}
-Make a figure with a bunch of extra outputs for the state at Niwot Ridge (CUE, fraction NPP nuptake, BTRAN)
-
-Make SI figures with the high leaf run (state for Niwot?)
+Maybe include the delta in N expenditure.
 

### 2018-11-19 figures.tex
Updates from ShareLaTeX

@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1_2012.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
      \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}

### 2018-11-19 main.tex

@@
 \begin{document}
-
+%tentative author list...
 \authors{Rosie Fisher\affil{1},
@@
 Chonggang Xu\affil{4},
-Ashehad Ali\affil{1},
 Joshua Fisher\affil{5},
@@
 For this experiment we utilized six different experimental sites.
-Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest) **tbd**.
+One in tropical forests (Caxiuan\~a), one in alpine forest (Niwot Ridge), one in temperate forest (Harvard Forest) and one in a seasonally dry evergreen forest (Metolius Forest).
 
@@
 Therefore,
-we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
+we focus on parameters of particular https://www.overleaf.com/project/5b05c31494d9960223796bc3relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
 The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.
@@
 
-\subsubsection{Parameter Choices}
+\subsubsection{Parameters Selected for PPE}
 
@@
 
-\subsubsection{Parameters not used}
+\subsubsection{Parameters not included in PPE}
 
@@
 We did not investigate any parameters relating to the crop model (Lombardozzi et al.
-in prep) nor to plant harvest, fire, dust, or soil decomposition, (the latter is largely unaltered from CLM4.5 as per \cite{koven2013}.
+in prep) nor to wood harvest, fire, dust, or soil carbon decomposition, (the latter is largely unaltered from CLM4.5 as per \cite{koven2013}.
 
@@
 \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}.
-The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood.
+Reasonable ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood.
 Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization (thus, these results should not be interpreted as covering a constrained range of variation).
-The same issues apply to $N_{costs}$.
+The same issues apply to $N_{costs}$ parameters.
 
@@
 In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
-(for Caxiuana, 3.0 v.s.
+(for Caxiuan\~a, 3.0 v.s.
 5.3 \cite{fisher2007}, for Niwot Ridge, 1.7 vs.
@@
 1.5 (\cite{spadavecchia2011}).
-This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect biases between the use of reanalysis products used to drive the model and site-specific meteorological data.
+This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.
 
 To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
-These indicated that, with the resulting higher LAI's, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAIs.
+These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAI levels.
 For example, at Harvard Forest, the high leaf allocation  simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.43, 1.66, 1.47 for GPP, NPP and LAI in the control vs.
@@
 Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control (unfertilized) parameter ensemble.
-Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of prognostic leaf water potential, which impacts gas exchange processes in a similar manner to the $B_{tran}$ parameter in CLM4.5, (in that it is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$).
+Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of prognostic leaf water potential, which impacts gas exchange processes in a similar manner to the $B_{tran}$ parameter in CLM4.5 (in that it is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$).
 Because leaf water potential changes diurnally, we now use the minimum term to calculate relative water stress between sites.
@@
 
-The Caxiuana site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress.
+The Caxiuan\~a site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress.
 Niwot Ridge indicated substantial amounts of drought ($B_{tran,mn}$ ~0.5), and N limitation, spending 27\% of carbon on nutrient acquisition.
@@
 
-These differences in the initial conditions at the sites have implications for the parameteric responses, as documented for each parameter in turn in the section below.
+These differences in the initial conditions at the sites have implications for the parametric responses, as documented for each parameter in turn in the section below.
 
@@
 
-For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area and leaves, reducing photosynthetic capacity.
-The impact of SLA on NPP can however be negative (e.g.
-Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance.
+For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area of leaves which reduces photosynthetic capacity.
+The impact of SLA on NPP can however be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance.
 At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.
@@
 Impact of \emph{froot\_leaf}\\
-Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) affects the uptake of water via root biomass and 4) of nutrients via root-mediated n uptake costs.
+Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation???
+through root-mediated n uptake costs.
 The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.
 Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios.
-Harvard forest and Caxiuana both show increases in all variables.
+Harvard forest and Caxiuan\~a both show increases in all variables.
 In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at the low productivity sites.
@@
 
-For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result.
+For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and low carbon use efficiency as a result.
 At Niwot, an optimum leaf CN ratio was apparent.
@@
 The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (including on leaf temperature and water stress).
-At the Caxiuana site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem.
+At the Caxiuan\~a site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem.
 In figure
 \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses).
-It is worth noting that these same simulations typically had extremely low photosynthesis, particularly at Caxiuana and Metolius.
+It is worth noting that these same simulations typically had extremely low photosynthesis, particularly at Caxiuan\~a and Metolius.
 
@@
 The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization.
-At Caxiuana, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.
+At Caxiuan\~a, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.
 
@@
 \emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details).
-Thus, low values of \emph{cn\_flex\_a}  results in reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.
+Thus, low values of \emph{cn\_flex\_a}  result in reductions in expenditure on N uptake when N is too expensive, and therefore in lower tissue N content, and vice-versa.
 For the range investigated here, this parameter had limited impact on the system state.
@@
 Impact of \emph{cn\_flex\_c}\\
- \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue CN values near to their target when environmental costs are varying.
+ \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying.
 For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.
@@
 For LAI (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) growth at the Metolius site.
-Leaf N concentrations also dropped the furthest at this site (0.92; CAX, 0.93; NWR, 0.93 HVF, 0.90 MET) suggesting that expenditure on N uptake might be limited by environmental costs at this site (thus leaf CN deviates further from the target).
+Leaf N concentrations also dropped the furthest at this site (0.92; CAX, 0.93; NWR, 0.93 HVF, 0.90 MET) suggesting that expenditure on N uptake might be limited by environmental costs at this site (thus leaf C:N deviates further from the target).
 
-For Nitrogen responses, impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET) (albeit lower for the drought-stressed Caxiuana and Niwot sites).
-For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET) increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts as some sites, but impacts on LAI were all positive, reflecting lower costs of N uptake and a greater capacity to acquire N to build tissues and retain carbon for growth (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET)
+For Nitrogen responses, impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET) (albeit lower for the drought-stressed Caxiuan\~a and Niwot sites).
+For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET), increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts on NPP at some sites, even though impacts on LAI were all positive, reflecting lower costs of N uptake and a greater capacity to acquire N to build tissues and retain carbon for growth (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET)
 
 \subsection{Impact of parameters}
-Across the four sites, parametric impacts were very varied, when compared to the relatively consistant impact of parameters on the system state.
+Across the four sites, parametric impacts were very varied, when compared to the relatively consistent impact of parameters on the system state.
 A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses, and the range of their impacts varied over an order of magnitude between sites.
@@
 Impact of \emph{froot\_leaf}\\
-Despite it’s  impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses.
+Despite its strong impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses.
 At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP is less pronounced.
@@
 Impact of \emph{stem\_leaf}\\}
-Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuana and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and\ref{LAI CO2 and N respones 2001}).
+Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuan\~a and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and\ref{LAI CO2 and N respones 2001}).
 Initial LAI was larger in these cases, explaining the decease in fertilization response as above.
@@
 
- At Caxiuana, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
+ At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
 There was also a slight increase in the CO$_{2}$ fertilization response at high \emph{N\_costs}, perhaps as this high N limitation was alleviated by extra assimilated carbon.
 
- For very low \emph{N\_costs}, at Caxiuana and Niwot, N responses increased, counter-intuitively.
-In these cases, plants appear to make leaves with higher N content (Figure \ref{Leaf N CO2 and N respones 2001}), as per the algorithm which allows a higher $N_{area}$ to be achieved at low N cost.
-However, the total amount of N in the system (the balance of N inputs and outputs) remains the same with low N costs, and so these cases do not have higher leaf area.
+ For very low \emph{N\_costs}, at Caxiuan\~a and Niwot, N responses increased, counter-intuitively.
+In these cases, plants appeared to have leaves with higher N content in the initial state (Figure \ref{NR1 state}), as per the algorithm which encourages a higher $N_{area}$ to be at lower N costs.
+This appears to restrict the total LAI in the initial state, given that, the total amount of N in the system (the balance of N inputs and outputs) is not changed by \emph{N\_costs}, and so these cases do not have higher leaf area.
 When more total N is added, the leaf area can expand, and thus the N concentration declines.
@@
 Impact of \emph{frac\_fixers}\\
-While nitrogen uptake costs were important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available, in many cases.
+While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.
+
 The fraction of N fixers in the system has a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
-In line with expectations, for GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$.
+In line with expectations, for GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP CO2 and N respones 2001}, \ref{LAI CO2 and N respones 2001}).
 
-For NPP, this result is muted for the Caxiuana and Metolius sites, and indeed at Caxiuana the low N fixer cases had a very negative (-8\%) NPP response (despite a positive 7.6\% GPP response), on account of the model increasing the N content of the leaves by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}).
+For NPP, N deposition responses were muted at Metolius, and reversed in sign at Caxiuan\~a, where the low N fixer cases had a very negative (-8\%) NPP response (despite a +7.6\% GPP response), on account of the model increasing the N content of the leaves by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}).
 Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.
 
-At Niwot Ridge and Metolius, the changing fixation capacity had a very large impact on carbon fertilization, but not on nitrogen response, particularly for GPP.
-At Metolius, this impact persisted, but at Niwot the fixation capacity did have an impact on LAI.
-Interestingly, many variables increased NPP at this site, but only \emph{fracfixers} increased LAI, suggesting a limitation in the total available N suppressed the impact of all the other case where NPP was increased, and also that water limitation prevented any greater assimilation of carbon.
+At Niwot Ridge the changing fixation capacity had no impact on the response of GPP to N fixation, despite its influence on LAI, suggesting that water limitation was a greater constraint on total GPP than leaf area.
+
+At Metolius, again the changing fixation capacity had no impact on the response of GPP to N deposition, but in this case, it had no impact on LAI either.
+Only SLA had a large impact on N fertilization at this site, suggesting that only the higher LAI runs were able to respond to fertilization.
 
@@
 The impact of target \emph{leafcn} on fertilization responses was complex (recalling that this was a significant control over the system state at all sites).
-t Caxiuana, high values \emph{leafcn} (low leaf  $N_{area}$) had much stronger responses to N fertilization, indicating the dominant impact of the carboxylation capacity at this tropical site.
-At Niwot Ridge, low values of \emph{leafcn} (low leaf  $N_{area}$) had much stronger responses to CO$_{2}$ fertilization (the strongest responses at this site for GPP), potentially since the low stomatal conductance at this water-limited site could benefit most from additional CO$_{2}$ where carboxylation capacity was high.
+
+At Caxiuan\~a, high values of \emph{leafcn} (low leaf $N_{area}$) had much stronger responses to N fertilization, indicating the dominant impact of carboxylation rate limitation at this tropical site.
+
+At Niwot Ridge, low values of \emph{leafcn} (high leaf $N_{area}$) had much stronger responses to CO$_{2}$ fertilization (the strongest response at this site for GPP).
+These cases suffered from high water limitation, given their large fluxes (Figure \ref{btran state}) and thus the CO$_{2}$ fertilization would alleviate this problem.
 
 Impact of \emph{grperc}\\
-\emph{grperc}, or the degree to which growth is 'taxed' is constant and not responsive to any of the physiological processes upstream that depend on fertilization.
-Thus, its effects are all likely due to indirect impacts on the initial state.
-Two main impacts were observed.
-At Niwot Ridge low values of \emph{grperc} caused higher responses to CO$_{2}$ fertilization for GPP and NPP, but not LAI.
-These values corresponded to high initial LAI (on account of larger amounts of C being available for N uptake) and therefore allowed greater leverage of the increasing CO$_{2}$.
-At Harvard forest, high \emph{grperc} had caused low levels of LAI, resulting in increased capacity for CO$_{2}$ fertilization at that site.
+\emph{grperc}, or the degree to which growth is 'taxed' is constant and not responsive to any of the physiological processes upstream that depend on fertilization, thus, at most sites, impacts are minimal.
+At Niwot Ridge low values of \emph{grperc} caused higher responses to CO$_{2}$ fertilization for GPP and NPP, and LAI.
+These values corresponded to high initial LAI  and therefore allowed greater leverage of the increasing CO$_{2}$.
 
 Impact of \emph{medlyn\_slope}\\
-At Niwot Ridge, low values of the Medlyn Slope parameter (with high water use efficiency) were able to respond very strongly to increasing Nitrogen.
-These cases also had higher GPP in the initial state, and their higher WUE enabled them to take advantage of the increasing N availability.
-These cases also ended up with higher vegetation carbon.
+At Niwot Ridge, low values of the Medlyn Slope parameter (with high water use efficiency) were able to respond very strongly to increasing Nitrogen (Figure \ref{NR1 ndep}, given their lower initial water stress (Figure \ref{btran state}).
 
-Surpriingly, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization reponse, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.
+Surprisingly however, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.
 
@@
 
-Impact of \emph{cn\_flex\_a}, \emph{cn\_flex\_b} and \emph{cn\_flex\_c}\\
-As with it's impact on the state, there was limited impact of \emph{cn\_flex\_a} on the fertilization responses, indicating that this parametr is
+Impact of \emph{cn\_flex\_a}\\
+As with it's impact on the state, there was limited impact of \emph{cn\_flex\_a} on the fertilization responses.
 
 Impact of \emph{cn\_flex\_b}
-Low values of \emph{cn\_flex\_b}, where the leafCN concentration can increase in response to fertilization, showed higher responses to n fertilization, particularly at Harvard Forest.
-Small values of  \emph{cn\_flex\_b} should  decrease the amount of carbon spent on N uptake when NCosts are high, thus, a relaxation of the cost of N will increase the amount spent on uptake, and subsequently $N_{leaf}$ and all fluxes downstream.
-Thus, these changes result from the relaxation of an over-suppression of leaf N content.
+At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_[2}$.
+Leaf N concentration followed suit, increasing substantially in the N deposition experiment.
+
+Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when \emph{N\_costs} are high, therefore potentially suppressing leaf N content.
+Adding N overcomes this suppression, leading to greater fertilization responses.
+(n.b.
+I'm not sure about this, need to come back to it again).
 
 Impact of \emph{cn\_flex\_c}\\
-The impact of \emph{cn\_flex\_c} (the paremter controlling the degree to which plant CN ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon at Harvard Forest and Niwot Ridge (and Caxiuana, to a smaller extent).
-These impacts do not mirror those on leaf area index or NPP.
-At Harvard Forest and Caxiuana, there is a relative increase in the leaf N content, while at Metolius and Niwot, this parameter has a dominant control over the impacts of leaf N to CO$_{2}$ fertilization.
-The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high, shown clearly at Niwot Ridge and Metolius (e.g.
+The impact of \emph{cn\_flex\_c} (the parameter controlling the degree to which plant C:N ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon (Figure \ref{VEGC CO2 and N respones 2001})  at Harvard Forest and Niwot Ridge (and Caxiuan\~a, to a smaller extent).
+These impacts do not mirror those on leaf area index or NPP, and overall the response to this parameter appears complex.
+
+At Metolius and Niwot, this parameter has a dominant control over the impacts of CO$_{2}$ fertilization on $N_{leaf}$ (Figure \ref{Leaf N CO2 and N respones 2001}).
+The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high (e.g.
 Figure \ref{NR1 celev}, \ref{MET celev}).
-Of course, if N concentrations are allowed to decline too much in response to rising CO$_{2}$, then productivity declines, and other complext feedbacks occur, as shown at Harvard Forest and Niwot Ridge (Figure \ref{HRV celev}, \ref{CAX celev}).
-In terms of the overall impact, we can see that the degree of flexibility of the tissue CN ratio acts as an orthoganal control on fertilization response to the total fixation capacity.
+Low  $N_{leaf}$ values permit higher leaf area and vegetation carbon, but reduce assimilation rates, as illustrated at Harvard Forest and Niwot Ridge (Figure \ref{HRV celev}, \ref{CAX celev}).
 
@@
 The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).
-In all of these activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
+Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
 At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes.
-In particular, as expected, parameters that relate most strongly to the acquisition of N, (\emph{frac\_fixers}, \emph{Ncosts}) and to the flexibility of plant tissues in their C:N ratio (\emph{leafcn}, \emph{cn\_flex\_b}, \emph{cn\_flex\_c}), have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
-It has long been understood that uncertainties in the representation of Nitrogen are large(\cite{zaehle2014}), but previous model-model comparison studies have struggled to illustrate the exact cause of differences between models predictions, given the complexity of the structural and parametric differences involved.
+
+In particular, parameters that relate most strongly to the acquisition of N, (\emph{frac\_fixers}, \emph{Ncosts}) and to the flexibility of plant tissues in their C:N ratio (\emph{leafcn}, \emph{cn\_flex\_b}, \emph{cn\_flex\_c}), have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.
+It has long been understood that uncertainties in the representation of Nitrogen are large (\cite{zaehle2014}), but previous model-model comparison studies have struggled to illustrate the exact cause of differences between model predictions, given the complexity of the structural and parametric differences involved.
 The new CLM5 model has a number of explicit parameterizations, many of which were hard-coded or assumed to be 0 or 1 in previous versions of the model.
@@
 
-\subsection{Limitations}
+\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
+Many model development efforts are motivated under the auspices of 'reducing uncertainty'.
+Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 are partly motivated by the need to move the model behaviour closer to several key dynamics features of the system.
+Notably, the dynamics of leaf Nitrogen content with respect to environmental scarcity (\cite{zaehle2014}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen.
+The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.
+We argue that all of these developments generate model emergent behaviour that is closer in essence to reality.
+However, in doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.
+By investigating the critical roles of these processes in the model dynamics, we hope to highlight a research need, and to stimulate research on and improvements in the representation of these processes.
+
+In analyzing the results of this study, for example, we found numerous instances where the model reduced carbon use efficiency by changing tissue C:N ratios in response to forcing.
+In principle, there exists an optimal leaf C:N ratio that maximizes leaf C export under a given set of conditions.
+Despite its optimal allocation of leaf N \emph{within} the leaf (to photosynthetic enzymes), CLM5 does not include any scheme that ensures that overall tissue CN ratios conform to any physiological optima.
+Instead, a heuristic model adjusts tissue CN ratios round the target ratio (an input parameter).
+Numerous models have been proposed along these lines in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{a,nten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}).
+We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in the LUNA model (Xu.
+C.
+pers comm), but future nitrogen cycling model development efforts could re-visit this topic.
+
+Another area requiring focus is the dominance of the 'fraction of fixers' over fertilization responses.
+(\emph{frac\_fixers}) is a proxy for the community level composition of Nitrogen fixing plants.
+Technically, because plant density is not represented in the CLM5, this parameter really represents the fraction of net assimilated carbon ($gpp$ - $R_{auto}$) that can be used for symbiotic N fixation.
+Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would response to changes in environmental conditions and the relative competitive abilities of N fixing plants.
+In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (references**), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).
+
+\subsection{ Methodological Limitations}
+
 \subsubsection{Limitations of step-change experiments}
-Here investigate the parameter space of the model using step-change experiments.
-However, such experiments necessarily involve unrealistically rapid changes in environmental boundary condition, and cannot necessarily be considered as analagous to either real-world or transient future scenario conditions.
-Further, we expect that the carbon and nitrogen cycles mayb take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.
+Here we investigate the parameter space of the model using step-change experiments.
+However, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions.
+Further, we expect that the carbon and nitrogen cycles may take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.
 
 To test the impact of time on the fertilization responses, we also analyzed the impact of fertilization at 5 and 50 years into the elevated CO$_{2}$ and N experiments.
-At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, result not shown) to those at 15 years, but at 5 years, in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent x~2.06 impact on NPP, compared to ~0.95 in the 2015 analysis at NVR), indicating that the results described above had already been subject to progressive N limitation and were closer to an equilibrated state.
+At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, (result not shown, but I could include them, the figures are done**)) to those at 15 years.
+However, at 5 years into the fertilization in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent (a x~2.06 impact on NPP, compared to ~0.95 in the year 15 analysis at NVR), indicating that the results described above had already been subject to progressive N limitation, and were closer to an equilibrated state.
 
@@
 The analysis presented here deliberately does not focus on the relative fit of the model to the data collected at the individual sites.
-The fit at individual sites of the globally adjusted model cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, which is beyond the scope of this paper.
+The fit at individual sites of the globally adjusted model cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, and a focus on that would preclude the detailed analysis of the parameter space presented here.
 
-Instead, we chose a set of individual sites, rather than global analysis to extensively probe a set of alternative vegetation and climate conditions under realistic forcing scenarios.
 We also do not make any assessment of how well the model output fits observations across this parameter space, since the relevant datasets with which to test alternative model configurations are, we argue, the set of globally-relevant data products described by \cite{lawrence2018}.
@@
 \subsubsection{Parameter calibration of the default model, or absence thereof}
-
 During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
-We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces,  3) the tendency of the modeled ecosystem to die under past low-CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting pattern that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the different data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
-The parameterization used in the default version of the model utilized the mean trait values where possible, from trait datasets (\cite{lawrence2018}) in combination with targeted tuning of parameters   (including allocation parameters and Nitrogen costs)that were less well constrained by data, to avoid excessively spurious values for some model states.
+We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces,  3) the tendency of the modeled ecosystem to die under past low-CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting pattern that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
+The parameterization used in the default version of the model utilized the mean trait values where possible, from trait datasets (\cite{lawrence2018}) in combination with targeted tuning of parameters   (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states.
 
@@
 The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
-In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
+In principle, integrating more targeted datasets and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
 Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
@@
 In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water.
-The suite of parameters than in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implictly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.
+The suite of parameters that in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implicitly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.
 Properties more closely linked to water availability are associated with the hydrological components of the model as well as the PHS plant hydraulics scheme and will be investigated further by \cite{dagon} and \cite{kennedy}.
-
-\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
-Many model development efforts are motivated under the auspices of 'reducing uncertainty'.
-Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 are partly motivated by the need to move the model behaviour closer to several key observable features of the system.
-Notably, the dynamics of leaf Nitrogen content with respect to environmental scarcity (\cite{zaehle2014}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen.
-The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.
-We argue that all of these development generate model emergent behaviour that is closer in essence to reality.
-However, in doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.
-
-In this study, we found numerous instances where the model reduced carbon use efficiency by increasing tissue C:N ratios in response to forcing.
-In principle, there exists an optimal leaf CN ratio that maximizes leaf C export under a given set of conditions.
-Despite its optimal allocation of leaf N within the leaf (to photosynthetic enzymes), CLM5 does not include any scheme that ensures that overall tissue CN ratios conform to any physiological optima.
-Instead, a heuristic model adjusts tissue CN ratios round the target ratio (an input parameter).
-Numerous models have been proposed along these lines in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{a,nten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}).
-We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in the LUNA model (Xu.
-C.
-pers comm), but future nitrogen cycling developments might re-visit this topic.
-
-\subsubsection{Discourse on whether nitrogen fixation is realistic in CLM5}
-The 'fraction of fixers' (\emph{frac\_fixers}) is a proxy for the community level composition of Nitrogen fixing plants.
-Technically, because plant density is not represented in the CLM5, this parameter really represents the fraction of net assimilated carbon ($gpp$ - $R_{auto}$) that can be used for symbiotic N fixation.
-Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would response to changes in environmental conditions and the relative competitive abilities of N fixing plants.
-In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (references**), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).
-
-\subsection{Differences between sites}
-The overall response of the different sites to CO$_{2}$ and N fertilization, in terms of the total vegetation carbon, is to some extent a function of their initial limiting conditions.
-The Niwot Ridge site is limited by both N availablilty and water in the initial state, and thus demonstrates strong CO$_{2}$ and N responses.
 
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
-     \caption{Influence of parameteric variation on the impact of 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
+     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
 }

### 2018-11-29 main.tex
Updates from ShareLaTeX

@@
 %tentative author list...
-\authors{Rosie Fisher\affil{1},
-William Wieder\affil{1},
-Benjamin Sanderson\affil{2},
-Charles Koven\affil{3},
-Keith Oleson\affil{1},
-Chonggang Xu\affil{4},
-Joshua Fisher\affil{5},
-Mingjie Shi\affil{5},
-Anthony Walker\affil{6},
-Gordon Bonan\affil{1},
+\authors{Rosie A. Fisher\affil{1},
+William Wieder\affil{1,2},
+Benjamin M. Sanderson\affil{3},
+Charles D. Koven\affil{4},
+Keith W. Oleson\affil{1},
+Chonggang Xu\affil{5},
+Joshua Fisher\aff B.il{6},
+Mingjie Shi\affil{6},
+Anthony P. Walker\affil{7},
+Gordon B. Bonan\affil{1},
 David M. Lawrence\affil{1}
@@
 
-\affiliation{2}{CERFACS, Toulouse, France}
+\affiliation{2}{Institute of Arctic and Alpine Research, University of Colorado, Boulder Colorado, USA}
 
-\affiliation{3}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
+\affiliation{3}{Centre Europe'en Research et de Formation Avenc'ee en Calcul Scientifique, Toulouse, France}
 
-\affiliation{4}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
+\affiliation{4}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
 
-\affiliation{5}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
+\affiliation{5}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
 
-\affiliation{6}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
+\affiliation{6}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
+
+\affiliation{7}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
 
@@
 \begin{keypoints}
+\item The Community Land Model, version 5, contains numerous upgrades to carbon and nitrogen cycling.
 
-\item A suite of modifications to the Nitrogen cycle representation were added to the Community Land Model, version 5 (CLM5).
+\item Model state and responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.
 
-\item New features of the N model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of Nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios and dynamic leaf resorption of Nitrogen.
+\item Parameters controlling nitrogen fixation, uptake, and stoichiometric flexibilty are most important
+\end{keypoints}
 
-\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycling of Carbon and Nitrogen.
-We assess the dependence of the both the model state, and also the responses of the system to CO$_{2}$ and Nitrogen fertilization to parametric uncertainty.
+\begin{abstract}
+A suite of modifications to the nitrogen cycle representation were added to the Community Land Model, version 5.
+New features of the model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios and dynamic leaf resorption of nitrogen.
+
+Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycling of carbon and nitrogen.
+We assess the dependence of both the model state, and also the responses of the system to CO$_{2}$ and nitrogen fertilization to parametric uncertainty.
 Model runs are conducted at representative individual sites in temperate, tropical and boreal systems.
 
-\item The model responses to CO$_{2}$  and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, the costs of N in the environment and stoichiometric flexibility of plant tissues, all of which are subject to large empirical and theoretical uncertainty.
-
-\item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.
-\end{keypoints}
+The model responses to CO$_{2}$  and N fertilization are primarily linked to the representation of the fraction of plants that can fix nitrogen, the costs of N in the environment and stoichiometric flexibility of plant tissues, all of which are subject to large empirical and theoretical uncertainty.
+Identification of the primary factors driving ecosystem responses to fertilization assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.
+\end{abstract}
 
 \section{Introduction}
-
 Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}).
-In the CMIP5 multi-model inter-comparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle.
-The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{friedlingstein2014}, \cite{arora2013}).
+In the CMIP5 multi-model inter-comparison, only one land model, the Community Land Model v4 (CLM4), included an active nitrogen (N) cycle.
+The CLM4 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide ( \cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}).
 Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent global N dynamics using alternative structural assumptions have been developed (\cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}).
-Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehle2011}).
+Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehledalmonech2011}).
 
@@
 The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}).
-Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}.
-An overview of model developments is presented by \cite{lawrence2018}, a description of the global Nitrogen cycling by Wieder et al.
-(in prep), the crop model by(Lombardozzi et al.
+Version 5 of the CLM was released to the research community as part of the CESM2.0 release and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}.
+An overview of model developments is presented by \cite{lawrence2018}, a description of the global nitrogen cycling by Wieder et al.
+(2018), the crop model by(Lombardozzi et al.
 in prep), the land use components by (Lawrence P et al.
@@
 
-CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4.0 and CLM4.5.
-The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing.
+CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4 and CLM4.5 (\cite{lawrence2011}).
+The introduction of several new dynamical features into the CLM5 has implications for both parametric uncertainty and responses to environmental forcing.
 Wieder et al.
-(in prep) report the global-scale carbon cycle implications of the full model update and note that the CLM5 carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.
+(2018) report the global-scale carbon cycle implications of the full model update and note that, in CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.
 In this paper, we describe the new components of the biogeochemical cycling, and investigate the features of the new model that control the model state and biogeochemical forcing responses in depth, using single-site experiments and parametric sensitivity analyses.
@@
 \subsection{Model Updates}
+The CLM5 nitrogen cycle builds on the implementation of carbon-nitrogen interactions in the CLM4 model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).
+The new model integrates three major additional prognostic elements of the nitrogen cycle, including:\\
 
-The CLM5 nitrogen cycling builds on the implementation of carbon-nitrogen interactions in the CLM4.0(CN) model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).
-The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\
-
-1) The `LUNA' model (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}) \\
+1) The `LUNA' module (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}) \\
 
@@
 
-3) The `FLEXCN' N cycle implementation, adapted from \cite{ghimire2016}, which is used in CLM5 primarily to allow variation in tissue C:N ratios,  .\\
+3) The `FLEXCN' N cycle implementation, adapted from \cite{ghimire2016}, which is used in CLM5 primarily to allow variation in tissue C:N ratios.\\
 
@@
 
-The LUNA model predicts the optimal balance of of $V_{c,max}$ and $J_{max}$ for the prevailing time-averaged environmental conditions.
+Given a specific amount of leaf N and empirical environmental constraints, the LUNA model predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components, for the prevailing time-averaged environmental conditions.
 The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
-In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area $N_{area}$, (which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and prevailing environmental conditions (CO$_{2}$, temperature, humidity, soil moisture).
-The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
+In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area $N_{area}$, (which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and prevailing environmental conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length).
+The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the improved codes for global calibration and the geographical predictions of the model were described by \cite{ali2016}.
 
 \subsubsection{FUN}
-
-Under circumstances where insufficient N exists to match all of the carbon assimilated for a given C:N ratio, the CLM4.0 model reduced the gross photosynthetic flux down to the level at which growth could be supported by the assimilated N.
+Under circumstances where insufficient N uptake was available to match all of the carbon assimilated (for a given target C:N ratio), the CLM4 model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake.
 This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).
@@
 One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}).
-The FUN model operates on the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon.
+The FUN model (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}) addresses this issue by imposing the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}).
 Further, the cost of each uptake pathway is variable in both time and space.
-FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them.
-Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N.
-Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency).
+FUN hypothesizes that plants will take up nitrogen from pathways which are 'cheaper' to them.
+Where N is scarce in the environment, carbon that would otherwise have been used for plant growth (net primary productivity, NPP) can be deployed to acquire (pay for) more N.
+Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency- defined here as NPP divided by gross primary productivity, GPP).
 The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for nitrogen uptake, ($C_{nuptake}$).
@@
 \begin{equation}
-C_{growth}=C_{avail}+C_{nuptake}
+C_{growth}=C_{avail}-C_{nuptake}
 \end{equation}
 
-Further, the nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:
+Further, the nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant C:N ratio ($CN_{plant}$):
 
@@
 \begin{equation}
-C_{nuptake} =\frac{C_{avail}}{ ( (1.0+f_{gr})*(CN_{target} / N_{cost}) + 1) }
+C_{nuptake} =\frac{C_{avail}}{ ( (1.0+f_{gr})*(CN_{plant} / N_{cost}) + 1) }
 \end{equation}
 
-The average Nitrogen cost ($N_{cost}$) is derived from the uptake of N, weighted for it's cost across all uptake streams, while the plant CN ratio ($CN_{plant}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools.
+The average nitrogen cost ($N_{cost}$) is derived from the uptake of N, weighted for its cost across all uptake streams, while the plant CN ratio ($CN_{plant}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools.
 The FUN model is documented at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}
 
-One issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}).
-The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity ((\cite{wieder2015} illustrated the impacts of uncertainty in this function).
+One mechanism introduced by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}).
+The CLM4 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{cleveland1999}), and \cite{wieder2015} illustrated the impacts of uncertainty in this function.
 FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
 Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation.
-Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground.
+
+Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state (carbon and nitrogen pools in equilibrium) from bare ground.
 In most cases, N fixation is preferred when soil N is low, and so soil N builds up quickly.
@@
 
+The FUN model introduces the concept of carbon payments for N uptake, which are analagous to the carbon expended on root exudates, symbiotic relationships with soil fungi, enhanced expenditure on dynaic fine root growth or increased metabolic rates to support active uptake.
+A more comprehensive description of these processes might include explicit representations of these processes, but the model deployed here uses a bulk approach to account for the costs to plants of low environmental N availability, therefore, uncertainties associated with the cost functions, in particular, are large.
+
 \subsubsection{FLEXCN}
 The results of FACE model-data comparison exercises illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).
-These changes were not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle.
+These changes were not captured by the static tissue C:N ratios used in the CLM4 and CLM4.5 nitrogen cycle.
 For this reason, we introduce a flexible tissue C:N ratio into the CLM5.
 
-\cite{ghimire2016} implemented a suite of modifications to the CLM Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
-In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, thus the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen.
-In the new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates.
-The allocation scheme retains a fixed-fraction approach to allocation in general, and thus is capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an issue which must be taken into account during simulations under high productivity conditions.
+\cite{ghimire2016} implemented a suite of modifications to the CLM4.5 Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
+In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents nitrogen uptake, thus the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying N supply.
+In the new allocation algorithm, the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).
+The carbon allocation scheme uses a fixed-fraction approach to allocation in general, with constant ratios between, leaf, fine roots, stem and coarse root allocation.
+We note here that the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an issue which must be taken into account during simulations under high productivity conditions.
 
 \subsubsection{Adaptation of FUN to flexible CN ratios}
-
-The FUN model was originally conceived with a static tissue C:N ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target C:N ratio.
-In CLM5, one of the aims of the development process was to incorporate a varying C:N ratio.
-To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required).
-The degree to which tissue C:N content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments under N limited conditions (including \cite{zaehle2010},\cite{ghimire2016}).
-Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content.
+The FUN model was originally conceived with a static tissue C:N ratio, and will by default allocate an amount of carbon to N uptake that exactly tracks the target C:N ratio.
+In CLM5, one of the aims of the development process was to incorporate a varying C:N ratio, as above.
+To that end, it is necessary to modify the amount of C expended on N uptake (and thus the amount of N required).
+The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments under N limited conditions (including \cite{zaehle2010},\cite{ghimire2016}).
+Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content.
 When N is expensive in the environment, less C is spent on uptake, increasing C:N ratios where N is scarce.
@@
 \subsection{Content of this paper}
+In this paper, we assess in detail the sensitivity of the CLM5 model parameterization of ecosystem biogeochemistry.
+We use a range of single site experiments, combined with one-at-a-time parameter uncertainty analyses from the default state, as well as CO$_{2}$ and N fertilization experiments, to fully investigate the performance of the model under steady-state and step-change modifications in forcing.
 
-In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters.
-We use a range of single site experiments, combined with one-at-a-time parameter uncertainty analyses from the default state, as well as CO$_{2}$ fertilization experiments and Nitrogen addition experiments,
-to fully investigate the performance of the model under steady-state and step-change modifications in forcing.
+Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations, including the controls on global carbon fertilization `$\beta$' values  (the change in ecosystem carbon storage in response to CO$^{2}$ fertilization).
+We also hope to highlight areas of process and parameter uncertainty that appear to require greater attention in future model development and testing cycles.
+{CXU: Will it possible to propose some key hypotheses here for easier guidance of the paper?}
 
-Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect overall behaviour in larger-scale experiments, including the controls on global carbon fertilization `$\beta$' values.
-We also hope to highlight areas of process and parameter uncertainty that appear to require greater attention in future model development and testing cycles.
-
-Because of the increased number of prognostic processes with the CLM5, many alternative representations of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible.
-We thus consider that an exploration of the CLM5 parameter space is also equivalent to an exploration of a large part of the relevant model structural space.
+Because of the increased number of prognostic processes {CXU: probabily put the total number of parameter for CLM5 here} with the CLM5, many alternative representations of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation), variation in the degree to which to C:N ratios are flexible, and the ease with which plants may used carbon resources to more efficiently exploit available soil N.
+Thus, the parameter space explored here represents a reasonable fraction of the relevant model structural space.
 
 \section{Methods}
-
 \subsection{Sites}
-
-For this experiment we utilized six different experimental sites.
-One in tropical forests (Caxiuan\~a), one in alpine forest (Niwot Ridge), one in temperate forest (Harvard Forest) and one in a seasonally dry evergreen forest (Metolius Forest).
+For this experiment we utilized four different observational flux tower sites to capture a range of initial conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, -1.9$^{o}$N, -51.4$^o}$E),  sub-alpine evergreen forest (Niwot Ridge, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, 44.5$^{o}$N -121.6$^{o}$E).
+These sites are all flux tower sites, and all have previously been used for model development.
+Each site was driven with the available meteorological data (\textit{give date ranges of met data}).
+Soil texture, color, depth and, plant functional type composition were taken from previous model validation exercises undertaken at these sites (\cite{bonan2012} for Metolius and Harvard Forest, \cite{fisher2007} for Caxiuan\~a and ?
+for Niwot Ridge.
 
 \subsection{Simulation Protocol}
-We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the turnover of slower cycling carbon and nitrogen pools is accelerated, see (\cite{lawrence2018}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level half-hourly meteorological inputs.
-400 years was sufficient to bring all carbon and nitrogen pools into equilibrium.
+We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the turnover of slower cycling carbon and nitrogen pools is increased for the duration of the spin-up, and the pool sizes are modified accordingly at the end of the spin-up - see (\cite{lawrence2018}).
+We used pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level half-hourly meteorological inputs.
+400 years was sufficient to bring all soil and vegetation carbon and nitrogen pools into equilibrium.
 Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years for each site, parameter and value combination, with accelerated mode turned off.
@@
 
-For each site/parameter/value combination, we also ran one CO$_{2}$ and one Nitrogen deposition experiment.
+For each site/parameter/value combination, we also ran one CO$_{2}$ and one nitrogen deposition simulation.
 To allow comparison with \cite{wieder}, we increased CO$_{2}$ to 550ppm, starting in 1998.
-For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1998.
-In line with \cite{wieder}, we assessed the impact of changes after 15 years of fertilization in both cases.
+For Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.
+Similar to \cite{wieder}, we assessed the impact of changes after 15 years of fertilization in both cases.
 
 \subsection{Perturbed Physics Ensemble (PPE)}
-The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5.
-For this manuscript, we focus on parameters included in recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization.
-Therefore,
-we focus on parameters of particular https://www.overleaf.com/project/5b05c31494d9960223796bc3relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
+The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the model.
+For this manuscript, we focus on parameters included in developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization.
+Therefore, we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle.
 The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.
-
-Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model.
+Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth system model.
 However, we also report responses of the model state to perturbations.
 Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al.
-in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here.
+in prep) and into the plant hydraulics parameterizations (Kennedy et al in review), thus, we have not included the full range of those parameters here.
 We expect that further focused analyses will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.
@@
 
-\subsubsection{Parameters Selected for PPE}
-
+\subsubsection{Parameters selected for perturbation experiment}
 The parameters included in our sensitivity analysis are both detailed in Table \ref{table_parameters}, and described in more depth below. \\
@@
 5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}.
-This parameter is the fraction of the assimilation (gross primary productivity - autotrophic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.
+This parameter is the fraction of the assimilation (gross primary productivity - autotrophic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.
 \\
@@
 
-\subsubsection{Parameters not included in PPE}
-
-We conducted a set of tests on the parameters of the denitrification model, to assess the impacts of differing rates of N loss on the system.
-However, recent analysis has shown that the rates of overall denitrification in CLM5 are very low, overall (compared to plant uptake and immobilization \textbf{I've forgotten why this is?
-Can someone remind me}) and hence these parameters had very little impact.
-We did not conduct tests on leaf lifespan, since for evergreen trees in CLM5, leaf lifespan and specific leaf area have very similar impacts on overall achievable leaf area, and for deciduous trees, leaf lifespan is primarily determined by phenological triggers.
-We did not investigate any parameters relating to the crop model (Lombardozzi et al.
-in prep) nor to wood harvest, fire, dust, or soil carbon decomposition, (the latter is largely unaltered from CLM4.5 as per \cite{koven2013}.
-
 \subsubsection{Parameter Ranges}
-
 \emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY plant trait database(\url{www.try-db.org} and we take their values from the analysis by (\cite{kattge2011}) of their mean and distribution.
@@
 The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across the whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake).
-\emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM value (0.3).
+\emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM(4,4.5) value (0.3).
 \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}.
@@
 The same issues apply to $N_{costs}$ parameters.
-
+{CXU: it would be great if you could point out how you change the parameters..e.g., change the parameter based on random sample from the distributions or one at a time ....}
 \subsection{Analysis}\\
-We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the fertilized simulations 15 years after fertilization began.
+We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the fertilized simulations 15-20 years after fertilization began.
 For some parameter settings, the constraint on growth placed the system in a state close to death.
-Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in the relationship between parameters and  with  to environmental forcing.
-Therefore, for those parameter combinations which cause LAI to be less than 60\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plots, so as to focus on the responses of parameter combinations that give more reasonable initial vegetation conditions.
+Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in the relationship between parameters and environmental forcing.
+To focus on the responses of parameter combinations that give more reasonable initial vegetation conditions, for parameter combinations that caused LAI to be less than 60\% of the control simulation value, CO$_{2}$ and Nitrogen simulations were excluded from the plots.
 
@@
 Including the default simulations, this is 672 different run configurations.
-In this analysis, we look at the parametric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon).
+In this analysis, we examine the parametric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon).
 In the first section, we look at the impacts of parametric variation on the system state and secondly on the responses to carbon and nitrogen fertilization.
-We do not focus on the goodness of fit to observations at these sites, because our aim is to discern the primary parametric controls over CO$_{2}$ and N fertilization responses, and including extensive data-fitting and calibration experiments would detract from this goal.
+We do not focus on the goodness of fit to observations at these sites, because our aim is to discern the primary parametric controls over CO$_{2}$ and N fertilization responses, and including extensive data-fitting and calibration experiments (including on parameters mostly unrelated to C and N cycling) which would detract from this goal.
 
 \subsection{Overall system state}
-Figure \ref{NR1 state} shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Niwot Ridge site (which we choose to highlight as it shows some of the more complex parameter responses).
-The other site state variation figures are presented in the Supplementary Information (SI \ref{CAX state}, \ref{HVF state}, \ref{MET state}).
-
-\subsubsection{Low leaf area investigations}
-In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
-(for Caxiuan\~a, 3.0 v.s.
-5.3 \cite{fisher2007}, for Niwot Ridge, 1.7 vs.
-4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs.
-3.5 \cite{williams1996} and for Metolius, 0.5 (modeled) v.s.
-1.5 (\cite{spadavecchia2011}).
-This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.
-
-To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
-These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAI levels.
-For example, at Harvard Forest, the high leaf allocation  simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.43, 1.66, 1.47 for GPP, NPP and LAI in the control vs.
-1.18, 1.17 and 1.22 for the high LAI case) and a lower N response as well (1.1, 1.26, 1.21 for the control GPP, NPP, LAI, and 1.05, 1.20 and 1.15 for the high LAI case).
-Similar results were found at all other sites.
-Thus, by virtue of their lower LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology (analysis tbd).
-
+The one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content per unit area ($N_{area}$, gN m$^{-2}$) and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites , and Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control parameter ensemble.
+In each of these site figures, the y-axis  corresponds to the output variable and The x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in table \ref{table_ranges}.
+Line and symbols depict responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case  {CXU: this seems could be part of a caption of the figure}.
 \subsubsection{Water and N limitation across sites}
-Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control (unfertilized) parameter ensemble.
+Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control (unfertilized) parameter ensemble {CXU: I do not like start the paragraph with description of the figure.
+It would be better to start with a main message of the paragraph}.
 Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of prognostic leaf water potential, which impacts gas exchange processes in a similar manner to the $B_{tran}$ parameter in CLM4.5 (in that it is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$).
-Because leaf water potential changes diurnally, we now use the minimum term to calculate relative water stress between sites.
+Because leaf water potential changes diurnally, we now use the daily minimum term to calculate relative water stress between sites.
 For N limitation, the fraction of total NPP which is used to pay for nitrogen uptake, is a direct indicator of how N stress affects the carbon budget.
@@
 \subsection{Parametric control over system state}
+{CXU: I do not like the report of list of sensitivity. Instead it would be helpful to make a nice story based on hypothesis in the introductions to stich all these element together}
 
@@
 
-For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area of leaves which reduces photosynthetic capacity.
+For GPP, the impact of SLA is always positive, given increasing LAI and despite higher LAI causing lower N per unit area of leaves which reduces photosynthetic capacity.
 The impact of SLA on NPP can however be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance.
-At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.
+At Niwot Ridge, an optimal SLA (or local maximum at the second-largest parameter values) is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.
 
 Impact of \emph{froot\_leaf}\\
-Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation???
-through root-mediated n uptake costs.
+Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation through root-mediated n uptake costs.
 The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.
@@
 Low values of stem:leaf ratio typically resulted in much higher LAI at all sites, as well as declines in vegetation carbon, as expected from the lower stem carbon allocation.
-At Niwot Ridge, there is an optimum stem carbon observed, but this is due to the impacts of a too-high leaf area index, which actually appears to cause declines in NPP when LAI is above the default.
+At Niwot Ridge, there is an optimum (or local maximum) stem carbon observed, but this is due to the impacts of a too-high leaf area index, which actually appears to cause declines in NPP when LAI is above the default.
 At all other sites, the impact of increasing leaf allocation is to increase productivity.
@@
 Increasing \emph{Fracfixers} affects the maximum amount of carbon that plants can pay to symbiotically uptake Nitrogen.
-In all cases here, it has a relatively moderate and positive impact on the overall system state, in particular at the Niwot and Harvard forest sites.
+In all cases here, it has a relatively moderate and positive impact on all variables, in particular at the Niwot and Harvard forest sites.
 As is the case for $N_{costs}$, with high fixing capacity N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that change.
@@
 In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).
-We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
+We do not find much evidence of saturation in the GPP/leafCN relationships, in that GPP response does not level-off at high leaf N content.
+This corresponds to a dominance of carboxylation-limited photosynthesis in these simulations (as opposed to co-limitation by light), in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.
 
 For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and low carbon use efficiency as a result.
-At Niwot, an optimum leaf CN ratio was apparent.
+At Niwot, an optimum leaf CN ratio was apparent (or a local maximum in the parameteric response).
 For NPP this was around the default value, but for LAI it was 0.5s.d.
@@
 Impact of \emph{grperc}.\\
-The response to modifications in the growth respiration coefficient in line with first-order expectations, where the higher numbers (which correspond to the previously used value of 0.33 in CLM4.5) resulted in  NPP reductions, corresponding changes in LAI, impacting GPP at the largest values.
+The response to modifications in the growth respiration coefficient were in line with first-order expectations, where higher values (which correspond to the previously used value of 0.33 in CLM4.5) resulted in NPP reductions, corresponding changes in LAI, impacting GPP at the largest values.
 The impact varied between sites, reflecting different partitioning of the carbon budget into maintenance respiration, nitrogen uptake and growth.
@@
 Impact of \emph{medlyn\_slope}\\
-The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (including on leaf temperature and water stress).
+The change in the stomatal slope parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (potentially including on leaf temperature and water stress).
 At the Caxiuan\~a site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem.
@@
 
-At Niwot, a clear optimum was found in the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough.
+At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough.
 At Metolius and Harvard forest, only the lower values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.
@@
 \subsection{CO$_{2}$ and Ndep responses}
-
 To illustrate the relationships between N and CO$_{2}$ fertilization, and to reduce the dimensionality of the output, figures \ref{GPP CO2 and N respones 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the response for GPP, NPP and, LAI respectively.
@@
 
-For the default parameterization, the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.
-For LAI (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) growth at the Metolius site.
+For the default parameterization (the intersection of all lines in figures \ref{GPP CO2 and N respones 2001} to \ref{LAI CO2 and N respones 2001}), the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.
+LAI responses (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) reflected the NPP change, except for at Metolius.
 Leaf N concentrations also dropped the furthest at this site (0.92; CAX, 0.93; NWR, 0.93 HVF, 0.90 MET) suggesting that expenditure on N uptake might be limited by environmental costs at this site (thus leaf C:N deviates further from the target).
@@
 For Nitrogen responses, impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET) (albeit lower for the drought-stressed Caxiuan\~a and Niwot sites).
-For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET), increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts on NPP at some sites, even though impacts on LAI were all positive, reflecting lower costs of N uptake and a greater capacity to acquire N to build tissues and retain carbon for growth (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET)
+For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET), increasing $N_{leaf}$ (and by definition, respiration rates) (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts on NPP at some sites, even though impacts on LAI (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET) were all positive, reflecting lower costs of N uptake and greater capacity to acquire N to build tissues and retain carbon for growth
 
@@
 A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses, and the range of their impacts varied over an order of magnitude between sites.
-Thus, the parameter space of the model is complex with regard to these changing initial conditions and fertilization experiments.
+Thus, the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change.
 Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.
@@
 Impact of \emph{froot\_leaf}\\
-Despite its strong impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses.
+Despite its strong impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO$_{2}$ responsiveness and at Harvard, where low values decrease both responses.
 At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP is less pronounced.
@@
 
-For NPP, N deposition responses were muted at Metolius, and reversed in sign at Caxiuan\~a, where the low N fixer cases had a very negative (-8\%) NPP response (despite a +7.6\% GPP response), on account of the model increasing the N content of the leaves by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}).
+For NPP, N deposition responses were muted at Metolius, and reversed in sign at Caxiuan\~a, where the low N fixer cases had a very negative (-8\%) NPP response (despite a +7.6\% GPP response), on account of the model increasing the N content of the leaves (and thus respiration rates) by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}).
 Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.
@@
 Impact of \emph{cn\_flex\_b}
-At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_[2}$.
+At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$.
 Leaf N concentration followed suit, increasing substantially in the N deposition experiment.
@@
 
+\subsubsection{Low leaf area investigations}
+In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
+(for Caxiuan\~a, 3.0 (modeled) vs.
+5.3 (observed) \cite{fisher2007}, for Niwot Ridge, 1.7 vs.
+4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs.
+3.5 \cite{williams1996} and for Metolius, 0.5 vs.
+1.5 (\cite{spadavecchia2011}).
+This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.
+
+To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
+These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition.
+In the default parameterization CO$_{2}$ and N fertilization increased LAI, amplifying these magnitude of ecosystem responses to these perturbations.
+For example, at Harvard Forest, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.18, 1.17 and 1.22 for GPP, NPP and LAI, respectively, in the high LAI case vs.
+1.43, 1.66, 1.47 in the control) and a lower N response as well (1.05, 1.20 and 1.15 for GPP, NPP, and LAI, respectively, in the high LAI case vs.
+1.1, 1.26, 1.21 for the control).
+Similar results were found at all other sites.
+Thus, by virtue of their lower LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology (analysis tbd)
+
 \section{Discussion}
-The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).
+The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}, \cite{eyring2016}).
 Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
@@
 Many model development efforts are motivated under the auspices of 'reducing uncertainty'.
-Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 are partly motivated by the need to move the model behaviour closer to several key dynamics features of the system.
-Notably, the dynamics of leaf Nitrogen content with respect to environmental scarcity (\cite{zaehle2014}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen.
+Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 are partly motivated by the need to move the model behaviour closer to several key dynamic features of the system.
+Notably, the dynamics of leaf Nitrogen content with respect to environmental scarcity (\cite{zaehle2014}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen.
 The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.
@@
 Numerous models have been proposed along these lines in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{a,nten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}).
-We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in the LUNA model (Xu.
-C.
-pers comm), but future nitrogen cycling model development efforts could re-visit this topic.
+We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in the LUNA model, but future nitrogen cycling model development efforts could re-visit this topic.
 
@@
 \subsection{ Methodological Limitations}
-
 \subsubsection{Limitations of step-change experiments}
@@
 
-We recognize that this less objective calibration method is than is ideal, and ongoing studies (\cite{dagon}) will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated in a recent study by \cite{fer2018}, .
+We recognize that this less objective calibration method is less than ideal, and ongoing studies (\cite{dagon}) will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}, .
 
-The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.
+The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2018} for assessing model skill across a very broad range of model components.
 In principle, integrating more targeted datasets and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
@@
  n\_costs    & 6 parameters related to Nitrogen costs & g/g \\
- fracfixers  & Fraction of C availble for fixation & - \\
+ fracfixers  & Fraction of C available for fixation & - \\
   grperc  & Fraction of C used for growth respiration  & - \\
@@
       lmr\_intercept & Intercept of leaf N /respiration relationship & \\
-      frac\_ecto & Fraction of active uptake that is ectohycorrhizal & - \\
-      cn\_flex\_a & Parameter of N flexibilty algorithm& \\
-      cn\_flex\_b & Parameter of N flexibilty algorithm & \\
-      cn\_flex\_c & Parameter of N flexibilty algorithm & \\
+      frac\_ectomy\_fungi} & Fraction of active uptake that is ectohycorrhizal & - \\
+      cn\_flex\_a & Parameter of N flexibility algorithm& \\
+      cn\_flex\_b & Parameter of N flexibility algorithm & \\
+      cn\_flex\_c & Parameter of N flexibility algorithm & \\
 \hline
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/btran_npp.png}
+     \caption{Response of drought index ($B_{tran,mn}$ and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{btran state}
+ \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/btran_npp.png}
-     \caption{Response of drought index ($B_{tran,mn}$ and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{btran state}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Caxiuan\~a National Forest}
+     \label{CAX state}
+  \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Harvard Forest}
+     \label{HVF state}
+ \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest}
+     \label{MET state}
  \end{figure}

### 2018-11-29 figures.tex

--- 2018-11-16 figures.tex
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth,left]{matlab/figures/btran_npp.png}
+     \caption{Response of drought index ($B_{tran,mn}$) and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
+Legend colours as for Figure \ref{NR1 state}})
+     \label{btran state}
+ \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Niwot Ridge.}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Niwot Ridge.}
      \label{NR1 state}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Caxiuan\~a National Forest}
+     \label{CAX state}
+  \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Harvard Forest}
+     \label{HVF state}
+ \end{figure}
+
+ \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Metolius Forest}
+     \label{MET state}
+ \end{figure}
+
+  \begin{figure}[h]
+    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
      \label{NR1 celev}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
      \label{NR1 ndep}

### 2019-05-28 SIfigs.tex

@@
 \begin{document}
-
- \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Caxiuan\~a National Forest}
-     \label{CAX state}
-  \end{figure}
-
- \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Harvard Forest}
-     \label{HVF state}
- \end{figure}
-
- \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation at Metolius Forest}
-     \label{MET state}
- \end{figure}
 

### 2019-05-28 workflow2.tex
Updates from ShareLaTeX
Updates from Overleaf

@@
 \section{Spin-up}
+
+\subsection{Make base Case}
+/gpfs/fs1/work/rfisher/git/clm5.0_may19/cime/scripts/CLM5_POINTSIMS/CLM5.0_may19_basebuild
+Using the /gpfs/fs1/work/rfisher/git/clm5.0_may19/cime/scripts/CLM5_POINTSIMS/CLM5.0_may19_basebuild
+
 \subsection{Make 1850 spin-up default runs from scratch}
@@
 
+\subsection{Make spinup ensemble}
+/glade/work/rfisher/git/CLM5_FACE_EXPTS/case_scripts
+{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_mic_ensembles
+}}}\\
+
 \end{document}

### 2019-05-28 referee_responses.tex

\documentclass{article}
\usepackage{hyperref}
\usepackage[dvipsnames]{color}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}

\begin{document}
\title{Referee Responses}

Reviewer 1 Evaluations: 
Recommendation: Return to author for major revisions 
Significant (Required): There are major errors or gaps in the paper but it could still become significant with major changes, revisions, and/or additional data. 
Supported (Required): Mostly yes, but some further information and/or data are needed. 
Referencing (Required): Mostly yes, but some additions are necessary. 
Quality (Required): The organization of the manuscript and presentation of the data and results need some improvement. 
Data (Required): Yes 
Accurate Key Points (Required): Yes

Reviewer 1 (Formal Review for Authors (shown to authors)):

To be fully transparent I should declare that this review is written by Marin De Kauwe. I agreed to review the paper before I realised Anthony Walker was one of the authors (a potential conflict of interest). I subsequently informed the editor but they were keen for me to still review the manuscript.

In this paper Fisher et al. explore the sensitivity of CLM5's vegetation responses to assumed parameters in relation to the nitrogen cycle. Overall the paper is very thorough, clearly written and sets out to achieve its objectives. I do have a series of small suggestions (below) and a few larger suggestions. which I would outline here.

1. (Classic reviewer here - sorry) Whilst the paper does do what it said it would, I can't help thinking this paper could be more interesting if it presented the analysis in a slightly different way. Feel free to ignore all of this!

RR: Noted...

This paper documents three new assumptions (arguably four -> section 2.1.4 as well) to CLM5 related to N; however, it doesn't give much consideration to these hypotheses, instead focussing on a broader list of impacts related to parameter choice. After reading the methods, what I was expecting (hoping) to see, was what impact these different assumptions would have on carbon fluxes. I was then expecting to see the impact of the underlying params from *these schemes (assumptions)* being probed. That isn't what we get from this paper and (my bias) I feel like this is an opportunity missed. I found it hard to interpret the impact of parameters (semi)divorced from the underlying structural assumptions. Moreover, many of the params (in my eyes) had little to do with the new code additions. For example, a lot of text is devoted to exploring the sensitivity to the Medlyn slope parameter. To me, this feels out of scope for what this paper should(could) be.

RR: We appreciate these suggestions. One alternative approach would be to run the simulations with the various pieces of code turned off and returned to the defalt state. The problem with implementing that idea is that large modifications of that type tended to hugely modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the system state was moved to a more or less light saturated condition. We have to some extent tried to probe the sensitivity space of say, the FUN model by modifying the capacity of the plants to take up and fix nitrogen. \textit{I will look into the possibility of doing this.}

So, my suggestion here would be to show the impact of each of the 3(4) new assumptions in relation to how nutrients are treated with CLM5 and then to explore how sensitive those results were to the underlying params. Here, my expectation was that in some cases the results would be insensitive, but in others, it is clear that the notion of a single PFT value does not make sense. As I understand it, the authors varied the gs slope term based on the param range in my 2015 paper. This gives them some erroneous results, but that wouldn't be unexpected as using the C3 grass param (a "spendy" plant) in an arid environment wouldn't make biophysical sense?

RR: You are right, wrt. the range of variation across PFTs perhaps being excessively wide. \emph{Perhaps a lower range should be adopted?}

In my opinion the editor shouldn't impose such a change on the authors as it is potentially a lot of work and not what they wished their paper to be about. It is just my suggestion. Equally, I would have loved to have seen those N hypotheses tested at the Duke/Oak Ridge FACE experiments, perhaps that is for another paper :P.

RR: We actually ran the model for those sites, but got very low LAI predictions, and therefore very sensitive responses to fertilization. We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO2 and N forcing.

2. At the very least, I think the number of params considered should be streamlined. Many feel a bit irrelevant and there is a lot of text devoted to each. Whilst I appreciate the lengths the authors have gone to explaining the impact in the results, I found myself drifting because at times it felt out of scope. I think reducing the number considered (even moving N to the supplementary) would make the paper more readable.

RR: Fair point. It seems better to explore more parameters for the sake of completeness, but maybe some can be parsed out into SI to allow for a more readable paper.

3. The manuscript makes a number of citations to CLM papers in review. In many places these references add little to the paper as they provide no detail and I can't look the paper up (obviously at some point soon I will be able to). My sense is that the paper would benefit from either expanding the text in some places or removing the reference entirely. For example, the Kennedy plant hydraulics paper was referenced on five occasions, that seems excessive to me given the paper is on nitrogen. This might seem like I'm being pedantic, but honestly it breaks the readability of the text.

RR: The MS is part of a special issue, and so the paper will all become available at the same time (except the Kennedy paper which is already published). \emph{We will reduce depdance on these references.}

4. Whilst I wholeheartedly agree that assessing the model simulations against observations doesn't necessarily indicate skill, it does go a bit against the grain the way the authors have presented their results. The plots as presented are quite abstract and the authors may consider this point in revision. As a reviewer, it wouldn't bother me at all if the model seemingly performed worse with added "realism", particularly if the authors could explain why.

5. Finally, with the discussion I'm reading this as a non-CLM user and thinking what can I take away here? Here I struggled a little bit if I'm honest. There is some interesting text about objective calibration (great to be able to cite this :P), but again I'm wondering if this is really in the scope? Six paragraphs are devoted to apparent limitations - seems excessive, there is more to sell here than criticise? I think other modelling groups who haven't yet implemented these hypotheses in relation to the representation of the N cycle could learn a great deal from a paper like this. I'd ask the authors to reflect on what they think the messages they'd like to portray are and see if they feel those messages are in the discussion.

Minor stuff (Key points/Abstract/Intro/Methods): 
------------------------------------------------

- Could the second key point be written in more readable English? In the third key point N has not been defined. Also, fertilisation of what (presumably CO2? But it could be nitrogen...or anything). Finally, "other important parameters depend on the initial system state". This is hard to interpret and without context is limited as a take home point for the reader. I really suggest the authors revise the key points as I take little from them about this study.

RR: Fair poit. We have modified these points to read
\item The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle.

\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.

\item Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state.

- In general, when the authors talk about "model state" - both in the key points and the abstract - doesn't this require some time context? Are we talking about the equilibrium state? The state at 2019?

RR: We have changed this to 'pre-fertilization state'.

- Line 45: what would a passive nitrogen cycle look like?

RR: OK. We took out 'active'. and changed to "included nitrogen (N) cycling and limitation of plant growth."

- Line 46: why is it "might" the model either said that would happen or it didn't.

RR: Changed to "In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to ... "

- Line 52-3: Is that true? Unless that can be supported I would suggest omitting. 
RR: Changed from 'a majority of models' to 'many models'

- Line 64: It would be good not to overlook the numerous earlier studies that did actually explore these issues ... e.g. Rastetter et al. (1991) Tree Physiology 9, 101 -126.; Comins & McMurtrie (1993) Ecological Applications 3, 666-681.; Kirscbaum et al. (1994) Plant, Cell & Environment, 17: 1081-1099, etc. 
RR: Thanks for the references. I didn't know about those.

- Line 70 onwards. Does this paragraph add anything to the introduction? If anything it is more methods anyway ...

RR: OK, we will move this to the methods.

- Line 109: Is 1 really a hypothesis as stated?

RR: Maybe not. \emph{will think of an alternative}

- Line 152: I realise this is more about the LUNA model paper, but I do wonder about the assumption that day length drives variations in photosynthetic capacity, given that you already account for radiation. This just feels like a fudge for a gap in our knowledge as the mechanistic interpretation is less, clear. Anyway, irrelevant I'm sure. 
RR: I am inclined to agree, but there's not much that can be done for this paper, at this point.

- Line 163 onwards: In the description of FUN, it would be helpful to provide a little more detail about how much of the cost is parameterised vs emergent. The text is very clear, but I still found myself wanting to know a little more. Obviously I could read the documentation!

RR: OK, \emph{we will add more text on the emergence of the costs.}

- Line 243: How does the FATES model account for this issue? This seems like it might be an interesting point, but the lack of details means that this statement should either be omitted, or briefly expanded upon.

RR: \emph{Add description of FATES trimming}

- May I suggest (an optional) text reorganisation? Swap 2.13 and 2.12. Then 2.1.4 follows more logically after the text about FUN for the reader.

RR: \emph{swop text round}

- Line 291 - it would be hard to see a paper in review? Is it in open review? Otherwise more detail should be added, I have no issue with keeping the citation.

RR: RR: \emph{It is in review for the special issue at JAMES, and a copy should somehow be in the system to be accessed, although I don't know how referees get access to the other companion papers exactly}

- Line 303 - how is the analysis in this paper different to Wieder et al. in review? I would argue to remove this statement as it adds nothing. Similarly on line 314, reference to Kennedy. What relevance are these sentences?

RR: \emph{IAdd description of what wieder did.}

- Line 311 - it's

- Line 318: I think the approach used by the authors is perfectly legitimate and I see no need to expand words justifying why they did not do a global simulation. I would argue to remove these statements.

RR: \emph{Maybe remove.}

Results 
-------

- Line 423: To not look at soil outputs feels like an oversight to me. At the very least, linking the results to available plant N in the soil would be very interesting. Many of the changes have feedbacks for the soil and I think it is important to consider the responses of some of these variables too.

RR: \emph{Think about this.}

- Fig 2. I can't tell the difference between slatop and stem:leaf.

RR: One is red and the other is a sort of purple. \emph{Will modify the colormap to increase the purpleness.}

- Line 509: this is another example where i feel there is too much assumed knowledge by the reader. Having not read Lawrence et al. in review, it is hard to take much from this. Given I know who Lawrence is, I just end up confused how two CLM papers would contradict each other, but lacking the details.

RR: OK, \emph{we will remove this refernce to the Lawrence paper.}

- Overall, I struggled to see what much of the description around parameter sensitivity had to do with the various nutrient assumptions. For example, the text around the Medlyn model could easily be completely omitted. This is true of a number of these parameter paragraphs. Unless there is a link to the nutrient assumptions I fail to see why these details are included? I also think if you dropped some of these irrelevant params, then the plots (fig 6-8) would be easier to interpret.

- Lines 576: These response ratios are really quite large? GPP increased by 44\% for an increase of CO2 to 550? Bigger still for NPP, e.g. 1.66 at CAX, presumably the CO2 x water interaction? Are these the responses without a nutrient cycle? I must have missed this point if so. I see that the text on line 582 would imply this to be the case. It might be worth clarifying in the results at this point.

RR: OK, this does need clarifying. Actually, CLM5 can't be run without its N cycle. I agree the responses are high. They are in contrast to those found in the global analyses. \emph{Maybe this needs calling out directly in the paper. }

- Line 701: Or that global maps are a poor way to see how comparable models and observations are ... 
RR: Or that, yes.

Reviewer 2 Evaluations: 
Recommendation: Reject and encourage resubmission 
Significant (Required): There are major errors or gaps in the paper but it could still become significant with major changes, revisions, and/or additional data. 
Supported (Required): Mostly yes, but some further information and/or data are needed. 
Referencing (Required): Yes 
Quality (Required): The organization of the manuscript and presentation of the data and results need some improvement. 
Data (Required): Yes 
Accurate Key Points (Required): Yes

Reviewer 2 (Formal Review for Authors (shown to authors)):

This study presents an introduction to some new modifications to the N cycle representation in CLM5 model, investigates how the model state and the responses of the system to CO2 and N addition are sensitive to parameter choices, and showed that parameters controlling N fixation dominate fertilization responses.

I find it hard to convince myself what does it mean to change parameters for each site without proper calibration. I suppose that most of these parameters are calibratable. Therefore, how these parameters change within a given climate space is somewhat pre-defined. You already suggested that these parameters are influential to plant responses to CO2 and N. Once you calibrated them against a PFT, does it mean that you have significantly reduced degree of freedom in your current testing? The authors made some good points in the discussion (L743-752), for example, calibration does not address the issue of dynamic responses. I think these points are brought in too late. I'd suggest the authors to revise their introduction with these points incorporated more explicitly.

RR: We are unsure how to respond to this statement. \emph{We could reference the text earlier}

Additionally, I totally agree that it is important to identify how model behaviors are influenced by parameter decisions. However, I don't think that this message is new, and therefore I find the results and interpretations of the results not so informative. To start, I find the section headings in the result section very confusing. For example, within section 3.1, there is only one sub-section, i.e. 3.1.1. Additionally, section 3.3 seems to be just an empty section with no real result included. The result for section 3.3 seems to be in section 3.4. Then section 3.5 is extremely long, and with only one subsection 3.5.1. I suggest the authors to revise the result section entirely, for clearer presentations, and more logical storyline.

RR: We do not suggest that the message is new. Our model is a new version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project. In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.

RR: We admit we had not given much thought to balancing the length of the sections under their various subheadings. \emph{Will reorganize}

Additionally, I think the second half of the abstract (L33-39) not so informative. Instead, I think the author need to be more specific about their findings. Given the increased model complexity, how are model predictions affected by these individual new modifications? Moreover, if the output space is complex, and varied by sites and variables, can the authors identify a list of weaknesses and strengths that the new modifications add to the model? For example, can the authors give more specific recommendations as to what model representations are useful under what circumstances? And, what are the areas of further investigation, both in terms of further data collection to constrain the process, or more focused model sensitivity test to identify uncertainty? I totally understand that you have a tight space in the abstract, but the current form is too general, I think.

RR: \emph{This is quite vague. Needs thought. }

Moreover, I think the authors did not present their hypotheses correctly. The 1st hypothesis presented in this study is really not a hypothesis that is testable. The impact of parameters on what? I think you need to be more specific. Right now, there is really no testable hypothesis presented. How can you prove or disprove it? Given that in the method (L307-309) you have clearly stated that you picked 13 parameters that are influential to your results, you have already "proved" your hypothesis, right? Similar, you haven't defined what a model state means, and therefore, it is my opinion that the 2nd hypothesis is not explicitly testable, either.

RR: \emph{Maybe we should just ditch the hypotheses? }

Lastly, I am not convinced what this paper differs significantly from those that they included as "under reviews". Any parameter decision should have an impact to the model predictions, otherwise you wouldn't have it as a parameter in the model. This study is a sensitivity test paper, but the message is too complicated to make any generalizable take-home messages. I think some significant re-writing is needed.

RR: The paper differs substantially from all the paper in revision, in that it explicitly tests the parameter space of the model.  The remainder of the papers assess the responses of the default configuration. Our aim here is to be transparent about the uncertainties in the model and provide guidance on the interpretation of future results.

Specific comments

L 63 and L1082: De Kauwe. 
RR: Apologies for this Bibtex error.

L 88, CO2, not CO2

L 112: You haven't define what a model state you mean, and therefore, this hypothesis is not explicitly testable.

L 113 - 117: This is a very long sentence. And it's not easy to follow. Suggest rephrasing.

RR:\emph{ We will break it upas needed. }

L268-269: Here the author refers to the full CLM5 technical report for the description of the new algorithm. I find it not satisfactory. I think the author should give all the details relevant to the result in the paper, not a section within a technical report.

RR: We have chosen to document the model both in the Lawrence et al. paper and in the technical documentatin online for CLM. The full description of the FUN model and the FlexCN algorithm runs to many pages, which would ad substantially to this already quite long manuscript. An alternative would be to add these descriptions to the supplementry infomation, but this would not necessarily provide a greater utility to the reader than the two sources they are already available and would also consitute publishing the same material twice in different peer reviewed manuscripts.

L 307 - 309: I am confused here. So you identified a list of 13 parameters which you think should have direct mechanistic impacts on responses to CO2 and N fertilization. Then you performed sensitivity tests around them, and highlighted that the model indeed is sensitive to these parameter decisions. Isn't the answer obvious? 
RR: This

L341: You say it's fine root area per unit leaf area in bold, but its calculation is the ratio of fine root biomass to leaf area? What exactly is it?

L372: remove ).

L420: Not sure how you ended up with 672 runs. 4 sites x 13 parameters x 5 values * 3 treatments = 780. Did I miss something?

L 436: "The x-axis" to the x-axis.

L438: Table 1, not table 1.

L 447: remove ).

L458-461: Can you rephrase? I can't follow this sentence.

L528: Figure 1, not figure 1. 
RR: Done. Thanks for catching.

L560: the both the? 
RR: Done. Thanks for catching.

L 562: ))? 
RR: Done. Thanks for catching.

L567: Figures 6 to 8. 
RR: We are not sure what the suggestion is here.

L743 - 752: I think these lines should be more thoroughly discussed. This is clearly the key novelty of this research, but I am afraid that I don't see them identified, and discussed enough throughout the text. 
RR: \emph{Make this clearer}

L782: leaf nitrogen, not leaf Nitrogen. 
RR: Done. Thanks for catching.

L811: nitrogen.

L829: replace "(/citewieder2019" with proper text. 
RR: Done. Thanks for catching.

Table 1: What's default? Also, why some scenarios have the same multipliers?

\section{ideas for how to approach re write}
Modify to lose focus on N cycle updates.
OR
actually do simulations without LUNA etc.

\end{document}

### 2019-05-28 SIfigs.tex

@@
 
+\title{Supplementary Infomation for \emph{Parameteric controls on vegetation responses to biogeochemical forcing in the CLM5} by Fisher et al. (submitted to JAMES)}
+
+    \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{Leaf N CO2 and N respones 2001}
+  \end{figure}
+
+    \begin{figure}[h]
+     \centering
+     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{VEGC CO2 and N respones 2001}
+  \end{figure}
+
+\begin{figure}[h]
+     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the model response (fertilized/control) to 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
+     \label{NR1 celev}
+  \end{figure}
+
+  \begin{figure}[h]
+     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the model response (fertilized/control) to 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
+     \label{NR1 ndep}
+  \end{figure}
+
  \begin{figure}[h]
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Caxiuan\~a }
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the model response (fertilized/control) to 15 years of CO$_{2}$ fertilization (550ppm) at Caxiuan\~a }
      \label{CAX celev}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 uears of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Caxiuan\~a}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the model response (fertilized/control) to 15 uears of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Caxiuan\~a}
      \label{CAX ndep}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Harvard Forest}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the model response (fertilized/control) to 15 years of CO$_{2}$ fertilization (550ppm) at Harvard Forest}
      \label{HRV celev}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Harvard Forest}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the model response (fertilized/control) to 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Harvard Forest}
      \label{HRV ndep}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Metolius Forest}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the model response (fertilized/control) to 15 years of CO$_{2}$ fertilization (550ppm) at Metolius Forest}
      \label{MET celev}
@@
      \includegraphics[width=1.3\textwidth,left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Metolius Forest}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the model response (fertilized/control) to 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Metolius Forest}
      \label{MET ndep}
@@
 
-    \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{VEGC CO2 and N respones 2001}
-  \end{figure}
-
-  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{Leaf N CO2 and N respones 2001}
-  \end{figure}
 \end{document}

### 2019-05-28 figures.tex

@@
 
+\nocite{*}
+\bibliography{aguCLM5_bibtex}
+\clearpage
+
+\pagebreak
+\begin{table}
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| } 
+ \hline
+ Parameter Name & Meaning & Units \\
+  \hline
+ slatop & Specific Leaf Area & m2/g \\ 
+ froot\_leaf & Fine root:leaf ratio. & g/g \\
+ stem\_leaf  & Stem:leaf ratio. & g/g \\ 
+ n\_costs    & 6 parameters related to Nitrogen costs & g/g \\
+ fracfixers  & Fraction of C available for fixation & - \\
+  grperc  & Fraction of C used for growth respiration  & - \\
+  leafcn  & Leaf C:N ratio. & g/g \\
+     medlyn\_slope  & Slope of stomatal conductance curve & g/g \\
+      lmr\_intercept & Intercept of leaf N /respiration relationship & \\
+      frac\_ectomy\_fungi} & Fraction of active uptake that is ectohycorrhizal & - \\
+      cn\_flex\_a & Parameter of N flexibility algorithm& \\
+      cn\_flex\_b & Parameter of N flexibility algorithm & \\
+      cn\_flex\_c & Parameter of N flexibility algorithm & \\
+\hline
+\end{tabular}
+\end{center}
+\caption{CLM5 Parameters subject to sensitivity analysis in this study. }
+\label{table_parameters}
+\end{table}
+
+ \begin{table}
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| } 
+ \hline
+ Parameter Name & s1 &s2 & s3 & s4\\
+  \hline
+ slatop & 0.6761 & 0.8769 &1.2785 &1.4791\\ 
+ froot\_leaf &  0.342 &0.671 &1.05 & 1.1\\
+ stem\_leaf  &  0.3057 &0.61 &1 &1\\ 
+ n\_costs    & 10$^{-2}$ &10$^{-2}$&  10$^{1}$& 10$^{2}$\\
+ fracfixers  & 0 &0.5 & 2 & 4 \\
+  grperc  &  0 & 0.5& 2 & 3\\
+  leafcn  &0.7413 & 0.8932 & 1.1970  & 1.349\\
+
+     medlyn\_slope  &0.5258 & 0.7191 & 1.1057  &1.29\\
+      lmr\_intercept & 0.8539 & 0.9531 & 1.1512& 1.25\\
+      frac\_ecto & 0 &0.5& 1 & 1 \\
+      cn\_flex\_a &0 &0.5 & 2  &4\\
+      cn\_flex\_b & 0.1 & 0.5 & 2  & 4\\
+      cn\_flex\_c &0.001 & 0.1 & 10 & 100\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Multipliers on the original values in CLM5 used in this sensitivity analysis .}
+\label{table_ranges}
+\end{table}
+
 \begin{figure}[h]
@@
      \caption{Response of drought index ($B_{tran,mn}$) and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
-Legend colours as for Figure \ref{NR1 state}})
+Legend colours as for Figure \ref{NR1 state}}
      \label{btran state}
@@
     % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1.png}
+     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1_2012.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Metolius Forest}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-NR1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the model response (fertilized/control) to 15 years of CO$_{2}$ fertilization (550ppm) at Niwot Ridge}
      \label{NR1 celev}
@@
      \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-NR1_ens_MIC_p1.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table \ref{table_ranges}) on the model response (fertilized/control) to 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Niwot Ridge}
      \label{NR1 ndep}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
-     \caption{Influence of parameteric variation on the impact of 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
+     \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
 }
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_NPP1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on net primary productivity (NPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on net primary productivity (NPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
      \label{NPP CO2 and N respones 2001}
@@
      \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TLAI1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf area index (LAI), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on leaf area index (LAI), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
      \label{LAI CO2 and N respones 2001}

### 2019-05-28 main_revision.tex

\documentclass[draft,linenumbers]{agujournal}

\draftfalse

\usepackage{hyperref}
\hypersetup{
colorlinks=true,
linkcolor=blue,
filecolor=magenta,
urlcolor=cyan}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}

\begin{document}
%tentative author list...
\authors{Rosie A. Fisher\affil{1},
William R. Wieder\affil{1,2},
Benjamin M. Sanderson\affil{3},
Charles D. Koven\affil{4},
Keith W. Oleson\affil{1},
Chonggang Xu\affil{5},
Joshua Fisher\affil{6},
Mingjie Shi\affil{6},
Anthony P. Walker\affil{7},
David M. Lawrence\affil{1}
}

\title{Parametric controls on vegetation responses to biogeochemical forcing in the CLM5}
\author{Rosie Fisher}
\date{November 2018}

\affiliation{1}{National Center for Atmospheric Research, Table Mesa Drive, Boulder, Colorado, USA}

\affiliation{2}{Institute of Arctic and Alpine Research, University of Colorado, Boulder Colorado, USA}

\affiliation{3}{Centre Europ\'een Research et de Formation Avenc\'ee en Calcul Scientifique, Toulouse, France}

\affiliation{4}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}

\affiliation{5}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}

\affiliation{6}{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, California, USA}

\affiliation{7}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}

\correspondingauthor{Rosie Fisher}{rfisher@ucar.edu}

\begin{keypoints}
\item The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle.

\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.

\item Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state.

\end{keypoints}

\begin{abstract}
A suite of modifications to the nitrogen cycle representation were added to the Community Land Model, version 5.  New features of the model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon cost of nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios and dynamic leaf resorption of nitrogen. Here we assess the sensitivity of the model to a suite of thirteen parameters pertinent to the cycling of carbon and nitrogen by vegetation. We assess the parameter dependence of both the model state, and also the responses of the system to CO$_{2}$ and nitrogen fertilization at individual sites in temperate, tropical and boreal systems.

Model responses to CO$_{2}$ and N fertilization are strongly impacted by how much plant communities can increase their fixation of nitrogen. Model initial state, stoichiometric flexibility of plant tissues, and nitrogen uptake costs all have notable impacts, but in general the output space is complex, and varies between sites, output variables, and fertilization scenarios.  These responses can typically be understood in the context of the model structure and the functionality of individual parameters. Identification of the primary model features driving ecosystem responses to CO$_{2}$  and N fertilization is a necessary step in the interpretation of global carbon cycle feedback simulations, and in the identification of priorities for fieldwork and process updates. Further studies focused on how alternative parameters impact biogeochemical function would provide useful compliments to the ongoing CMIP6 process.
\end{abstract}

\section{Introduction}
Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}). In the CMIP5 multi-model inter-comparison, only one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth. In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}). Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent global N dynamics using alternative structural assumptions have been developed, with variable severity of N limitation in future high $CO_{2}$ scenarios (\cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}). In the CMIP6 inter-comparison process, many models are expected to contain active Nitrogen cycling.

Unlike, for example, representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not underscored by well-tested and widely accepted paradigms, and substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehle2011} \cite{zaehle2014}). Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{kauwe2013}, \cite{zaehle2014}, \cite{medlyn2015}). Few studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model (c.f. \cite{meyerholt2018}). Given the observed large impacts of the inclusion of such processes on responses to increasing CO$_{2}$ and N deposition, understanding the controls on the magnitude of biogeochemical feedbacks both within and between models is an important part of interpretation of their results.

The Community Land Model is the land surface representation of the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019} and plant hydraulics by \cite{kennedy2019}. Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.

The CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4 and CLM4.5 (\cite{lawrence2011}). The introduction of several new dynamical features into the CLM5 has implications for responses to environmental forcing. \cite{wieder2019} report the global-scale carbon cycle implications of the full model update and note that, in the default CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors, with both fertilization responses being closer to experimental results than the previous versions of the model.

Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global biogeochemical feedback analysis, including the controls on global carbon fertilization `$\beta$' values  (the change in ecosystem carbon storage in response to CO$^{2}$ fertilization) (\cite{arora2013}, \cite{friedlingstein2014}),
we consider it a worthwhile undertaking to illustrate how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations.

In this paper, to that end, we describe the new components of the biogeochemical cycling in CLM5, and assess in detail the sensitivity of the CLM5 model parameterization of ecosystem biogeochemistry. We use a range of single site simulations, combined with one-at-a-time parameter uncertainty analyses, and CO$_{2}$ and N fertilization experiments, to investigate the dynamics of the model under steady-state and step-change modifications in forcing. We also hope to highlight areas of process and parameteric uncertainty that appear to require greater attention in future model development and testing cycles.

Model uncertainty quantification efforts are often framed as an attempt to narrow the parameter space to a small range of 'important' parameters, as-determined by their impact on the model's mean-state predictions (e.g. \cite{dietze2014}, \cite{ricciuto2018}).  Predicted global biogeochemical feedbacks, however, are typically related to the response of the system to forcing, and not necessarily to those system properties that control the mean state. Narrowing the range of variables of interest to those that only impact control states and fluxes (and directing more resources to studying those variables) is thus likely to miss important aspects of transient responses under climate change scenarios.  Similarly, much effort is expended on calibration of models to present-day conditions (\cite{fer2018},\cite{lu2018}, when their responses to climate forcing might be relatively unconstrained by these efforts.  Thus, our motivating hypotheses here are:

1. The impact of parameters is complex, and their relative importance varies across site conditions, output variables, and model forcing.

2. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control model state.

The analysis herein does not focus on the relative fit of the model to the data collected at the individual sites. The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, a focus on that would preclude detailed analysis of the parameter space, as presented here. Assessment of the skill of the CLM5 model against a broad suite of globally-relevant data products and manipulation experiments is the focus of \cite{lawrence2018} and \cite{wieder2019}. Future studies will consider the impact of alternative parameterizations, informed by these analyses, on the global performance and skill of CLM5.

\section{Methods}

\subsection{Updates to Nitrogen Cycling in the CLM5}
The CLM5 nitrogen cycle builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}). The new model integrates three major additional prognostic elements of the nitrogen cycle, including:\\

1) The `LUNA' module (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}) \\

2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics and carbon economics of N acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}), and\\

3) The `FLEXCN' N cycle implementation, from \cite{ghimire2016}, which is used in CLM5 primarily to allow variation in tissue C:N ratios.\\

More details on each of these new components is given below. The full technical description of the CLM5 is an appendix to \cite{lawrence2018} and is also available online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html).}

\subsubsection{LUNA}
Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}). Photosynthetic capacity is a key model parameter (\cite{rogers2017}), and the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).

Given the current leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components, for the prevailing time-averaged environmental drivers. The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes. In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area $N_{area}$, (which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and prevailing meteorology. The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was demonstrated by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

\subsubsection{FUN}
Under circumstances where insufficient N uptake was available to match all of the carbon assimilated (for a given target C:N ratio), the CLM4 model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake. This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).

A major issue confronting N cycle models remains how to deal with this 'excess' carbon under N limiting conditions (\cite{dekauwe2014}). The FUN model (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}) addresses this issue by imposing the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, either directly from soil to roots or from mycorrhizal fungi to roots, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}). The cost of each uptake pathway is variable in time and space as a function of nutrient availability and temperature (in the case of the symbiotic N fixation pathway). FUN hypothesizes that plants will spend excess carbon on nitrogen uptake,  preferentially from pathways which are `cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to both reduced growth and higher expenditure of C on N uptake. The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for nitrogen uptake, ($C_{nuptake}$).

\begin{equation}
C_{avail}=C_{growth}+C_{nuptake}
\end{equation}

Further, the nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant C:N ratio ($CN_{plant}$):

\begin{equation}
N_{uptake}=C_{nuptake}/N_{cost}
=\frac{C_{growth}}{(CN_{plant}/(1.0+f_{gr})}
\end{equation}

The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$, which applies only to tissues that are constructed). Rearrangement of these constraints gives a predictive term for the amount of carbon that is expended on N uptake, as:

\begin{equation}
C_{nuptake} =\frac{C_{avail}}{ ((1.0+f_{gr})*(CN_{plant} / N_{cost}) + 1) }
\end{equation}

The average nitrogen cost ($N_{cost}$) is derived from the uptake of N, weighted for its cost across all uptake streams, while the plant CN ratio ($CN_{plant}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools. The FUN model is documented at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}

One mechanism introduced by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}). The CLM4 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{cleveland1999}), and \cite{wieder2015} illustrated the impacts of uncertainty in this function.  FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}). Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation,  typically when soil N availability is very low.

Note that while the CLM5 has no representation of biological succession (\cite{fisher2015}, \cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state (carbon and nitrogen pools in equilibrium) from bare ground. In most cases, N fixation is preferred when soil N is low, and so soil N builds up quickly. In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.

The FUN model introduces the concept of carbon payments for N uptake, which are intended to implicitly represent the carbon expended on root exudates, symbiotic relationships with soil fungi, enhanced expenditure on dynamic fine root growth or increased metabolic rates to support active uptake. A more comprehensive description of these processes might include explicit representations of these processes, but the model deployed here uses a bulk approach to account for the costs to plants of low environmental N availability. Therefore, uncertainties associated with the cost functions, in particular, are large.

\subsubsection{FLEXCN}

Plant trait databases (e.g. \cite{kattge2011}) typically indicate a wide variability in tissue N concentrations within a single plant functional type, and this is in part driven by climate (\cite{wright2005}).  To address the impact of environmental conditions on tissue N content, 
\cite{ghimire2016} implemented a suite of modifications to the CLM4.5 Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm. In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents nitrogen uptake, thus the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying N supply.

In the new allocation algorithm, the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).  The carbon allocation scheme uses a fixed-fraction approach to allocation in general, with constant ratios between, leaf, fine roots, stem and coarse root allocation. We note here that the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an feature which must be considered when analyzing simulations under high productivity conditions. Newer codes within the CLM framework (e.g. the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue.

FLEXCN allows a second set of degrees of freedom, in addition to those provided by LUNA, to the problem of maintaining stoichiometric relationships during photosynthesis and allocation in the presence of N limitation.  A distinction between the FUN and FLEXCN approaches is that, whereas the FUN logic links the carbon to nitrogen cycles synchronously within a given timestep and at the physiological point between GPP and NPP, FLEXCN links the two cycles asynchronously and at the physiological stage of GPP directly.  The relative roles of these two different approaches to carbon-nitrogen coupling can be determined parametrically, as discussed below.

\subsubsection{Adaptation of FUN to flexible CN ratios}
The FUN model was originally conceived with a static tissue C:N ratio, and will by default allocate an amount of carbon to N uptake that exactly tracks the target C:N ratio. In CLM5, one of the aims of the development process was to incorporate a varying C:N ratio, as above. To that end, it is necessary to modify the amount of C expended on N uptake (and thus the amount of N obtained) with the prevailing cost of N. The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include semi-heuristic and/or bounded representations of tissue N adjustments under N limited conditions (\cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content. When N is expensive in the environment, less C is spent on uptake, increasing C:N ratios where N is scarce. When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high. This new algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.

\subsection{Sites}
We utilized four observational flux tower sites to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, 44.5$^{o}$N -121.6$^{o}$E).  Each site was driven with the available site-specific meteorological data. Soil texture, color, depth and, plant functional type composition were taken from previous model validation exercises undertaken at these sites \cite{bonan2012} for Metolius and Harvard Forest, \cite{fisher2007} for Caxiuan\~a and \cite{bonan2013} for Niwot Ridge.   We note here that our use of a small number of single sites cannot fully describe the environmental conditions under which the model is applied. Instead, our aim is to provide detailed illustrations of how model parameter choices impact a model outputs under a range of realistic forcing conditions. Conducting and analysing these simulations globally is beyond the scope of this study, but investigations of the response of global biogeochemical feedback strength to the important dimensions of parameter space identified by this investigation will provide the basis of future studies.

\subsection{Simulation Protocol}
We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the decomposition of slower cycling carbon and nitrogen pools is increased for the duration of the spin-up, and the pool sizes are modified accordingly at the end of the spin-up - see (\cite{lawrence2018}). We used pre-industrial CO$_{2}$ concentrations (274ppm) and recycled the available site-level half-hourly meteorological inputs for the duration of the simulations. 400 years was sufficient to bring all soil and vegetation carbon and nitrogen pools into equilibrium. Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years for each site, parameter and value combination, with accelerated mode turned off. Restarting from the state at the end of the perturbed parameter spin-ups, we ran 'control' transient simulations for each site and parameter combination starting in 1760 and invoking historical transient CO$_{2}$ and N deposition until 2018.

For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation. Following \cite{wieder2019}, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998. Similar to \cite{wieder2019}, we assessed the impact of changes after 15-19 years of fertilization in both cases.

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the model.  Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization. These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not emnurate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in it's own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.  We focus here on biogeochemically important inputs. Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme will be discussed by \cite{kennedy2019}.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around the default state, which (compared to a global parameter sensitivity analysis) allows more straightforward visualization and interpretation of the results.  At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below. We recognize that a global parameter sensitivity analysis would also be of interest, particularly to identify interacting effects, and indeed have conducted many such analyses. We concluded that a OAAT analysis provided clearer insight into the functionality of the parameters with respect to biogeochemical forcing, and further that the more complex visualization and analysis of a global sensitivity study, across the entire domain of results presented here, would reduce the opportunity for understanding the individual parameter responses in depth, as is the goal of our study.

\subsubsection{Parameters selected for perturbation experiment}

The parameters included in our sensitivity analysis are listed in Table \ref{table_ranges}. Their impact on model functionality is described below. As before, further details can be found in the appendix to \cite{lawrence2018}. The shorthand version of the parameter name is given in italics.

\textbf{Specific Leaf Area}, \emph{`SLA'}, is a parameter common to almost all land surface models, which determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass. High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies and greater interception of light. In CLM5 high SLA leaves by definition also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.

\textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf area. In the CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow (\cite{kennedy2019}) and the cost of N acquisition in the FUN model, respectively. The impact of this parameter is thus of greater interest than its functionality as a simple 'tax' on plant growth in previous versions of the CLM. Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.

 \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody tissues, which in CLM5 is again a constant fraction of leaf allocation. Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass. Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter, a feature which will be resolved in future model versions that allow plants to compete for light resources (\cite{fisher2018}).

\textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The six \emph{'n\_costs'} parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.

\textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}, is the fraction of the net primary production (NPP) or gross primary productivity (GPP) minus autotrophic respiration (not including carbon spent on N uptake, $r_{auto}$) that can be used for symbiotic N fixation.  It is thus a proxy for the ecosystem-level fractional productivity of symbiotically N-fixing plants. The constant \emph{`fracfixers'} parameter here acts in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM.

\textbf{Growth respiration)}, \emph{`grperc'}, or the (fixed) fraction of carbon spent on respiration per unit of new tissue, was revised substantially in CLM5 to reflect new estimates that account for methodological issues with the observation of growth respiration (Atkin O., pers comm.) from 0.3 to 0.11. CLM5 further adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significantly increases maintenance respiration costs. Here we test the impacts of this change on overall carbon budgets.

\textbf{Leaf carbon:nitrogen target}, \emph{leafcn}. CLM5 uses the concept of a 'target' leaf C:N ratio, around which flexibility is allowed, given the environmental costs of Nitrogen. As with most land surface schemes, higher leaf N concentrations lead to both higher maintenance respiration costs and higher photosynthetic capacity (in this case via the LUNA model).

\textbf{Slope of the stomatal conductance model}, \emph{medlyn\_slope}. CLM5 replaced the Ball-Berry stomatal conductance scheme with the Medlyn stomatal conductance scheme (\cite{medlyn2011}). The slope parameter determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit, in the absence of plant hydraulics limitations. Thus, is it intrinsically linked to the response of ecosystems to CO$_{2}$.

\textbf{Leaf maintenance respiration intercept}, \emph{lmr\_intercept}. The new leaf maintenance respiration model, proposed by \cite{atkin2015} and adopted by CLM5, has it's intercept (but not slope) as a PFT dependant parameter, with variation between PFTs reported by \cite{atkin2015} and used herein. Here we test the impacts of the uncertainty in the observations of this parameter.

\textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is an ecosystem-level property defined by the FUN model which allows the proportions of the fungal symbiotic pathways to be modified. Default parameters are as provided by \cite{shi2016}.

\textbf{Parameters a,b and c of the flexible C:N model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.

\subsubsection{Parameter Ranges}
\emph{SLA} and \emph{leafcn} are readily observable and thus among the most common traits represented in the TRY plant trait database (\url{www.try-db.org}). We derive values from the analysis by (\cite{kattge2011}) for their mean and distribution. For both parameters, values are log-normally distributed. We here take the PFT average log-normalised standard deviation range from TRY and apply it across all PFTs (thus the lower bounds are closer to the mean than the higher bounds, Table \ref{table_ranges}). The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across their whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake).  \emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM(4, 4.5) value (0.3). \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}. Reasonable ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood. Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization (thus, these results should not be interpreted as covering a constrained range of variation). The same issues apply to $N_{costs}$ parameters.

\section{Results}
Our analysis used four sites, with thirteen parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations. Including the default simulations, the total comprises 672 different run configurations. We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$). We did not look at soil output variables as our parameter analysis is focused on vegetation responses.

In this section we assess first the impacts of parametric variation on the system state and second the responses to carbon and nitrogen fertilization. We do not focus on the goodness of fit to observations, because our aim is to discern the primary parametric controls over CO$_{2}$ and N fertilization responses, and including extensive data-fitting and calibration experiments (including on parameters mostly unrelated to C and N cycling) would detract from this goal.

\subsection{Overall system state}
The parameter responses of the site-level water and nitrogen limitation (see below) in the control (unfertilized) parameter ensemble are shown in figure \ref{btran state}. GPP, NPP, LAI, leaf N content per unit area and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites. In each of these figures, the y-axis  corresponds to the output variable and The x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in table \ref{table_ranges}. Lines of differing colors illustrate responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case.

\subsubsection{Water and N limitation across sites}
The four sites varied in the degree to which they are limited by water and N availability in the control state, with important impacts upon the parametric response characteristics (Figure \ref{btran state}). Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential (\cite{kennedy2019}). $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$). Given that leaf water potential and  $B_{tran}$ vary diurnally, we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure). For N limitation, the fraction of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake is a direct indicator of how N stress affects the carbon budget.

The Caxiuan\~a site experienced substantial drought stress in these simulations (with the mean $B_{tran,mn}$ of the default condition around $\sim$0.55), with little nutrient stress. The simulate forest at Niwot Ridge experienced substantial amounts of drought ($B_{tran,mn}$ $\sim$0.5), and N limitation, spending 27\% of net assimilated carbon on nutrient acquisition. Harvard Forest site had a moderate drought stress ($B_{tran,mn}$ $\sim$90\%), and moderate N limitation ($\sim$15\% of assimilation spent on uptake). The Metolius site had some drought stress ($B_{tran,mn}$ $\sim$86\%) and very little nutrient stress. The differences in the initial conditions at the sites have implications for the parametric responses, since some parameters, as limitations in one resource can affect the ability of the system to respond to fertilization and/or parameter modifications affecting other resources.

\subsection{Parametric control over system state}
At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected since it is used directly in the calculation of both of these quantities. For GPP, the impact of SLA is always positive, despite the higher LAI causing lower N per unit area of leaf (which reduces per-unit leaf area photosynthetic capacity). The impact of SLA on NPP can, however, be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI. At Niwot Ridge, the parameter range straddles an optimal SLA (or local maximum at the second-largest parameter values) for NPP, again reflecting the assimilation/respiration cost trade-off.

Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation through root-mediated N uptake costs.  The first two impacts decrease NPP with \emph{froot\_leaf} and the latter two increase it.  Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios, reflecting mechanisms 3 and 4.  Harvard forest and Caxiuan\~a both show increases in all variables with low \emph{froot\_leaf}.  In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at Metolius, in particular.

Low values of \emph{stem\_leaf} ratio typically resulted in much higher LAI at all sites, as well as declines in vegetation carbon, as expected from the lower stem carbon allocation. At Niwot Ridge, there is an optimum (or local maximum) stem carbon observed, due to the impacts of a too-high leaf area index, which caused declines in NPP at low \emph{stem\_leaf}. At all other sites, the impact of increasing leaf allocation is to increase NPP (almost doubling at Metolius). Decreases in stem:leaf ratio also caused a decline in leaf N per unit area, potentially on account of the higher demand for N uptake and thus higher N limitation. This effect is llustrated clearly for Niwot Ridge and Metolius in Figure \ref{btran state}).

The first-order impact of increasing \emph{Ncosts} was to increase the respiration losses associated with nitrogen uptake.  Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of \emph{Ncosts} on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low. Other sites showed more muted responses in the vegetation state.

Increasing \emph{fracfixers} affects the maximum amount of carbon that plants can pay towards symbiotic nitrogen uptake (N fixation). For all sites and variables here, it has a relatively moderate and positive impact on the control state, in particular at the Niwot and Harvard forest sites. As is the case for \emph{N\_costs}, with high fixing capacity, N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that LAI increase.

Changing \emph{leafcn} -the target leaf C:N ratio- over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.  In particular, the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).  We do not find much evidence of saturation in the GPP vs. \emph{leafcn} relationships, in that GPP response does not level off at high leaf N content, in line with findings in \cite{lawrence2018} that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP,  decreases in respiration rates caused by high \emph{leafcn}, and reduced carbon expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and respiration rates as a result. At Niwot, an optimum \emph{leafcn} ratio (or a local maximum in the parameteric response) was apparent. For NPP this was around the default value, but for LAI it was 0.5 standard deviations lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

Responses to modifications in \emph{grperc}, (the growth respiration coefficient) were in line with first-order expectations, where higher values resulted in NPP reductions and corresponding changes in LAI, and GPP. The impact varied between sites, reflecting different partitioning of the carbon budget into maintenance respiration, nitrogen uptake and growth.

Changes in \emph{medlyn\_slope}, the stomatal slope parameter, most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux. Water limitation is also modified as a result of altered transpiration. At Caxiuan\~a (Figure \ref{CAX state}), the very lowest values of  \emph{medlyn\_slope} caused a complete death of the ecosystem. In Figure \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses), presumably at the cost of limiting photosynthesis past the point at which positive plant growth could be maintained. At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance. Note that \cite{kennedy2019} hydrodynamic functioning of the model in greater depth and that  studies currently underway will discuss surface energy and plant.

The variance in \emph{lmr\_intercept} (across the range of reported intercepts in \cite{atkin2015}) also has important consequences for ecosystem productivity, broadly in line first-order expectations.  At Caxiuan\~a, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states. Whether the full range of values reported by \cite{atkin2015} is an appropriate reflection of the range of uncertainty for all PFTs is debatable, and a narrower range might be discernible with finer interpretation of their data.

\emph{frac\_ecto\_fungi}, the fraction of ectomycorhizzal vs arbuscular fungi had limited impact here, suggesting that the parameterization of the varying costs between these uptake pathways did not have a large impact on N uptake. This suggests that this difference is overridden by other parts of the dynamics parameterized in FUN.

\emph{cn\_flex\_a} is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details). Thus, low values of \emph{cn\_flex\_a} result in reductions in expenditure on N uptake when N is too expensive, and therefore in lower tissue N content, and vice-versa.   For the range of parameters and the sites investigated here, this parameter had limited impact on the system state.  Increasing \emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, and also has limited impacts on the system state here. \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying. For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.

\subsection{CO$_{2}$ and Nitrogen deposition responses}
Figures \ref{GPP CO2 and N respones 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the responses to both N and to CO$_{2}$ fertilization simultaneously, for GPP, NPP and, LAI. Corresponding figures for vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1 and 2). Figures for each site showing the individual N and CO$_{2}$ responses are also included in the SI (Figures 3 to 10).

We calculated the impacts of the CO$_{2}$ and N deposition by assessing the ratio of the fertilized to control simulations, averaging years 15-19 after fertilization began. For some parameter values, the constraint on growth placed the control system in a state close to or at death (zero live carbon pools). Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass was already zero, leading to some large non-linearities in the relationship between parameters and environmental forcing. To focus on the responses of parameter combinations that give more reasonable initial vegetation conditions, parameter combinations that caused LAI to be less than 60\% of the control simulation value were excluded from the fertilization plots.

\subsection{Differences across sites}
The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.

For the default parameterization (the intersection of all lines in figures \ref{GPP CO2 and N respones 2001} to \ref{LAI CO2 and N respones 2001}), the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.  LAI responses (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) reflected the NPP change, except for at Metolius where the canopy did not respond to the higher NPP.

For Nitrogen responses, while impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET), they were lower for the drought-stressed Caxiuan\~a and Niwot sites.  For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET), increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused increasing respiration and negative impacts on NPP at some sites, even though impacts on LAI (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET) were all positive, reflecting the lower costs of N uptake and greater capacity to acquire N to build tissues and retain carbon for growth.

\subsection{Impact of parameters}
Across the four sites, parametric impacts on fertilization responses were very varied, when compared to the relatively consistent impact of parameters on the initial system state. A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses across sites, and the range of their impacts varied over an order of magnitude. Thus, the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change. Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.

The strong impact of \emph{slatop} had important implications for the initial state in many cases, and for several of the lower values, we screened out the results of the fertilization on account of very low LAIs (which resulted in erratic fertilization responses). The impact of \emph{slatop} on fertilization was different at every site. For example, at Metolius, high  \emph{slatop} was related to high N responses, plausibly since these cheaper leaves could produce a greater LAI/GPP feedback in response to the N fertilization. At Niwot Ridge, high SLA was related to strong GPP responses to CO$_{2}$. These simulations have higher water limitation (on account of their larger canopies, figure \ref{btran state}) and so respond more vigarously to the increased water use efficiency of high CO$_{2}$ growth conditions (while reducing their NPP and LAI with N fertilization, on account of higher respiration costs). These contrasting results are illustrative of the complexity of the interactions between model initial state, environmental forcing and parameterization, and of the need to understand how multiple simultaneous growth limitations can influence responses to climate forcing.

Despite its strong impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO$_{2}$ responsiveness, and at Harvard, where low values decrease both responses. At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP (which is a non-linear response related to the saturation of light interception).

Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuan\~a and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001}). Initial LAI was larger in these cases, explaining the decrease in fertilization response as above (lower increases in GPP for a unit change in LAI).  At Niwot Ridge, low values of \emph{stem\_leaf} were associated with very high LAI and high water limitation (Figure \ref{btran state}), and CO$_{2}$ fertilization thus had a large impact on GPP and NPP (as for \emph{slatop}), but the high NPP has no impact on LAI (SI Figure 3), reflecting the still-high N limitation.  Correspondingly, adding Nitrogen in the low \emph{stem\_leaf} runs increases LAI substantially, but only has a very limited impact on GPP and NPP, given water limitations.

The \emph{N\_costs} parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}). At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).  These high \emph{N\_costs} scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.   Low values of \emph{N\_costs} were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.  While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.

The fraction of N fixers (\emph{frac\_fixers}) has a particularly dominant role over the model response to step changes in N and to CO$_{2}$. For GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP CO2 and N respones 2001}, \ref{LAI CO2 and N respones 2001}).  For NPP, N deposition responses were muted at Metolius, and reversed in sign at Caxiuan\~a, where the low N fixer cases had a very negative (-8\%) NPP response (despite a +7.6\% GPP response), on account of the model increasing the N content of the leaves (and thus respiration rates) by up to 10\% in the lowest N cost case (SI Figure 6). Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.  At Niwot Ridge the changing fixation capacity had no impact on the response of GPP to N fixation, despite its influence on LAI, suggesting that water limitation was a greater constraint on total GPP than leaf area.  At Metolius, again the changing fixation capacity had no impact on the response of GPP to N deposition (Figure \ref{GPP CO2 and N respones 2001}) but in this case, it had no impact on LAI either. Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization at this site, suggesting fundamental metabolic limitations to canopy carbon economics (e.g. a short growing season) that render only higher LAI simulations sensitive to fertilization.

The impact of target \emph{leafcn} on fertilization responses was not particularly dominant (recalling that this was a significant control over the system state at all sites), an exception being Niwot Ridge, where low values of \emph{leafcn} had very strong to CO$_{2}$ fertilization (the strongest response at this site for GPP). These cases suffered from high water limitation, given their high $N\_{leaf}$ values and high gas exchange fluxes (Figure \ref{btran state}) and thus CO$_{2}$ fertilization alleviates this problem. At Caxiuan\~a, high \emph{leafcn} cases had stronger responses to N fertilization, but responses to N fertilization were low at this site in general.

\emph{grperc}, or the degree to which growth is 'taxed', is constant and not responsive to any of the physiological processes upstream that depend on fertilization, thus, at most sites, its impacts are minimal, or indirect.

At Niwot Ridge, runs with low values of the \emph{medlyn\_slope} parameter (and thus high water use efficiency) were able to respond very strongly to increasing Nitrogen (SI Figure 4), given their lower initial water stress (Figure \ref{btran state}). Surprisingly, however, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.

In general the impacts of the maintenance respiration intercept \emph{lmr\_intercept} were small, and limited to the impact of large values increasing fertilization responses to CO$_{2}$ and N, potentially on account of the low LAI and thus greater feedback between LAI and GPP, and in line with its limited impact on the system state, \emph{frac\_ectomy\_fungi} had no discernible impact on fertilization responses.

There was limited impact of \emph{cn\_flex\_a} on the fertilization responses, as per it's impact on the state.  At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$. Leaf N concentration followed suit, increasing substantially in the N deposition experiment. Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when \emph{N\_costs} are high, therefore potentially suppressing leaf N content. Adding N overcomes this suppression, leading to greater fertilization responses.

The impact of \emph{cn\_flex\_c} (the parameter controlling the degree to which plant C:N ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon (SI Figure 2)  at Harvard Forest and Niwot Ridge (and Caxiuan\~a, to a smaller extent). These impacts do not mirror those on leaf area index or NPP, and overall the response to this parameter appears complex.  At Metolius and Niwot, \emph{cn\_flex\_c} has a dominant control over the impacts of CO$_{2}$ fertilization on $N_{leaf}$ (SI Figure 1).  The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high (e.g. SI Figure 3).  Low  $N_{leaf}$ values permit higher leaf area and vegetation carbon, but reduce assimilation rates, as illustrated at Harvard Forest and Niwot Ridge (SI Figure 5 \& 7).

\subsubsection{Low leaf area investigations}
In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites. (for Caxiuan\~a, 3.0 (modeled) vs. 5.3 (observed) \cite{fisher2007}, for Niwot Ridge, 1.7 vs. 4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs. 3.5 \cite{williams1996} and for Metolius, 0.5 vs. 1.5 (\cite{spadavecchia2011}). This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.

To interrogate this further, we ran an entire separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition. In the default parameterization CO$_{2}$ and N fertilization increased LAI, amplifying these magnitude of ecosystem responses to these perturbations. For example, at Harvard Forest, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.18, 1.17 and 1.22 for GPP, NPP and LAI, respectively, in the high LAI case vs. 1.43, 1.66, 1.47 in the control) and a lower N response as well (1.05, 1.20 and 1.15 for GPP, NPP, and LAI, respectively, in the high LAI case vs. 1.1, 1.26, 1.21 for the control). Similar results were found at all other sites.  Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology, and this may explain the large fertilization responses found here compared to those of \cite{wieder2019}.

\section{Discussion}

The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.

At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (Wieder et al. 2018).  The trade-off between responses to N fertilization and CO$_{2}$ response when the fraction of N fixing plants varies mirrors the change in fertilization response between CLM4, CLM4.5 and CLM5, and the CLM5 response is consistant with the capacity of the model to increase fixation rates in response to higher carbon dioxide levels. The number of degrees of freedom illustrated here precludes, however, proof that this is the only contributing factor. This analysis indicates that global experiments with lower or zero \emph{frac\_fixers} would help to ascertain whether this is indeed the case.

The CLM5 model has a number of new prognostic elements, which allow predictions in elements of the C and N cycles which were assumed to be constant in previous versions of the model. Exploration of the parameter space that controls these new elements allows a clear exploration of the sensitivities of the output space of the model to critical assumptions.

The model output space is notable complex, in line with our hypothesis 1. In many cases the sign of the impact of the different parameters changes between sites, and trade-offs in the influence of carbon and nitrogen economic parameters interact with both the initial state and environmental forcings. That the parameter space of CLM5 is complex is not surprising, given its relatively high process fidelity, but a) this type of sensitivity analysis,  illustrating how different parameterizations respond to different growth conditions and environmental forcings, is relatively rare for models of this kind,  b) the large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and c) the results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites. Sensitivity tests across broader geographical ranges would help to eludicate more clearly how changes in baseline water and N limitation impact parameter responses and patterns of responses to fertilization.

Further, our results indicate that, in line with our initial hypothesis 2, model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization, since for each site there was substantial lack of overlap between the parameters controlling the default state and the fertilization responses.  We did observe, however, many instances where simulations with high leaf area index responded less strongly to fertilization, and thus the numerous parameters impacting this feature of the initial model state (dominated by tissue allocation ratios and \emph{slatop}) would also have an indirect impact on fertilization.  Our results suggest a more general point, that high LAI systems should respond less to fertilization. We did not find a clear statistical link between initial LAI and model output across the whole parameter space, but interpretation of global patterns in terrestrial cardon storage should likely take this mechanism into account.

In the CMIP6 generation of land surface models, it is expected that a much greater fraction of model submissions will contain active representations of the terrestrial nitrogen cycle. We propose here that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained for all models, and that these types of analysis would allow a deeper understanding of land surface model outputs and behaviour. Many further analyses are possible, including global feedback analyses with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constrains of the parameter ranges used. We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.

\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 were primarily motivated by the need to move the model behaviour closer to several key dynamic features of the system. Notably, the dynamics of leaf nitrogen content with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}). The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.  In doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.  By investigating the critical roles of these processes in the model dynamics, we hope to highlight a research need, and to stimulate research on and improvements in the representation of these processes.

In analyzing the results of this study, for example, we found numerous instances where the model reduced carbon use efficiency (CUE, defined here as net primary productivity, NPP, divided by gross primary productivity, GPP, where the respiratory carbon used for N uptake is not counted within NPP and thus lowers CUE) by changing tissue C:N ratios in response to forcing. In principle, there exists an optimal leaf C:N ratio that maximizes leaf C export under a given set of conditions. Despite its optimal allocation of leaf N \emph{within} the leaf (to photosynthetic enzymes) via the LUNA model, CLM5 does not include any scheme that ensures that overall tissue CN ratios conform to any physiological optima. Instead, a heuristic model adjusts tissue CN ratios round the target ratio (an input parameter).  Numerous models of this type have been proposed in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{anten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}). We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in  LUNA, but future nitrogen cycling model development efforts could re-visit this topic.

Another area requiring focus is the dominance of the 'fraction of fixers' over fertilization responses. \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants. Technically, because plant density is not represented in the CLM5, this parameter represents the fraction of net assimilated carbon ($gpp$ - $r_{auto}$) that can be used for symbiotic N fixation. Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants. In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}). The 0-1 range of \emph{frac\_fixers} used here is ecologically unlikely, given the costs borne by plants that can fix N give them a competitive disadvantage in systems no longer limited by N. However, it suggest that a high CO$_{2}$ world would likely provide much greater benefits to N fixation, and that those benefits would be ameliorated by ongoing increases in N deposition.

\subsection{ Methodological Limitations}
\subsubsection{Limitations of step-change experiments}
Here we investigate the parameter space of the model using step-change experiments. These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions. Further, we expect that the carbon and nitrogen cycles may take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.

To test the impact of time on the fertilization responses, we also analyzed the impact of fertilization at 5 and 50 years into the elevated CO$_{2}$ and N experiments. At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, (result not shown) to those at 15 years.  However, at 5 years into the fertilization in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent (a x~2.06 impact on NPP, compared to ~0.95 in the year 15 analysis at Niwot Ridge, indicating that the results described above had already been subject to progressive N limitation, and were closer to an equilibrated state.

\subsubsection{Objective calibration of the default model, or absence thereof}
During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products. We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystem to die under lower preindustrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting effect that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets. The parameterization used in the default version of the model consequentially utilized the mean trait values where possible from trait datasets (\cite{lawrence2018}) in combination with targeted manual tuning of parameters (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states.

We recognize that this more subjective calibration method is less than ideal, and ongoing studies will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}.

The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2018}, \cite{lawrence2018}) for assessing model skill across a very broad range of model components.  In principle, integrating more targeted datasets and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic. Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.

Lastly, we recognize the importance of other limiting nutrients, particularly phosphorus, in controlling the influence of CO$_{2}$ fertilization (\cite{reed2015}). Phosphorus dynamics were not integrated into this version of the CLM, but alternative configurations of the CLM exist (e.g. \cite{yang2014}, \cite{zhu2016}) that will provide contrasting biogeochemical predictions in future studies.

\subsubsection{Temperature and water responses}
In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model to transient climate forcing has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water. The suite of parameters that in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implicitly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.

\section{Conclusions}
In this paper we present an analysis of a suite of parameters relevant to vegetation-mediate biogeochemical feedbacks in the extensively modified nitrogen cycle of the Community Land Model, version 5. We illustrate how responses of the model to elevated carbon dioxide and nitrogen deposition might be impacted by parameter choice. We find, in particular, that parameters controlling the degree to which ecosystems can fix additional nitrogen have a dominant effect over responses to fertilization, but limited impacts on the control state. Further, we find that the impact of most parameters is complex and affected by the balance of water, nutrient and other limitations to growth specific to individual sites

It is common practice in land surface modeling literature to present studies that focus on the modification of single aspects of model structure, process representation or calibration, and to highlight the significance of individual modifications when scaled up to global land-mediated climate feedback processes. In this analysis, we illustrate the degree to which a multitude of individual parameters can potentially have important and interacting impacts on model results, how these results differ across sites and output variables and with initial vegetation conditions.  Future studies of process and parameter modification in CLM5 might use these results as a guide to the relative (rather than absolute) importance of individual focus parameters and processes. We thus hope to move forwards from studies that modify individual processes in isolation to those that take a more holistic view of the complexity inherent in modeling coupled land surface systems.

We have attempted here to illustrate and understand the connections between the structure and parameterization of the CLM5 and its biogeochemical functionality. This necessarily complex exercise represents one small part of the assessment of CMIP6-era representations of the land surface needed to understand how advances in model structures alter our predictions of land surface functionality.  The challenges of comprehending, measuring and constraining the parametric and structural uncertainty of increasingly complex process-based and land surface models are daunting. We argue that in order to make progress on this task, the scientific community should move beyond the paradigm of assessing single points in model parameter space in great depth, ignoring all other plausible instantiations. In atmospheric sciences, for example, 'internal variability' caused by initial conditions is now routinely sampled, given its major impact on even long timescale forecasts (\cite{kay2015}). \cite{bonan2018} illustrate how initial conditions are unimportant for carbon cycle prediction on land, relative to structural and parametric uncertainty. Given this, sampling the important dimensions of parametric uncertainty in global Earth system model predictions should be taken seriously during future assessments of biogeochemical feedback strength.

\section{Acknowledgements}
RAF, WRW, BMS, KWO, and DML were supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977. The CESM project is supported primarily by the National Science Foundation.  Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR. WRW was supported by the US Department of Agriculture NIFA Award number 2015-67003-23485 and NASA Interdisciplinary Science Program award number NNX17AK19G. DML was supported in part by the RUBISCO Scientific Focus Area (SFA), which is sponsored by the Regional and Global Climate Modeling (RGCM) Program in the Climate and Environmental Sciences Division (CESD) of the Office of Biological and Environmental Research in the U.S. Department of Energy Office of Science. The material from JBF is based upon work supported by the U.S. Department of Energy, Office of Science, Office of Biological and Environmental Research, terrestrial Ecosystem Science under Award Numbers DE-SC0008317 and DE-SC0016188. JBF carried out part of the work at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration. California Institute of Technology. Government sponsorship acknowledged. JBF was supported in part by the NASA CARBON program. AW is supported by Oak Ridge National Laboratory, which is operated by UT-Battelle, LLC, under contract DE-AC05-00OR22725 to the United States Department of Energy. CX is supported by the United States Department of Energy (US DOE) Office of Science Next Generation Ecosystem Experiment: Arctic project and the UC-Lab Fees Research Program Award 237285.

\section{Data Availability}
Data from these simulations will be made available, prior to publication, on the NCAR-DASH data repository, \url{https://www2.cisl.ucar.edu/dash}

\nocite{*}
\bibliography{aguCLM5_bibtex}
\clearpage

\begin{table}
\begin{center}
\begin{tabular}{ |c|c|c|c|c|c| } 
 \hline
 Parameter Name & units & s1 &s2 & s3 & s4\\
  \hline
 slatop & m2/g & 0.6761 & 0.8769 &1.2785 &1.4791\\ 
 froot\_leaf & gC/gC &  0.342 &0.671 &1.05 & 1.1\\
 stem\_leaf  & gC/gC & 0.3057 &0.61 &1 &1\\ 
 n\_costs    & gC/gN &10$^{-2}$ &10$^{-2}$&  10$^{1}$& 10$^{2}$\\
 fracfixers  & - & 0 &0.5 & 2 & 4 \\
  grperc  &  - &0 & 0.5& 2 & 3\\
  leafcn  & gC/gN &0.7413 & 0.8932 & 1.1970  & 1.349\\

     medlyn\_slope  &- &0.5258 & 0.7191 & 1.1057  &1.29\\
      lmr\_intercept &log(nmol CO$_{2}$ m$^{2}$ s$^{-1}$) &0.8539 & 0.9531 & 1.1512& 1.25\\
      frac\_ecto &- & 0 &0.5& 1 & 1 \\
      cn\_flex\_a & &0 &0.5 & 2  &4\\
      cn\_flex\_b & &0.1 & 0.5 & 2  & 4\\
      cn\_flex\_c & &0.001 & 0.1 & 10 & 100\\
\hline
\end{tabular}
\end{center}
\caption{Parameters, units and ranges (as multipliers of the original values) in this sensitivity analysis .}
\label{table_ranges}
\end{table}

\section{Figures}
\begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/btran_npp.png}
     \caption{Response of drought index ($B_{tran,mn}$) and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites. Legend colours as for Figure \ref{NR1 state}}
     \label{btran state}
 \end{figure}

 \begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Niwot Ridge.}
     \label{NR1 state}
 \end{figure}

 \begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Caxiuan\~a National Forest}
     \label{CAX state}
  \end{figure}

 \begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Harvard Forest}
     \label{HVF state}
 \end{figure}

 \begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1_2012.png}
     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Metolius Forest}
     \label{MET state}
 \end{figure}

 \begin{figure}[h]
     \includegraphics[width=1.35\textwidth]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
     \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites. }
     \label{GPP CO2 and N respones 2001}
  \end{figure}

 \begin{figure}[h]
     \includegraphics[width=1.35\textwidth]{matlab/figures/NOVc_CNdep_NPP1__p2012.png}
     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on net primary productivity (NPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
     \label{NPP CO2 and N respones 2001}
  \end{figure}

  \begin{figure}[h]
     \includegraphics[width=1.35\textwidth]{matlab/figures/NOVc_CNdep_TLAI1__p2012.png}
     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on leaf area index (LAI), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
     \label{LAI CO2 and N respones 2001}
  \end{figure}

\end{document}

### 2019-06-20 referee_responses.tex
Updates from Overleaf

@@
 \begin{document}
-\title{Referee Responses}
 
-Reviewer 1 Evaluations: 
-Recommendation: Return to author for major revisions 
-Significant (Required): There are major errors or gaps in the paper but it could still become significant with major changes, revisions, and/or additional data. 
-Supported (Required): Mostly yes, but some further information and/or data are needed. 
-Referencing (Required): Mostly yes, but some additions are necessary. 
-Quality (Required): The organization of the manuscript and presentation of the data and results need some improvement. 
-Data (Required): Yes 
-Accurate Key Points (Required): Yes
+\textsf{Reviewer 1}
 
-Reviewer 1 (Formal Review for Authors (shown to authors)):
-
-To be fully transparent I should declare that this review is written by Marin De Kauwe.
+To be fully transparent I should declare that this review is written by Martin De Kauwe.
 I agreed to review the paper before I realised Anthony Walker was one of the authors (a potential conflict of interest).
 I subsequently informed the editor but they were keen for me to still review the manuscript.
+
+\textsf{AR: Fair enough}
 
@@
 
+\textsf{AR: Thanks for the positive comments. We hope this version is somewhat easier to digest}
+
 1.
@@
 
-RR: Noted...
+\textsf{AR: Noted...}
 
-This paper documents three new assumptions (arguably four -> section 2.1.4 as well) to CLM5 related to N; however, it doesn't give much consideration to these hypotheses, instead focussing on a broader list of impacts related to parameter choice.
+This paper documents three new assumptions (arguably four- section 2.1.4 as well) to CLM5 related to N; however, it doesn't give much consideration to these hypotheses, instead focussing on a broader list of impacts related to parameter choice.
 After reading the methods, what I was expecting (hoping) to see, was what impact these different assumptions would have on carbon fluxes.
@@
 
-RR: We appreciate these suggestions.
-One alternative approach would be to run the simulations with the various pieces of code turned off and returned to the defalt state.
-The problem with implementing that idea is that large modifications of that type tended to hugely modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the system state was moved to a more or less light saturated condition.
-We have to some extent tried to probe the sensitivity space of say, the FUN model by modifying the capacity of the plants to take up and fix nitrogen.
-\textit{I will look into the possibility of doing this.}
+\textsf{AR: We appreciate the disconnect between the parameter choice being seemingly unrelated to the lengthy discussion of the new nitrogen cycling elements of the model.
+On further reflection,  our overall aim is to illustrate how the C and N responses of the model, as determined for the global simulations by the Wieder et al.
+study, are impacted by parameter choice, given their large implications for carbon cycle feedback analysis.
+We think this is of slightly greater interest than specifically probing the differences between the CLM4.5 and CLM5 N cycle, although the two are related.
+While many of the elements that affect the CO2 and N responses are linked to the new model structure, there are several parameters that have large impacts which are not.
+In the new version of the manuscript, we have chosen to reduce the amount of text devoted to the specific changes to the N cycle, and instead to shift the focus slightly such that we are now looking specifically at how parameters affect the fertilization responses.
+Parameters of the new model components are included specifically to illustrate their impact.
+}
+
+\textsf{An alternative approach would be to run the simulations with the various pieces of code turned off and returned to the default state.
+The problem with implementing that idea is that large changes of that type tended to hugely modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the system state was moved to a more or less light saturated condition.
+We have to some extent tried to probe the sensitivity space of say, the FUN model by modifying the capacity of the plants to take up and fix nitrogen.}
 
@@
 
-RR: You are right, wrt. the range of variation across PFTs perhaps being excessively wide. \emph{Perhaps a lower range should be adopted?}
+\textsf{AR: We agree with this idea wrt.
+the range of variation across PFTs perhaps being excessively wide.
+We have now modified the stomatal slope range to use the standard error reported in Table 2 of de Kauwe et al.
+2015}}
 
@@
 
-RR: We actually ran the model for those sites, but got very low LAI predictions, and therefore very sensitive responses to fertilization.
-We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO2 and N forcing.
+\textsf{AR: We actually ran the model for those sites, but got very low LAI predictions, and therefore very sensitive responses to fertilization.
+We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO$_{2}$ and N forcing.
+}
 
@@
 
-RR: Fair point. It seems better to explore more parameters for the sake of completeness, but maybe some can be parsed out into SI to allow for a more readable paper.
+\textsf{AR: Fair point. It seems better to explore more parameters for the sake of completeness, but \textbf{maybe some can be parsed out into SI to allow for a more readable paper.} }
 
@@
 
-RR: The MS is part of a special issue, and so the paper will all become available at the same time (except the Kennedy paper which is already published).
-\emph{We will reduce depdance on these references.}
+\textsf{AR: The MS is part of a special issue, and so the paper will all become available at the same time (except the Kennedy paper which is already published).
+\textbf{We will reduce depdance on these references.} }
 
@@
 As a reviewer, it wouldn't bother me at all if the model seemingly performed worse with added "realism", particularly if the authors could explain why.
+
+\textsf{AR:  \textbf{needs work}}
 
@@
 
-RR: Fair poit. We have modified these points to read
-\item The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle.
+\textsf{AR: Fair point. We have modified these points to read 
+"The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle."}
 
-\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.
+\textsf{The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.}
 
-\item Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state.
+\textsf{Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state.}
 
@@
 
-RR: We have changed this to 'pre-fertilization state'.
+\textsf{AR: We have changed this to 'pre-fertilization state'. }
 
@@
 
-RR: OK. We took out 'active'. and changed to "included nitrogen (N) cycling and limitation of plant growth."
+\textsf{AR:  We took out 'active'. and changed to "included nitrogen (N) cycling and limitation of plant growth."}
 
@@
 
-RR: Changed to "In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to ... "
+\textsf{AR: Changed to "In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to ... "}
 
 - Line 52-3: Is that true? Unless that can be supported I would suggest omitting. 
-RR: Changed from 'a majority of models' to 'many models'
+\textsf{AR: Changed from 'a majority of models' to 'many models'}
 
@@
 Rastetter et al.
-(1991) Tree Physiology 9, 101 -126.; Comins & McMurtrie (1993) Ecological Applications 3, 666-681.; Kirscbaum et al.
-(1994) Plant, Cell & Environment, 17: 1081-1099, etc.
-RR: Thanks for the references. I didn't know about those.
+(1991) Tree Physiology 9, 101 -126.; Comins \& McMurtrie (1993) Ecological Applications 3, 666-681.; Kirscbaum et al.
+(1994) Plant, Cell \& Environment, 17: 1081-1099, etc.
+\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications}
 
@@
 
-RR: OK, we will move this to the methods.
+\textsf{AR: OK, we will move this to the methods. }
 
@@
 
-RR: Maybe not. \emph{will think of an alternative}
+\textsf{AR: We have deleted this hypothesis and replaced with "2.
+There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization."}
 
@@
 Anyway, irrelevant I'm sure.
-RR: I am inclined to agree, but there's not much that can be done for this paper, at this point.
+\textsf{AR: I am inclined to agree, but there's not much that can be done for this paper, at this point.}
 
@@
 
-RR: OK, \emph{we will add more text on the emergence of the costs.}
+\textsf{AR: OK, \textbf{we will add more text on the emergence of the costs.}}
 
@@
 
-RR: \emph{Add description of FATES trimming}
+\textsf{AR: Add description of FATES trimming}
 
@@
 
-RR: \emph{swop text round}
+\textsf{AR:swop text round}
 
@@
 
-RR: RR: \emph{It is in review for the special issue at JAMES, and a copy should somehow be in the system to be accessed, although I don't know how referees get access to the other companion papers exactly}
+\textsf{AR: It is in review for the special issue at JAMES, and a copy should somehow be in the system to be accessed, although I don't know how referees get access to the other companion papers exactly}
 
@@
 
-RR: \emph{IAdd description of what wieder did.}
+\textsf{AR: Add description of what wieder did.}
 
@@
 
-RR: \emph{Maybe remove.}
+AR: \textsf{Maybe remove.}
 
@@
 
-RR: \emph{Think about this.}
+\textsf{AR: Think about this.}
 
@@
 
-RR: One is red and the other is a sort of purple. \emph{Will modify the colormap to increase the purpleness.}
+\textsf{AR: One is red and the other is a sort of purple. Will modify the colormap to increase the purpleness.}
 
@@
 
-RR: OK, \emph{we will remove this refernce to the Lawrence paper.}
+\textsf{AR: We will remove this refernce to the Lawrence paper.}
 
@@
 
-RR: OK, this does need clarifying.
-Actually, CLM5 can't be run without its N cycle.
-I agree the responses are high.
-They are in contrast to those found in the global analyses.
-\emph{Maybe this needs calling out directly in the paper.
-}
+\textsf{AR: OK, this does need clarifying. Actually, CLM5 can't be run without its N cycle. I agree the responses are high. They are in contrast to those found in the global analyses. }
 
 - Line 701: Or that global maps are a poor way to see how comparable models and observations are ... 
-RR: Or that, yes.
+\textsf{AR: Or that, yes. }
 
-Reviewer 2 Evaluations: 
-Recommendation: Reject and encourage resubmission 
-Significant (Required): There are major errors or gaps in the paper but it could still become significant with major changes, revisions, and/or additional data. 
-Supported (Required): Mostly yes, but some further information and/or data are needed. 
-Referencing (Required): Yes 
-Quality (Required): The organization of the manuscript and presentation of the data and results need some improvement. 
-Data (Required): Yes 
-Accurate Key Points (Required): Yes
-
-Reviewer 2 (Formal Review for Authors (shown to authors)):
+Reviewer 2
 
@@
 
-RR: We are unsure how to respond to this statement. \emph{We could reference the text earlier}
+\textsf{AR: We are unsure how to respond to this statement. }
 
@@
 
-RR: We do not suggest that the message is new.
+\textsf{AR: We do not suggest that this particular message is new.
 Our model is a new version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project.
-In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.
+In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.}
 
-RR: We admit we had not given much thought to balancing the length of the sections under their various subheadings. \emph{Will reorganize}
+\textsf{AR: We admit that we had not given much thought to balancing the length of the sections under their various subheadings. \textbf{Will reorganize}}
 
@@
 
-RR: \emph{This is quite vague. Needs thought. }
+\textsf{AR: \textbf{This is quite vague. Needs thought. }}
 
@@
 
-RR: \emph{Maybe we should just ditch the hypotheses? }
+\textsf{AR: \textbf{Maybe we should just ditch the hypotheses? }}
 
@@
 
-RR: The paper differs substantially from all the paper in revision, in that it explicitly tests the parameter space of the model.
+\textsf{AR: The paper differs substantially from all the paper in revision, in that it explicitly tests the parameter space of the model.
 The remainder of the papers assess the responses of the default configuration.
-Our aim here is to be transparent about the uncertainties in the model and provide guidance on the interpretation of future results.
+Our aim here is to be transparent about the uncertainties in the model and provide guidance on the interpretation of future results.}
 
@@
 L 63 and L1082: De Kauwe. 
-RR: Apologies for this Bibtex error.
+\textsf{AR: Fixed.}
 
@@
 
-RR:\emph{ We will break it upas needed. }
+\textsf{AR:\textbf{ We will break it up as needed. }}
 
@@
 
-RR: We have chosen to document the model both in the Lawrence et al.
+\textsf{AR: We have chosen to document the model both in the Lawrence et al.
 paper and in the technical documentatin online for CLM.
-The full description of the FUN model and the FlexCN algorithm runs to many pages, which would ad substantially to this already quite long manuscript.
-An alternative would be to add these descriptions to the supplementry infomation, but this would not necessarily provide a greater utility to the reader than the two sources they are already available and would also consitute publishing the same material twice in different peer reviewed manuscripts.
+The Lawrence paper is within the same AGU special issue on the CLM5, and we have already obtained editorial support for this strategy.
+Further, the full description of the FUN model and the FlexCN algorithm runs to many pages, which would ad substantially to this already quite long manuscript.
+An alternative would be to add these descriptions to the supplementery infomation, but this would not necessarily provide a greater utility to the reader than the two sources they are already available and would also constitute publishing the same material twice in different peer reviewed manuscripts.
+We have nonetheless reframed the manuscript so that it is more generally concerned with the CO2 and N responses, and not with the N updates per-se, so this is less necessary than before.
+}
 
@@
 Isn't the answer obvious?
-RR: This
+
+\textsf{AR: Our objective here is not simply to illustrate that the parameters have an impact, but to provide an illustrate of how the different parameters impact the model and under what conditions, (as per hypothesis 1).
+Thus, there is value in doing the simulations subsequent to our narrowing of the parameter set to those elements with a direct impact on fertilization responses.
+}
 
@@
 
+\textsf{AR: Thanks for finding this inconsistancy. It is fine root mass to leaf mass. We will correct accordingly}.
+
 L372: remove ). 
+\textsf{AR: Done}
 
@@
 
+\textsf{AR: We miscalculated this.
+It's actually 4*13*3*4 +4*3 = 636, the former term being sites * parameters * experiments * iterations and the latter being the number of control runs at the default (sites * experiments.}
+
 L 436: "The x-axis" to the x-axis. 
+\textsf{AR: Modified accordingly}
 
 L438: Table 1, not table 1. 
+\textsf{AR: Modified accordingly}
 
 L 447: remove ). 
+\textsf{AR:Done}
 
@@
 
+\textsf{AR: We have re-phrased this sentence as "Differences in the initial conditions at the sites have implications for the parametric responses.
+Underlying limitations by various boundary conditions (water, nutrients, light, temperature) can affect both the ability of the system to respond to fertilization and/or parameter modifications affecting other resources, as we describe further in the following sections."}
+
 L528: Figure 1, not figure 1. 
-RR: Done. Thanks for catching.
+\textsf{: Done. Thanks for catching. }
 
 L560: the both the? 
-RR: Done. Thanks for catching.
+\textsf{AR: Done. Thanks for catching.}
 
 L 562: ))? 
-RR: Done. Thanks for catching.
+\textsf{AR: Done. Thanks for catching.}
 
 L567: Figures 6 to 8. 
-RR: We are not sure what the suggestion is here.
+\textsf{AR: We are not sure what the suggestion is here. }
 
@@
 This is clearly the key novelty of this research, but I am afraid that I don't see them identified, and discussed enough throughout the text.
-RR: \emph{Make this clearer}
+\textsf{AR: \textbf{Make this clearer}}
 
 L782: leaf nitrogen, not leaf Nitrogen. 
-RR: Done. Thanks for catching.
+\textsf{AR: Done..}
 
 L811: nitrogen. 
+\textsf{AR: Done..}
 
 L829: replace "(/citewieder2019" with proper text. 
-RR: Done. Thanks for catching.
+\textsf{AR: Done.}
 
@@
 
-\section{ideas for how to approach re write}
-Modify to lose focus on N cycle updates.
-OR
-actually do simulations without LUNA etc.
+\textsf{AR:\textbf{this needs some work} }
 

### 2019-06-24 main_revision.tex

@@
 
-\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.
+\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice, but parameter responses are varied across sites and experiments.
 
 \item Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. 
-
 \end{keypoints}
@@
 \begin{abstract}
-A suite of modifications to the nitrogen cycle representation were added to the Community Land Model, version 5.
-New features of the model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon cost of nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios and dynamic leaf resorption of nitrogen.
-Here we assess the sensitivity of the model to a suite of thirteen parameters pertinent to the cycling of carbon and nitrogen by vegetation.
-We assess the parameter dependence of both the model state, and also the responses of the system to CO$_{2}$ and nitrogen fertilization at individual sites in temperate, tropical and boreal systems.
+Future projections of land carbon uptake in Earth System Models are affected by land surface model responses to both CO$_{2}$ and nitrogen fertilization.
+The Community Land Model, version 5, contains a suite of modifications to carbon and nitrogen cycle representation, and globally, has a lower nitrogen response and higher CO$_{2}$ response than it's predecessors.
+To improve our understanding of the controls over the fertilization responses of the new model, we assess the sensitivity of the model to a suite of thirteen parameters pertinent to the cycling of carbon and nitrogen by vegetation.
+We assess the parameter dependence of the model state, and also the responses of the system to CO$_{2}$ and nitrogen fertilization at individual sites in temperate, tropical and boreal systems.
 
@@
 Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}).
-In the CMIP5 multi-model inter-comparison, only one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth.
+In the CMIP5 multi-model inter-comparison, one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth.
 In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}).
-Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent global N dynamics using alternative structural assumptions have been developed, with variable severity of N limitation in future high $CO_{2}$ scenarios (\cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}).
-In the CMIP6 inter-comparison process, many models are expected to contain active Nitrogen cycling.
-
-Unlike, for example, representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not underscored by well-tested and widely accepted paradigms, and substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehle2011} \cite{zaehle2014}).
-Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{kauwe2013}, \cite{zaehle2014}, \cite{medlyn2015}).
-Few studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model (c.f.
-\cite{meyerholt2018}).
-Given the observed large impacts of the inclusion of such processes on responses to increasing CO$_{2}$ and N deposition, understanding the controls on the magnitude of biogeochemical feedbacks both within and between models is an important part of interpretation of their results.
 
@@
 
-Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global biogeochemical feedback analysis, including the controls on global carbon fertilization `$\beta$' values  (the change in ecosystem carbon storage in response to CO$^{2}$ fertilization) (\cite{arora2013}, \cite{friedlingstein2014}),
-we consider it a worthwhile undertaking to illustrate how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations.
+Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global biogeochemical feedback analysis, including the controls on global carbon fertilization `$\beta$' values (the change in ecosystem carbon storage in response to CO$^{2}$ fertilization) (\cite{arora2013}, \cite{friedlingstein2014}), we consider the illustration of how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations to be an important feature of the documentation of the new model.
 
@@
 
-1. The impact of parameters is complex, and their relative importance varies across site conditions, output variables, and model forcing.
+1. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control model state.
 
-2. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control model state.
+2. There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization.
 
@@
 
-\subsection{Updates to Nitrogen Cycling in the CLM5}
-The CLM5 nitrogen cycle builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).
-The new model integrates three major additional prognostic elements of the nitrogen cycle, including:\\
+\subsection{Updates to biogeochemistry Cycling in the CLM5}
+The CLM5 biogeochemistry scheme builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).
 
-1) The `LUNA' module (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}) \\
+ A full technical description of the CLM5 is beyond the scope of this paper and is available as an appendix to \cite{lawrence2018} and also online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}.
 
-2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics and carbon economics of N acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}), and\\
+In this paper, we focus on the response to CO$_{2}$ and nitrogen deposition, which are primarily mediated by the biogeochemical processes within the model.
+The primary changes to the biogeochemical cycling in CLM5 are within the vegetation nitrogen cycle, wherein the model contains three new prognostic elements that are likely to impact responses to fertilization.
 
-3) The `FLEXCN' N cycle implementation, from \cite{ghimire2016}, which is used in CLM5 primarily to allow variation in tissue C:N ratios.\\
-
-More details on each of these new components is given below.
-The full technical description of the CLM5 is an appendix to \cite{lawrence2018} and is also available online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html).}
-
-\subsubsection{LUNA}
-Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
-There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
-Photosynthetic capacity is a key model parameter (\cite{rogers2017}), and the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
-
-Given the current leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components, for the prevailing time-averaged environmental drivers.
-The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes.
-In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area $N_{area}$, (which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and prevailing meteorology.
+Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}).
+Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}), which appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
+There is significant evidence that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
+LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for a given leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.
+$V_{c,max}$ is thus primarily controlled by the N per unit leaf area $N_{area}$ and the prevailing meteorology.
 The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was demonstrated by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
 
-\subsubsection{FUN}
-Under circumstances where insufficient N uptake was available to match all of the carbon assimilated (for a given target C:N ratio), the CLM4 model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake.
+Secondly, the `FUN' (Fixation and Uptake of Nitrogen) model was incorporated into CLM5.
+FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun} and \cite{brzostek2014} and \cite{shi2016}).
+A major issue confronting N cycle models is how to manage 'excess' carbon under conditions where carbon assimilation is faster than that which can be supported by nutrient uptake (\cite{dekauwe2014}).
+Previous versions of CLM model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake.
 This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).
-
-A major issue confronting N cycle models remains how to deal with this 'excess' carbon under N limiting conditions (\cite{dekauwe2014}).
-The FUN model (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}) addresses this issue by imposing the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, either directly from soil to roots or from mycorrhizal fungi to roots, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}).
-The cost of each uptake pathway is variable in time and space as a function of nutrient availability and temperature (in the case of the symbiotic N fixation pathway).
-FUN hypothesizes that plants will spend excess carbon on nitrogen uptake,  preferentially from pathways which are `cheaper' to them.
+The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}).
+Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them.
 Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N.
 Thus, N limitation leads to both reduced growth and higher expenditure of C on N uptake.
-The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for nitrogen uptake, ($C_{nuptake}$).
+The switching of N uptake to cheaper pathways results in the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}), whereas CLM4 and CLM4.5 predicted N fixation as a linear function of net primary productivity (\cite{cleveland1999}; \cite{wieder2015}).
+The FUN model is documented within the CLM technical note at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}.
 
-\begin{equation}
-C_{avail}=C_{growth}+C_{nuptake}
-\end{equation}
-
-Further, the nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant C:N ratio ($CN_{plant}$):
-
-\begin{equation}
-N_{uptake}=C_{nuptake}/N_{cost}
-=\frac{C_{growth}}{(CN_{plant}/(1.0+f_{gr})}
-\end{equation}
-
-The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$, which applies only to tissues that are constructed).
-Rearrangement of these constraints gives a predictive term for the amount of carbon that is expended on N uptake, as:
-
-\begin{equation}
-C_{nuptake} =\frac{C_{avail}}{ ((1.0+f_{gr})*(CN_{plant} / N_{cost}) + 1) }
-\end{equation}
-
-The average nitrogen cost ($N_{cost}$) is derived from the uptake of N, weighted for its cost across all uptake streams, while the plant CN ratio ($CN_{plant}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools.
-The FUN model is documented at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}
-
-One mechanism introduced by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}).
-The CLM4 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{cleveland1999}), and \cite{wieder2015} illustrated the impacts of uncertainty in this function.
-FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}).
-Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation,  typically when soil N availability is very low.
-
-Note that while the CLM5 has no representation of biological succession (\cite{fisher2015}, \cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state (carbon and nitrogen pools in equilibrium) from bare ground.
-In most cases, N fixation is preferred when soil N is low, and so soil N builds up quickly.
-In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.
-
-The FUN model introduces the concept of carbon payments for N uptake, which are intended to implicitly represent the carbon expended on root exudates, symbiotic relationships with soil fungi, enhanced expenditure on dynamic fine root growth or increased metabolic rates to support active uptake.
-A more comprehensive description of these processes might include explicit representations of these processes, but the model deployed here uses a bulk approach to account for the costs to plants of low environmental N availability.
-Therefore, uncertainties associated with the cost functions, in particular, are large.
-
-\subsubsection{FLEXCN}
-
+Thirdly, CLM5 contains the capacity to modify tissue C:N ratios with spatial and temporal variation in the cost of N uptake.
 Plant trait databases (e.g.
-\cite{kattge2011}) typically indicate a wide variability in tissue N concentrations within a single plant functional type, and this is in part driven by climate (\cite{wright2005}).
-To address the impact of environmental conditions on tissue N content,
-\cite{ghimire2016} implemented a suite of modifications to the CLM4.5 Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm.
-In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents nitrogen uptake, thus the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying N supply.
-
-In the new allocation algorithm, the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).
-The carbon allocation scheme uses a fixed-fraction approach to allocation in general, with constant ratios between, leaf, fine roots, stem and coarse root allocation.
-We note here that the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an feature which must be considered when analyzing simulations under high productivity conditions.
-Newer codes within the CLM framework (e.g.
-the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue.
-
-FLEXCN allows a second set of degrees of freedom, in addition to those provided by LUNA, to the problem of maintaining stoichiometric relationships during photosynthesis and allocation in the presence of N limitation.
-A distinction between the FUN and FLEXCN approaches is that, whereas the FUN logic links the carbon to nitrogen cycles synchronously within a given timestep and at the physiological point between GPP and NPP, FLEXCN links the two cycles asynchronously and at the physiological stage of GPP directly.
-The relative roles of these two different approaches to carbon-nitrogen coupling can be determined parametrically, as discussed below.
-
-\subsubsection{Adaptation of FUN to flexible CN ratios}
-The FUN model was originally conceived with a static tissue C:N ratio, and will by default allocate an amount of carbon to N uptake that exactly tracks the target C:N ratio.
-In CLM5, one of the aims of the development process was to incorporate a varying C:N ratio, as above.
-To that end, it is necessary to modify the amount of C expended on N uptake (and thus the amount of N obtained) with the prevailing cost of N.
+\cite{kattge2011}) typically indicate a wide variability in tissue N concentrations within a single plant functional type (\cite{wright2005}).
+\cite{ghimire2016} introduced the capacity for variable tissue C:N ratios, whereby the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).
+The FUN model used in CLM5 was originally conceived in the context of a static tissue C:N ratio.
+Carbon is allocated between growth and N uptake in a manner that exactly tracks the target C:N ratio.
+To allow a variable C:N ratio in the context of FUN, it is necessary to modify how much C is expended on N uptake vs on growth from this default calculation.
 The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include semi-heuristic and/or bounded representations of tissue N adjustments under N limited conditions (\cite{zaehle2010},\cite{ghimire2016}).
-Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content.
-When N is expensive in the environment, less C is spent on uptake, increasing C:N ratios where N is scarce.
-When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high.
-This new algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.
+Here we include a placeholder algorithm that modifies C:N ratios with 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content.
+When N is scarce in the environment, less C is spent on uptake, increasing C:N ratios.
+When plants become N limited and C:N ratios increase, however, expenditure increases.
+This algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.
 
 \subsection{Sites}
-We utilized four observational flux tower sites to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, 44.5$^{o}$N -121.6$^{o}$E).
+We utilized four observational flux tower sites, in order to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, 44.5$^{o}$N -121.6$^{o}$E).
 Each site was driven with the available site-specific meteorological data.
@@
 Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization.
-These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not emnurate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in it's own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.
+These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in it's own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.
 We focus here on biogeochemically important inputs.
@@
 
-\textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf area.
+\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf biomass.
 In the CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow (\cite{kennedy2019}) and the cost of N acquisition in the FUN model, respectively.
@@
 Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.
+The carbon allocation  in CLM5 uses a fixed-fraction approach to allocation in general, with constant ratios between, leaf, fine roots, stem and coarse root allocation.
+We note here that the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an feature which must be considered when analyzing simulations under high productivity conditions.
+Newer codes within the CLM framework (e.g.
+the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue by removing leaves in negative carbon balance.
 
- \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody tissues, which in CLM5 is again a constant fraction of leaf allocation.
+ \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody tissues, which in CLM5 is again a constant fraction of leaf allocation.
 Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass.
@@
 \textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.
+FUN introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake.
+Therefore, uncertainties associated with the cost functions, in particular, are large.
 These six parameters were defined and calibrated by \cite{brzostek2014}.
@@
 GPP, NPP, LAI, leaf N content per unit area and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites.
-In each of these figures, the y-axis  corresponds to the output variable and The x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in table \ref{table_ranges}.
+In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}.
 Lines of differing colors illustrate responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case.
@@
 Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential (\cite{kennedy2019}).
-$B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$).
+$B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$.
 Given that leaf water potential and  $B_{tran}$ vary diurnally, we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure).
@@
 The Metolius site had some drought stress ($B_{tran,mn}$ $\sim$86\%) and very little nutrient stress.
-The differences in the initial conditions at the sites have implications for the parametric responses, since some parameters, as limitations in one resource can affect the ability of the system to respond to fertilization and/or parameter modifications affecting other resources.
+Differences in the initial conditions at the sites have implications for the parametric responses.
+Underlying limitations by various boundary conditions (water, nutrients, light, temperature) can affect both the ability of the system to respond to fertilization and/or parameter modifications affecting other resources, as we describe further in the following sections.
 

### 2019-06-24 referee_responses.tex
Updates from Overleaf

@@
 \textsf{AR: We actually ran the model for those sites, but got very low LAI predictions, and therefore very sensitive responses to fertilization.
-We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO$_{2}$ and N forcing.
-}
+We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO$_{2}$ and N forcing.}
 

### 2019-06-24 referee_responses.tex
Updates from Overleaf

@@
 \textsf{An alternative approach would be to run the simulations with the various pieces of code turned off and returned to the default state.
-The problem with implementing that idea is that large changes of that type tended to hugely modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the system state was moved to a more or less light saturated condition.
-We have to some extent tried to probe the sensitivity space of say, the FUN model by modifying the capacity of the plants to take up and fix nitrogen.}
+The problem with implementing that idea is that large changes of that type tended to modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the initial system state was moved to conditions that were more or less light saturated, and more or less limited by water or nitrogen.
+}
 

### 2019-06-25 referee_responses.tex

@@
 The problem with implementing that idea is that large changes of that type tended to modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the initial system state was moved to conditions that were more or less light saturated, and more or less limited by water or nitrogen.
-}
+Further, with the exception of the LUNA code, technically it is quite difficult to turn off and on the FUN and FlexCN options without invoking other more poorly documented post CLm4.5 modifications, thus we have taken the approach of exploring the behaviour of the new models by modifying their major controlling parameters.}
 

### 2019-06-26 referee_responses.tex

@@
 \textsf{AR: Noted...}
-
 This paper documents three new assumptions (arguably four- section 2.1.4 as well) to CLM5 related to N; however, it doesn't give much consideration to these hypotheses, instead focussing on a broader list of impacts related to parameter choice.
@@
 
-\textsf{AR:  \textbf{needs work}}
+\textsf{AR: Fair point, but 1) you may be an atypical reviewer and 2) we still think this would distract from the main purpose of the paper, which is to document and explain the differential sensitivities of the parameter ensemble to CO2 and N fertilization.
+If we added in specific site data, readers would almost certainly focus on the model fit to particular sites, which will no doubt be explored in great detail in the wider literature.}
 

### 2019-06-26 referee_responses.tex
Updates from Overleaf

@@
 \textsf{AR: The MS is part of a special issue, and so the paper will all become available at the same time (except the Kennedy paper which is already published).
-\textbf{We will reduce depdance on these references.} }
+We have reduced the number of citations for the kennedy paper down to two.
+The primary resource we cite is the CLM5 tech note, which is already online but will also become an appendix to the Lawrence et al.
+paper when these papers are all published simultaneously.
+This is our overall strategy to documenting the entirety of the model, and we wanted to avoid duplicating large amounts of the technical text in different places.}
 
@@
 I'd ask the authors to reflect on what they think the messages they'd like to portray are and see if they feel those messages are in the discussion.
+
+\textsf{AR: Needs reflection **}
 
@@
 
-- In general, when the authors talk about "model state" - both in the key points and the abstract - doesn't this require some time context?
+-In general, when the authors talk about "model state" - both in the key points and the abstract - doesn't this require some time context?
 Are we talking about the equilibrium state?
@@
 (1994) Plant, Cell \& Environment, 17: 1081-1099, etc.
-\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications}
+
+\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications **}
 
@@
 
-\textsf{AR: OK, we will move this to the methods. }
+\textsf{AR: OK, we added this part to the methods section. }
 
@@
 \textsf{AR: We have deleted this hypothesis and replaced with "2.
-There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization."}
+There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization.
+** reconsider hypotheses"}
 
@@
 Anyway, irrelevant I'm sure.
-\textsf{AR: I am inclined to agree, but there's not much that can be done for this paper, at this point.}
+
+\textsf{AR: I am inclined to agree, but there's not much that can be done at this point.}
 
@@
 
-\textsf{AR: OK, \textbf{we will add more text on the emergence of the costs.}}
+\textsf{AR: We added a new sentence to describe the dependance of costs on the environment: \emph{The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation.
+Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them.
+Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N.
+}}
 
@@
 
-\textsf{AR: Add description of FATES trimming}
+\textsf{AR: Changed this to: \emph{Newer codes within the CLM framework (e.g.
+the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue by removing leaves in negative carbon balance.}}
 
@@
 
-\textsf{AR:swop text round}
+\textsf{AR: We have deleted these sections, and reorganized the grouping of the more generic text that replaces them.}
 
@@
 
-\textsf{AR: It is in review for the special issue at JAMES, and a copy should somehow be in the system to be accessed, although I don't know how referees get access to the other companion papers exactly}
+\textsf{AR: It is in review for the special issue at JAMES to which this paper is also submitted, and a copy should somehow be in the system to be accessed, although I don't know how referees get access to the other companion papers exactly}
 
@@
 
-\textsf{AR: Add description of what wieder did.}
+\textsf{AR: Modified the Wieder reference to read: \emph{ollowing \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.}}
 
 - Line 311 - it's 
+\textsf{AR: done}
 
@@
 
-AR: \textsf{Maybe remove.}
+AR: \textsf{We have reduced the amount of text devoted to this idea.}
 
@@
 
-\textsf{AR: Think about this.}
+\textsf{AR: Think about this. **}
 
@@
 
-\textsf{AR: One is red and the other is a sort of purple. Will modify the colormap to increase the purpleness.}
+\textsf{AR: One is red and the other is a purple.  We have modified the color map to distinguish better between these two colours.}
 
@@
 
-\textsf{AR: We will remove this refernce to the Lawrence paper.}
+\textsf{AR: The point we're trying to make here is that we find that there is no saturation with increasing leafN content, which is consistent with an analysis in the Lawrence paper that finds low Vcmax compared to observations.
+We have changed the text to try and make it clearer that this is a point of consensus not contradiction.}
 
@@
 I also think if you dropped some of these irrelevant params, then the plots (fig 6-8) would be easier to interpret.
+
+\textsf{AR: This point is well taken.
+As we discussed in response to a previous point, we decided to modify the focus of the manuscript to instead illustrate the parameters that are important for the C and N fertilization responses.
+**)
 
@@
 
-\textsf{AR: OK, this does need clarifying. Actually, CLM5 can't be run without its N cycle. I agree the responses are high. They are in contrast to those found in the global analyses. }
+\textsf{AR: OK, this does need clarifying.
+Actually, CLM5 can't be run without its N cycle.
+We have found in subsequent analyses that the increase in LAI with either CO2 or n can have large impacts on the response ratios, with the instantaneous response differing substantially from the longer term response on account of increases in leaf area.
+At Caxiuana there is a 50 \% increase in LAI on account of the CO2 fertilization, which is on top of the 33\% increase in GPP in the first year, for example.
+In the new version of the paper, we focus more closely on the time dependance and LAI interactions of the results.
+**}
 
 - Line 701: Or that global maps are a poor way to see how comparable models and observations are ... 
-\textsf{AR: Or that, yes. }
+\textsf{AR: Or that, yes. We don't really have enough here to say that the global maps are off though, so I'm going to desist from stating that.}
 
@@
 
-\textsf{AR: We are unsure how to respond to this statement. }
+\textsf{AR: The referee raises a number of different points here.
+The first point to note is that, as we discussed in L743 onwards, we did not in fact calibrate the model either at this site, nor using a global inverse calibration method, on account of the numerous difficulties encountered in that exercise.
+We are pleased that the referee considers these points to be in scope and have moved the discussion of the inverse calibration exercise to the methods section (within the discussion of the perturbed parameter ensemble)}
 
@@
 
-\textsf{AR: We do not suggest that this particular message is new.
-Our model is a new version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project.
+\textsf{AR: The referee suggests that the section headings are poorly balanced, and as such we have modified the organization so that section 3.1 deals with the system state with 3.1.1.
+for site differences and 3.1.2 for parametric differences.
+Section 3.2 deals with the fertilization responses, with 3.2.1 for overall responses and 3.2.2 for parametric controls.
+
+\textsf{AR: The referee further questions whether this message is new, and also suggests that it is important to identify how model behaviours are influenced by parameter decisions, which is the objective of this paper.
+We do not suggest that this particular message is new.
+Our model is a very heavily modified version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project - all using the default configuration.
 In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.}
-
-\textsf{AR: We admit that we had not given much thought to balancing the length of the sections under their various subheadings. \textbf{Will reorganize}}
 
@@
 
-\textsf{AR: \textbf{This is quite vague. Needs thought. }}
+\textsf{AR: We are not sure how it follows that the model output space being complex should lead to a list of strengths and weaknesses of the model.
+Indeed, assessing the validity of the process descriptions of the model is something we specifically aim to not do.
+We agree that the abstract is perhaps too general, **}}
 
@@
 
-\textsf{AR: \textbf{Maybe we should just ditch the hypotheses? }}
+\textsf{AR: \textbf{Maybe we should just ditch the hypotheses?** }}
 
@@
 
-\textsf{AR: The paper differs substantially from all the paper in revision, in that it explicitly tests the parameter space of the model.
+\textsf{AR: The paper differs substantially from the papers in revision, in that it explicitly tests the parameter space of the model.
 The remainder of the papers assess the responses of the default configuration.
@@
 L 88, CO2, not CO2 
+\textsf{AR: Fixed.}
 
 L 112: You haven't define what a model state you mean, and therefore, this hypothesis is not explicitly testable. 
+\textsf{AR: **}
 
 L 113 - 117: This is a very long sentence. And it's not easy to follow. Suggest rephrasing.
-
-\textsf{AR:\textbf{ We will break it up as needed. }}
+\textsf{AR: Changed this to \emph{ The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics.
+A focus on model fit at individual sites would preclude detailed consideration of how parameter variation impacts fertilization responses, as presented here}}
 
@@
 
-\textsf{AR: We have chosen to document the model both in the Lawrence et al.
-paper and in the technical documentatin online for CLM.
-The Lawrence paper is within the same AGU special issue on the CLM5, and we have already obtained editorial support for this strategy.
-Further, the full description of the FUN model and the FlexCN algorithm runs to many pages, which would ad substantially to this already quite long manuscript.
-An alternative would be to add these descriptions to the supplementery infomation, but this would not necessarily provide a greater utility to the reader than the two sources they are already available and would also constitute publishing the same material twice in different peer reviewed manuscripts.
-We have nonetheless reframed the manuscript so that it is more generally concerned with the CO2 and N responses, and not with the N updates per-se, so this is less necessary than before.
-}
+\textsf{AR: We have chosen to document the model within the Lawrence et al.
+paper and in the technical documentation online.
+The Lawrence paper is within the same AGU special issue on the CLM5 to which we have submitted this manuscript, and we have previously obtained editorial support for this strategy.
+Indeed, the need for these papers to cross-reference one-another is the main purpose of having them in a special issue.
+Further, the full description of just the new nutrient algorithm runs to many pages, which would add substantially to this already quite long manuscript, and in principle, many other parts of the model are relevant to this analysis.
+If we were to include these descriptions in a way that would allow a complete understanding of the technical details, they would no doubt have to reside as a supplement to this paper, and so,  this would not necessarily provide a greater utility to the reader than the two sources where are already available, Lastly, this would also constitute publishing the same material twice in different peer reviewed manuscripts.
+We have nonetheless re-framed the manuscript so that it is more focused on the CO$_{2}$ and N responses, and not on the nutrient cycling updates, so we feel that this is less necessary than before.}
 
@@
 
-\textsf{AR: Our objective here is not simply to illustrate that the parameters have an impact, but to provide an illustrate of how the different parameters impact the model and under what conditions, (as per hypothesis 1).
-Thus, there is value in doing the simulations subsequent to our narrowing of the parameter set to those elements with a direct impact on fertilization responses.
+\textsf{AR: Our objective here is not simply to illustrate that the parameters do have an impact (we agree that this would be self-evident) but to provide an illustrate of \emph{how} the different parameters impact the model and under what conditions.
+Thus, there is value in doing the simulations subsequent to our narrowing of the parameter set to those elements we considered to have a direct impact on fertilization responses.
 }
@@
 
-\textsf{AR: Thanks for finding this inconsistancy. It is fine root mass to leaf mass. We will correct accordingly}.
+\textsf{AR: Thanks for finding this inconsistency. It is fine root mass to leaf mass. We will correct accordingly}.
 
@@
 L 447: remove ). 
-\textsf{AR:Done}
+\textsf{AR: Done}
 
@@
 This is clearly the key novelty of this research, but I am afraid that I don't see them identified, and discussed enough throughout the text.
-\textsf{AR: \textbf{Make this clearer}}
+\textsf{AR: We have moved the key points here into the abstract, which highlights these issues earlier and makes the abstract less generic, as per an earlier point.}
 
@@
 
-\textsf{AR:\textbf{this needs some work} }
+\textsf{AR: ** }
 

### 2019-06-26 main_revision.tex

@@
 
+ The impact of fertilization is dependant both on parametric variation explored here varies with the balance of limiting factors (water, temperature, nutrients, light) in the pre-fertilization model state.
+The model parameters that impact the pre-fertilization state are in general not the same as those that control fertilization responses, meaning that calibration to present-day conditions will not necessarily constrain future projections.
+Where pre-fertilization state has low leaf area, fertilization induced increases in leaf production amplify the model response to the initial fertilization.
 Model responses to CO$_{2}$ and N fertilization are strongly impacted by how much plant communities can increase their fixation of nitrogen.
-Model initial state, stoichiometric flexibility of plant tissues, and nitrogen uptake costs all have notable impacts, but in general the output space is complex, and varies between sites, output variables, and fertilization scenarios.
-These responses can typically be understood in the context of the model structure and the functionality of individual parameters.
-Identification of the primary model features driving ecosystem responses to CO$_{2}$  and N fertilization is a necessary step in the interpretation of global carbon cycle feedback simulations, and in the identification of priorities for fieldwork and process updates.
 Further studies focused on how alternative parameters impact biogeochemical function would provide useful compliments to the ongoing CMIP6 process.
@@
 In the CMIP5 multi-model inter-comparison, one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth.
-In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}).
+In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}), leading to speculation that the addition of nitrogen limitation as a more general feature of models might reduce the capacity for vegetation to respond to CO2 fertilization and hence reduce the degree to which the terrestrial biosphere can mitigate increasing atmospheric CO$_{2}$.
 
-The Community Land Model is the land surface representation of the Community Earth System Model (CESM, \cite{hurrell2013}).
-Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}.
-An overview of model developments is presented by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019} and plant hydraulics by \cite{kennedy2019}.
-Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.
-
-The CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4 and CLM4.5 (\cite{lawrence2011}).
-The introduction of several new dynamical features into the CLM5 has implications for responses to environmental forcing.
-\cite{wieder2019} report the global-scale carbon cycle implications of the full model update and note that, in the default CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors, with both fertilization responses being closer to experimental results than the previous versions of the model.
+The newly-released CLM version 5  (CLM5,\cite{lawrence2018}) includes numerous major updates relative to CLM4 and CLM4.5 (\cite{lawrence2011}).
+The introduction of several new dynamical features, including substantial modification of the vegetation nitrogen cycling, as well as the introduction of plant hydrodynamics, altered surface hydrology and dynamic crop representation, has implications for responses to environmental forcing.
+\cite{wieder2019} report the global carbon cycle implications of these changes and note that, in CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors, with both fertilization responses being closer to experimental results than the previous versions of the model.
 
@@
 
-1. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control model state.
+1. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control the pre-fertilization model state.
 
-2. There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization.
+2. There is a trade-off between model parameterizations that allow strong CO$_{2}$ fertilization and those that allow strong N fertilization.
 
 The analysis herein does not focus on the relative fit of the model to the data collected at the individual sites.
-The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, a focus on that would preclude detailed analysis of the parameter space, as presented here.
+The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics.
+A focus on model fit at individual sites would preclude detailed consideration of how parameter variation impacts fertilization responses, as presented here.
 Assessment of the skill of the CLM5 model against a broad suite of globally-relevant data products and manipulation experiments is the focus of \cite{lawrence2018} and \cite{wieder2019}.
@@
 \section{Methods}
+
+\subsection{CLM5}
+ Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}.
+An overview of model developments is presented within this special issue by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019}/.
+The plant hydraulics model is documented by \cite{kennedy2019}.
+Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.
 
@@
 Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}).
-Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}), which appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).
+Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
+This method, however, appears to be a poor means of capturing variation in photosynthetic capacity in time and space and responses to climatic change (\cite{walker2017}).
 There is significant evidence that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
@@
 This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).
-The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}).
+The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation.
 Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them.
@@
 For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation.
-Following \cite{wieder2019}, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.
-Similar to \cite{wieder2019}, we assessed the impact of changes after 15-19 years of fertilization in both cases.
+Following \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.
 
@@
 Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization.
-These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in it's own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.
+These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in its own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.
 We focus here on biogeochemically important inputs.
-Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme will be discussed by \cite{kennedy2019}.
+Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme are discussed in \cite{kennedy2019}.
 
@@
 At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below.
-We recognize that a global parameter sensitivity analysis would also be of interest, particularly to identify interacting effects, and indeed have conducted many such analyses.
-We concluded that a OAAT analysis provided clearer insight into the functionality of the parameters with respect to biogeochemical forcing, and further that the more complex visualization and analysis of a global sensitivity study, across the entire domain of results presented here, would reduce the opportunity for understanding the individual parameter responses in depth, as is the goal of our study.
+We recognize that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses.
+We concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing.
+
+\subsection{Default parameter determination}
+During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
+We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystem to die under lower pre-industrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting effect that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
+The parameterization used in the default version of the model consequentially utilized the mean trait values where possible from trait datasets (\cite{lawrence2018}) in combination with targeted manual tuning of parameters (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states.
+We recognize that this more subjective calibration method is less than ideal, and ongoing studies will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}.
 
@@
 Their impact on model functionality is described below.
-As before, further details can be found in the appendix to \cite{lawrence2018}.
 The shorthand version of the parameter name is given in italics.
@@
 High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies and greater interception of light.
-In CLM5 high SLA leaves by definition also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.
+In CLM5 high SLA leaves also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.
 
 \textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf biomass.
-In the CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow (\cite{kennedy2019}) and the cost of N acquisition in the FUN model, respectively.
-The impact of this parameter is thus of greater interest than its functionality as a simple 'tax' on plant growth in previous versions of the CLM.
+In CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow, and the cost of N acquisition in the FUN model, respectively.
 Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.
-The carbon allocation  in CLM5 uses a fixed-fraction approach to allocation in general, with constant ratios between, leaf, fine roots, stem and coarse root allocation.
-We note here that the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an feature which must be considered when analyzing simulations under high productivity conditions.
+The carbon allocation in CLM5 uses a fixed-fraction approach, with constant ratios between, leaf, fine roots, stem and coarse root allocation.
+As such, the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance.
 Newer codes within the CLM framework (e.g.
-the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue by removing leaves in negative carbon balance.
+the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}m \cite{fisher2018vegetation}) account for this issue by removing leaves in negative carbon balance.
 
- \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody tissues, which in CLM5 is again a constant fraction of leaf allocation.
+ \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the fraction of biomass to stem tissues as a fraction of the leaf allocation.
 Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass.
-Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter, a feature which will be resolved in future model versions that allow plants to compete for light resources (\cite{fisher2018}).
+Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter).
 
 \textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.
-FUN introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake.
+The FUN model introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake.
 Therefore, uncertainties associated with the cost functions, in particular, are large.
@@
 Our analysis used four sites, with thirteen parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.
-Including the default simulations, the total comprises 672 different run configurations.
+Including the default simulations, the total comprises 636 different run configurations.
 We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$).
@@
 The four sites varied in the degree to which they are limited by water and N availability in the control state, with important impacts upon the parametric response characteristics (Figure \ref{btran state}).
-Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential (\cite{kennedy2019}).
+Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential.
 $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$.
@@
 
-\subsection{Parametric control over system state}
+\subsubsection{Parametric control over system state}
 At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected since it is used directly in the calculation of both of these quantities.
@@
 We do not find much evidence of saturation in the GPP vs.
-\emph{leafcn} relationships, in that GPP response does not level off at high leaf N content, in line with findings in \cite{lawrence2018} that V${_c,max}$ numbers in CLM5 might be low compared to observations.
+\emph{leafcn} relationships, in that GPP response does not level off at high leaf N content.
+This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.
 For NPP,  decreases in respiration rates caused by high \emph{leafcn}, and reduced carbon expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and respiration rates as a result.
@@
 At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.
-Note that \cite{kennedy2019} hydrodynamic functioning of the model in greater depth and that  studies currently underway will discuss surface energy and plant.
 
@@
 
-\subsection{Differences across sites}
+\subsubsection{Differences across sites}
 The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.
@@
 
-\subsection{Impact of parameters}
+\subsubsection{Impact of parameters}
 Across the four sites, parametric impacts on fertilization responses were very varied, when compared to the relatively consistent impact of parameters on the initial system state.
@@
 
-\subsubsection{Low leaf area investigations}
+\subsection{Low leaf area investigations}
 In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
@@
 
-The model output space is notable complex, in line with our hypothesis 1.
-In many cases the sign of the impact of the different parameters changes between sites, and trade-offs in the influence of carbon and nitrogen economic parameters interact with both the initial state and environmental forcings.
-That the parameter space of CLM5 is complex is not surprising, given its relatively high process fidelity, but a) this type of sensitivity analysis,  illustrating how different parameterizations respond to different growth conditions and environmental forcings, is relatively rare for models of this kind,  b) the large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and c) the results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites.
+The model output space is notable complex, and in many cases the sign of the impact of the different parameters changes between sites, and trade-offs in the influence of carbon and nitrogen economic parameters interact with both the initial state and environmental forcing.
+That the parameter space of CLM5 is complex is not surprising, given its relatively high process fidelity, but a) this type of sensitivity analysis,  illustrating how different parameterizations respond to different growth conditions and environmental forcing, is relatively rare for models of this kind,  b) the large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and c) the results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites.
 Sensitivity tests across broader geographical ranges would help to eludicate more clearly how changes in baseline water and N limitation impact parameter responses and patterns of responses to fertilization.
@@
 However, at 5 years into the fertilization in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent (a x~2.06 impact on NPP, compared to ~0.95 in the year 15 analysis at Niwot Ridge, indicating that the results described above had already been subject to progressive N limitation, and were closer to an equilibrated state.
-
-\subsubsection{Objective calibration of the default model, or absence thereof}
-During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
-We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystem to die under lower preindustrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting effect that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
-The parameterization used in the default version of the model consequentially utilized the mean trait values where possible from trait datasets (\cite{lawrence2018}) in combination with targeted manual tuning of parameters (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states.
-
-We recognize that this more subjective calibration method is less than ideal, and ongoing studies will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}.
 

### 2019-06-27 main_revision.tex

@@
 
-\textbf{Growth respiration)}, \emph{`grperc'}, or the (fixed) fraction of carbon spent on respiration per unit of new tissue, was revised substantially in CLM5 to reflect new estimates that account for methodological issues with the observation of growth respiration (Atkin O., pers comm.) from 0.3 to 0.11.
-CLM5 further adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significantly increases maintenance respiration costs.
-Here we test the impacts of this change on overall carbon budgets.
-
 \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}.
@@
 
-\textbf{Leaf maintenance respiration intercept}, \emph{lmr\_intercept}.
-The new leaf maintenance respiration model, proposed by \cite{atkin2015} and adopted by CLM5, has it's intercept (but not slope) as a PFT dependant parameter, with variation between PFTs reported by \cite{atkin2015} and used herein.
-Here we test the impacts of the uncertainty in the observations of this parameter.
+\textbf{C:N flexibility }, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c},  as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.
+Only the thid parameter (c) had an important impact in this sensitivity analysis, and so we focus here on the impact of that parameter.
 
-\textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is an ecosystem-level property defined by the FUN model which allows the proportions of the fungal symbiotic pathways to be modified.
-Default parameters are as provided by \cite{shi2016}.
-
-\textbf{Parameters a,b and c of the flexible C:N model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.
+Other variables we tested in earlier versions of the analysis included rate constant controlling the growth and maintenance respiration rates, the fraction of ectomycorrhizal fungi, and parameters 'a' and 'b' of the flexible C:N ratio model.
+None of these had important impacts on the fertilization responses, and growth and respiration rates affected the model state in a linear and predictable fashion, thus they are omitted here but can be accessed via the available simulations.
 
@@
 We here take the PFT average log-normalised standard deviation range from TRY and apply it across all PFTs (thus the lower bounds are closer to the mean than the higher bounds, Table \ref{table_ranges}).
-The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across their whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake).
-\emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM(4, 4.5) value (0.3).
-\emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}.
+The range of \emph{fracfixers} and was set to vary from 0 to 100\% of ecosystem productivity being available for fixation.
+\emph{medlyn\_slope} varies across the stndard errors reported in \cite{dekauwe2015}.
 Reasonable ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood.
-Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization (thus, these results should not be interpreted as covering a constrained range of variation).
+Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization.
 The same issues apply to $N_{costs}$ parameters.
@@
 \section{Results}
-Our analysis used four sites, with thirteen parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.
-Including the default simulations, the total comprises 636 different run configurations.
+Our analysis used four sites, with eight parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.
 We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$).
@@
 
-The variance in \emph{lmr\_intercept} (across the range of reported intercepts in \cite{atkin2015}) also has important consequences for ecosystem productivity, broadly in line first-order expectations.
-At Caxiuan\~a, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.
-Whether the full range of values reported by \cite{atkin2015} is an appropriate reflection of the range of uncertainty for all PFTs is debatable, and a narrower range might be discernible with finer interpretation of their data.
-
-\emph{frac\_ecto\_fungi}, the fraction of ectomycorhizzal vs arbuscular fungi had limited impact here, suggesting that the parameterization of the varying costs between these uptake pathways did not have a large impact on N uptake.
-This suggests that this difference is overridden by other parts of the dynamics parameterized in FUN.
-
-\emph{cn\_flex\_a} is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details).
-Thus, low values of \emph{cn\_flex\_a} result in reductions in expenditure on N uptake when N is too expensive, and therefore in lower tissue N content, and vice-versa.
-For the range of parameters and the sites investigated here, this parameter had limited impact on the system state.
-Increasing \emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, and also has limited impacts on the system state here.
  \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying.

### 2019-06-27 main_revision.tex
Updates from Overleaf

@@
 
-Other variables we tested in earlier versions of the analysis included rate constant controlling the growth and maintenance respiration rates, the fraction of ectomycorrhizal fungi, and parameters 'a' and 'b' of the flexible C:N ratio model.
-None of these had important impacts on the fertilization responses, and growth and respiration rates affected the model state in a linear and predictable fashion, thus they are omitted here but can be accessed via the available simulations.
+Other variables we tested in earlier versions of this analysis included rate constants controlling the growth and maintenance respiration rates, the fraction of ectomycorrhizal fungi, and parameters 'a' and 'b' of the flexible C:N ratio model.
+None of these had important impacts on the fertilization responses, and growth and respiration rates affected the model state in a linear and predictable fashion.
+They are omitted here for the sake of simplification, but data can be accessed via the available simulations.
 

### 2019-06-27 main_revision.tex

@@
 $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$.
-Given that leaf water potential and  $B_{tran}$ vary diurnally, we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure).
+Leaf water potential and  $B_{tran}$ vary diurnally, so we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure).
 For N limitation, the fraction of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake is a direct indicator of how N stress affects the carbon budget.
@@
 The Caxiuan\~a site experienced substantial drought stress in these simulations (with the mean $B_{tran,mn}$ of the default condition around $\sim$0.55), with little nutrient stress.
-The simulate forest at Niwot Ridge experienced substantial amounts of drought ($B_{tran,mn}$ $\sim$0.5), and N limitation, spending 27\% of net assimilated carbon on nutrient acquisition.
+The simulated forest at Niwot Ridge experienced substantial amounts of drought ($B_{tran,mn}$ $\sim$0.5), and N limitation, spending 27\% of net assimilated carbon on nutrient acquisition.
 Harvard Forest site had a moderate drought stress ($B_{tran,mn}$ $\sim$90\%), and moderate N limitation ($\sim$15\% of assimilation spent on uptake).
@@
 \subsubsection{Parametric control over system state}
-At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected since it is used directly in the calculation of both of these quantities.
+At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities.
 For GPP, the impact of SLA is always positive, despite the higher LAI causing lower N per unit area of leaf (which reduces per-unit leaf area photosynthetic capacity).
@@
 
-Responses to modifications in \emph{grperc}, (the growth respiration coefficient) were in line with first-order expectations, where higher values resulted in NPP reductions and corresponding changes in LAI, and GPP.
-The impact varied between sites, reflecting different partitioning of the carbon budget into maintenance respiration, nitrogen uptake and growth.
-
 Changes in \emph{medlyn\_slope}, the stomatal slope parameter, most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux.
@@
  \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying.
-For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.
+For the state impacts here, the role of this parameter was limited, but for the step-change fertilization it was much more important, so we will discuss it's role further in the section below.
 
 \subsection{CO$_{2}$ and Nitrogen deposition responses}
+
+\subsubsection{Timescale dependance of responses to fertilization}
+
+The initial changes in carbon economy induced by the fertilization experiments are, where conditions permit, amplified by increases in the leaf area index.
+Figure \ref{timescales} illustrates, for the default parameter set, the evolution through time of output variables (GPP, NPP, LAI, $B_{tran,mn}$, NPP spent on N uptake).
+These results illustrate the complexities of the response.
+
+For example, the initial stimulation of GPP by high CO$_{2}$ ($\sim$1.37) lead to a $\sim$50\% increase in LAI.
+This, however, lead to increase drought stress (lower $B_{tran,mn}$), and hence no large increase in GPP.
+At Metolious, in contrast, initial GPP stimulation was similar, the resulting growth in LAI was much slower, given the lower overall carbon turnover at this colder site.
+The 50\% increase in  carbon at 80 years had, however, a much larger impact on GPP, giving a stimulation of ** at 80 years.
+ At Harvard forest, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (as illustrated by the lower response to decreasing leaf C:N ratios in figure ** than other figures).
+At all sites, the fraction of NPP spent on N uptake increased, but most dramatically at Metolius, where initial N expenditure was lowest initially on account of low growth rates.
+
+As expected from the initial conditions, Niwot Ridge responded most strongly to nutrient additions.
+An initial NPP stimulation of 1.3 for NPP was magnified by subsequent increases in LAI to nearly 1.8 times the control value, causing a stimulation of 1.7 in NPP.
+None of the other sites exhibited such a dramatic change.
+In particular, Caxiuana, which was heavily water limited
+
 Figures \ref{GPP CO2 and N respones 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the responses to both N and to CO$_{2}$ fertilization simultaneously, for GPP, NPP and, LAI.

### 2019-06-27 main_revision.tex

@@
 None of the other sites exhibited such a dramatic change.
-In particular, Caxiuana, which was heavily water limited
+In particular, Caxiuana, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%.
+
+The time-dependence of the responses generates an interesting issue for interpretation, since where LAI increases a lot, the responses tend to be dominated by this feature of the response.
+The first year responses, by way of contrast, tend to be much more similar across sites, parti.
 

### 2019-06-28 main_revision.tex

@@
 Leaf water potential and  $B_{tran}$ vary diurnally, so we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure).
-For N limitation, the fraction of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake is a direct indicator of how N stress affects the carbon budget.
+For N limitation, the percentage of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake (PNU) is a direct indicator of how N stress affects the carbon budget.
 
-The Caxiuan\~a site experienced substantial drought stress in these simulations (with the mean $B_{tran,mn}$ of the default condition around $\sim$0.55), with little nutrient stress.
-The simulated forest at Niwot Ridge experienced substantial amounts of drought ($B_{tran,mn}$ $\sim$0.5), and N limitation, spending 27\% of net assimilated carbon on nutrient acquisition.
-Harvard Forest site had a moderate drought stress ($B_{tran,mn}$ $\sim$90\%), and moderate N limitation ($\sim$15\% of assimilation spent on uptake).
-The Metolius site had some drought stress ($B_{tran,mn}$ $\sim$86\%) and very little nutrient stress.
+The Caxiuan\~a site (CAX) experienced substantial drought stress and little nutrient stress ($B_{tran,mn}$ $\sim$0.6, PNU$\sim$13\%).
+Niwot Ridge (NWR) experienced substantial amounts of drought and N limitation, ($B_{tran,mn}$, $\sim$0.55, PNU$\sim$21\%).
+Harvard Forest (HVF) site had very little drought stress, and moderate N limitation ($B_{tran,mn}$ $\sim$0.96, PNU$\sim$15\%).
+Metolius (MET)  had some drought stress and the least nutrient stress ($B_{tran,mn}$ $\sim$0.86, PNU$\sim$10\%).
+Thus, the sites chosen span a range of initial limiting conditions.
+
 Differences in the initial conditions at the sites have implications for the parametric responses.
@@
 In particular, Caxiuana, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%.
+Stimulation at Metolious was higher than Caxiuana, despite it's lower initial N limitation, given it's lack of limitation by water.
 
 The time-dependence of the responses generates an interesting issue for interpretation, since where LAI increases a lot, the responses tend to be dominated by this feature of the response.
-The first year responses, by way of contrast, tend to be much more similar across sites, parti.
+The first year responses, by way of contrast, tend to be much more similar across sites.
 

### 2019-06-28 main_revision.tex
Updates from Overleaf

@@
 Thus, the sites chosen span a range of initial limiting conditions.  
-
 Differences in the initial conditions at the sites have implications for the parametric responses.
@@
 For GPP, the impact of SLA is always positive, despite the higher LAI causing lower N per unit area of leaf (which reduces per-unit leaf area photosynthetic capacity).
-The impact of SLA on NPP can, however, be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.
+The impact of SLA on NPP can, however, be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.
 At Niwot Ridge, the parameter range straddles an optimal SLA (or local maximum at the second-largest parameter values) for NPP, again reflecting the assimilation/respiration cost trade-off.
@@
 \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/btran_npp.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_MEANCOND_-r200_1999.png}
      \caption{Response of drought index ($B_{tran,mn}$) and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1_2012.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1999.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Niwot Ridge.}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1_2012.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1999.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Caxiuan\~a National Forest}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1_2012.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1999.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Harvard Forest}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/NOVc_STATE_1CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1_2012.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1999.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Metolius Forest}

### 2019-06-28 main_revision.tex

@@
 Metolius (MET)  had some drought stress and the least nutrient stress ($B_{tran,mn}$ $\sim$0.86, PNU$\sim$10\%).
+
 Thus, the sites chosen span a range of initial limiting conditions.  
@@
 At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities.
-For GPP, the impact of SLA is always positive, despite the higher LAI causing lower N per unit area of leaf (which reduces per-unit leaf area photosynthetic capacity).
-The impact of SLA on NPP can, however, be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.
-At Niwot Ridge, the parameter range straddles an optimal SLA (or local maximum at the second-largest parameter values) for NPP, again reflecting the assimilation/respiration cost trade-off.
+The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf.
+The impact of SLA on NPP can, however, sometimes be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.
 
-Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation through root-mediated N uptake costs.
-The first two impacts decrease NPP with \emph{froot\_leaf} and the latter two increase it.
+Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in tissue allocation fractions , 2) an indirect effect on NPP, via changes in maintenance costs of fine root pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation through root-mediated N uptake costs.
+The first two impacts decrease NPP with higher \emph{froot\_leaf} costs and the latter two increase it via \emph{froot\_leaf} benefits.
+CAX and NWR exhibit higher LAI with fewer roots, HVF exhibits lower LAI with fewer roots, and MET
+
 Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios, reflecting mechanisms 3 and 4.

### 2019-06-28 main_revision.tex

@@
 The first two impacts decrease NPP with higher \emph{froot\_leaf} costs and the latter two increase it via \emph{froot\_leaf} benefits.
-CAX and NWR exhibit higher LAI with fewer roots, HVF exhibits lower LAI with fewer roots, and MET
+CAX and HVF exhibit higher LAI with fewer roots.
+NWR has large reduction in LAI (and GPP & NPP) with fewer roots.
+MET exhibits a mixed response crossing an optimal value.
+\emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}).
 
-Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios, reflecting mechanisms 3 and 4.
-Harvard forest and Caxiuan\~a both show increases in all variables with low \emph{froot\_leaf}.
-In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at Metolius, in particular.
-
-Low values of \emph{stem\_leaf} ratio typically resulted in much higher LAI at all sites, as well as declines in vegetation carbon, as expected from the lower stem carbon allocation.
+Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon.
 At Niwot Ridge, there is an optimum (or local maximum) stem carbon observed, due to the impacts of a too-high leaf area index, which caused declines in NPP at low \emph{stem\_leaf}.

### 2019-07-01 main_revision.tex

@@
 
+\subsubsection{Responses to forcing}
+We calculated the impacts of the CO$_{2}$ and N deposition by assessing the ratio of the fertilized to control simulations, averaging years 15-19 after fertilization began.
+For some parameter values, the constraint on growth placed the control system in a state close to or at death (zero live carbon pools).
+Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass was already zero, leading to some large non-linearities in the relationship between parameters and environmental forcing.
+To focus on the responses of parameter combinations that give more reasonable initial vegetation conditions, parameter combinations that caused LAI to be less than 60\% of the control simulation value were excluded from the fertilization plots.
+
 \section{Results}
@@
 Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon.
-At Niwot Ridge, there is an optimum (or local maximum) stem carbon observed, due to the impacts of a too-high leaf area index, which caused declines in NPP at low \emph{stem\_leaf}.
+At NWR, there is an optimum (or local maximum) stem carbon observed, due to the impacts of a too-high leaf area index, which caused declines in NPP at low \emph{stem\_leaf}.
 At all other sites, the impact of increasing leaf allocation is to increase NPP (almost doubling at Metolius).
@@
 
-The first-order impact of increasing \emph{Ncosts} was to increase the respiration losses associated with nitrogen uptake.
-Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of \emph{Ncosts} on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low.
+The first-order impact of increasing \emph{N\_{costs} was to increase the respiration losses associated with nitrogen uptake.
+Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of \emph{N\_{costs} on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low.
 Other sites showed more muted responses in the vegetation state.
@@
 For all sites and variables here, it has a relatively moderate and positive impact on the control state, in particular at the Niwot and Harvard forest sites.
-As is the case for \emph{N\_costs}, with high fixing capacity, N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that LAI increase.
+As is the case for \emph{N\_{costs}, with high fixing capacity, N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that LAI increase.
 
@@
 
-\subsubsection{Timescale dependance of responses to fertilization}
-
+\subsubsection{Timescale dependence of responses to fertilization}
 The initial changes in carbon economy induced by the fertilization experiments are, where conditions permit, amplified by increases in the leaf area index.
@@
 The time-dependence of the responses generates an interesting issue for interpretation, since where LAI increases a lot, the responses tend to be dominated by this feature of the response.
-The first year responses, by way of contrast, tend to be much more similar across sites.
+The first year responses, by way of contrast, tend to be much more similar across sites, but do not reflect the overall response of the system.
+
+\subsubsection{Overall responses to fertilization}
+
+\subsubsection{Differences across sites}
 
@@
 
-We calculated the impacts of the CO$_{2}$ and N deposition by assessing the ratio of the fertilized to control simulations, averaging years 15-19 after fertilization began.
-For some parameter values, the constraint on growth placed the control system in a state close to or at death (zero live carbon pools).
-Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass was already zero, leading to some large non-linearities in the relationship between parameters and environmental forcing.
-To focus on the responses of parameter combinations that give more reasonable initial vegetation conditions, parameter combinations that caused LAI to be less than 60\% of the control simulation value were excluded from the fertilization plots.
-
-\subsubsection{Differences across sites}
 The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.
@@
 
-The \emph{N\_costs} parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
-At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
-These high \emph{N\_costs} scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.
-Low values of \emph{N\_costs} were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.
+The \emph{N\_{costs} parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
+At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_{costs} resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
+These high \emph{N\_{costs} scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.
+Low values of \emph{N\_{costs} were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.
 While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_US-NR1_ens_MIC_y1999.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_NWR_y2000.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Niwot Ridge.}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_Br-cax_ens_MIC_y1999.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_CAX_y2000.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Caxiuan\~a National Forest}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_US_Ha1_ens_MIC_y1999.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_HVF_y2000.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Harvard Forest}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_1-r200_CLM50defpft_trans_1x1pt_US-Me2_ens_MIC_y1999.png}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_MET_y2000.png}
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Metolius Forest}

### 2019-07-01 main_revision.tex

@@
 Our analysis used four sites, with eight parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.
-We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$).
+We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen per unit area ($N_{leaf}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$).
 We did not look at soil output variables as our parameter analysis is focused on vegetation responses.
@@
 The parameter responses of the site-level water and nitrogen limitation (see below) in the control (unfertilized) parameter ensemble are shown in figure \ref{btran state}.
-GPP, NPP, LAI, leaf N content per unit area and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites.
+GPP, NPP, LAI, $N_{leaf}$ and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites.
 In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}.
@@
 \subsubsection{Water and N limitation across sites}
-The four sites varied in the degree to which they are limited by water and N availability in the control state, with important impacts upon the parametric response characteristics (Figure \ref{btran state}).
-Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential.
+The four sites varied in the degree to which they are limited by water and N availability in the control state (Figure \ref{btran state}).
+Water limitation in CLM5 can be expressed via a model property termed $B_{tran}$, which is calculated as a function of the (prognostic) leaf water potential.
 $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$.
-Leaf water potential and  $B_{tran}$ vary diurnally, so we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure).
+Leaf water potential and $B_{tran}$ vary diurnally, so we use the average daily minimum ($B_{tran,mn}$) here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure).
 For N limitation, the percentage of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake (PNU) is a direct indicator of how N stress affects the carbon budget.
@@
 Thus, the sites chosen span a range of initial limiting conditions.
-Differences in the initial conditions at the sites have implications for the parametric responses.
-Underlying limitations by various boundary conditions (water, nutrients, light, temperature) can affect both the ability of the system to respond to fertilization and/or parameter modifications affecting other resources, as we describe further in the following sections.
+Underlying limitations by different boundary conditions (water, nutrients, light, temperature) affect the ability of the system to respond to fertilization and/or parameter modifications, as we describe further in the following sections.
 
 \subsubsection{Parametric control over system state}
-At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities.
+In this section, we discuss the impacts of the parameter variation on the initial model state.
+Many parameters have both first-order direct impacts, and then further impacts via other parts of the model.
+Here we attempt to illustrate how these mechanisms operate in CLM5 within the parameter space explored here.
+
+At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on $N_{leaf}$ (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities.
 The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf.
@@
 
-Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in tissue allocation fractions , 2) an indirect effect on NPP, via changes in maintenance costs of fine root pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation through root-mediated N uptake costs.
-The first two impacts decrease NPP with higher \emph{froot\_leaf} costs and the latter two increase it via \emph{froot\_leaf} benefits.
+The root:leaf ratio (\emph{froot\_leaf}) has numerous impacts, including 1) on LAI, via tissue allocation fractions , 2) on NPP, via changes in maintenance costs of fine root pools, 3) on the uptake of water (via root length density) and 4) on nutrient limitation through root-mediated N uptake costs.
+The first two impacts represent the costs of more roots, and the latter two their benefits.
 CAX and HVF exhibit higher LAI with fewer roots.
-NWR has large reduction in LAI (and GPP & NPP) with fewer roots.
+NWR has large reduction in LAI (and GPP \& NPP) with fewer roots.
 MET exhibits a mixed response crossing an optimal value.
-\emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}).
+\emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}) indicating the impact of mechanism '3' here, for very low root densities.
 
 Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon.
-At NWR, there is an optimum (or local maximum) stem carbon observed, due to the impacts of a too-high leaf area index, which caused declines in NPP at low \emph{stem\_leaf}.
-At all other sites, the impact of increasing leaf allocation is to increase NPP (almost doubling at Metolius).
-Decreases in stem:leaf ratio also caused a decline in leaf N per unit area, potentially on account of the higher demand for N uptake and thus higher N limitation.
-This effect is llustrated clearly for Niwot Ridge and Metolius in Figure \ref{btran state}).
+At NWR, the parameter range crosses a local maximum for vegetation carbon, since increases in NPP are non-linear with respect to leaf investment, resulting in lower vegetation biomass overall.
+At all other sites, the impact of increasing leaf allocation is to invoke large increases in LAI and NPP.
+Lower stem allocation and higher LAI increase the N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).
 
-The first-order impact of increasing \emph{N\_{costs} was to increase the respiration losses associated with nitrogen uptake.
-Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of \emph{N\_{costs} on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low.
-Other sites showed more muted responses in the vegetation state.
+Increasing $N_{costs}$ increases the respiration losses associated with nitrogen uptake.
+All sites showed a moderate decline in LAI, NPP, GPP and vegetation carbon when costs were high, but lesser impacts when costs were reduced (Figures \ref{NR1 state}, \ref{CAX state} \ref{HVF state}, \ref{MET state}).
+In contrast to the allocation parameters, $N_{costs}$ does not appear to be a first-order control over the pre-fertilization system state.
 
 Increasing \emph{fracfixers} affects the maximum amount of carbon that plants can pay towards symbiotic nitrogen uptake (N fixation).
-For all sites and variables here, it has a relatively moderate and positive impact on the control state, in particular at the Niwot and Harvard forest sites.
-As is the case for \emph{N\_{costs}, with high fixing capacity, N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that LAI increase.
+For all sites and variables here, it has a relatively moderate and positive impact on the control state.
+As is the case for $N_{costs}$, \emph{fracfixers} does not appear to be a first-order control over the pre-fertilization system state.
 
-Changing \emph{leafcn} -the target leaf C:N ratio- over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites.
-Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.
-In particular, the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).
-We do not find much evidence of saturation in the GPP vs.
-\emph{leafcn} relationships, in that GPP response does not level off at high leaf N content.
+Changing \emph{leafcn} over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites.
+Higher \emph{leafcn} values decrease $N_{leaf}$, decreasing photosynthetic capacity and GPP.
+In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero Figures \ref{HVF state}, \ref{MET state}).
+GPP continues to increase almost linearly with declining \emph{leafcn} (increasing $N_{leaf}$), indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.
 This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.
-For NPP,  decreases in respiration rates caused by high \emph{leafcn}, and reduced carbon expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and respiration rates as a result.
-At Niwot, an optimum \emph{leafcn} ratio (or a local maximum in the parameteric response) was apparent.
-For NPP this was around the default value, but for LAI it was 0.5 standard deviations lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.
+For NPP, respiration costs of modified $N_{leaf}$ and altered PNU (Figure \ref{btran state}) buffered the larger changes in GPP.
+At CAX, for example, the lowest \emph{leafcn} had a GPP 225 gC/m^{2}/y higher, while the corresponding NPP increase was only 52 gC/m^{2}/y.
 
-Changes in \emph{medlyn\_slope}, the stomatal slope parameter, most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux.
+Changes in \emph{medlyn\_slope}, the stomatal slope parameter, across the range shown here, have limited impacts on the pre-fertilized system state, with the exception of NWR, where lower values of the slope parameter generated higher photosynthesis, consistent with this site experiencing severe drought stress and thus benefiting from increased water efficiency.
+
+most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux.
 Water limitation is also modified as a result of altered transpiration.
@@
 
-The \emph{N\_{costs} parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
-At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_{costs} resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
-These high \emph{N\_{costs} scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.
-Low values of \emph{N\_{costs} were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.
+The $N_{costs}$ parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
+At Caxiuan\~a, Niwot and Harvard forest, high  $N_{costs}$ resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
+These high $N_{costs}$ scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.
+Low values of $N_{costs}$ were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.
 While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.
@@
 At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$.
-Leaf N concentration followed suit, increasing substantially in the N deposition experiment.
-Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when \emph{N\_costs} are high, therefore potentially suppressing leaf N content.
+$N_{leaf}$ followed suit, increasing substantially in the N deposition experiment.
+Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when N costs are high, therefore potentially suppressing $N_{leaf}$.
 Adding N overcomes this suppression, leading to greater fertilization responses.
@@
 Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 were primarily motivated by the need to move the model behaviour closer to several key dynamic features of the system.
-Notably, the dynamics of leaf nitrogen content with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}).
+Notably, the dynamics of $N_{leaf}$ with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}).
 The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.

### 2019-07-01 main_revision.tex

@@
 
-most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux.
-Water limitation is also modified as a result of altered transpiration.
-At Caxiuan\~a (Figure \ref{CAX state}), the very lowest values of  \emph{medlyn\_slope} caused a complete death of the ecosystem.
-In Figure \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses), presumably at the cost of limiting photosynthesis past the point at which positive plant growth could be maintained.
-At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough.
-At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.
-
 \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying.

### 2019-07-02 main_revision.tex
Updates from Overleaf

@@
 For NPP, respiration costs of modified $N_{leaf}$ and altered PNU (Figure \ref{btran state}) buffered the larger changes in GPP.
-At CAX, for example, the lowest \emph{leafcn} had a GPP 225 gC/m^{2}/y higher, while the corresponding NPP increase was only 52 gC/m^{2}/y.
+At CAX, for example, the lowest \emph{leafcn} had a GPP 225 gC/m$^{2}$/y higher, while the corresponding NPP increase was only 52 gC/$m^{2}$/y.
 
@@
 \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying.
-For the state impacts here, the role of this parameter was limited, but for the step-change fertilization it was much more important, so we will discuss it's role further in the section below.
+For the pre-fertilization state impacts here, the role of this parameter was limited, but for the step-change fertilization it was much more important, so we discuss its role further in the section below.
 

### 2019-07-02 referee_responses.tex

@@
 As we discussed in response to a previous point, we decided to modify the focus of the manuscript to instead illustrate the parameters that are important for the C and N fertilization responses.
-**)
+Further, in light of the need to simplify the discussion and analysis, we have removed 5 of the 13 parameters from the initial version.
+The reduced range of variation for the Medlyn slope parameter further reduced the impact of that parameter on the output, and therefore less text is expended on its discussion.)
 
@@
 Our model is a very heavily modified version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project - all using the default configuration.
-In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.}
+In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.
+
+We agree, however, that the results section in the first version of the manuscript was dense and difficult to follow.
+We have attempted to remedy this both by reducing the number of parameters considered from 13 to 8, and also by attempting to streamline the text reporting of the results to focus on the features of the output of highest importance for understanding model sensitivity.
+We hope this is an easier read than the prior version.
+}
 
@@
 L 112: You haven't define what a model state you mean, and therefore, this hypothesis is not explicitly testable. 
-\textsf{AR: **}
+\textsf{AR: Through the new version of the manuscript, we have used 'pre-fertilization state' to describe this condition. }
 

### 2019-07-02 main_revision.tex

@@
-\documentclass[draft,linenumbers]{agujournal}
+\usepackage{}\documentclass[draft,linenumbers]{agujournal}
 
@@
 We concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing.
-
 \subsection{Default parameter determination}
@@
 \subsubsection{Parameters selected for perturbation experiment}
-
 The parameters included in our sensitivity analysis are listed in Table \ref{table_ranges}.
@@
 Higher \emph{leafcn} values decrease $N_{leaf}$, decreasing photosynthetic capacity and GPP.
-In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero Figures \ref{HVF state}, \ref{MET state}).
-GPP continues to increase almost linearly with declining \emph{leafcn} (increasing $N_{leaf}$), indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.
+In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero (Figures \ref{HVF state}, \ref{MET state}).
+GPP continues to increase almost linearly with declining \emph{leafcn}  (increasing $N_{leaf}$) at all sites except HVF, indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.
 This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.
@@
 \subsubsection{Timescale dependence of responses to fertilization}
-The initial changes in carbon economy induced by the fertilization experiments are, where conditions permit, amplified by increases in the leaf area index.
-Figure \ref{timescales} illustrates, for the default parameter set, the evolution through time of output variables (GPP, NPP, LAI, $B_{tran,mn}$, NPP spent on N uptake).
-These results illustrate the complexities of the response.
+The initial changes in carbon economy induced by the fertilization experiments are, where conditions permit, amplified by increases in LAI.
+Figure \ref{timescales} illustrates, for the default parameter set, the evolution through time of GPP, NPP, LAI, $B_{tran,mn}$ and PNU, illustrating the complexities of the response.
 
-For example, the initial stimulation of GPP by high CO$_{2}$ ($\sim$1.37) lead to a $\sim$50\% increase in LAI.
-This, however, lead to increase drought stress (lower $B_{tran,mn}$), and hence no large increase in GPP.
-At Metolious, in contrast, initial GPP stimulation was similar, the resulting growth in LAI was much slower, given the lower overall carbon turnover at this colder site.
-The 50\% increase in  carbon at 80 years had, however, a much larger impact on GPP, giving a stimulation of ** at 80 years.
- At Harvard forest, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (as illustrated by the lower response to decreasing leaf C:N ratios in figure ** than other figures).
+For CO$_{2}$ fertilization at CAX the initial stimulation of GPP ($\sim$1.37) lead to an almost 50\% increase in LAI.
+At this site, water limitations (lower $B_{tran,mn}$), prevented any further increase in GPP.
+
+ At MET, in contrast, initial GPP stimulation was lower (26\%) on account of the lesser water stress.
+This resulted in slow but prolonged LAI increase, given the lower overall carbon turnover at this colder site.
+The 50\% increase in carbon at 80 years had, however, a much larger impact on GPP and NPP giving a stimulation of 53-69\% and 49-71\% respectively, (varying through the cycling climate forcing data).
+Thus, the overall impact of fertilization appears much greater than is typically reported in elevated CO$_{2}$ experiments, but \cite{king2005} report both large response to $sim$530 ppm CO$_{2}$ in Aspen stands, with increases in LAI and biomass of around 60\%.
+
+ At HVF, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (HVF has a smaller response to decreasing higher leaf N ratios, Fig.
+\ref{HVF state}).
+NWR had moderate (25\%) increases in initial GPP and LAI ($sim$30\%).
 At all sites, the fraction of NPP spent on N uptake increased, but most dramatically at Metolius, where initial N expenditure was lowest initially on account of low growth rates.
 
-As expected from the initial conditions, Niwot Ridge responded most strongly to nutrient additions.
-An initial NPP stimulation of 1.3 for NPP was magnified by subsequent increases in LAI to nearly 1.8 times the control value, causing a stimulation of 1.7 in NPP.
+N addition most strongly affects carbon expenditure on N uptake, and thus NPP, rather than GPP directly.
+At NWR an initial NPP stimulation of 1.3 for NPP was magnified by subsequent increases in LAI to nearly 1.8 times the control value, causing a stimulation of 1.7 in NPP, reflected the high N limitation at this site in the pre-fertilization state.
 None of the other sites exhibited such a dramatic change.
-In particular, Caxiuana, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%.
-Stimulation at Metolious was higher than Caxiuana, despite it's lower initial N limitation, given it's lack of limitation by water.
+In particular, for CAX, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%.
 
-The time-dependence of the responses generates an interesting issue for interpretation, since where LAI increases a lot, the responses tend to be dominated by this feature of the response.
-The first year responses, by way of contrast, tend to be much more similar across sites, but do not reflect the overall response of the system.
+The time-dependence of the responses generates an interesting issue for interpretation.
+The first year responses are illustrative of the short-term immediate impacts of change, whereas the longer-term responses are heavily dominated by the feedback responses.
+Further, depiction of absolute vs.
+relative differences can serve to highlight different aspects of the model response.
+Here we report values from 15-20 years after fertilization began, to be consistent with the methodology used by \cite{wieder2019} in their analysis of global fertilization of the default model, and to capture the dynamics of the whole system feedbacks captured in Fig.
+\ref{timescales}.
 
-\subsubsection{Overall responses to fertilization}
-
-\subsubsection{Differences across sites}
-
-Figures \ref{GPP CO2 and N respones 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the responses to both N and to CO$_{2}$ fertilization simultaneously, for GPP, NPP and, LAI.
+Figures \ref{GPP CO2 and N responses 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the responses to both N and to CO$_{2}$ fertilization simultaneously, for GPP, NPP and, LAI.
 Corresponding figures for vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1 and 2).
 Figures for each site showing the individual N and CO$_{2}$ responses are also included in the SI (Figures 3 to 10).
-
-The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.
-
-For the default parameterization (the intersection of all lines in figures \ref{GPP CO2 and N respones 2001} to \ref{LAI CO2 and N respones 2001}), the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.
-LAI responses (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) reflected the NPP change, except for at Metolius where the canopy did not respond to the higher NPP.
-
-For Nitrogen responses, while impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET), they were lower for the drought-stressed Caxiuan\~a and Niwot sites.
-For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET), increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused increasing respiration and negative impacts on NPP at some sites, even though impacts on LAI (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET) were all positive, reflecting the lower costs of N uptake and greater capacity to acquire N to build tissues and retain carbon for growth.
 
@@
   \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/NOVc_CNdep_GPP1__p2012.png}
-     \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on gross primary productivity (GPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
-}
-     \label{GPP CO2 and N respones 2001}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_CNtime_defpft.png}
+     \caption{Fertilization responses of the default model though time since the (step-change) in CO$_{2}$ (top row) and Nitrogen deposition (bottom row) for CAX (blue), NWR (red), HVF (yellow), MET (purple).}
+     \label{MET state}
  \end{figure}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/NOVc_CNdep_NPP1__p2012.png}
-     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on net primary productivity (NPP), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{NPP CO2 and N respones 2001}
+     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_GPP_y2013.png}
+     \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on gross primary productivity (GPP). }
+     \label{GPP_CN_2016}
   \end{figure}
@@
  \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/NOVc_CNdep_TLAI1__p2012.png}
-     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of CO$_{2}$ and N fertilization on leaf area index (LAI), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{LAI CO2 and N respones 2001}
+     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_GPP_y2013.png}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on net primary productivity (NPP).}
+     \label{NPP_CN_2016}
+  \end{figure}
+
+  \begin{figure}[h]
+     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_GPP_y2013.png}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on leaf area index (LAI).}
+     \label{LAI_CN_2016}
   \end{figure}

### 2019-07-02 main_revision.tex

@@
 \ref{timescales}.
-
-Figures \ref{GPP CO2 and N responses 2001}, \ref{NPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001} illustrate the responses to both N and to CO$_{2}$ fertilization simultaneously, for GPP, NPP and, LAI.
-Corresponding figures for vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1 and 2).
+Thus, figures \ref{GPP_CN} and \ref{LAI_CN} illustrate the responses to 15-20 years of CO$_{2}$ and N  fertilization simultaneously, for GPP and LAI.
+Corresponding figures for NPP, vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1-3).
 Figures for each site showing the individual N and CO$_{2}$ responses are also included in the SI (Figures 3 to 10).
@@
 \subsubsection{Impact of parameters}
-Across the four sites, parametric impacts on fertilization responses were very varied, when compared to the relatively consistent impact of parameters on the initial system state.
-A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses across sites, and the range of their impacts varied over an order of magnitude.
-Thus, the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change.
+ the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change.
 Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.
@@
      \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on gross primary productivity (GPP). }
-     \label{GPP_CN_2016}
+     \label{GPP_CN}
   \end{figure}
@@
      \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on net primary productivity (NPP).}
-     \label{NPP_CN_2016}
+     \label{NPP_CN}
   \end{figure}
@@
   \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_GPP_y2013.png}
+     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_TLAI_y2013.png}
      \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on leaf area index (LAI).}
-     \label{LAI_CN_2016}
+     \label{LAI_CN}
   \end{figure}

### 2019-07-02 main_revision.tex

@@
 Thus, figures \ref{GPP_CN} and \ref{LAI_CN} illustrate the responses to 15-20 years of CO$_{2}$ and N  fertilization simultaneously, for GPP and LAI.
+Note that each figure is a different scale to best illustrate the parametric impacts.
 Corresponding figures for NPP, vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1-3).
@@
 \subsubsection{Impact of parameters}
- the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change.
-Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.
+The parameter response space of the model is complex, and parameter impacts on fertilization responses vary with changing initial conditions and with treatment.
+Here, we discuss the major impacts of each parameter on the responses and how they relate to expectations based on model structure.
 
-The strong impact of \emph{slatop} had important implications for the initial state in many cases, and for several of the lower values, we screened out the results of the fertilization on account of very low LAIs (which resulted in erratic fertilization responses).
 The impact of \emph{slatop} on fertilization was different at every site.
-For example, at Metolius, high  \emph{slatop} was related to high N responses, plausibly since these cheaper leaves could produce a greater LAI/GPP feedback in response to the N fertilization.
-At Niwot Ridge, high SLA was related to strong GPP responses to CO$_{2}$.
-These simulations have higher water limitation (on account of their larger canopies, figure \ref{btran state}) and so respond more vigarously to the increased water use efficiency of high CO$_{2}$ growth conditions (while reducing their NPP and LAI with N fertilization, on account of higher respiration costs).
-These contrasting results are illustrative of the complexity of the interactions between model initial state, environmental forcing and parameterization, and of the need to understand how multiple simultaneous growth limitations can influence responses to climate forcing.
+For MET, \emph{slatop} was the most important control over N responses, with cheaper leaves allowing a greater LAI/GPP feedback in response to the N fertilization.
+At NWR, the opposite behaviour was found, with high SLA reducing the impact of N in cases where the pre-fertilization LAI meant that a saturating response of GPP was found.
 
@@
  \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_GPP_y2013.png}
+     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_GPP_y2013.png}
      \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on gross primary productivity (GPP). }
@@
   \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_GPP_y2013.png}
-     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on net primary productivity (NPP).}
-     \label{NPP_CN}
-  \end{figure}
-
-  \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_relCNdep_defpft_TLAI_y2013.png}
-     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on leaf area index (LAI).}
+     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_TLAI_y2013.png}
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on leaf area index (LAI).
+Note changes in scale between plots to show detail.}
      \label{LAI_CN}

### 2019-07-02 main_revision.tex

@@
 The impact of \emph{slatop} on fertilization was different at every site.
-For MET, \emph{slatop} was the most important control over N responses, with cheaper leaves allowing a greater LAI/GPP feedback in response to the N fertilization.
+For MET, \emph{slatop} was the most important control over N responses, with higher values allowing a greater LAI/GPP feedback in response to the N fertilization.
 At NWR, the opposite behaviour was found, with high SLA reducing the impact of N in cases where the pre-fertilization LAI meant that a saturating response of GPP was found.
+At HVF, higher values of \emph{slatop} slightly reduced both CO$_{2}$ and N fertilization for the same reason (although the range of variation at HVF is only 5\% between the highest and lowest values, as opposed to 40\% at NWR).
 

### 2019-07-02 main_revision.tex

@@
 
-The impact of \emph{slatop} on fertilization was different at every site.
+The impact of \emph{slatop} on fertilization was different at every site (figs.
+\ref{GPP_CN}, figures \ref{LAI_CN}).
 For MET, \emph{slatop} was the most important control over N responses, with higher values allowing a greater LAI/GPP feedback in response to the N fertilization.
@@
 
-Despite its strong impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO$_{2}$ responsiveness, and at Harvard, where low values decrease both responses.
-At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP (which is a non-linear response related to the saturation of light interception).
+\emph{froot\_leaf} had contrasting impacts at different sites.
+At MET,  where low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions.
+At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the respnse to \emph{slatop}.
 
@@
 At Caxiuan\~a, high \emph{leafcn} cases had stronger responses to N fertilization, but responses to N fertilization were low at this site in general.
-
-\emph{grperc}, or the degree to which growth is 'taxed', is constant and not responsive to any of the physiological processes upstream that depend on fertilization, thus, at most sites, its impacts are minimal, or indirect.
 
@@
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_NWR_y2000.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Niwot Ridge.}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_y1999.png}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) for GPP, NPP, LAI and N$+{leaf}$ at the CAX, NWR, HVF and MET sites}
      \label{NR1 state}
- \end{figure}
-
- \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_CAX_y2000.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Caxiuan\~a National Forest}
-     \label{CAX state}
-  \end{figure}
-
- \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_HVF_y2000.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Harvard Forest}
-     \label{HVF state}
- \end{figure}
-
- \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_-r200_MET_y2000.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) at Metolius Forest}
-     \label{MET state}
  \end{figure}

### 2019-07-02 main_revision.tex
Updates from Overleaf

@@
 
-Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuan\~a and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and \ref{LAI CO2 and N respones 2001}).
-Initial LAI was larger in these cases, explaining the decrease in fertilization response as above (lower increases in GPP for a unit change in LAI).
-At Niwot Ridge, low values of \emph{stem\_leaf} were associated with very high LAI and high water limitation (Figure \ref{btran state}), and CO$_{2}$ fertilization thus had a large impact on GPP and NPP (as for \emph{slatop}), but the high NPP has no impact on LAI (SI Figure 3), reflecting the still-high N limitation.
+Low \emph{stem\_leaf} ratio caused lower responses to N fertilization CAX, NWR and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the GPP-LAI non-linearity issue.
+At NWR,  N limitation restricted CO$_{2}$ responses in general, but
+
  Correspondingly, adding Nitrogen in the low \emph{stem\_leaf} runs increases LAI substantially, but only has a very limited impact on GPP and NPP, given water limitations.

### 2019-07-03 main_revision.tex
Updates from Overleaf

@@
 At HVF, higher values of \emph{slatop} slightly reduced both CO$_{2}$ and N fertilization for the same reason (although the range of variation at HVF is only 5\% between the highest and lowest values, as opposed to 40\% at NWR).
-
 \emph{froot\_leaf} had contrasting impacts at different sites.
@@
 
-Low \emph{stem\_leaf} ratio caused lower responses to N fertilization CAX, NWR and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the GPP-LAI non-linearity issue.
-At NWR,  N limitation restricted CO$_{2}$ responses in general, but
+Low \emph{stem\_leaf} ratio caused lower responses to N fertilization CAX, NWR and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the NPP-LAI non-linearity issue.
+At NWR,  N limitation restricted CO$_{2}$ responses, but at CAX and NWR the same type of response was seen.
+At MET, lower \emph{stem\_leaf} allowed greater responses, plausibly from increasing the speed of the NPP/LAI feedback.
 
- Correspondingly, adding Nitrogen in the low \emph{stem\_leaf} runs increases LAI substantially, but only has a very limited impact on GPP and NPP, given water limitations.
+At all site, higher $N_{costs}$ increases the responsiveness of the system to N, and to a lesser extent CO$_{2}$.
 
-The $N_{costs}$ parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
+parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
 At Caxiuan\~a, Niwot and Harvard forest, high  $N_{costs}$ resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
@@
      \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) for GPP, NPP, LAI and N$+{leaf}$ at the CAX, NWR, HVF and MET sites}
-     \label{NR1 state}
+     \label{state}
  \end{figure}

### 2019-07-03 main_revision.tex
Updates from Overleaf

@@
 At HVF, higher values of \emph{slatop} slightly reduced both CO$_{2}$ and N fertilization for the same reason (although the range of variation at HVF is only 5\% between the highest and lowest values, as opposed to 40\% at NWR).
+
 \emph{froot\_leaf} had contrasting impacts at different sites.
@@
 
-At all site, higher $N_{costs}$ increases the responsiveness of the system to N, and to a lesser extent CO$_{2}$.
-
-parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}).
-At Caxiuan\~a, Niwot and Harvard forest, high  $N_{costs}$ resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).
-These high $N_{costs}$ scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.
-Low values of $N_{costs}$ were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.
-While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.
+At CAX, NWT and HVF, higher $N_{costs}$ increases the responsiveness of the system to N, and to a lesser extent CO$_{2}$.
+The magnitude of the impact on N fertilization appeared to be related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).
+Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF.
 
 The fraction of N fixers (\emph{frac\_fixers}) has a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
-For GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP CO2 and N respones 2001}, \ref{LAI CO2 and N respones 2001}).
-For NPP, N deposition responses were muted at Metolius, and reversed in sign at Caxiuan\~a, where the low N fixer cases had a very negative (-8\%) NPP response (despite a +7.6\% GPP response), on account of the model increasing the N content of the leaves (and thus respiration rates) by up to 10\% in the lowest N cost case (SI Figure 6).
-Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.
-At Niwot Ridge the changing fixation capacity had no impact on the response of GPP to N fixation, despite its influence on LAI, suggesting that water limitation was a greater constraint on total GPP than leaf area.
-At Metolius, again the changing fixation capacity had no impact on the response of GPP to N deposition (Figure \ref{GPP CO2 and N respones 2001}) but in this case, it had no impact on LAI either.
-Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization at this site, suggesting fundamental metabolic limitations to canopy carbon economics (e.g.
-a short growing season) that render only higher LAI simulations sensitive to fertilization.
+For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP_CN}, \ref{LAI_CN}).
+At MET,  the changing fixation capacity had no impact on the response of GPP or LAI to N fertilization.
+Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization for this site (which had the smallest PNU in the first instance), suggesting fundamental metabolic limitations to carbon economics and growth (e.g.
+a short growing season).
 
-The impact of target \emph{leafcn} on fertilization responses was not particularly dominant (recalling that this was a significant control over the system state at all sites), an exception being Niwot Ridge, where low values of \emph{leafcn} had very strong to CO$_{2}$ fertilization (the strongest response at this site for GPP).
-These cases suffered from high water limitation, given their high $N\_{leaf}$ values and high gas exchange fluxes (Figure \ref{btran state}) and thus CO$_{2}$ fertilization alleviates this problem.
-At Caxiuan\~a, high \emph{leafcn} cases had stronger responses to N fertilization, but responses to N fertilization were low at this site in general.
+The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state.
+The primary impacts were at, an exception being Niwot Ridge, where low values of \emph{leafcn} had a low response to N fertilization These cases suffered from high water limitation, given their high $N\_{leaf}$ values and high gas exchange fluxes (Figure \ref{btran state}) and thus CO$_{2}$ fertilization alleviates this problem.
+At CAX, high \emph{leafcn} cases had stronger responses to N fertilization, reflecting the higher PNU in those cases prior to fertilizaion.
 

### 2019-07-03 main_revision.tex

@@
 
-At CAX, NWT and HVF, higher $N_{costs}$ increases the responsiveness of the system to N, and to a lesser extent CO$_{2}$.
-The magnitude of the impact on N fertilization appeared to be related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).
-Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF.
+Higher $N_{costs}$ increase the responsiveness of the system to N at CAX, NWT and HVF.
+The magnitude of the impact on N fertilization ism as might be expected, related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).
+Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF **.
 
-The fraction of N fixers (\emph{frac\_fixers}) has a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
+\emph{frac\_fixers} (the maximum fraction of NPP available for N fixation) had a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
 For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP_CN}, \ref{LAI_CN}).
@@
 The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state.
-The primary impacts were at, an exception being Niwot Ridge, where low values of \emph{leafcn} had a low response to N fertilization These cases suffered from high water limitation, given their high $N\_{leaf}$ values and high gas exchange fluxes (Figure \ref{btran state}) and thus CO$_{2}$ fertilization alleviates this problem.
-At CAX, high \emph{leafcn} cases had stronger responses to N fertilization, reflecting the higher PNU in those cases prior to fertilizaion.
+Complex responses were illustrated, reflecting the differential costs and benefits of Ni on the system.
+For N fertilization, at CAX and NWR, lower values of \emph{leafcn} had smaller responses to N fertilization, reflecting the higher initial LAI of these configurations and the non-linear GPP feedback discusses earlier.
+At CAX, the highest \emph{leafcn} had smaller CO$_{2}$ responses.
+(Figure \ref{btran state} illustrates that this configuration had less water stress, given that both the lower LAI and lower photosynthetic capacity should lead to smaller stomatal conductances and evapotranspiration.
+and thus had less water stress to be alleviated by the high CO$_{2}$.
 
-At Niwot Ridge, runs with low values of the \emph{medlyn\_slope} parameter (and thus high water use efficiency) were able to respond very strongly to increasing Nitrogen (SI Figure 4), given their lower initial water stress (Figure \ref{btran state}).
-Surprisingly, however, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.
+Surprisingly, given the importance of \emph{medlyn\_slope} on stomatal responses to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response.
+Earlier studies using a less constrained range of \emph{medlyn\_slope} values indicated that it can have a potentially large impact on system state, (results not shown) but that the fertilization response is typically limited.
+The linear response of the stomatal conductance term to $CO_{2}$ mean that the direct impact of \emph{medlyn\_slope} is likely to be restricted to changes in the pre-fertilization state.
 
-In general the impacts of the maintenance respiration intercept \emph{lmr\_intercept} were small, and limited to the impact of large values increasing fertilization responses to CO$_{2}$ and N, potentially on account of the low LAI and thus greater feedback between LAI and GPP, and in line with its limited impact on the system state, \emph{frac\_ectomy\_fungi} had no discernible impact on fertilization responses.
-
-There was limited impact of \emph{cn\_flex\_a} on the fertilization responses, as per it's impact on the state.
-At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$.
-$N_{leaf}$ followed suit, increasing substantially in the N deposition experiment.
-Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when N costs are high, therefore potentially suppressing $N_{leaf}$.
-Adding N overcomes this suppression, leading to greater fertilization responses.
-
-The impact of \emph{cn\_flex\_c} (the parameter controlling the degree to which plant C:N ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon (SI Figure 2)  at Harvard Forest and Niwot Ridge (and Caxiuan\~a, to a smaller extent).
-These impacts do not mirror those on leaf area index or NPP, and overall the response to this parameter appears complex.
-At Metolius and Niwot, \emph{cn\_flex\_c} has a dominant control over the impacts of CO$_{2}$ fertilization on $N_{leaf}$ (SI Figure 1).
-The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high (e.g.
-SI Figure 3).
-Low  $N_{leaf}$ values permit higher leaf area and vegetation carbon, but reduce assimilation rates, as illustrated at Harvard Forest and Niwot Ridge (SI Figure 5 \& 7).
+The impact of \emph{cn\_flex\_c} was small compared to the more dominant parameters.
+Higher values of \emph{cn\_flex\_c} spend less carbon on N uptake when tissue N is lower than the target amount, allowing N ratios to vary more.
+Impacts of this parameter are thus manifested through their impact on $N_{leaf}$ (Figure SI **).
+Smaller values of \emph{cn\_flex\_c} corresponded in all cases to smaller variations in $N_{leaf}$ (values closer to 1.0) under both CO$_{2}$ and N fertilization (For N deposition, $N_{leaf}$ always increases and for CO$_{2}$ it always declines.) Lower values of \emph{cn\_flex\_c} also corresponded to larger CO$_{2}$ stimulation of GPP, illustrating the direct impact of this parameter, as leaves maintain higher photosynthetic capacity under more N limiting conditions.
+The impact on NPP and LAI was less clear, given the corresponding impacts of higher respiration both for maintenance and for N uptake.
+At CAX, NWR and HVF,  the highest values of \emph{cn\_flex\_c} \emph{also} had a higher response to CO$_{2}$ than the default.
+The reasons for this discontinuity are unclear, however, the impact of \emph{cn\_flex\_c} in the CLM5 code is subject to at least two 'caps', (whereby plants cannot spend less than 0.5 or more than 1.0 of the carbon that would retain the target LAI), potentially leading to complex feedbacks once these caps are reached.
 
@@
 Similar results were found at all other sites.
-Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology, and this may explain the large fertilization responses found here compared to those of \cite{wieder2019}.
+Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology.
 
 \section{Discussion}
-
 The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}).
@@
 
-At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (Wieder et al.
+At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (CLM4 and CLM4.5, Wieder et al.
 2018).
-The trade-off between responses to N fertilization and CO$_{2}$ response when the fraction of N fixing plants varies mirrors the change in fertilization response between CLM4, CLM4.5 and CLM5, and the CLM5 response is consistant with the capacity of the model to increase fixation rates in response to higher carbon dioxide levels.
-The number of degrees of freedom illustrated here precludes, however, proof that this is the only contributing factor.
-This analysis indicates that global experiments with lower or zero \emph{frac\_fixers} would help to ascertain whether this is indeed the case.
+Given the dominant impact of N fixation on the model fertilization responses, it seems likely that the capacity for the model to fix addition N in response to CO$_{2}$ fertilization is at least partly responsible for the change in the global responses, and also that any changes in this parameter would be likely to have large impacts on this result.
+Intuitively, it seems likely that an ecosystem that could more rapidly shift a large fraction of resources to N uptake would
 
@@
      \caption{Fertilization responses of the default model though time since the (step-change) in CO$_{2}$ (top row) and Nitrogen deposition (bottom row) for CAX (blue), NWR (red), HVF (yellow), MET (purple).}
-     \label{MET state}
+     \label{timescales}
  \end{figure}

### 2019-07-04 main_revision.tex
Updates from Overleaf

@@
-\usepackage{}\documentclass[draft,linenumbers]{agujournal}
+\documentclass[draft,linenumbers]{agujournal}
 
@@
 \emph{froot\_leaf} had contrasting impacts at different sites.
-At MET,  where low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions.
+At MET,  low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions.
 At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the respnse to \emph{slatop}.
@@
 \subsection{Low leaf area investigations}
-In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites.
-(for Caxiuan\~a, 3.0 (modeled) vs.
-5.3 (observed) \cite{fisher2007}, for Niwot Ridge, 1.7 vs.
-4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs.
-3.5 \cite{williams1996} and for Metolius, 0.5 vs.
+In general, we found that LAI predicted by the default CLM5 were often low  compared to published estimates for these sites.
+(CAX; 3.0 (modeled) vs.
+5.3 (observed) \cite{fisher2007},  NWR; 1.7 vs.
+4.2 \cite{bowling2009}, HVF; 2.6 vs.
+3.5 \cite{williams1996}  MET; 0.5 vs.
 1.5 (\cite{spadavecchia2011}).
@@
 
-To interrogate this further, we ran an entire separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
-These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition.
-In the default parameterization CO$_{2}$ and N fertilization increased LAI, amplifying these magnitude of ecosystem responses to these perturbations.
-For example, at Harvard Forest, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.18, 1.17 and 1.22 for GPP, NPP and LAI, respectively, in the high LAI case vs.
-1.43, 1.66, 1.47 in the control) and a lower N response as well (1.05, 1.20 and 1.15 for GPP, NPP, and LAI, respectively, in the high LAI case vs.
-1.1, 1.26, 1.21 for the control).
+To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
+These simulations indicated that, with the resulting higher (and more consistent with observed) LAI values, the model was less sensitive to both CO$_{2}$ and N deposition.
+For example, at HVF, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response of  (1.18, vs 1.43 for the control)  and a lower N response as well (1.20 vs 1.26 for the control).
 Similar results were found at all other sites.
-Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology.
 
@@
 Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
+Interpretation of these simulations should be informed by solid comprehension of which model properties dominate the size of the feedback strength both within and between models, and we intend these illustrative simulations to provide initial guidance of this process.
 
-At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (CLM4 and CLM4.5, Wieder et al.
+At the time of writing, we do not yet know the carbon-climate feedback strength of the default version of CLM5/CESM2.
+In offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ and a weaker response to N fertilization than previous versions of the model (CLM4 and CLM4.5, Wieder et al.
 2018).
-Given the dominant impact of N fixation on the model fertilization responses, it seems likely that the capacity for the model to fix addition N in response to CO$_{2}$ fertilization is at least partly responsible for the change in the global responses, and also that any changes in this parameter would be likely to have large impacts on this result.
-Intuitively, it seems likely that an ecosystem that could more rapidly shift a large fraction of resources to N uptake would
+Further, the impact of both types of fertilization was found, in all versions of the model, to be highly variable with PFT (from 0\% to 110\% for N, -10\% to +55\% for CO$_{2}$).
+In this study we illustrate how variations in the pre-fertilization state, water and N limitation, and parametric variation combine to generate variation in the responses to forcing.
+While we set out primarily to to determine responses to model parameters, findings concerning the impact of model state on forcing sensitivity will have important implications for the interpretation of simulations with the default model.
 
-The CLM5 model has a number of new prognostic elements, which allow predictions in elements of the C and N cycles which were assumed to be constant in previous versions of the model.
-Exploration of the parameter space that controls these new elements allows a clear exploration of the sensitivities of the output space of the model to critical assumptions.
+Given the dominant impact of N fixation (\emph{fracfixers}) on the model fertilization responses, it seems likely that the new capacity, introduced in CLM5, for plants to use carbon to pay for the fixation of addition nitrogen under CO$_{2}$ fertilization is at least partly responsible for the change in the global responses, and also that any changes in this parameter would be likely to have large impacts on this result.
 
-The model output space is notable complex, and in many cases the sign of the impact of the different parameters changes between sites, and trade-offs in the influence of carbon and nitrogen economic parameters interact with both the initial state and environmental forcing.
-That the parameter space of CLM5 is complex is not surprising, given its relatively high process fidelity, but a) this type of sensitivity analysis,  illustrating how different parameterizations respond to different growth conditions and environmental forcing, is relatively rare for models of this kind,  b) the large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and c) the results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites.
-Sensitivity tests across broader geographical ranges would help to eludicate more clearly how changes in baseline water and N limitation impact parameter responses and patterns of responses to fertilization.
+Aside from the straightforward direct impacts of the fixation rate on model predictions, however, the model output space was notably complex.
+For example,  in some cases the sign of the impact of individual parameters changes between sites, while other parameters can demonstrate non-linear and counter-intuitive impacts as the result of numerous interacting feedbacks.
+To summarize our findings, we note a variety of key outcomes of our analysis on the factors controlling responses to CO$_{2}$ and N fertilization in the CLM5, including:
 
-Further, our results indicate that, in line with our initial hypothesis 2, model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization, since for each site there was substantial lack of overlap between the parameters controlling the default state and the fertilization responses.
-We did observe, however, many instances where simulations with high leaf area index responded less strongly to fertilization, and thus the numerous parameters impacting this feature of the initial model state (dominated by tissue allocation ratios and \emph{slatop}) would also have an indirect impact on fertilization.
-Our results suggest a more general point, that high LAI systems should respond less to fertilization.
-We did not find a clear statistical link between initial LAI and model output across the whole parameter space, but interpretation of global patterns in terrestrial cardon storage should likely take this mechanism into account.
+1.
+The initial model state has a dominant effect on  responses to CO$_{2}$ and N fertilization.
+The initial leaf area index, and whether the system is on a linear or saturated part of the LAI vs GPP relationship, has a large impact on any fertilization response.
+Parameter impacts in more productive systems are likely to generate smaller LAI/GPP feedbacks and lower parametric variation as a result.
+Many of the direct impacts of the parameters assessed here were masked by their large impacts on the initial model state.
+The clearest examples of this were the \emph{slatop} and \emph{stem\_leaf} impacts, but this impact was illustrated for many model configurations.
+
+2.
+The degree of water or nutrient limitation in the initial state is also a dominant factor over the model responses to CO$_{2}$ fertilization and N deposition.
+Once again, the direct impact of parameter perturbation is frequently masked by how changes affect the initial limiting conditions.
+Examples of this include how low \emph{froot\_leaf} increases water limitation at NWR, increasing responsiveness to CO$_{2}$ and how high \emph{leafcn} reduces water stress and thus decreases the CO$_{2}$ response at CAX.
+
+3.
+Other parameters with a direct impact on fertilization responses included those related to the acquisition of nitrogen (\emph{N\_costs}) and the stoichiometric flexibility of the system (\emph{cn\_flex\_c}) illustrate the sensitivity of the model to relatively unconstrained processes within the nitrogen cycle.
+
+4.
+Many of the parameters that are dominant over the initial state (\emph{slatop}, \emph{stem\_leaf} \emph{leafcn}) either have limited impacts on fertilization responses, or have impacts that are largely traceable to their impacts on the initial model state.
+In contrast, numerous parameters that more directly and uniformly impact fertilization responses, (\emph{frac\_fixers} \emph{N\_costs}) are less important for the initial model state.
+We thus suggest that model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization.
+
+5.
+The results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites.
+Given the complexity of the model, this helps generate confidence that the model outcomes are derived from the intended process mechanisms, and are not subject to significant deviations from the intent of the model structure (as opposed to .
+
+6.
+The large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and  Sensitivity tests across broader geographical ranges would help to eludicate more clearly how changes in baseline water and N limitation impact parameter responses and patterns of responses to fertilization.
+
+\subsection{Applicability of results to other land surface models}
+Our finding that Nitrogen fixation plays a dominant role over responses to $CO_{2}$ fertilization concurs with earlier studies.
+\cite{wieder2015} investigated the role of two alternative linear models of N fixation within the CLM4.5, while \cite{meyerholt2015} investigate a range of alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}.
 
@@
 
-\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
-Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 were primarily motivated by the need to move the model behaviour closer to several key dynamic features of the system.
-Notably, the dynamics of $N_{leaf}$ with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}).
-The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.
-In doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.
-By investigating the critical roles of these processes in the model dynamics, we hope to highlight a research need, and to stimulate research on and improvements in the representation of these processes.
-
-In analyzing the results of this study, for example, we found numerous instances where the model reduced carbon use efficiency (CUE, defined here as net primary productivity, NPP, divided by gross primary productivity, GPP, where the respiratory carbon used for N uptake is not counted within NPP and thus lowers CUE) by changing tissue C:N ratios in response to forcing.
+\subsection{Model improvements highlighted by this analysis}
+While there are many potential improvements that could be made to any land surface model, this particular assessment of the parameter space of the CLM5 in this study revealed several dynamics which highlight parts of the representation of biogeochemistry which might be improved in future versions.
+Firstly, the response of NPP to increasing \emph{slatop} was often negative, despite increasing GPP, as the fixed-fraction allocation scheme in the model will generate leaves that respire more than they assimilate under very high productivity conditions.
+This highlights the need for either more realistic tissue allocation, or for a scheme which optimizes leaf allocation to prevent leaf layers in negative carbon balance (\cite{fisher2010}).
+Secondly, in other cases, the heuristic C:N stoichiometry algorithm can increase tissue N content in response to environmental forcing in a manner which, again, can act to reduce net assimilation by modifying the balance between assimilation and respiration.
 In principle, there exists an optimal leaf C:N ratio that maximizes leaf C export under a given set of conditions.
-Despite its optimal allocation of leaf N \emph{within} the leaf (to photosynthetic enzymes) via the LUNA model, CLM5 does not include any scheme that ensures that overall tissue CN ratios conform to any physiological optima.
-Instead, a heuristic model adjusts tissue CN ratios round the target ratio (an input parameter).
-Numerous models of this type have been proposed in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{anten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}).
-We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in  LUNA, but future nitrogen cycling model development efforts could re-visit this topic.
+Numerous models of this type have been proposed in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{anten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}) and future nitrogen cycling model development efforts could re-visit this topic.
 
@@
 Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants.
-In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).
-The 0-1 range of \emph{frac\_fixers} used here is ecologically unlikely, given the costs borne by plants that can fix N give them a competitive disadvantage in systems no longer limited by N.
-However, it suggest that a high CO$_{2}$ world would likely provide much greater benefits to N fixation, and that those benefits would be ameliorated by ongoing increases in N deposition.
+In general, Nitrogen fixers are known to primarily exist in early-successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).
+
+Lastly, we recognize the importance of other limiting nutrients, particularly phosphorus, in controlling the influence of CO$_{2}$ fertilization (\cite{reed2015}).
+Phosphorus dynamics were not integrated into this version of the CLM, but alternative configurations of the CLM exist (e.g.
+\cite{yang2014}, \cite{zhu2016}) that will provide contrasting biogeochemical predictions in future studies.
 
 \subsection{ Methodological Limitations}
-\subsubsection{Limitations of step-change experiments}
 Here we investigate the parameter space of the model using step-change experiments.
 These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions.
-Further, we expect that the carbon and nitrogen cycles may take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.
-
-To test the impact of time on the fertilization responses, we also analyzed the impact of fertilization at 5 and 50 years into the elevated CO$_{2}$ and N experiments.
-At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, (result not shown) to those at 15 years.
-However, at 5 years into the fertilization in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent (a x~2.06 impact on NPP, compared to ~0.95 in the year 15 analysis at Niwot Ridge, indicating that the results described above had already been subject to progressive N limitation, and were closer to an equilibrated state.
 
@@
 
-Lastly, we recognize the importance of other limiting nutrients, particularly phosphorus, in controlling the influence of CO$_{2}$ fertilization (\cite{reed2015}).
-Phosphorus dynamics were not integrated into this version of the CLM, but alternative configurations of the CLM exist (e.g.
-\cite{yang2014}, \cite{zhu2016}) that will provide contrasting biogeochemical predictions in future studies.
-
-\subsubsection{Temperature and water responses}
 In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5.

### 2019-07-05 referee_responses.tex

@@
 
-\textsf{AR: Fair enough}
+\textsf{AR: Fair enough.}
 
@@
 \textsf{AR: Noted...}
+
 This paper documents three new assumptions (arguably four- section 2.1.4 as well) to CLM5 related to N; however, it doesn't give much consideration to these hypotheses, instead focussing on a broader list of impacts related to parameter choice.
@@
 \textsf{AR: We appreciate the disconnect between the parameter choice being seemingly unrelated to the lengthy discussion of the new nitrogen cycling elements of the model.
-On further reflection,  our overall aim is to illustrate how the C and N responses of the model, as determined for the global simulations by the Wieder et al.
-study, are impacted by parameter choice, given their large implications for carbon cycle feedback analysis.
-We think this is of slightly greater interest than specifically probing the differences between the CLM4.5 and CLM5 N cycle, although the two are related.
-While many of the elements that affect the CO2 and N responses are linked to the new model structure, there are several parameters that have large impacts which are not.
+On further reflection,  we determined that our overall aim is to illustrate how the C and N responses of the model are impacted by parameter choice, given their large implications for carbon cycle feedback analysis.
+We think this is of slightly greater interest to a wider audience than specifically probing the differences between the CLM4.5 and CLM5 N cycle, although the two are related.
+While many of the elements that affect the CO$_{2}$ and N responses are linked to the new model structure, there are several parameters that have large impacts which are not.
 In the new version of the manuscript, we have chosen to reduce the amount of text devoted to the specific changes to the N cycle, and instead to shift the focus slightly such that we are now looking specifically at how parameters affect the fertilization responses.
@@
 \textsf{An alternative approach would be to run the simulations with the various pieces of code turned off and returned to the default state.
-The problem with implementing that idea is that large changes of that type tended to modify the system state, and therefore (as is one of the conclusions of the analysis) impact upon the fertilization responses just because the initial system state was moved to conditions that were more or less light saturated, and more or less limited by water or nitrogen.
-Further, with the exception of the LUNA code, technically it is quite difficult to turn off and on the FUN and FlexCN options without invoking other more poorly documented post CLm4.5 modifications, thus we have taken the approach of exploring the behaviour of the new models by modifying their major controlling parameters.}
+The problem with implementing that idea is that large changes of that type tended to modify the system state.
+Therefore there are impacts on the fertilization responses solely because the initial system state was moved to conditions that were more or less light saturated, and more or less limited by water or nitrogen.
+(indeed, this is one of the main conclusions of the analysis).
+Further, with the exception of the LUNA code, technically it is quite difficult to turn off and on the FUN and FlexCN options without invoking other more poorly documented post CLM4.5 modifications, thus we have taken the approach of exploring the behaviour of the new models by modifying their major controlling parameters.}
 
@@
 We have now modified the stomatal slope range to use the standard error reported in Table 2 of de Kauwe et al.
-2015}}
+2015.
+This does notably reduce the impact of the Medlyn slope parameter on the analysis, and remove some of the more interesting feedbacks on water limitation that the wider range had illustrated, but maybe that's just how it has to be.}
 
@@
 
-\textsf{AR: We actually ran the model for those sites, but got very low LAI predictions, and therefore very sensitive responses to fertilization.
+\textsf{AR: For the sake of full disclosure, we did actually run the model for those sites, but for reasons that are unclear, we got very low LAI predictions, and therefore extremely sensitive responses to fertilization.
 We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO$_{2}$ and N forcing.}
@@
 
-\textsf{AR: Fair point. It seems better to explore more parameters for the sake of completeness, but \textbf{maybe some can be parsed out into SI to allow for a more readable paper.} }
+\textsf{AR: We agree here, and have as a result removed 5 of the 13 parameters that had limited impacts on the fertilization responses.
+We have also made efforts to streamline the remaining results text, and we hope the new text is not such a trial to get through (concurring that the first version was challenging in this respect).
+}
 
@@
 
-\textsf{AR: The MS is part of a special issue, and so the paper will all become available at the same time (except the Kennedy paper which is already published).
-We have reduced the number of citations for the kennedy paper down to two.
+\textsf{AR: The MS is part of a special issue, and so all the papers should become available at the same time (except the Kennedy paper which is already published).
+We have reduced the number of citations for the Kennedy paper down to two.
 The primary resource we cite is the CLM5 tech note, which is already online but will also become an appendix to the Lawrence et al.
 paper when these papers are all published simultaneously.
-This is our overall strategy to documenting the entirety of the model, and we wanted to avoid duplicating large amounts of the technical text in different places.}
+Across the special issue, this is our overall strategy to documenting the entirety of the model, and we wanted to avoid duplicating large amounts of the technical text in different places.}
 
@@
 
-\textsf{AR: Fair point, but 1) you may be an atypical reviewer and 2) we still think this would distract from the main purpose of the paper, which is to document and explain the differential sensitivities of the parameter ensemble to CO2 and N fertilization.
-If we added in specific site data, readers would almost certainly focus on the model fit to particular sites, which will no doubt be explored in great detail in the wider literature.}
+\textsf{AR: Fair point, but 1) you may be an atypical reviewer and 2) we still think this would distract from the main purpose of the paper, which is to document and explain the differential sensitivities of the parameter ensemble to CO$_{2}$ and N fertilization.
+If we added in specific site data, readers would almost certainly focus on the model fit to particular sites, which we anticipate will be explored in future literature.
+Nonetheless, the plots are a little abstract.
+We have tried to improve the presentation of the results by consolidating the number of plots, putting all of the state outputs into the same figure, adding a new figure focused on the time-dependant responses of the default model, and }
 
@@
 
-\textsf{AR: Needs reflection **}
+\textsf{AR: We concur that this aspect was difficult to ascertain from the original text, and so have re-written the discussion section with a view to including clearer digestion of the findings, and also a specific section on which parts of the analysis may and may not be applicable to models other than CLM5.
+We also reduced the scope of the limitations section to balance out weighting of the discussion.}
 
@@
 
-\textsf{AR: Fair point. We have modified these points to read 
-"The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle."}
+\textsf{AR: Fair point. We have modified these points to read\\
+"The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle.}
 
-\textsf{The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice.}
+\textsf{The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice, but parameter responses are varied between sites that differ in their pre-fertilization water and nitrogen limitation.
+}
 
-\textsf{Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state.}
+\textsf{Parameters controlling nitrogen fixation rates dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. "}
 
@@
 
-\textsf{AR: We have changed this to 'pre-fertilization state'. }
+\textsf{AR: We have changed this to 'pre-fertilization state' throughout. }
 
@@
 
-\textsf{AR:  We took out 'active'. and changed to "included nitrogen (N) cycling and limitation of plant growth."}
+\textsf{AR:  We took out "active". and changed to "included nitrogen (N) cycling and limitation of plant growth."}
 
@@
 - Line 52-3: Is that true? Unless that can be supported I would suggest omitting. 
-\textsf{AR: Changed from 'a majority of models' to 'many models'}
+\textsf{AR: We have deleted this section,}
 
@@
 
-\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications **}
+\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications.**}
 
@@
 
-\textsf{AR: We have deleted this hypothesis and replaced with "2.
-There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization.
-** reconsider hypotheses"}
+\textsf{AR: We have modified the hypotheses to be "motivating questions", to re-frame the results and discussion to encompass our intended aims.}
 
@@
 Anyway, irrelevant I'm sure.
-
-\textsf{AR: I am inclined to agree, but there's not much that can be done at this point.}
+\textsf{AR: I am inclined to agree, but there's not much that can be done about this issue at this point.}
 
@@
 
-\textsf{AR: We added a new sentence to describe the dependance of costs on the environment: \emph{The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation.
+\textsf{AR: We added a new sentence to describe the dependence of costs on the environment: \emph{The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, re-translocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (Bloom et al.
+[1985], Jiang et al.
+[2017], Terrer et al.
+[2018]) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation.
 Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them.
@@
 
-\textsf{AR: Changed this to: \emph{Newer codes within the CLM framework (e.g.
+\textsf{AR: We changed this to: \emph{Newer codes within the CLM framework (e.g.
 the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue by removing leaves in negative carbon balance.}}
@@
 
-\textsf{AR: We have deleted these sections, and reorganized the grouping of the more generic text that replaces them.}
+\textsf{AR: We have deleted these sections, and reorganized the grouping of the more high-level text that replaces them.}
 
@@
 Further, in light of the need to simplify the discussion and analysis, we have removed 5 of the 13 parameters from the initial version.
-The reduced range of variation for the Medlyn slope parameter further reduced the impact of that parameter on the output, and therefore less text is expended on its discussion.)
+The reduced range of variation for the Medlyn slope parameter further reduced the impact of that parameter on the output, and therefore less text is expended on its discussion.}
 

### 2019-07-05 main_revision.tex

@@
 
-\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice, but parameter responses are varied across sites and experiments.
+\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice, but parameter responses are varied between sites that differ in their pre-fertilization water and nitrogen limitation.
 
-\item Parameters controlling nitrogen fixation dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. 
+\item Parameters controlling nitrogen fixation rates dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. 
 \end{keypoints}
@@
 
-Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global biogeochemical feedback analysis, including the controls on global carbon fertilization `$\beta$' values (the change in ecosystem carbon storage in response to CO$^{2}$ fertilization) (\cite{arora2013}, \cite{friedlingstein2014}), we consider the illustration of how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations to be an important feature of the documentation of the new model.
+Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global biogeochemical feedback analysis, including the controls on global carbon fertilization `$\beta$' values (the change in ecosystem carbon storage in response to CO$_{2}$ fertilization) (\cite{arora2013}, \cite{friedlingstein2014}), we consider the illustration of how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations to be an important feature of the documentation of the new model.
 
@@
 Narrowing the range of variables of interest to those that only impact control states and fluxes (and directing more resources to studying those variables) is thus likely to miss important aspects of transient responses under climate change scenarios.
-Similarly, much effort is expended on calibration of models to present-day conditions (\cite{fer2018},\cite{lu2018}, when their responses to climate forcing might be relatively unconstrained by these efforts.
-Thus, our motivating hypotheses here are:
+Similarly, much effort is expended on calibration of models to present-day conditions (\cite{fer2018}, \cite{lu2018}, when their responses to climate forcing might be relatively unconstrained by these efforts.
+Thus, our motivating questions here are:
 
-1. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control the pre-fertilization model state.
+1. Are parameters that dominate responses to CO$_{2}$ and nitrogen fertilization also dominant over the pre-fertilization model state?
 
-2. There is a trade-off between model parameterizations that allow strong CO$_{2}$ fertilization and those that allow strong N fertilization.
+2. How do fertilization impacts and parameter sensitivities vary with site conditions, as expressed via water and nutrient limitations.
 
@@
 
-\subsection{Applicability of results to other land surface models}
-Our finding that Nitrogen fixation plays a dominant role over responses to $CO_{2}$ fertilization concurs with earlier studies.
-\cite{wieder2015} investigated the role of two alternative linear models of N fixation within the CLM4.5, while \cite{meyerholt2015} investigate a range of alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}.
-
-In the CMIP6 generation of land surface models, it is expected that a much greater fraction of model submissions will contain active representations of the terrestrial nitrogen cycle.
-We propose here that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained for all models, and that these types of analysis would allow a deeper understanding of land surface model outputs and behaviour.
-Many further analyses are possible, including global feedback analyses with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constrains of the parameter ranges used.
-We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.
-
 \subsection{Model improvements highlighted by this analysis}
@@
 This highlights the need for either more realistic tissue allocation, or for a scheme which optimizes leaf allocation to prevent leaf layers in negative carbon balance (\cite{fisher2010}).
+
 Secondly, in other cases, the heuristic C:N stoichiometry algorithm can increase tissue N content in response to environmental forcing in a manner which, again, can act to reduce net assimilation by modifying the balance between assimilation and respiration.
@@
 
-Another area requiring focus is the dominance of the 'fraction of fixers' over fertilization responses.
+Thirdly, requiring focus is the dominance of the 'fraction of fixers' over fertilization responses.
 \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants.
@@
 
-\subsection{ Methodological Limitations}
+\subsection{Applicability of results to other land surface models}
+Our finding that Nitrogen fixation plays a dominant role over responses to $CO_{2}$ fertilization concurs with earlier studies.
+Notably, \cite{meyerholt2015} investigate six alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}).
+They emphasize that the formulation which, as in CLM5, switches N sources depending on their relative costs, represent a promising means by which N fixation might be simulated.
+Interestingly, their model also includes a maximum rate of fixation per unit (root) biomass, but analysis to it's sensitivity is not shown.
+Further, given the difficulty of matching observations of complex N cycle processes with more abstract 'big leaf' models such as CLM5, and the ecological complexities governing the abundance their abundance, \cite{meyerholt2015} also advocate for the dynamic simulation of the distribution of N fixing plants within ESMs, as opposed to the assumption that maximum N fixing capacity if static in space and time.
+
+Given that the response of GPP is almost by definition non-linear with respect to LAI, (given the saturating properties of light absorbtion in increasingly dense canopies) we anticipate that the impacts of initial state will be similar in most land surface models.
+While it is perhaps not a novel point, the degree of water limitation should act to modulate CO$_{2}$ fertilization while N fertilization will of course be stronger in areas that are more N limited in the first instance.
+
+We find that the model parameters with substantial control over fertilization responses are dissimilar to those with greatest control over the initial model state.
+However, the CLM5 utilizes a highly simplified fixed-fraction plant allocation scheme, wherein alternative parameterizations can generate very highly divergent ecosystem properties, and further, the allocation fractions are not easily constrained by real-world estimates which vary with plant size.
+Other models with alternative allocation schemes might not be so sensitive to these parameterizations.
+Nonetheless, the majority of the parameterizations affecting the transient responses had limited impact on the pre-fertilization state.
+This suggests that model behaviour in response to rapid change might indeed be time-scale dependent, and perhaps requires further analysis.
+
+\subsection{Caveats and future directions}
+Many kinds of further analyses are of course possible, including ramping transient simulations,  including global feedback analyses with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constrains of the parameter ranges used.
+We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.
+
 Here we investigate the parameter space of the model using step-change experiments.
 These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions.
-
-The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2018}, \cite{lawrence2018}) for assessing model skill across a very broad range of model components.
-In principle, integrating more targeted datasets and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
-Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
 
@@
 
+In principle, integrating more targeted datasets and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
+The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2018}, \cite{lawrence2018}) for assessing model skill across a very broad range of model components.
+Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
+
 \section{Conclusions}
-In this paper we present an analysis of a suite of parameters relevant to vegetation-mediate biogeochemical feedbacks in the extensively modified nitrogen cycle of the Community Land Model, version 5.
+In this paper we present an analysis of a suite of parameters relevant to vegetation-mediated biogeochemical feedbacks in the extensively modified Community Land Model, version 5.
 We illustrate how responses of the model to elevated carbon dioxide and nitrogen deposition might be impacted by parameter choice.
@@
 In this analysis, we illustrate the degree to which a multitude of individual parameters can potentially have important and interacting impacts on model results, how these results differ across sites and output variables and with initial vegetation conditions.
-Future studies of process and parameter modification in CLM5 might use these results as a guide to the relative (rather than absolute) importance of individual focus parameters and processes.
+Future studies of process and parameter modification in CLM5 might use these results as a guide to the relative importance of individual focus parameters and processes.
 We thus hope to move forwards from studies that modify individual processes in isolation to those that take a more holistic view of the complexity inherent in modeling coupled land surface systems.
@@
 We have attempted here to illustrate and understand the connections between the structure and parameterization of the CLM5 and its biogeochemical functionality.
-This necessarily complex exercise represents one small part of the assessment of CMIP6-era representations of the land surface needed to understand how advances in model structures alter our predictions of land surface functionality.
+This necessarily complex exercise represents one small part of the assessment of CMIP6-era representations of the land surface required to understand how advances in model structures alter our predictions of land surface functionality.
 The challenges of comprehending, measuring and constraining the parametric and structural uncertainty of increasingly complex process-based and land surface models are daunting.
@@
 In atmospheric sciences, for example, 'internal variability' caused by initial conditions is now routinely sampled, given its major impact on even long timescale forecasts (\cite{kay2015}).
-\cite{bonan2018} illustrate how initial conditions are unimportant for carbon cycle prediction on land, relative to structural and parametric uncertainty.
-Given this, sampling the important dimensions of parametric uncertainty in global Earth system model predictions should be taken seriously during future assessments of biogeochemical feedback strength.
+\cite{bonan2018} illustrate how initial conditions are relatively unimportant for carbon cycle prediction on land, relative to structural and parametric uncertainty.
+Given this, sampling the important dimensions of parametric uncertainty that govern biospheric feedbacks in the climate system should be taken seriously when planning future assessments of biogeochemical feedback strength.
 

### 2019-07-05 main_revision.tex

@@
 %tentative author list...
-\authors{Rosie A. Fisher\affil{1},
-William R. Wieder\affil{1,2},
+\authors{Rosie A. Fisher\affil{1,1},
+William R. Wieder\affil{1,3},
 Benjamin M. Sanderson\affil{3},
@@
 \author{Rosie Fisher}
-\date{November 2018}
+\date{June 2019}
 
@@
 
-\affiliation{2}{Institute of Arctic and Alpine Research, University of Colorado, Boulder Colorado, USA}
+\affiliation{2}{Centre Europ\'een Research et de Formation Avenc\'ee en Calcul Scientifique, Toulouse, France}
 
-\affiliation{3}{Centre Europ\'een Research et de Formation Avenc\'ee en Calcul Scientifique, Toulouse, France}
+\affiliation{3}{Institute of Arctic and Alpine Research, University of Colorado, Boulder Colorado, USA}
 
@@
 \begin{keypoints}
-\item The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle.
+\item The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycles.\\
 
-\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice, but parameter responses are varied between sites that differ in their pre-fertilization water and nitrogen limitation.
+\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition.
+are sensitive to parameter choice, and parameter responses further vary with pre-fertilization water and nitrogen limitation.
+\\
 
@@
 Future projections of land carbon uptake in Earth System Models are affected by land surface model responses to both CO$_{2}$ and nitrogen fertilization.
-The Community Land Model, version 5, contains a suite of modifications to carbon and nitrogen cycle representation, and globally, has a lower nitrogen response and higher CO$_{2}$ response than it's predecessors.
-To improve our understanding of the controls over the fertilization responses of the new model, we assess the sensitivity of the model to a suite of thirteen parameters pertinent to the cycling of carbon and nitrogen by vegetation.
-We assess the parameter dependence of the model state, and also the responses of the system to CO$_{2}$ and nitrogen fertilization at individual sites in temperate, tropical and boreal systems.
+The Community Land Model (CLM), version 5, contains a suite of modifications to carbon and nitrogen cycle representation.
+Globally, the CLM5 has a larger CO$_{2}$ responses and smaller nitrogen response than its predecessors.
+To improve our understanding of the controls over the fertilization responses of the new model, we assess sensitivity to eight parameters pertinent to the cycling of carbon and nitrogen by vegetation, both under present-day conditions and with CO$_{2}$ and nitrogen fertilization.
 
- The impact of fertilization is dependant both on parametric variation explored here varies with the balance of limiting factors (water, temperature, nutrients, light) in the pre-fertilization model state.
-The model parameters that impact the pre-fertilization state are in general not the same as those that control fertilization responses, meaning that calibration to present-day conditions will not necessarily constrain future projections.
-Where pre-fertilization state has low leaf area, fertilization induced increases in leaf production amplify the model response to the initial fertilization.
-Model responses to CO$_{2}$ and N fertilization are strongly impacted by how much plant communities can increase their fixation of nitrogen.
-Further studies focused on how alternative parameters impact biogeochemical function would provide useful compliments to the ongoing CMIP6 process.
+ The impact of fertilization varies with both model parameters and with the balance of limiting factors (water, temperature, nutrients, light) in the pre-fertilization model state.
+The model parameters that impact the pre-fertilization state are in general not the same as those that control fertilization responses, meaning that goodness of fit to present-day conditions does not necessarily imply a constraint on future transient projections.
+Where pre-fertilization state has low leaf area, fertilization-induced increases in leaf production amplify the model response to the initial fertilization via further increases in photosynthesis.
+Model responses to CO$_{2}$ and N fertilization are strongly impacted by how much plant communities can increase their rates of nitrogen fixation, and also directly affected by costs of N extraction from soil and stoichiometric flexibility.
+Illustration of how parameteric flexibility impacts model outputs should help inform the interpretation of carbon-climate feedbacks estimated by in fully-coupled Earth system model simulations.
 \end{abstract}
@@
 \section{Introduction}
-Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}).
-In the CMIP5 multi-model inter-comparison, one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth.
-In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}), leading to speculation that the addition of nitrogen limitation as a more general feature of models might reduce the capacity for vegetation to respond to CO2 fertilization and hence reduce the degree to which the terrestrial biosphere can mitigate increasing atmospheric CO$_{2}$.
+The response of the terrestrial biosphere to carbon dioxide enrichment and increasing nitrogen (N) deposition represents a major source of uncertainty in predictions of the evolution of the climate system (\cite{friedlingstein2006},  \cite{arora2013}, \cite{friedlingstein2014}).
+In the CMIP5 multi-model inter-comparison, the smallest response to CO$_{2}$ forcing was predicted by the Community Land Model v4 (CLM4, \cite{thornton2007}, \cite{lawrence2011}), the only model in that comparison that represented the impacts of nitrogen availability on plant growth, leading to suggesting that the addition of nitrogen limitation might substantially reduce the capacity for vegetation to respond to CO$_{2}$ fertilization (and hence reduce the degree to which the terrestrial biosphere can mitigate increasing atmospheric CO$_{2}$.
 
-The newly-released CLM version 5  (CLM5,\cite{lawrence2018}) includes numerous major updates relative to CLM4 and CLM4.5 (\cite{lawrence2011}).
-The introduction of several new dynamical features, including substantial modification of the vegetation nitrogen cycling, as well as the introduction of plant hydrodynamics, altered surface hydrology and dynamic crop representation, has implications for responses to environmental forcing.
+The newly-released CLM version 5  (CLM5, \cite{lawrence2018}) includes numerous updates relative to the CLM4 and CLM4.5  (\cite{oleson2013} ).
+The introduction of several new dynamical features, including substantial modification of the vegetation nitrogen cycling, a new plant hydrodynamics scheme (\cite{kennedy2019}), altered surface hydrology, snow physics and dynamic crop representation, has implications for responses to environmental forcing.
 \cite{wieder2019} report the global carbon cycle implications of these changes and note that, in CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors, with both fertilization responses being closer to experimental results than the previous versions of the model.
 
-Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global biogeochemical feedback analysis, including the controls on global carbon fertilization `$\beta$' values (the change in ecosystem carbon storage in response to CO$_{2}$ fertilization) (\cite{arora2013}, \cite{friedlingstein2014}), we consider the illustration of how parametric choices in the CLM5 might affect and be responsible for overall behaviour in larger-scale simulations to be an important feature of the documentation of the new model.
+Large quantities of analysis and computing resources will likely be devoted to assessing the impacts of transient CO$_{2}$ and N on CMIP class models, such as CLM5, in the context of global biogeochemical feedback analysis (e.g.
+\cite{arora2013}, \cite{friedlingstein2014}).
+Land surface models, in particular the elements related to biogeochemical cycling, are nonetheless subject to very high degrees of parametric and structural uncertainty (\cite{bonan2018}, \cite{zaehle2014}) Given this,  we consider that the illustration of how parametric choices affect the responses to biogeochemical forcing in such models is an important feature of model documentation.
+
+might affect and be responsible for overall behaviour in larger-scale simulations to be an important feature of the documentation of the new model.
 

### 2019-07-05 referee_responses.tex

@@
 
-\textsf{AR: Modified the Wieder reference to read: \emph{ollowing \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.}}
+\textsf{AR: Modified the Wieder reference to read: \emph{Following Wieder et al.
+[2019], who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998."} The main point here being that the Wieder paper only uses the default model, whereas here we want to probe the parameter space around the default model responses.
+}
 
@@
 
-\textsf{AR: One is red and the other is a purple.  We have modified the color map to distinguish better between these two colours.}
+\textsf{AR: On reducing the number of parameters,  we were able to modify the color map to distinguish the colors better.}
 
@@
 \textsf{AR: The point we're trying to make here is that we find that there is no saturation with increasing leafN content, which is consistent with an analysis in the Lawrence paper that finds low Vcmax compared to observations.
-We have changed the text to try and make it clearer that this is a point of consensus not contradiction.}
+We have changed the text to try and make it clearer that this is a point of consensus not contradiction.
+It now reads \emph{"GPP continues to increase almost linearly with declining \emph{leafcn}  (increasing $N_{leaf}$) at all sites except HVF, indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.
+This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.
+"}}
 
@@
 \textsf{AR: OK, this does need clarifying.
-Actually, CLM5 can't be run without its N cycle.
-We have found in subsequent analyses that the increase in LAI with either CO2 or n can have large impacts on the response ratios, with the instantaneous response differing substantially from the longer term response on account of increases in leaf area.
-At Caxiuana there is a 50 \% increase in LAI on account of the CO2 fertilization, which is on top of the 33\% increase in GPP in the first year, for example.
-In the new version of the paper, we focus more closely on the time dependance and LAI interactions of the results.
-**}
+The large response ratios are generally to do with the LAI/GPP feedback, which, as we now discuss at much greater length, increases the fertilization effect over the initial stimulation.
+At Caxiuana, for example, there is a 50 \% increase in LAI on account of the CO2 fertilization, which is on top of the 33\% increase in GPP in the first year, for example.
+Because many of these plots have quite low LAI in the first instance, this is quite notable in our simulations.
+We now discuss this point in greater depth with reference to the new time-series figure.
+CLM5 cannot be run without an N cycle at this point.
+}
 
-- Line 701: Or that global maps are a poor way to see how comparable models and observations are ... 
+- Line 701: Or that global maps are a poor way to see how comparable models and observations are...
+
 \textsf{AR: Or that, yes. We don't really have enough here to say that the global maps are off though, so I'm going to desist from stating that.}
@@
 
-This study presents an introduction to some new modifications to the N cycle representation in CLM5 model, investigates how the model state and the responses of the system to CO2 and N addition are sensitive to parameter choices, and showed that parameters controlling N fixation dominate fertilization responses.
+This study presents an introduction to some new modifications to the N cycle representation in CLM5 model, investigates how the model state and the responses of the system to CO$_{2}$ and N addition are sensitive to parameter choices, and showed that parameters controlling N fixation dominate fertilization responses.
 
@@
 Therefore, how these parameters change within a given climate space is somewhat pre-defined.
-You already suggested that these parameters are influential to plant responses to CO2 and N.
+You already suggested that these parameters are influential to plant responses to CO$_{2}$ and N.
 Once you calibrated them against a PFT, does it mean that you have significantly reduced degree of freedom in your current testing?
@@
 The first point to note is that, as we discussed in L743 onwards, we did not in fact calibrate the model either at this site, nor using a global inverse calibration method, on account of the numerous difficulties encountered in that exercise.
-We are pleased that the referee considers these points to be in scope and have moved the discussion of the inverse calibration exercise to the methods section (within the discussion of the perturbed parameter ensemble)}
+We are pleased that the referee considers these points to be in scope and have moved the discussion of the inverse calibration exercise to the methods section (within the discussion of the perturbed parameter ensemble).}
 
@@
 
-\textsf{AR: The referee suggests that the section headings are poorly balanced, and as such we have modified the organization so that section 3.1 deals with the system state with 3.1.1.
+\textsf{AR: The referee correctly suggests that the results section sub headings were poorly balanced.
+As such we have modified the organization so that section 3.1 deals with the system state with 3.1.1.
 for site differences and 3.1.2 for parametric differences.
 Section 3.2 deals with the fertilization responses, with 3.2.1 for overall responses and 3.2.2 for parametric controls.
+Further, we have modified the }
 
-\textsf{AR: The referee further questions whether this message is new, and also suggests that it is important to identify how model behaviours are influenced by parameter decisions, which is the objective of this paper.
-We do not suggest that this particular message is new.
-Our model is a very heavily modified version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project - all using the default configuration.
-In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.
+\textsf{The referee further questions whether the message in the paper is new, suggesting that it is somewhat self-evident that parameter changes will impact the output of land surface models.
+We concur that this would indeed not be a novel point, but also argue that the dominant purpose of our paper is to explore \emph{how} model behaviours are influenced by parameter decisions, how these effects interact with model initial conditions, and whether parameters affecting the system state will also have an impact on fertilization responses.
+These additional insights are the focus of this paper.
+}
 
-We agree, however, that the results section in the first version of the manuscript was dense and difficult to follow.
+\textsf{The CLM5 model is a very heavily modified version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project - all using the default configuration.
+In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.}
+
+\textsf{We concede, however, that the results section in the first version of the manuscript was dense and difficult to follow.
 We have attempted to remedy this both by reducing the number of parameters considered from 13 to 8, and also by attempting to streamline the text reporting of the results to focus on the features of the output of highest importance for understanding model sensitivity.
-We hope this is an easier read than the prior version.
-}
+We hope this it is now easier to digest than the prior version.}
 
@@
 
-\textsf{AR: We are not sure how it follows that the model output space being complex should lead to a list of strengths and weaknesses of the model.
-Indeed, assessing the validity of the process descriptions of the model is something we specifically aim to not do.
-We agree that the abstract is perhaps too general, **}}
+\textsf{AR: We agree that the abstract is perhaps too general, and have re-written the majority of the text therein to focus on the major points that emerge in the discussion.
+Further, we have added a section in the discussion which focuses on the model deficiencies that are specifically revealed by the senstivity analysis itself.
+(Model improvements highlighted by this analysis)}
 
@@
 
-\textsf{AR: \textbf{Maybe we should just ditch the hypotheses?** }}
+\textsf{AR: We agree that the initial hypotheses were vaguely worded.
+We have changed the format of this section to include "motivating questions" as opposed to hypotheses, to reflect the multi-faceted nature of a sensitivity analysis of a complex model.
+The questions are\\ \emph{"1.
+Are parameters that dominate responses to CO$_{2}$ and nitrogen fertilization also dominant over the pre-fertilization model state?\\
+2. How do fertilization impacts and parameter sensitivities vary with site conditions, as expressed via water and nutrient limitations." }}
 
@@
 
-\textsf{AR: The paper differs substantially from the papers in revision, in that it explicitly tests the parameter space of the model.
-The remainder of the papers assess the responses of the default configuration.
-Our aim here is to be transparent about the uncertainties in the model and provide guidance on the interpretation of future results.}
+\textsf{AR: The paper does  differs substantially from the papers in revision, (indeed, the Wieder paper is now online at \url{https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/2018GB006141}.
+That paper looks at the global response to step changes in CO$_{2}$ and N, whereas our manuscript looks at the parameter response of those same step changes, therefore while they investigate similar topics, they are specifically intended to be complimentary.
+All of the other CLM5 papers cited here only consider the default configuration.
+Our aim here is to be transparent about the uncertainties in the model and provide guidance on the interpretation of future results.
+}
+
+\textsf{AR: We agree, nonetheless, that the results and discussion section was in need of simplification and summary, and hope that this version of the paper has improved in this respect.}
 
@@
 L 63 and L1082: De Kauwe.
+
 \textsf{AR: Fixed.}
@@
 L 88, CO2, not CO2
+
 \textsf{AR: Fixed.}
@@
 L 112: You haven't define what a model state you mean, and therefore, this hypothesis is not explicitly testable.
+
 \textsf{AR: Through the new version of the manuscript, we have used 'pre-fertilization state' to describe this condition. }
@@
 L 113 - 117: This is a very long sentence. And it's not easy to follow. Suggest rephrasing.
+
 \textsf{AR: Changed this to \emph{ The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics.
@@
 Indeed, the need for these papers to cross-reference one-another is the main purpose of having them in a special issue.
-Further, the full description of just the new nutrient algorithm runs to many pages, which would add substantially to this already quite long manuscript, and in principle, many other parts of the model are relevant to this analysis.
-If we were to include these descriptions in a way that would allow a complete understanding of the technical details, they would no doubt have to reside as a supplement to this paper, and so,  this would not necessarily provide a greater utility to the reader than the two sources where are already available, Lastly, this would also constitute publishing the same material twice in different peer reviewed manuscripts.
+Further, the full description of just the new nutrient algorithm runs to many tens of pages, which would add substantially to this already quite long manuscript.
+In principle, many other parts of the model are relevant to this analysis.
+If we were to include these descriptions in a way that would allow a complete understanding of the technical details, they would no doubt have to reside as a supplement to this paper.
+Thus, we do not think that this approach would not necessarily provide a greater utility to the reader than the two sources where are already available, Lastly, this would also constitute publishing the same material twice in different peer reviewed manuscripts.
 We have nonetheless re-framed the manuscript so that it is more focused on the CO$_{2}$ and N responses, and not on the nutrient cycling updates, so we feel that this is less necessary than before.}
@@
 L 307 - 309: I am confused here.
-So you identified a list of 13 parameters which you think should have direct mechanistic impacts on responses to CO2 and N fertilization.
+So you identified a list of 13 parameters which you think should have direct mechanistic impacts on responses to CO$_{2}$ and N fertilization.
 Then you performed sensitivity tests around them, and highlighted that the model indeed is sensitive to these parameter decisions.
@@
 \textsf{AR: Our objective here is not simply to illustrate that the parameters do have an impact (we agree that this would be self-evident) but to provide an illustrate of \emph{how} the different parameters impact the model and under what conditions.
-Thus, there is value in doing the simulations subsequent to our narrowing of the parameter set to those elements we considered to have a direct impact on fertilization responses.
-}
+Thus, there is still value in actually doing the simulations subsequent to our narrowing of the parameter set to those elements we considered to have a direct impact on fertilization responses.
+As we describe in the text, it is common practice for land surface modeling studies to modify a single element of a land surface scheme and report the often large modifications to the global carbon/water/nutrient budgets that result.
+One of our aims herein is to place future modifications within the context of how a wide range of parameters impact the output of the CLM.}
 
@@
 L372: remove ).
+
 \textsf{AR: Done}
 
-L420: Not sure how you ended up with 672 runs. 4 sites x 13 parameters x 5 values * 3 treatments = 780. Did I miss something?
+L420: Not sure how you ended up with 672 runs. 4 sites x 13 parameters x 5 values x 3 treatments = 780. Did I miss something?
 
@@
 L 436: "The x-axis" to the x-axis.
+
 \textsf{AR: Modified accordingly}
@@
 L438: Table 1, not table 1.
+
 \textsf{AR: Modified accordingly}
@@
 L 447: remove ).
+
 \textsf{AR: Done}
@@
 L528: Figure 1, not figure 1.
+
 \textsf{: Done. Thanks for catching. }
@@
 L560: the both the?
+
 \textsf{AR: Done. Thanks for catching.}
@@
 L 562: ))?
+
 \textsf{AR: Done. Thanks for catching.}
@@
 L567: Figures 6 to 8.
-\textsf{AR: We are not sure what the suggestion is here. }
+
+\textsf{AR: We were not sure what the suggestion is here. }
 
@@
 This is clearly the key novelty of this research, but I am afraid that I don't see them identified, and discussed enough throughout the text.
+
 \textsf{AR: We have moved the key points here into the abstract, which highlights these issues earlier and makes the abstract less generic, as per an earlier point.}
@@
 L782: leaf nitrogen, not leaf Nitrogen.
+
 \textsf{AR: Done..}
@@
 L811: nitrogen.
+
 \textsf{AR: Done..}
@@
 L829: replace "(/citewieder2019" with proper text.
+
 \textsf{AR: Done.}

### 2019-07-05 main_revision.tex

@@
 
-might affect and be responsible for overall behaviour in larger-scale simulations to be an important feature of the documentation of the new model.
+In this paper, to that end, we describe the new components of the biogeochemical cycling in CLM5, and assess the sensitivity of the CLM5 model parameterization of ecosystem biogeochemistry.
+We use a range of single site simulations, combined with one-at-a-time parameter uncertainty analyses, and CO$_{2}$ and N fertilization experiments, to investigate the dynamics of the model under steady-state and step-change modifications in forcing.
+We further highlight areas of process and parametric uncertainty that appear to require greater attention in future model development and testing cycles.
 
-In this paper, to that end, we describe the new components of the biogeochemical cycling in CLM5, and assess in detail the sensitivity of the CLM5 model parameterization of ecosystem biogeochemistry.
-We use a range of single site simulations, combined with one-at-a-time parameter uncertainty analyses, and CO$_{2}$ and N fertilization experiments, to investigate the dynamics of the model under steady-state and step-change modifications in forcing.
-We also hope to highlight areas of process and parameteric uncertainty that appear to require greater attention in future model development and testing cycles.
-
-Model uncertainty quantification efforts are often framed as an attempt to narrow the parameter space to a small range of 'important' parameters, as-determined by their impact on the model's mean-state predictions (e.g.
-\cite{dietze2014}, \cite{ricciuto2018}).
-Predicted global biogeochemical feedbacks, however, are typically related to the response of the system to forcing, and not necessarily to those system properties that control the mean state.
-Narrowing the range of variables of interest to those that only impact control states and fluxes (and directing more resources to studying those variables) is thus likely to miss important aspects of transient responses under climate change scenarios.
-Similarly, much effort is expended on calibration of models to present-day conditions (\cite{fer2018}, \cite{lu2018}, when their responses to climate forcing might be relatively unconstrained by these efforts.
+The majority of land surface model sensitivity tests, uncertainty quantification efforts and calibration exercises are conducted under present-day conditions (\cite{fer2018}, \cite{lu2018}, \cite{dietze2014}, \cite{ricciuto2018}).
+Predicted global biogeochemical feedbacks, however, are in principle related to the response of the system to forcing, and not necessarily to those system properties that control the mean state.
 Thus, our motivating questions here are:

### 2019-07-05 main_revision.tex
Updates from Overleaf

@@
 \subsection{CLM5}
- Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}.
-An overview of model developments is presented within this special issue by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019}/.
+ Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and ongoing development are housed at \url{https://github.com/ESCOMP/ctsm/}.
+An overview of model developments is presented (within this special issue) by \cite{lawrence2018}, a description of the global nitrogen cycling responses to fertilizaion by \cite{wieder2019}.
 The plant hydraulics model is documented by \cite{kennedy2019}.
@@
 \subsection{Updates to biogeochemistry Cycling in the CLM5}
-The CLM5 biogeochemistry scheme builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}),  and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).
-
- A full technical description of the CLM5 is beyond the scope of this paper and is available as an appendix to \cite{lawrence2018} and also online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}.
+The CLM5 biogeochemistry scheme builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}, \cite{lawrence2011}),  and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{bonan2012}, \cite{koven2013}).
+The full technical description of the CLM5 is available as an appendix to \cite{lawrence2018} and also online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}.
+We refer to this document here, in lieu of repeating large sections of the documentation an as appendix.
 
 In this paper, we focus on the response to CO$_{2}$ and nitrogen deposition, which are primarily mediated by the biogeochemical processes within the model.
-The primary changes to the biogeochemical cycling in CLM5 are within the vegetation nitrogen cycle, wherein the model contains three new prognostic elements that are likely to impact responses to fertilization.
+The primary changes to the biogeochemical cycling in CLM5 are within the vegetation nitrogen cycle.
+In particular, the new model contains three new prognostic elements that are likely to impact responses to fertilization.
 
 Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}).
-Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}).
-This method, however, appears to be a poor means of capturing variation in photosynthetic capacity in time and space and responses to climatic change (\cite{walker2017}).
-There is significant evidence that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
-LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for a given leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.
+Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}) but do not capture observed variation in photosynthetic capacity in response to CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).
+LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for given leaf nitrogen per unit area ($N_{leaf}$) and environmental boundary conditions, assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.
 $V_{c,max}$ is thus primarily controlled by the N per unit leaf area $N_{area}$ and the prevailing meteorology.
-The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was demonstrated by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.
 
 Secondly, the `FUN' (Fixation and Uptake of Nitrogen) model was incorporated into CLM5.
-FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun} and \cite{brzostek2014} and \cite{shi2016}).
+FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun}, \cite{brzostek2014} \& \cite{shi2016}).
 A major issue confronting N cycle models is how to manage 'excess' carbon under conditions where carbon assimilation is faster than that which can be supported by nutrient uptake (\cite{dekauwe2014}).
-Previous versions of CLM model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake.
+Previous versions of CLM managed this by reducin the gross photosynthetic flux down to the level at which growth could be supported by N uptake.
 This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).
-The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation.
+The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{rastetter2001},  \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation.
 Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them.
@@
 \cite{ghimire2016} introduced the capacity for variable tissue C:N ratios, whereby the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).
-The FUN model used in CLM5 was originally conceived in the context of a static tissue C:N ratio.
-Carbon is allocated between growth and N uptake in a manner that exactly tracks the target C:N ratio.
-To allow a variable C:N ratio in the context of FUN, it is necessary to modify how much C is expended on N uptake vs on growth from this default calculation.
-The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include semi-heuristic and/or bounded representations of tissue N adjustments under N limited conditions (\cite{zaehle2010},\cite{ghimire2016}).
-Here we include a placeholder algorithm that modifies C:N ratios with 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content.
+The FUN model used in CLM5, however, was originally conceived in the context of a static tissue C:N ratio.
+Carbon is initially allocated between growth and N uptake in a manner that exactly tracks a target C:N ratio.
+To allow a variable C:N ratio in the context of FUN, it was necessary to modify how much C is expended on N uptake vs on growth from this default calculation.
+The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include semi-heuristic and/or bounded representations of tissue N adjustments under N limited conditions (\cite{zaehle2010}, \cite{ghimire2016}).
+Here we include a placeholder algorithm that modifies C:N ratios as a function of 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content.
 When N is scarce in the environment, less C is spent on uptake, increasing C:N ratios.
 When plants become N limited and C:N ratios increase, however, expenditure increases.
-This algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.
+This algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters the sensitivity of which is investigated here.
 
 \subsection{Sites}
-We utilized four observational flux tower sites, in order to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, 44.5$^{o}$N -121.6$^{o}$E).
+We utilized four observational flux tower sites, in order to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, CAX -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, NWR, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, HVF, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, MET, 44.5$^{o}$N -121.6$^{o}$E).
 Each site was driven with the available site-specific meteorological data.
-Soil texture, color, depth and, plant functional type composition were taken from previous model validation exercises undertaken at these sites \cite{bonan2012} for Metolius and Harvard Forest, \cite{fisher2007} for Caxiuan\~a and \cite{bonan2013} for Niwot Ridge.
-We note here that our use of a small number of single sites cannot fully describe the environmental conditions under which the model is applied.
+Soil texture, color, depth and, plant functional type composition were taken from previous model validation exercises undertaken at these sites \cite{bonan2012} for MET and HVF,  \cite{fisher2007} for CAX and \cite{bonan2013} for NWR.
+Conducting and analysing these simulations globally is beyond the scope of this study.
 Instead, our aim is to provide detailed illustrations of how model parameter choices impact a model outputs under a range of realistic forcing conditions.
-Conducting and analysing these simulations globally is beyond the scope of this study, but investigations of the response of global biogeochemical feedback strength to the important dimensions of parameter space identified by this investigation will provide the basis of future studies.
 
@@
 Restarting from the state at the end of the perturbed parameter spin-ups, we ran 'control' transient simulations for each site and parameter combination starting in 1760 and invoking historical transient CO$_{2}$ and N deposition until 2018.
-
 For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation.
@@
 Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization.
-These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in its own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.
+These parameters are narrowed down from the much larger complete set of CLM5 parameters using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.
 We focus here on biogeochemically important inputs.
-Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme are discussed in \cite{kennedy2019}.
+Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme are discussed in \cite{kennedy2019}.
 
@@
 At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below.
-We recognize that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses.
-We concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing.
+We note that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses, but concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing.
+
 \subsection{Default parameter determination}
-During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products.
-We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystem to die under lower pre-industrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting effect that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets.
-The parameterization used in the default version of the model consequentially utilized the mean trait values where possible from trait datasets (\cite{lawrence2018}) in combination with targeted manual tuning of parameters (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states.
-We recognize that this more subjective calibration method is less than ideal, and ongoing studies will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}.
+During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model parameter space using neural-network derived emulators.
+We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the parameter space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystems to die under lower pre-industrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, resulting in marginal areas with many dead grid cells, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed.
+The parameterization used in the default version of the model consequentially was derived from the mean observed trait values where possible, in combination with targeted manual tuning of parameters (including allocation parameters and nitrogen costs) that were less well constrained by data, where excessively unrealistic values for simulated states were detected.
+We recognize that this more subjective calibration method is less than ideal.
+Ongoing studies will focus on more limited-scope calibration attempts to overcome the difficulties described (e.g.
+\cite{fer2018}).
 
@@
 
-\textbf{Specific Leaf Area}, \emph{`SLA'}, is a parameter common to almost all land surface models, which determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass.
-High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies and greater interception of light.
-In CLM5 high SLA leaves also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.
+\textbf{Specific Leaf Area}, \emph{`SLA'} determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass.
+High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies.
+In CLM5, high SLA leaves also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.
 
-\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf biomass.
-In CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow, and the cost of N acquisition in the FUN model, respectively.
-Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.
-The carbon allocation in CLM5 uses a fixed-fraction approach, with constant ratios between, leaf, fine roots, stem and coarse root allocation.
-As such, the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance.
-Newer codes within the CLM framework (e.g.
-the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}m \cite{fisher2018vegetation}) account for this issue by removing leaves in negative carbon balance.
+\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf biomass.
+In CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow, and the cost of N acquisition, respectively.
+Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition while higher values imply greater construction and maintenance costs.
 
@@
 Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass.
-Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter).
+Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter.
 
 \textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.
-The FUN model introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake.
-Therefore, uncertainties associated with the cost functions, in particular, are large.
+FUN uses a bulk approach to account for the various costs to plants of low environmental N availability, including carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake.
+This abstract means that, uncertainties associated with the cost functions, in particular, are large.
 These six parameters were defined and calibrated by \cite{brzostek2014}.
 In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}.
-The six \emph{'n\_costs'} parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
+The six \emph{'n\_costs'} parameters include, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.
 
@@
 CLM5 uses the concept of a 'target' leaf C:N ratio, around which flexibility is allowed, given the environmental costs of Nitrogen.
-As with most land surface schemes, higher leaf N concentrations lead to both higher maintenance respiration costs and higher photosynthetic capacity (in this case via the LUNA model).
+As with most land surface schemes, higher leaf N concentrations lead to both higher maintenance respiration costs and higher photosynthetic capacity.
 
@@
 The slope parameter determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit, in the absence of plant hydraulics limitations.
-Thus, is it intrinsically linked to the response of ecosystems to CO$_{2}$.
+It is included here on account of it's critical impact on stomatal responses to CO$_{2}$.
 
-\textbf{C:N flexibility }, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c},  as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.
-Only the thid parameter (c) had an important impact in this sensitivity analysis, and so we focus here on the impact of that parameter.
+\textbf{C:N flexibility }, the \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, parameters  as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.
+Only the third parameter (c) had an important impact in this sensitivity analysis, and so we focus here on the impact of that parameter.
 
@@
 \section{Results}
-Our analysis used four sites, with eight parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.
-We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen per unit area ($N_{leaf}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$).
-We did not look at soil output variables as our parameter analysis is focused on vegetation responses.
-
-In this section we assess first the impacts of parametric variation on the system state and second the responses to carbon and nitrogen fertilization.
-We do not focus on the goodness of fit to observations, because our aim is to discern the primary parametric controls over CO$_{2}$ and N fertilization responses, and including extensive data-fitting and calibration experiments (including on parameters mostly unrelated to C and N cycling) would detract from this goal.
 
 \subsection{Overall system state}
-The parameter responses of the site-level water and nitrogen limitation (see below) in the control (unfertilized) parameter ensemble are shown in figure \ref{btran state}.
-GPP, NPP, LAI, $N_{leaf}$ and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites.
+In this section we assess first the impacts of parametric variation on the pre-fertilization system state.
+The parameter responses of the site-level water and nitrogen limitation (see below) in the pre-fertilization parameter ensemble are shown in Fig.
+\ref{btran state}.
+GPP, NPP, LAI and $N_{leaf}$ in Fig.\ref{state}.
 In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}.
-Lines of differing colors illustrate responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case.
+Lines of differing colors illustrate responses to variation in the target parameters, with the central crossover point in each figure corresponding to the default case.
 
@@
 In this section, we discuss the impacts of the parameter variation on the initial model state.
-Many parameters have both first-order direct impacts, and then further impacts via other parts of the model.
+Many parameters have both first-order direct impacts, and potential indirect effects via modification of the initial state of the system or on other elements of the model.
 Here we attempt to illustrate how these mechanisms operate in CLM5 within the parameter space explored here.
 
-At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on $N_{leaf}$ (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities.
+Specific leaf area (\emph{slatop}) has a large first-order impact on LAI , and on $N_{leaf}$ (Figure \ref{state}), at all sites.
+\emph{slatop} it is used directly in the calculation of both of these quantities.
 The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf.
@@
 
-The root:leaf ratio (\emph{froot\_leaf}) has numerous impacts, including 1) on LAI, via tissue allocation fractions , 2) on NPP, via changes in maintenance costs of fine root pools, 3) on the uptake of water (via root length density) and 4) on nutrient limitation through root-mediated N uptake costs.
-The first two impacts represent the costs of more roots, and the latter two their benefits.
-CAX and HVF exhibit higher LAI with fewer roots.
-NWR has large reduction in LAI (and GPP \& NPP) with fewer roots.
+Root:leaf ratio (\emph{froot\_leaf}) has numerous impacts, including 1) on LAI, via tissue allocation fractions, 2) on NPP, via changes in maintenance costs of fine root pools, 3) on the uptake of water, via root length density and 4) on nutrient limitation, via root-mediated N uptake costs.
+The first two impacts represent the costs of roots, and the latter two their benefits.
+CAX and HVF exhibit higher LAI with lower root allocation, whereas NWR has large reduction in LAI (and GPP \& NPP) with lower root allocation.
 MET exhibits a mixed response crossing an optimal value.
-\emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}) indicating the impact of mechanism '3' here, for very low root densities.
+\emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}) illustrating the impact of low root density on water uptake capacity.
 
-Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon.
-At NWR, the parameter range crosses a local maximum for vegetation carbon, since increases in NPP are non-linear with respect to leaf investment, resulting in lower vegetation biomass overall.
-At all other sites, the impact of increasing leaf allocation is to invoke large increases in LAI and NPP.
-Lower stem allocation and higher LAI increase the N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).
+Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon (.
+The impact of increasing leaf allocation is to invoke large increases in LAI and NPP.
+Lower stem allocation and higher LAI also increase N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).
 

### 2019-08-30 SIfigs.tex

--- 2019-05-28 SIfigs.tex
@@
 
+\title{Supplementary Information for \emph{Parametric controls on vegetation responses to biogeochemical forcing in the CLM5} by Fisher et al. (submitted to JAMES)}
+
     \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Caxiuan\~a }
-     \label{CAX celev}
+     \centering
+     \includegraphics[width=1.0\textwidth, left]{matlab/figures/MAY19jp_MEANCOND_-r200_2013TOTSOMN.png}
+     \caption{Influence of parametric variation   (over the range tested: -1 to +1, Table \ref{table_ranges}) on the model pre-fertilization state for Nitrogen fixation and total soil N content across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
+     \label{MEANSOM}
 \end{figure}
@@
 \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_Br-cax_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 uears of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Caxiuan\~a}
-     \label{CAX ndep}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_TOTVEGC_y2013.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table 1) on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5 Kg m$^{-2}$ y$^{-1}$ N fertilization, for net primary production (NPP)}
+     \label{CN_NPP}
   \end{figure}
@@
     \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Harvard Forest}
-     \label{HRV celev}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_TOTVEGC_y2013.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table 1) on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5 Kg m$^{-2}$ y$^{-1}$ N fertilization, for total vegetation carbon)}
+     \label{CN_TOTVEGC}
   \end{figure}
@@
 \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US_Ha1_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Harvard Forest}
-     \label{HRV ndep}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_NFIX_y2013.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table 1) on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5 Kg m$^{-2}$ y$^{-1}$ N fertilization, for nitrogen fixation rates (NFIX, gN m^{-2})}
+     \label{CN_NFIX}
   \end{figure}
@@
   \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth, left]{matlab/figures/NOVc_FERT_1CLM50defpft_celev_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of CO$_{2}$ fertilization (550ppm) at Metolius Forest}
-     \label{MET celev}
+     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_LEAFN_y2013.png}
+     \caption{Influence of parametric variation (over the range tested: -1 to +1, Table 1) on the model response (fertilized/control) to to 15 years of 500ppm CO$_{2}$ and +5 Kg m$^{-2}$ y$^{-1}$ N fertilization, for leaf nitrogen ($N_{leaf}$)}
+     \label{CN_LEAFN}
   \end{figure}
 
-  \begin{figure}[h]
-    % \centering\includegraphics[width=0.5\textwidth, right]{lion-lo9
-     \includegraphics[width=1.3\textwidth,left]{matlab/figures/NOVc_FERT_1CLM50defpft_ndep_1x1pt_US-Me2_ens_MIC_p1_2012.png}
-     \caption{Influence of parametric variation (over the range tested: -1 to +1) on the influence of 15 years of Nitrogen fertilization (+5 kgN m$^{2}$ y$^{-1}$) at Metolius Forest}
-     \label{MET ndep}
-  \end{figure}
-
-    \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_TOTVEGC1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on vegetation carbon, across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{VEGC CO2 and N respones 2001}
-  \end{figure}
-
-  \begin{figure}[h]
-     \centering
-     \includegraphics[width=1.55\textwidth, left]{matlab/figures/NOVc_CNdep_LEAFN1__p2012.png}
-     \caption{Influence of parametric variation on the impact of 15 years of CO$_{2}$ and N fertilization on leaf nitrogen content ($N_{area}$), across the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.}
-     \label{Leaf N CO2 and N respones 2001}
-  \end{figure}
 \end{document}

### 2019-08-30 main_revision.tex

@@
 
-\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition.
-are sensitive to parameter choice, and parameter responses further vary with pre-fertilization water and nitrogen limitation.
-\\
+\item The control state, and responses to CO$_{2}$ and nitrogen, are sensitive to parameter choice, and to baseline water and nitrogen limitation. \\
 
-\item Parameters controlling nitrogen fixation rates dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. 
+\item Parameters controlling nitrogen fixation rates dominate fertilization responses, highlighting the need for increased scrutiny of this process.  
 \end{keypoints}
@@
 
-Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon (.
-The impact of increasing leaf allocation is to invoke large increases in LAI and NPP.
-Lower stem allocation and higher LAI also increase N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).
+Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon (SI Fig.
+3).
+The impact of increasing LAI is typically to increase light interception and hence GPP, thus these responses are typically quite large for most sites.
+Lower stem allocation and higher LAI also increase N demand, reducing $N_{leaf}$ and increase PNU at NWR and MET (Fig.
+\ref{btran state}).
 
 Increasing $N_{costs}$ increases the respiration losses associated with nitrogen uptake.
-All sites showed a moderate decline in LAI, NPP, GPP and vegetation carbon when costs were high, but lesser impacts when costs were reduced (Figures \ref{NR1 state}, \ref{CAX state} \ref{HVF state}, \ref{MET state}).
+All sites showed a moderate decline in LAI, NPP, GPP and vegetation carbon when costs were high, but lesser impacts when costs were reduced (Fig.
+\ref{state}).
 In contrast to the allocation parameters, $N_{costs}$ does not appear to be a first-order control over the pre-fertilization system state.
@@
 As is the case for $N_{costs}$, \emph{fracfixers} does not appear to be a first-order control over the pre-fertilization system state.
+Interestingly, SI Fig.
+1 illustrates that while \emph{fracfixers} had a large impact on the N fixation rates, this did not actually affect the overall soil Nitrogen content, implying that other factors were dominant over the equilibrium soil nutrient budget.
 
@@
 Higher \emph{leafcn} values decrease $N_{leaf}$, decreasing photosynthetic capacity and GPP.
-In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero (Figures \ref{HVF state}, \ref{MET state}).
+In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero (Fig.
+\ref{state}).
 GPP continues to increase almost linearly with declining \emph{leafcn}  (increasing $N_{leaf}$) at all sites except HVF, indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.
@@
 
-For CO$_{2}$ fertilization at CAX the initial stimulation of GPP ($\sim$1.37) lead to an almost 50\% increase in LAI.
-At this site, water limitations (lower $B_{tran,mn}$), prevented any further increase in GPP.
+ At MET, for example, the initial GPP stimulation by CO$_{2}$ (26\%) resulted in a slow but prolonged LAI increase, (given the lower overall carbon turnover at this colder site).
+At 80 years, LAI had increased by 50\%, amplifying the much GPP and NPP stimulation to 53-69\% and 49-71\% respectively, (varying through the cycling climate forcing data).
+For CO$_{2}$ fertilization at CAX, in contrast, the initial stimulation of GPP ($\sim$1.37) lead to an almost 50\% increase in LAI, but at this site, water limitations (lower $B_{tran,mn}$), prevented any further increase in GPP.
+At HVF, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (HVF has a saturating response to decreasing \emph{leafcn},  Fig.
+\ref{state}).
+NWR had moderate (25\%) increases in initial GPP and LAI ($sim$30\%).
+ The stimulation of GPP via LAI here is typically greater than is typically reported in elevated CO$_{2}$ experiments, but \cite{king2005} report a large response to $sim$530 ppm CO$_{2}$ in Aspen stands, with increases in LAI and biomass of around 60\%.
 
- At MET, in contrast, initial GPP stimulation was lower (26\%) on account of the lesser water stress.
-This resulted in slow but prolonged LAI increase, given the lower overall carbon turnover at this colder site.
-The 50\% increase in carbon at 80 years had, however, a much larger impact on GPP and NPP giving a stimulation of 53-69\% and 49-71\% respectively, (varying through the cycling climate forcing data).
-Thus, the overall impact of fertilization appears much greater than is typically reported in elevated CO$_{2}$ experiments, but \cite{king2005} report both large response to $sim$530 ppm CO$_{2}$ in Aspen stands, with increases in LAI and biomass of around 60\%.
-
- At HVF, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (HVF has a smaller response to decreasing higher leaf N ratios, Fig.
-\ref{HVF state}).
-NWR had moderate (25\%) increases in initial GPP and LAI ($sim$30\%).
-At all sites, the fraction of NPP spent on N uptake increased, but most dramatically at Metolius, where initial N expenditure was lowest initially on account of low growth rates.
+ At all sites, the percentage of NPP spent on N uptake (PNU) increased, but most dramatically at MET, where initial N expenditure was lowest initially on account of low growth rates.
+Further, N fixation rates increased markedly at all sites (Fig.
+\ref{timescales}) from their default values as shown in SI Fig 1.
+An increase of 2-3 was found at HVF and NWR.
+At MET, the very low initial rates (0.01 gNm$^{-2}$y$^{-1}$) mean that the amplification was extremely large and thus omitted from the plot.
+At CAX, fixation rates initially increased and then declined through time in tandem with increases in soil Nitrogen.
+Soil N in general was not in equilibrium after 85 years of simulation time, but this did not appear to greatly affect the fixation rates at NWR and HVF.
+The soil N dynamics observed raise interesting questions about potential transient responses of CLM5 which will be covered in future analyses.
 
@@
 In particular, for CAX, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%.
+At all sites, addition of N caused a radical drop in the rates of N fixation, and gradual, non-equilibrated increases in the soil N content.
 
@@
 Note that each figure is a different scale to best illustrate the parametric impacts.
-Corresponding figures for NPP, vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1-3).
-Figures for each site showing the individual N and CO$_{2}$ responses are also included in the SI (Figures 3 to 10).
+Corresponding figures for NPP, vegetation carbon, N fixation and $N_{area}$ are in the supplemental information (SI Figures 2-5)..
 
@@
 
-The impact of \emph{slatop} on fertilization was different at every site (figs.
-\ref{GPP_CN}, figures \ref{LAI_CN}).
-For MET, \emph{slatop} was the most important control over N responses, with higher values allowing a greater LAI/GPP feedback in response to the N fertilization.
-At NWR, the opposite behaviour was found, with high SLA reducing the impact of N in cases where the pre-fertilization LAI meant that a saturating response of GPP was found.
+The impact of \emph{slatop} on fertilization was different at every site (Figs.
+\ref{GPP_CN} \&  \ref{LAI_CN}).
+For MET, \emph{slatop} was the most important control over N responses, with higher values allowing a greater (or faster) LAI/GPP feedback in response to N fertilization.
+At NWR, the opposite behaviour was found, where high SLA reduces the impact of N.
+In such cases, high pre-fertilization LAI meant that the response of GPP was more saturated, and so the feedback from increased leaf area was dampened.
 At HVF, higher values of \emph{slatop} slightly reduced both CO$_{2}$ and N fertilization for the same reason (although the range of variation at HVF is only 5\% between the highest and lowest values, as opposed to 40\% at NWR).
@@
 \emph{froot\_leaf} had contrasting impacts at different sites.
-At MET,  low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions.
-At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the respnse to \emph{slatop}.
+At MET,  low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions (Fig.
+\ref{btran state}).
+At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the response to \emph{slatop}.
 
-Low \emph{stem\_leaf} ratio caused lower responses to N fertilization CAX, NWR and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the NPP-LAI non-linearity issue.
-At NWR,  N limitation restricted CO$_{2}$ responses, but at CAX and NWR the same type of response was seen.
-At MET, lower \emph{stem\_leaf} allowed greater responses, plausibly from increasing the speed of the NPP/LAI feedback.
+Low \emph{stem\_leaf} ratio caused lower responses to N and CO$_{2}$ fertilization CAX and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the NPP-LAI non-linearity issue.
+At NWR,  N limitation restricted CO$_{2}$ responses, nullifying the impact of \emph{stem\_leaf}.
+In contrast, MET, lower \emph{stem\_leaf} allowed greater responses to N plausibly from increasing the speed of the NPP/LAI feedback.
 
 Higher $N_{costs}$ increase the responsiveness of the system to N at CAX, NWT and HVF.
-The magnitude of the impact on N fertilization ism as might be expected, related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).
-Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF **.
+The magnitude of the impact on N fertilization is tracks the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).
+Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF.
+These configurations potentially are limited in their ability to direct carbon to fixation pathways, since estimated soil N uptake costs remain low.
 
 \emph{frac\_fixers} (the maximum fraction of NPP available for N fixation) had a particularly dominant role over the model response to step changes in N and to CO$_{2}$.
-For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP_CN}, \ref{LAI_CN}).
+For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Fig.
+\ref{GPP_CN} \& \ref{LAI_CN}).
 At MET,  the changing fixation capacity had no impact on the response of GPP or LAI to N fertilization.
@@
 The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state.
-Complex responses were illustrated, reflecting the differential costs and benefits of Ni on the system.
-For N fertilization, at CAX and NWR, lower values of \emph{leafcn} had smaller responses to N fertilization, reflecting the higher initial LAI of these configurations and the non-linear GPP feedback discusses earlier.
+Complex responses were illustrated, reflecting the differential costs and benefits of \emph{leafcn} on the system.
+For N fertilization, at CAX and NWR, lower values of \emph{leafcn} had smaller responses to N fertilization, reflecting the higher initial LAI of these configurations and the non-linear GPP feedback discussed earlier.
 At CAX, the highest \emph{leafcn} had smaller CO$_{2}$ responses.
-(Figure \ref{btran state} illustrates that this configuration had less water stress, given that both the lower LAI and lower photosynthetic capacity should lead to smaller stomatal conductances and evapotranspiration.
-and thus had less water stress to be alleviated by the high CO$_{2}$.
+(Fig.
+\ref{btran state} illustrates that this configuration had less water stress, (both lower LAI and lower photosynthetic capacity should lead to smaller stomatal conductances and evapotranspiration) and thus had less water stress to be alleviated by the high CO$_{2}$.
 
 Surprisingly, given the importance of \emph{medlyn\_slope} on stomatal responses to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response.
-Earlier studies using a less constrained range of \emph{medlyn\_slope} values indicated that it can have a potentially large impact on system state, (results not shown) but that the fertilization response is typically limited.
-The linear response of the stomatal conductance term to $CO_{2}$ mean that the direct impact of \emph{medlyn\_slope} is likely to be restricted to changes in the pre-fertilization state.
+Earlier studies using a less constrained range of \emph{medlyn\_slope} values indicated that it can have a potentially large impact on system state via impacts on water status, (results not shown) but that its capacity to modify the fertilization response is limited.
+The linear response of the stomatal conductance term to $CO_{2}$ means that the direct impact of \emph{medlyn\_slope} is likely to be restricted to changes in the pre-fertilization state.
 
@@
 Higher values of \emph{cn\_flex\_c} spend less carbon on N uptake when tissue N is lower than the target amount, allowing N ratios to vary more.
-Impacts of this parameter are thus manifested through their impact on $N_{leaf}$ (Figure SI **).
-Smaller values of \emph{cn\_flex\_c} corresponded in all cases to smaller variations in $N_{leaf}$ (values closer to 1.0) under both CO$_{2}$ and N fertilization (For N deposition, $N_{leaf}$ always increases and for CO$_{2}$ it always declines.) Lower values of \emph{cn\_flex\_c} also corresponded to larger CO$_{2}$ stimulation of GPP, illustrating the direct impact of this parameter, as leaves maintain higher photosynthetic capacity under more N limiting conditions.
-The impact on NPP and LAI was less clear, given the corresponding impacts of higher respiration both for maintenance and for N uptake.
-At CAX, NWR and HVF,  the highest values of \emph{cn\_flex\_c} \emph{also} had a higher response to CO$_{2}$ than the default.
+Impacts of this parameter are manifested through their impact on $N_{leaf}$ (SI Fig.
+5).
+For N deposition, $N_{leaf}$ always increases and for CO$_{2}$ it always declines.
+Smaller values of \emph{cn\_flex\_c} corresponded in all cases to smaller variations in $N_{leaf}$ under both CO$_{2}$ and N fertilization.
+Lower values of \emph{cn\_flex\_c} corresponded to larger CO$_{2}$ stimulation of GPP, as these leaves maintain higher photosynthetic capacity under more N limiting conditions.
+The impact on NPP and LAI was less clear, given the consequence of high $N_{leaf}$ for both respiration and N uptake costs.
+At CAX, NWR and HVF,  the very highest values of \emph{cn\_flex\_c} \emph{also} had a higher response to CO$_{2}$ than the default.
 The reasons for this discontinuity are unclear, however, the impact of \emph{cn\_flex\_c} in the CLM5 code is subject to at least two 'caps', (whereby plants cannot spend less than 0.5 or more than 1.0 of the carbon that would retain the target LAI), potentially leading to complex feedbacks once these caps are reached.
 
-\subsection{Low leaf area investigations}
-In general, we found that LAI predicted by the default CLM5 were often low  compared to published estimates for these sites.
-(CAX; 3.0 (modeled) vs.
-5.3 (observed) \cite{fisher2007},  NWR; 1.7 vs.
-4.2 \cite{bowling2009}, HVF; 2.6 vs.
-3.5 \cite{williams1996}  MET; 0.5 vs.
-1.5 (\cite{spadavecchia2011}).
-This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.
-
-To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters).
-These simulations indicated that, with the resulting higher (and more consistent with observed) LAI values, the model was less sensitive to both CO$_{2}$ and N deposition.
-For example, at HVF, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response of  (1.18, vs 1.43 for the control)  and a lower N response as well (1.20 vs 1.26 for the control).
-Similar results were found at all other sites.
-
 \section{Discussion}
-The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}).
+The responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intensive investigations, and vast quantities of computing time will be expended on the release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}).
 Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.
-Interpretation of these simulations should be informed by solid comprehension of which model properties dominate the size of the feedback strength both within and between models, and we intend these illustrative simulations to provide initial guidance of this process.
-
-At the time of writing, we do not yet know the carbon-climate feedback strength of the default version of CLM5/CESM2.
-In offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ and a weaker response to N fertilization than previous versions of the model (CLM4 and CLM4.5, Wieder et al.
+Interpretation of these simulations should be informed by solid comprehension of which model properties dominate the size of the feedback strengths,  both within and between models, and ideally, a sense of how well constrained these properties are.
+We intend these illustrative simulations to provide initial guidance of this process.
+At the time of writing, we do not yet know the carbon-climate feedback strength of the default version of CLM5/CESM2 model.
+In offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ and a weaker response to N fertilization than previous versions of the model (Wieder et al.
 2018).
 Further, the impact of both types of fertilization was found, in all versions of the model, to be highly variable with PFT (from 0\% to 110\% for N, -10\% to +55\% for CO$_{2}$).
-In this study we illustrate how variations in the pre-fertilization state, water and N limitation, and parametric variation combine to generate variation in the responses to forcing.
-While we set out primarily to to determine responses to model parameters, findings concerning the impact of model state on forcing sensitivity will have important implications for the interpretation of simulations with the default model.
+In this study we set out to illustrate how variations in parametric space give rise to responses to forcing.
+While we set out primarily to to determine responses to model parameters, findings concerning the impact of pre-fertilizaion model state, water and N limitation will have important implications for the interpretation of simulations with the default model.
 
@@
 
-Aside from the straightforward direct impacts of the fixation rate on model predictions, however, the model output space was notably complex.
-For example,  in some cases the sign of the impact of individual parameters changes between sites, while other parameters can demonstrate non-linear and counter-intuitive impacts as the result of numerous interacting feedbacks.
+Aside from the straightforward direct impacts of the fixation rate on model predictions, interpretation of other configurations is less straightforward.
+In some cases the sign of the impact of individual parameters changes between sites, while other parameters can demonstrate non-linear and counter-intuitive impacts as the result of numerous interacting feedbacks.
 To summarize our findings, we note a variety of key outcomes of our analysis on the factors controlling responses to CO$_{2}$ and N fertilization in the CLM5, including:
@@
 The degree of water or nutrient limitation in the initial state is also a dominant factor over the model responses to CO$_{2}$ fertilization and N deposition.
-Once again, the direct impact of parameter perturbation is frequently masked by how changes affect the initial limiting conditions.
 Examples of this include how low \emph{froot\_leaf} increases water limitation at NWR, increasing responsiveness to CO$_{2}$ and how high \emph{leafcn} reduces water stress and thus decreases the CO$_{2}$ response at CAX.
@@
 3.
-Other parameters with a direct impact on fertilization responses included those related to the acquisition of nitrogen (\emph{N\_costs}) and the stoichiometric flexibility of the system (\emph{cn\_flex\_c}) illustrate the sensitivity of the model to relatively unconstrained processes within the nitrogen cycle.
+Other parameters with a direct impact on fertilization responses included those related to the acquisition of nitrogen (\emph{N\_costs}) and the stoichiometric flexibility of the system (\emph{cn\_flex\_c}) illustrating the sensitivity of the model to relatively unconstrained processes within the nitrogen cycle.
 
@@
 Many of the parameters that are dominant over the initial state (\emph{slatop}, \emph{stem\_leaf} \emph{leafcn}) either have limited impacts on fertilization responses, or have impacts that are largely traceable to their impacts on the initial model state.
-In contrast, numerous parameters that more directly and uniformly impact fertilization responses, (\emph{frac\_fixers} \emph{N\_costs}) are less important for the initial model state.
-We thus suggest that model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization.
+In contrast, numerous parameters that more directly and uniformly impact fertilization responses, (\emph{frac\_fixers}, \emph{N\_costs}) are less important for the initial model state.
+We thus suggest that model calibration efforts focused on parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization.
 
@@
 The results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites.
-Given the complexity of the model, this helps generate confidence that the model outcomes are derived from the intended process mechanisms, and are not subject to significant deviations from the intent of the model structure (as opposed to .
+Given the complexity of the model, this helps generate confidence that the model outcomes are derived from the intended process mechanisms, and are not subject to significant deviations from the intent of the model structure.
 
@@
 While there are many potential improvements that could be made to any land surface model, this particular assessment of the parameter space of the CLM5 in this study revealed several dynamics which highlight parts of the representation of biogeochemistry which might be improved in future versions.
-Firstly, the response of NPP to increasing \emph{slatop} was often negative, despite increasing GPP, as the fixed-fraction allocation scheme in the model will generate leaves that respire more than they assimilate under very high productivity conditions.
+
+Firstly, response (e.g.
+CAX) of NPP to increasing \emph{slatop} was often negative, despite increasing GPP.
+The fixed-fraction allocation scheme in the model will under some circumstances generate leaves that respire more than they assimilate under very high productivity conditions.
 This highlights the need for either more realistic tissue allocation, or for a scheme which optimizes leaf allocation to prevent leaf layers in negative carbon balance (\cite{fisher2010}).
@@
 
-Thirdly, requiring focus is the dominance of the 'fraction of fixers' over fertilization responses.
+Thirdly, one issue requiring focus is the dominance of the 'fraction of fixers' over fertilization responses.
 \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants.
 Technically, because plant density is not represented in the CLM5, this parameter represents the fraction of net assimilated carbon ($gpp$ - $r_{auto}$) that can be used for symbiotic N fixation.
-Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants.
+Ideally, \emph{frac\_fixers} would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants.
 In general, Nitrogen fixers are known to primarily exist in early-successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).
@@
 \subsection{Applicability of results to other land surface models}
-Our finding that Nitrogen fixation plays a dominant role over responses to $CO_{2}$ fertilization concurs with earlier studies.
+Our finding that Nitrogen fixation plays a dominant role over responses to CO$_{2}$ fertilization concurs with earlier studies.
 Notably, \cite{meyerholt2015} investigate six alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}).
-They emphasize that the formulation which, as in CLM5, switches N sources depending on their relative costs, represent a promising means by which N fixation might be simulated.
-Interestingly, their model also includes a maximum rate of fixation per unit (root) biomass, but analysis to it's sensitivity is not shown.
+They conclude that their optimizing algorithm, which, as in CLM5, switches N sources depending on their relative costs, represents a promising means by which N fixation might be simulated.
+Interestingly, this model also includes a maximum rate of fixation per unit (root) biomass parameter, but analysis to it's sensitivity is not shown.
 Further, given the difficulty of matching observations of complex N cycle processes with more abstract 'big leaf' models such as CLM5, and the ecological complexities governing the abundance their abundance, \cite{meyerholt2015} also advocate for the dynamic simulation of the distribution of N fixing plants within ESMs, as opposed to the assumption that maximum N fixing capacity if static in space and time.
 
-Given that the response of GPP is almost by definition non-linear with respect to LAI, (given the saturating properties of light absorbtion in increasingly dense canopies) we anticipate that the impacts of initial state will be similar in most land surface models.
-While it is perhaps not a novel point, the degree of water limitation should act to modulate CO$_{2}$ fertilization while N fertilization will of course be stronger in areas that are more N limited in the first instance.
+Given that the response of GPP is by definition non-linear with respect to LAI, (given the saturating properties of light absorbtion in increasingly dense canopies) we anticipate that the impacts of initial state will be similar in most land surface models.
+While it is perhaps not a novel point, water limitation should act to modulate CO$_{2}$ fertilization, while N fertilization will of course be stronger in areas that are more N limited in the first instance.
 
@@
 
-\subsection{Caveats and future directions}
-Many kinds of further analyses are of course possible, including ramping transient simulations,  including global feedback analyses with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constrains of the parameter ranges used.
-We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.
-
-Here we investigate the parameter space of the model using step-change experiments.
-These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions.
-
+\subsection{Limitations and future directions}
 In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5.
@@
 
+Many kinds of further analyses are of course possible, including ramping transient simulations,  global simulations with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constraints of the parameter ranges used.
+We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.
+
 In principle, integrating more targeted datasets and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic.
@@
 Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.
+
+It is common practice in land surface modeling literature to present studies that focus on the modification of a single aspect of model structure, process representation or calibration, and to highlight the significance of individual modifications when scaled up to global land-mediated climate feedback processes.
+In this analysis, we illustrate the degree to which a multitude of individual parameters can potentially have important and interacting impacts on model results, how these results differ across sites and output variables and with initial vegetation conditions.
+Future studies of process and parameter modification in CLM5 might use these results as a guide to the relative importance of individual focus parameters and processes.
+We thus hope to move forwards from studies that modify individual processes in isolation to those that take a more holistic view of the complexity inherent in modeling coupled land surface systems.
 
@@
 We find, in particular, that parameters controlling the degree to which ecosystems can fix additional nitrogen have a dominant effect over responses to fertilization, but limited impacts on the control state.
-Further, we find that the impact of most parameters is complex and affected by the balance of water, nutrient and other limitations to growth specific to individual sites
-
-It is common practice in land surface modeling literature to present studies that focus on the modification of single aspects of model structure, process representation or calibration, and to highlight the significance of individual modifications when scaled up to global land-mediated climate feedback processes.
-In this analysis, we illustrate the degree to which a multitude of individual parameters can potentially have important and interacting impacts on model results, how these results differ across sites and output variables and with initial vegetation conditions.
-Future studies of process and parameter modification in CLM5 might use these results as a guide to the relative importance of individual focus parameters and processes.
-We thus hope to move forwards from studies that modify individual processes in isolation to those that take a more holistic view of the complexity inherent in modeling coupled land surface systems.
+Further, we find that the impact of most parameters is complex and affected by the balance of water, nutrient and other limitations to growth specific to individual sites.
 
@@
 We argue that in order to make progress on this task, the scientific community should move beyond the paradigm of assessing single points in model parameter space in great depth, ignoring all other plausible instantiations.
-In atmospheric sciences, for example, 'internal variability' caused by initial conditions is now routinely sampled, given its major impact on even long timescale forecasts (\cite{kay2015}).
-\cite{bonan2018} illustrate how initial conditions are relatively unimportant for carbon cycle prediction on land, relative to structural and parametric uncertainty.
-Given this, sampling the important dimensions of parametric uncertainty that govern biospheric feedbacks in the climate system should be taken seriously when planning future assessments of biogeochemical feedback strength.
+In the atmospheric  sciences, for example, the 'internal variability' caused by initial conditions is now routinely sampled (\cite{kay2015}, \cite{sanderson2018}), in light of its impact on seasonal to decadal timescale predictions.
+\cite{bonan2018}, however, illustrate how initial conditions are relatively unimportant for carbon cycle prediction on land, relative to structural model uncertainty.
+Given this,  and the finding here that responses to forcing are sensitive to parametric variation, sampling the important dimensions of parametric uncertainty that govern biospheric feedbacks in the climate system should be taken seriously when planning future assessments of biogeochemical feedback strength.
 
@@
  froot\_leaf & gC/gC &  0.342 &0.671 &1.05 & 1.1\\
- stem\_leaf  & gC/gC & 0.3057 &0.61 &1 &1\\ 
- n\_costs    & gC/gN &10$^{-2}$ &10$^{-2}$&  10$^{1}$& 10$^{2}$\\
+ stem\_leaf  & gC/gC & 0.50 &0.75 &1.125 &1.25\\
+ n\_costs    & gC/gN &0.1 &0.5&  5& 10\\
  fracfixers  & - & 0 &0.5 & 2 & 4 \\
-  grperc  &  - &0 & 0.5& 2 & 3\\
  leafcn  & gC/gN &0.7413 & 0.8932 & 1.1970  & 1.349\\
-
-     medlyn\_slope  &- &0.5258 & 0.7191 & 1.1057  &1.29\\
-      lmr\_intercept &log(nmol CO$_{2}$ m$^{2}$ s$^{-1}$) &0.8539 & 0.9531 & 1.1512& 1.25\\
-      frac\_ecto &- & 0 &0.5& 1 & 1 \\
-      cn\_flex\_a & &0 &0.5 & 2  &4\\
-      cn\_flex\_b & &0.1 & 0.5 & 2  & 4\\
-      cn\_flex\_c & &0.001 & 0.1 & 10 & 100\\
+ medlyn\_slope  &- &-1 se & -0.5 se & +0.5 se  & +1 se\\
+ cn\_flex\_c &-&0.25 & 0.5 & 1.25 & 1.5\\
 \hline
@@
 \end{center}
-\caption{Parameters, units and ranges (as multipliers of the original values) in this sensitivity analysis .}
+\caption{Parameters, units and ranges (as multipliers of the original values) in this sensitivity analysis.
+For the Medlyn slope parameters, PFT-specific multipliers of the standard errors (se) from \cite{dekauwe2014} were used, where the standard errors were 0.25 for evergreen needleleaf, 0.09 for evergreen broadleaf, 0.25 for deciduous needleleaf and 0.36 for deciduous broadleaf trees.
+}
 \label{table_ranges}
@@
 
-\section{Figures}
+\section{Figure Captions}
 \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_MEANCOND_-r200_1999.png}
-     \caption{Response of drought index ($B_{tran,mn}$) and the \% of NPP spent on Nitrogen uptake to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
-Legend colours as for Figure \ref{NR1 state}}
+     \caption{Response of drought index ($B_{tran,mn}$) and the \% of NPP spent on Nitrogen uptake (PNU) to parameter perturbation across -1 to +1 range of parameter variation at the Caxiuan\~a (CAX), Niwot Ridge (NWR), Harvard Forest (HVF) and Metolious MET) sites.
+Legend colours as for Figure \ref{state}}
      \label{btran state}
  \end{figure}
-
  \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_STATE_y1999.png}
-     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) for GPP, NPP, LAI and N$+{leaf}$ at the CAX, NWR, HVF and MET sites}
+     \caption{Response of control model states and fluxes to parameter perturbation across -1 to +1 range of parameter variation (Table \ref{table_ranges}) for GPP, NPP, LAI and N$_{leaf}$ at the CAX, NWR, HVF and MET sites}
      \label{state}
  \end{figure}
-
   \begin{figure}[h]
-     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_CNtime_defpft.png}
      \caption{Fertilization responses of the default model though time since the (step-change) in CO$_{2}$ (top row) and Nitrogen deposition (bottom row) for CAX (blue), NWR (red), HVF (yellow), MET (purple).}
@@
  \end{figure}
-
  \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_GPP_y2013.png}
-     \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on gross primary productivity (GPP). }
+     \caption{Influence of parameteric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg m$^{-2}$ y$^{-1}$ N fertilization on gross primary productivity (GPP).
+Note changes in scale between plots to show detail.}
      \label{GPP_CN}
   \end{figure}
-
   \begin{figure}[h]
-     \includegraphics[width=1.35\textwidth]{matlab/figures/MAY19jp_at_relCNdep_defpft_TLAI_y2013.png}
-     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5Kg/m$^{-2}$/y N fertilization on leaf area index (LAI).
+     \caption{Influence of parametric variation on the model response (fertilized/control) to 15 years of 500ppm CO$_{2}$ and +5 Kg m$^{-2}$ y$^{-1}$ N fertilization on leaf area index (LAI).
 Note changes in scale between plots to show detail.}

### 2019-08-30 referee_responses.tex

@@
 
-\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications.**}
+\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications.}
 
@@
 
-\textsf{AR: Think about this. **}
+\textsf{AR: Thanks for the suggestion. We added the N fixation rates and the soil N content to the plots and the discussion.s}
 
@@
 
-\textsf{AR: ** }
+\textsf{AR: The defaults are complex to illustrate, as some vary by PFT, others do not, etc.
+We have now removed the duplicate range parameters, which were set as such on account of limited constraints of their ranges.
+}
 
